### 2024-09-19 paper/paper.qmd
Add files via upload

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.

The remainder of this paper is structured as follows. @sec-mydatasection
# Data {#sec-mydatasection}

Some of our data is of penguins (@fig-marriage), from @palmerpenguins.

```{r}
#| label: fig-marriage
#| fig-cap: Bills of penguins
#| echo: false

data |>
  ggplot(aes(x=date,
             y=marriage_licenses)) +
  geom_point()

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

\newpage

# References

### 2024-09-19 scripts/00-simulate_data.R

@@ -1,0 +1,22 @@
+#### Preamble ####
+# Purpose: Simulates data
+# Author: Yunkyung Ko
+# Date: 19 September 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Simulate data ####
+
+set.seed(304)
+data <- tibble(
+  date = seq.Date(from = as.Date('2020-01-01'), by = 'day', length.out = 100), 
+  number_of_marriage = rpois(n = 100, lambda = 10)
+)
+
+write_csv(data, file="data/raw_data/simulated.csv")

### 2024-09-24 paper/paper.qmd
Add files via upload

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Outbreaks in Toronto Healthcare Institutions After COVID-19"
+subtitle: "Analysis of the Causative Agents and Trend of Fluctuations"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Yunkyung Ko
+thanks: "Code and data are available at: https://github.com/koyunkyung/paper1_draft.git."
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "In Canada, outbreak incidence has been slowly increasing from June 2024 up to the beginning of the accumulating data period on August 10, 2024.
+(ADD CITATION https://health-infobase.canada.ca/covid-19/) Using the 'Outbreaks in Toronto Healthcare Institutions' dataset, we show how the trends of outbreak occurrence has been over the 5 years after COVID-19.
+(specify headline result).
+(implications)."
 format: pdf
@@
 library(tidyverse)
+library(opendatatoronto)
+library(ggplot2)
+library(dplyr)
+library(readr)
+library(lubridate)
 
 data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
 ```
@@
 
-You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.
+You can and should cross-reference sections and sub-sections. We use @citeR, @opendatatoronto and @rohan.
 
-The remainder of this paper is structured as follows. @sec-mydatasection
-# Data {#sec-mydatasection}
+The remainder of this paper is structured as follows. @sec-data....
 
-Some of our data is of penguins (@fig-marriage), from @palmerpenguins.
+# Data {#sec-data}
+
+@fig-outbreak shows the total number of outbreaks reported in Toronto healthcare institutions over the time period of 2020 to 2024.
+We can see that the numbers seem to decrease by the end of 2021, but exponentially increase from the start of 2022.
+A trend of relatively high outbreak occurrence rate continues till 2024.
 
 ```{r}
-#| label: fig-marriage
-#| fig-cap: Bills of penguins
+#| label: fig-outbreak
+#| fig-cap: Total Outbreaks Count per 6-month
 #| echo: false
@@
 data |>
-  ggplot(aes(x=date,
-             y=marriage_licenses)) +
-  geom_point()
+  ggplot(mapping = aes(x = date_of_outbreak)) +
+  geom_bar() +
+  theme_minimal()+
+  labs(x = "Date Period", y = "Number of Outbreaks Reported")
 
@@
 
-Talk more about it.
+Contrary to the expectation that the number of outbreaks would have increased the most rapidly immediately after COVID-19, the analysis with actual data shows that the number of outbreaks is on the rise as we approach 2022.
+Since there are various variables that can affect the result above, we would like to consider this one by one and proceed with an accurate analysis of the outbreak occurrence situation.
+### NEED TO MODIFY
+# check for factors such as comparing the number of outbreaks as a whole vs. only in healthcare institutions
+# would the _begin, _end sorting of timeline affect the shape of the graph above?
+# search for background information of healthcare environments in Canada that could relate to the data above
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+To dive deeper into the causations of this trend in outbreak reports, we show the frequency trend of top-3 causative agents respectively over the same time period (@fig-causations).
+```{r}
+# Process the data to get the top 3 causative agents, excluding 'NA' and 'Unable to identify'
+top_causative_agents <- cleaned_data |>
+  janitor::clean_names() |>
+
+  # Filter out rows where causative_agent is NA or 'Unable to identify'
+  filter(!is.na(causative_agent) & causative_agent != "Unable to identify") |>
+
+  # Group by causative agent and count occurrences
+  group_by(causative_agent) |>
+  summarise(total_cases = n()) |>
+
+  # Arrange in descending order of total cases
+  arrange(desc(total_cases)) |>
+
+  # Select the top 3 causative agents
+  slice_max(order_by = total_cases, n = 3)
+
+# View the top 3 causative agents
+top_causative_agents
+```
+
+```{r}
+#| label: fig-causations
+#| fig-cap: Outbreaks Count per 6-month, by Top 3 Causative Agent
+#| echo: false
+#| warning: false
+#| message: false
+
+### NEED A LINE PLOT WITH THREE LINES, EACH INDICATING THE TOP 3 CAUSATIVE AGENTS.
+
+```
+
+Talk way more about it.
 
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+analyze the total outbreak frequency trend over the time period given.
+(in relation to the social/political background in healthcare)
 
@@
 
+figure out what might be the main causative agent for the rapid increase in outbreak reports starting from 2022.
+acknowledging the causative agent that we should be aware of might help us prevent further outbreak increase from happening
+-> know what is causing this rise in outbreak reports and research about that specific disease settings
+
 ## Third discussion point
+
+from the results of the data analysis above, discuss a way we could prevent further increase of outbreak occurrence.
+in largely three dimensions: personal/ society/ government
 
@@
 
-Weaknesses and next steps should also be included.
+Dividing a year into just two could influence the analysis result. (ex. weather affecting the outbreak occurrence)
+Many 'N/A', 'Unable to identify' data in Causative Agent.
+
+For further inspection, taking the place information in consideration might also have been meaningful.
 
@@
 # Appendix {-}
+INFO OF RAW DATA, CONDITIONS OF DATA COLLECTION, ETC
 
 # Additional data details
+NEED TO MODIFY
 
@@
 # References
+NEED TO ADD REFERENCES

### 2024-09-24 scripts/00-simulate_data.R

@@ -1,0 +1,40 @@
+#### Preamble ####
+# Purpose: Simulates data
+# Author: Yunkyung Ko
+# Date: 22 September 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(lubridate)
+
+#### Simulate data ####
+
+# Set seed for reproducibility
+set.seed(123)
+
+# Define the start and end dates
+start_date <- as.Date("2020-01-01")
+end_date <- as.Date("2024-06-01")
+
+# Generate a sequence of monthly dates between start_date and end_date
+monthly_dates <- seq(from = start_date, to = end_date, by = "month")
+
+# Format dates to "YYYY-MM" without including the day
+formatted_dates <- format(monthly_dates, "%Y-%m")
+
+# Generate random Poisson distributed numbers with lambda = 10 for each month
+lambda <- 10
+poisson_values <- rpois(length(formatted_dates), lambda)
+
+# Create a data frame to hold the results
+simulated_data <- data.frame(Date = formatted_dates, Number_of_Outbreaks = poisson_values)
+
+
+
+### Write CSV
+write_csv(simulated_data, file="data/raw_data/simulated.csv")

### 2024-09-24 scripts/01-download_data.R

@@ -1,0 +1,31 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from Open Data Toronto
+# Author: Yunkyung Ko
+# Date: 22 September 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(opendatatoronto)
+
+
+#### Download data ####
+package <- show_package("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
+resources <- list_package_resources("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+resources_list <- filter(datastore_resources, row_number() %in% 1:5)
+datasets <- list()
+for (i in seq_len(nrow(resources_list))) {
+  datasets[[i]] <- get_resource(resources_list[i,])
+}
+
+#### Save data ####
+# change datasets to whatever name you assigned when you downloaded it.
+combined_data <- bind_rows(datasets)
+write_csv(combined_data, "data/raw_data/raw_data.csv")
+
+         

### 2024-09-24 scripts/02-data_cleaning.R

@@ -1,0 +1,44 @@
+#### Preamble ####
+# Purpose: Cleans the raw healthcare data into an analysis dataset
+# Author: Yunkyung Ko
+# Date: 22 September 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to have downloaded the data.
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read_csv("data/raw_data/raw_data.csv")
+
+cleaned_data <- raw_data |>
+  janitor::clean_names() |>
+  select(causative_agent_1, date_outbreak_began) |>
+  
+  # Rename the column causative_agent_1 to causative_agent
+  rename(causative_agent = causative_agent_1) |>
+  
+  mutate(
+    # Convert the date to "YYYY-MM" format
+    date_outbreak_began = format(as.Date(date_outbreak_began), "%Y-%m"),
+    
+    # Extract year and month
+    year = substr(date_outbreak_began, 1, 4),
+    month = as.numeric(substr(date_outbreak_began, 6, 7)),
+    
+    # Classify into "YYYY_first_half" or "YYYY_second_half"
+    date_of_outbreak = case_when(
+      month %in% 1:6 ~ paste0(year, "_begin"),
+      month %in% 7:12 ~ paste0(year, "_end")
+    )
+  ) |>
+  
+  # Filter out the rows with '2024_second_half'
+  filter(date_of_outbreak != "2024_end") |>
+  
+  select(causative_agent, date_of_outbreak)
+
+#### Save data ####
+write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-24 scripts/03-test_data.R

@@ -1,0 +1,21 @@
+#### Preamble ####
+# Purpose: Sanity check of the data
+# Author: Yunkyung Ko
+# Date: 22 September 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to have simulated data
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Test data ####
+data <- read_csv("data/raw_data/simulated.csv")
+
+# Test for negative numbers
+data$Number_of_Outbreaks |> min() >= 0
+
+# Test for NAs
+all(is.na(data$Number_of_Outbreaks |> min() >= 0))

### 2024-09-24 paper/paper.qmd
Update paper.qmd

@@
 title: "Outbreaks in Toronto Healthcare Institutions After COVID-19"
-subtitle: "Analysis of the Causative Agents and Trend of Fluctuations"
+subtitle: "Analysis of the Frequency Trends and the Causative Agent"
 author: 

### 2024-09-26 scripts/00-simulate_data.R
change the date range of the data

@@ -18,8 +18,9 @@ library(lubridate)
 set.seed(123)
 
 # Define the start and end dates
-start_date <- as.Date("2020-01-01")
-end_date <- as.Date("2024-06-01")
+
+start_date <- as.Date("2015-01-01")
+end_date <- as.Date("2025-12-31")
 
 # Generate a sequence of monthly dates between start_date and end_date
 monthly_dates <- seq(from = start_date, to = end_date, by = "month")
@@ -28,13 +29,11 @@ monthly_dates <- seq(from = start_date, to = end_date, by = "month")
 formatted_dates <- format(monthly_dates, "%Y-%m")
 
 # Generate random Poisson distributed numbers with lambda = 10 for each month
-lambda <- 10
+lambda <- 15
 poisson_values <- rpois(length(formatted_dates), lambda)
 
 # Create a data frame to hold the results
 simulated_data <- data.frame(Date = formatted_dates, Number_of_Outbreaks = poisson_values)
 
-
-
 ### Write CSV
-write_csv(simulated_data, file="data/raw_data/simulated.csv")
+write_csv(simulated_data, file="data/raw_data/simulated_total.csv")

### 2024-09-26 scripts/01-download_data.R

@@ -12,12 +12,13 @@
 library(tidyverse)
 library(opendatatoronto)
 
-
 #### Download data ####
 package <- show_package("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
 resources <- list_package_resources("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
 datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-resources_list <- filter(datastore_resources, row_number() %in% 1:5)
+
+resources_list <- filter(datastore_resources, row_number() %in% 1:9)
+
 datasets <- list()
 for (i in seq_len(nrow(resources_list))) {
   datasets[[i]] <- get_resource(resources_list[i,])
@@ -27,5 +28,3 @@ for (i in seq_len(nrow(resources_list))) {
 # change datasets to whatever name you assigned when you downloaded it.
 combined_data <- bind_rows(datasets)
 write_csv(combined_data, "data/raw_data/raw_data.csv")
-
-         

### 2024-09-26 scripts/02-data_cleaning.R

@@ -15,30 +15,16 @@ raw_data <- read_csv("data/raw_data/raw_data.csv")
 
 cleaned_data <- raw_data |>
   janitor::clean_names() |>
-  select(causative_agent_1, date_outbreak_began) |>
   
-  # Rename the column causative_agent_1 to causative_agent
-  rename(causative_agent = causative_agent_1) |>
+  select(outbreak_setting, type_of_outbreak, date_outbreak_began) |>
   
   mutate(
     # Convert the date to "YYYY-MM" format
-    date_outbreak_began = format(as.Date(date_outbreak_began), "%Y-%m"),
-    
+    date = format(as.Date(date_outbreak_began), "%Y-%m"),
     # Extract year and month
     year = substr(date_outbreak_began, 1, 4),
     month = as.numeric(substr(date_outbreak_began, 6, 7)),
-    
-    # Classify into "YYYY_first_half" or "YYYY_second_half"
-    date_of_outbreak = case_when(
-      month %in% 1:6 ~ paste0(year, "_begin"),
-      month %in% 7:12 ~ paste0(year, "_end")
     )
-  ) |>
-  
-  # Filter out the rows with '2024_second_half'
-  filter(date_of_outbreak != "2024_end") |>
-  
-  select(causative_agent, date_of_outbreak)
 
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")

### 2024-09-26 paper/paper.qmd
update required graphs/plots in the analysis

@@
   - Yunkyung Ko
-thanks: "Code and data are available at: https://github.com/koyunkyung/paper1_draft.git."
+thanks: "Code and data are available at: https://github.com/koyunkyung/toronto_outbreaks"
 date: today
@@
 ```{r}
+
+```
+
+```{r}
 #| label: fig-outbreak
-#| fig-cap: Total Outbreaks Count per 6-month
+#| fig-cap: Total Outbreaks Count in Toronto Healthcare Institutions
 #| echo: false
@@
 data |>
-  ggplot(mapping = aes(x = date_of_outbreak)) +
+  ggplot(mapping = aes(x = year)) +
   geom_bar() +
-  theme_minimal()+
-  labs(x = "Date Period", y = "Number of Outbreaks Reported")
+  theme_minimal() +
+  labs(x = "Date Period", y = "Number of Outbreaks Reported",
+       title = "Total Number of Outbreaks Reported in Toronto Healthcare Institutions") +
+  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
 
+```
+
+```{r}
+#| label: fig-setting
+#| fig-cap: Outbreaks Count by Outbreak Settings
+#| echo: false
+
+library(dplyr)
+library(ggplot2)
+
+outbreak_counts <- data %>%
+  count(outbreak_setting)
+
+ggplot(outbreak_counts, aes(x = "", y = n, fill = outbreak_setting)) +
+  geom_bar(stat = "identity", width = 1) +
+  coord_polar("y", start = 0) + 
+  theme_minimal() +
+  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
+                               "Retirement Home" = "#B3FFC6", 
+                               "Hospital-Chronic Care" = "#BA55D3", 
+                               "Hospital-Acute Care" = "#DDA0DD",
+                               "Hospital-Psychiatric" = "grey",
+                               "Shelter" = "brown",
+                               "Transitional Care" = "black")) +
+  labs(x = NULL, y = NULL, fill = "Outbreak Setting", 
+       title = "Number of Outbreaks Reported by Setting") +
+  theme(axis.text.x = element_blank(),  
+        axis.ticks = element_blank(),  
+        panel.grid = element_blank())  
 ```
@@
 
+# Result
+
 To dive deeper into the causations of this trend in outbreak reports, we show the frequency trend of top-3 causative agents respectively over the same time period (@fig-causations).
 ```{r}
-# Process the data to get the top 3 causative agents, excluding 'NA' and 'Unable to identify'
-top_causative_agents <- cleaned_data |>
-  janitor::clean_names() |>
 
-  # Filter out rows where causative_agent is NA or 'Unable to identify'
-  filter(!is.na(causative_agent) & causative_agent != "Unable to identify") |>
-
-  # Group by causative agent and count occurrences
-  group_by(causative_agent) |>
-  summarise(total_cases = n()) |>
-
-  # Arrange in descending order of total cases
-  arrange(desc(total_cases)) |>
-
-  # Select the top 3 causative agents
-  slice_max(order_by = total_cases, n = 3)
-
-# View the top 3 causative agents
-top_causative_agents
 ```
@@
 ```{r}
-#| label: fig-causations
-#| fig-cap: Outbreaks Count per 6-month, by Top 3 Causative Agent
+#| label: fig-outbreaks count by types
+#| fig-cap: Outbreaks Count in Toronto Healthcare Institutions, by Types of Outbreak
 #| echo: false
@@
 
-### NEED A LINE PLOT WITH THREE LINES, EACH INDICATING THE TOP 3 CAUSATIVE AGENTS.
+# Summarizing the data to get the number of outbreaks per month and type of outbreak
+annual_data <- data %>%
+  group_by(year, type_of_outbreak) %>%
+  summarise(outbreak_count = n()) %>%
+  ungroup()
+
+# Convert the year column to a factor with levels in chronological order
+annual_data$year <- factor(annual_data$year, levels = unique(annual_data$year))
+
+# Create the line plot
+ggplot(annual_data, aes(x = year, y = outbreak_count, group = type_of_outbreak, color = type_of_outbreak)) +
+  geom_line(size = 1) +
+  geom_point(size = 2) +
+  scale_color_manual(values = c("Respiratory" = "red", "Enteric" = "blue", "Other" = "black")) +
+  labs(title = "Trend of Outbreak Counts by Type (2016-2024)",
+       x = "Year", y = "Outbreak Count", color = "Type of Outbreak") +
+  theme_minimal() +
+  scale_x_discrete(drop = FALSE) # Ensure all years are shown on the x-axis
 
@@
 
+```{r}
+#| label: fig-respiratory outbreaks count by outbreak settings
+#| fig-cap: Respiratory Outbreaks Count, by Outbreak Settings
+#| echo: false
+#| warning: false
+#| message: false
+
+library(dplyr)
+library(ggplot2)
+
+# Filter for 'Respiratory' outbreaks
+respiratory_data <- data %>%
+  filter(type_of_outbreak == "Respiratory")
+
+# Aggregate data by year and outbreak setting
+annual_counts <- respiratory_data %>%
+  group_by(year, outbreak_setting) %>%
+  summarise(outbreak_count = n(), .groups = 'drop')
+
+# Create a stacked bar plot
+ggplot(annual_counts, aes(x = year, y = outbreak_count, fill = outbreak_setting)) +
+  geom_col(position = "stack") +  # Stack bars based on outbreak settings
+  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
+                               "Retirement Home" = "#B3FFC6", 
+                               "Hospital-Chronic Care" = "#BA55D3", 
+                               "Hospital-Acute Care" = "#DDA0DD",
+                               "Hospital-Psychiatric" = "grey",
+                               "Shelter" = "brown",
+                               "Transitional Care" = "black")) +
+  labs(title = "Annual Respiratory Outbreak Counts by Setting",
+       x = "Year",
+       y = "Outbreak Count",
+       fill = "Outbreak Setting") +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
+```
+
+```{r}
+#| label: fig-enteric outbreaks count by outbreak settings
+#| fig-cap: Enteric Outbreaks Count, by Outbreak Settings
+#| echo: false
+#| warning: false
+#| message: false
+
+library(dplyr)
+library(ggplot2)
+
+# Filter for 'Enteric' outbreaks
+enteric_data <- data %>%
+  filter(type_of_outbreak == "Enteric")
+
+# Aggregate data by year and outbreak setting
+annual_counts <- enteric_data %>%
+  group_by(year, outbreak_setting) %>%
+  summarise(outbreak_count = n(), .groups = 'drop')
+
+# Create a stacked bar plot
+ggplot(annual_counts, aes(x = year, y = outbreak_count, fill = outbreak_setting)) +
+  geom_col(position = "stack") +  # Stack bars based on outbreak settings
+  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
+                               "Retirement Home" = "#B3FFC6", 
+                               "Hospital-Chronic Care" = "#BA55D3", 
+                               "Hospital-Acute Care" = "#DDA0DD",
+                               "Hospital-Psychiatric" = "grey",
+                               "Shelter" = "brown",
+                               "Transitional Care" = "black")) +
+  labs(title = "Annual Enteric Outbreak Counts by Setting",
+       x = "Year",
+       y = "Outbreak Count",
+       fill = "Outbreak Setting") +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
+```
 Talk way more about it.

### 2024-09-27 scripts/00-simulate_data.R
update scripts for generating simulation data

@@ -12,13 +12,14 @@
 library(tidyverse)
 library(lubridate)
 
+
 #### Simulate data ####
 
+## Total Outbreak Counts Reported in Toronto Healthcare Institutions
 # Set seed for reproducibility
 set.seed(123)
 
 # Define the start and end dates
-
 start_date <- as.Date("2015-01-01")
 end_date <- as.Date("2025-12-31")
 
@@ -33,7 +34,89 @@ lambda <- 15
 poisson_values <- rpois(length(formatted_dates), lambda)
 
 # Create a data frame to hold the results
-simulated_data <- data.frame(Date = formatted_dates, Number_of_Outbreaks = poisson_values)
+sim_date_total <- data.frame(Date = formatted_dates, Number_of_Outbreaks = poisson_values)
 
 ### Write CSV
-write_csv(simulated_data, file="data/raw_data/simulated_total.csv")
+write_csv(sim_date_total, file="data/raw_data/simulated_data/simulated_date_total.csv")
+
+
+
+
+
+## Number of Outbreaks Reported by Setting
+# Set seed for reproducibility
+set.seed(123)
+
+# Simulated data with random numbers
+sim_setting_total <- data.frame(
+  Setting = c("Hospital-Acute Care", "Hospital-Chronic Care", "Hospital-Psychiatric", 
+              "LTCH", "Retirement Home", "Shelter", "Transitional Care"),
+  Count = round(runif(7, min = 1, max = 1000)) # Generate random numbers between 1 and 1000
+)
+### Write CSV
+write_csv(sim_setting_total, file="data/raw_data/simulated_data/simulated_setting_total.csv")
+
+
+
+
+
+## Trend of Outbreak Counts by Type
+# Set seed for reproducibility
+set.seed(123)
+
+## Simulated data for outbreak counts by type and year
+years <- 2015:2025
+
+# Generate random outbreak counts for each type of outbreak
+respiratory <- round(runif(length(years), min = 200, max = 1000)) # Respiratory outbreaks
+enteric <- round(runif(length(years), min = 0, max = 100))        # Enteric outbreaks
+other <- round(runif(length(years), min = 0, max = 50))           # Other outbreaks
+
+# Combine data into a dataframe
+sim_date_type <- data.frame(
+  Year = rep(years, 3),
+  Count = c(respiratory, enteric, other),
+  Type = factor(rep(c("Respiratory", "Enteric", "Other"), each = length(years)))
+)
+### Write CSV
[…]

### 2024-09-27 scripts/03-test_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Sanity check of the data
+# Purpose: Sanity check of the data, Hypotheis verification of how the simulated data would unfold
 # Author: Yunkyung Ko
 # Date: 22 September 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
@@ -11,11 +11,118 @@
 #### Workspace setup ####
 library(tidyverse)
 
+
 #### Test data ####
-data <- read_csv("data/raw_data/simulated.csv")
-
+date_total <- read_csv("data/raw_data/simulated_data/simulated_date_total.csv")
 # Test for negative numbers
-data$Number_of_Outbreaks |> min() >= 0
-
+date_total$Number_of_Outbreaks |> min() >= 0
 # Test for NAs
-all(is.na(data$Number_of_Outbreaks |> min() >= 0))
+all(is.na(date_total$Number_of_Outbreaks |> min() >= 0))
+
+# Create the bar plot
+p <- ggplot(date_total, aes(x = factor(Date), y = Number_of_Outbreaks)) +
+  geom_bar(stat = "identity", fill = "black") +
+  labs(title = "Total Outbreak Counts (Simulation)",
+       x = "Date",
+       y = "Number of Outbreaks Reported") +
+  theme_minimal()
+
+# Save the plot as a PNG file
+ggsave("data/raw_data/simulated_results/simulated_total_outbreak_count.png", plot = p, width = 10, height = 8, dpi = 300)
+
+
+
+
+#### Test data ####
+setting_total <- read_csv("data/raw_data/simulated_data/simulated_setting_total.csv")
+# Test for negative numbers
+setting_total$Count |> min() >= 0
+# Test for NAs
+all(is.na(setting_total$Count |> min() >= 0))
+
+# Create a pie chart
+p <- ggplot(setting_total, aes(x = "", y = Count, fill = Setting)) +
+  geom_bar(stat = "identity", width = 1) +
+  coord_polar("y") +
+  labs(title = "Outbreak Counts by Setting (Simulation)", fill = "Outbreak Setting") +
+  theme_void() +
+  scale_fill_manual(values = c("#d3a9d3", "#a455a4", "#b3b3b3", "#a2d6f9", "#b6f5d3", "#963232", "#000000"))
+
+# Save the plot as a PNG file
[…]

### 2024-09-27 scripts/00-simulate_data.R
code linting and styling

@@ -37,7 +37,7 @@ poisson_values <- rpois(length(formatted_dates), lambda)
 sim_date_total <- data.frame(Date = formatted_dates, Number_of_Outbreaks = poisson_values)
 
 ### Write CSV
-write_csv(sim_date_total, file="data/raw_data/simulated_data/simulated_date_total.csv")
+write_csv(sim_date_total, file = "data/raw_data/simulated_data/simulated_date_total.csv")
 
 
 
@@ -49,12 +49,11 @@ set.seed(123)
 
 # Simulated data with random numbers
 sim_setting_total <- data.frame(
-  Setting = c("Hospital-Acute Care", "Hospital-Chronic Care", "Hospital-Psychiatric", 
-              "LTCH", "Retirement Home", "Shelter", "Transitional Care"),
+  Setting = c("Hospital-Acute Care", "Hospital-Chronic Care", "Hospital-Psychiatric", "LTCH", "Retirement Home", "Shelter", "Transitional Care"),
   Count = round(runif(7, min = 1, max = 1000)) # Generate random numbers between 1 and 1000
 )
 ### Write CSV
-write_csv(sim_setting_total, file="data/raw_data/simulated_data/simulated_setting_total.csv")
+write_csv(sim_setting_total, file = "data/raw_data/simulated_data/simulated_setting_total.csv")
 
 
 
@@ -69,8 +68,8 @@ years <- 2015:2025
 
 # Generate random outbreak counts for each type of outbreak
 respiratory <- round(runif(length(years), min = 200, max = 1000)) # Respiratory outbreaks
-enteric <- round(runif(length(years), min = 0, max = 100))        # Enteric outbreaks
-other <- round(runif(length(years), min = 0, max = 50))           # Other outbreaks
+enteric <- round(runif(length(years), min = 0, max = 100)) # Enteric outbreaks
+other <- round(runif(length(years), min = 0, max = 50)) # Other outbreaks
 
 # Combine data into a dataframe
 sim_date_type <- data.frame(
@@ -79,7 +78,7 @@ sim_date_type <- data.frame(
   Type = factor(rep(c("Respiratory", "Enteric", "Other"), each = length(years)))
 )
 ### Write CSV
-write_csv(sim_date_type, file="data/raw_data/simulated_data/simulated_date_type.csv")
+write_csv(sim_date_type, file = "data/raw_data/simulated_data/simulated_date_type.csv")
 
 
 
@@ -90,15 +89,14 @@ set.seed(123)
 
[…]

### 2024-09-27 scripts/01-download_data.R

@@ -15,16 +15,22 @@ library(opendatatoronto)
 #### Download data ####
 package <- show_package("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
 resources <- list_package_resources("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+datastore_resources <- filter(resources, tolower(format) %in% c("csv", "geojson"))
 
 resources_list <- filter(datastore_resources, row_number() %in% 1:9)
 
 datasets <- list()
 for (i in seq_len(nrow(resources_list))) {
-  datasets[[i]] <- get_resource(resources_list[i,])
+  datasets[[i]] <- get_resource(resources_list[i, ])
 }
 
 #### Save data ####
 # change datasets to whatever name you assigned when you downloaded it.
 combined_data <- bind_rows(datasets)
 write_csv(combined_data, "data/raw_data/raw_data.csv")
+
+# Code linting and styling
+library(lintr)
+library(styler)
+style_file(path = "scripts/01-download_data.R")
+lint(filename = "scripts/01-download_data.R")

### 2024-09-27 scripts/02-data_cleaning.R

@@ -15,16 +15,20 @@ raw_data <- read_csv("data/raw_data/raw_data.csv")
 
 cleaned_data <- raw_data |>
   janitor::clean_names() |>
-  
   select(outbreak_setting, type_of_outbreak, date_outbreak_began) |>
-  
   mutate(
     # Convert the date to "YYYY-MM" format
     date = format(as.Date(date_outbreak_began), "%Y-%m"),
     # Extract year and month
     year = substr(date_outbreak_began, 1, 4),
     month = as.numeric(substr(date_outbreak_began, 6, 7)),
-    )
+  )
 
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")
+
+# Code linting and styling
+library(lintr)
+library(styler)
+style_file(path = "scripts/02-data_cleaning.R")
+lint(filename = "scripts/02-data_cleaning.R")

### 2024-09-27 scripts/03-test_data.R

@@ -22,9 +22,11 @@ all(is.na(date_total$Number_of_Outbreaks |> min() >= 0))
 # Create the bar plot
 p <- ggplot(date_total, aes(x = factor(Date), y = Number_of_Outbreaks)) +
   geom_bar(stat = "identity", fill = "black") +
-  labs(title = "Total Outbreak Counts (Simulation)",
-       x = "Date",
-       y = "Number of Outbreaks Reported") +
+  labs(
+    title = "Total Outbreak Counts (Simulation)",
+    x = "Date",
+    y = "Number of Outbreaks Reported"
+  ) +
   theme_minimal()
 
 # Save the plot as a PNG file
@@ -67,10 +69,12 @@ all(is.na(date_type$Count |> min() >= 0))
 p <- ggplot(date_type, aes(x = Year, y = Count, color = Type, group = Type)) +
   geom_line(size = 1) +
   geom_point(size = 2) +
-  labs(title = "Outbreak Counts by Type (Simulation)",
-       x = "Year",
-       y = "Outbreak Count",
-       color = "Type of Outbreak") +
+  labs(
+    title = "Outbreak Counts by Type (Simulation)",
+    x = "Year",
+    y = "Outbreak Count",
+    color = "Type of Outbreak"
+  ) +
   scale_color_manual(values = c("Respiratory" = "red", "Enteric" = "blue", "Other" = "black")) +
   theme_minimal()
 
@@ -91,10 +95,12 @@ all(is.na(date_repository$Count |> min() >= 0))
 # Create the stacked bar chart
 p <- ggplot(date_repository, aes(x = Year, y = Count, fill = Setting)) +
   geom_bar(stat = "identity") +
-  labs(title = "Respiratory Outbreak Counts by Setting (Simulation)",
-       x = "Year",
-       y = "Outbreak Count",
-       fill = "Outbreak Setting") +
+  labs(
+    title = "Respiratory Outbreak Counts by Setting (Simulation)",
+    x = "Year",
+    y = "Outbreak Count",
+    fill = "Outbreak Setting"
+  ) +
   scale_fill_manual(values = c("#d3a9d3", "#a455a4", "#b3b3b3", "#a2d6f9", "#b6f5d3", "#963232", "#000000")) +
   theme_minimal()
 
@@ -113,16 +119,22 @@ date_enteric$Count |> min() >= 0
 all(is.na(date_enteric$Count |> min() >= 0))
 
[…]

### 2024-09-27 scripts/02-data_cleaning.R
remove unnecessary columns

@@ -17,11 +17,8 @@ cleaned_data <- raw_data |>
   janitor::clean_names() |>
   select(outbreak_setting, type_of_outbreak, date_outbreak_began) |>
   mutate(
-    # Convert the date to "YYYY-MM" format
-    date = format(as.Date(date_outbreak_began), "%Y-%m"),
-    # Extract year and month
+    # Extract year
     year = substr(date_outbreak_began, 1, 4),
-    month = as.numeric(substr(date_outbreak_began, 6, 7)),
   )
 
 #### Save data ####

### 2024-09-27 paper/paper.qmd
analysis data with unnecessary columns removed

@@
 ---
-title: "Outbreaks in Toronto Healthcare Institutions After COVID-19"
-subtitle: "Analysis of the Frequency Trends and the Causative Agent"
+title: "Disease Outbreaks in Toronto Healthcare Institutions"
+subtitle: "Analysis of Outbreak Patterns Based on Its Type and Settings"
 author: 
@@
 date-format: long
-abstract: "In Canada, outbreak incidence has been slowly increasing from June 2024 up to the beginning of the accumulating data period on August 10, 2024.
-(ADD CITATION https://health-infobase.canada.ca/covid-19/) Using the 'Outbreaks in Toronto Healthcare Institutions' dataset, we show how the trends of outbreak occurrence has been over the 5 years after COVID-19.
-(specify headline result).
-(implications)."
+abstract: "(specify the general area of the paper and encourage the reader) (specify the dataset and methods at a general level) (specify headline result). (implications)."
 format: pdf
@@
 bibliography: references.bib
+toc: true
+toc-title: "Table of Contents"
+toc-depth: 2
+toc-location: left
 ---
@@
 
+raw_data <- read_csv(here::here("data/raw_data/raw_data.csv"))
 data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+
 ```
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR, @opendatatoronto and @rohan.
+As of September 9 2024, 1,016 infections of pertussis, a contagious respiratory disease also known as whopping cough, were reported in Ontario [@nowtoronto_whooping_cough].
+In all of Toronto, the figure is more than triple the five-year, pre-pandemic average of 31 cases for the same period [@publichealth_ontario].
 
-The remainder of this paper is structured as follows. @sec-data....
+Unexpected disease outbreaks significantly disrupt society by inducing widespread fear, anxiety, and economic losses [@frontiers_outbreaks].
+Especially in Canada, where people continue to face some of the longest medical wait times for treatment in the developed world [@torontosun_wait_times], costs are likely to increase exponentially in an instant.
+Before Toronto overspends the cost of medical care, this paper is intending to analyze the trends of outbreaks reported in Toronto healthcare institutions.
 
-# Data {#sec-data}
+(high-level results, focuesd on the one key result that is the main part of the story)
+(broadly discuss next steps in a sentence or two)
 
-@fig-outbreak shows the total number of outbreaks reported in Toronto healthcare institutions over the time period of 2020 to 2024.
-We can see that the numbers seem to decrease by the end of 2021, but exponentially increase from the start of 2022.
-A trend of relatively high outbreak occurrence rate continues till 2024.
+The paper starts with analyzing the overall pattern of outbreak occurrence frequency over the time period of 2016 to 2024.
+Then, it sorts the outbreaks into two different types, respiratory and enteric, to reiterate the analysis of outbreak frequency patterns.
+Diving deeper into the analysis, outbreak settings would also come into consideration.
+Finally, discussions about how specific healthcare institutions should reallocate their resources for particular outbreak types would proceed.
 
-```{r}
+# Data
 
-```
+## Data Overview
+
+(after clearly describing the dataset, use tables to display relevant cross-tabs and graphs to contrast groups)
+(thoroughly discuss the variables in the dataset we are using. graph the analysis data and table for summary stats)
+
+The 'Outbreaks in Toronto Healthcare Institutions' dataset [@toronto_outbreaks_dataset], obtained from City of Toronto Open Data Portal, was used for the trend analysis of outbreak occurrence.
+Published by Toronto Public Health under the requirement to monitor symptoms of infections, this dataset contains reports of suspected and confirmed outbreaks of gastroenteric (e.g., nausea, vomiting, diarrhea, fever) and respiratory (e.g., cough, runny nose, sore throat, fever) illness.
+An outbreak is defined as a localized increase in the rate of infection or illness, above that which is expected.
+Please note that this dataset was last updated on September 24, 2024, and contains only the information up to then.
+
+In this paper, R for statistical computing [@R] is used to handle data.
+In particular, packages like `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `lubridate` [@lubridate] were utilized to analyze data from the City of Toronto [@opendatatoronto].
+Data from year 2016 to 2024 were combined, and several variables were removed in the filtering process for analysis data.
+
+## Results
 
@@
   labs(x = "Date Period", y = "Number of Outbreaks Reported",
-       title = "Total Number of Outbreaks Reported in Toronto Healthcare Institutions") +
+       title = "Total Outbreak Counts Reported in Toronto Healthcare Institutions") +
   scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
@@
 
-Contrary to the expectation that the number of outbreaks would have increased the most rapidly immediately after COVID-19, the analysis with actual data shows that the number of outbreaks is on the rise as we approach 2022.
-Since there are various variables that can affect the result above, we would like to consider this one by one and proceed with an accurate analysis of the outbreak occurrence situation.
-### NEED TO MODIFY
-# check for factors such as comparing the number of outbreaks as a whole vs. only in healthcare institutions
-# would the _begin, _end sorting of timeline affect the shape of the graph above?
-# search for background information of healthcare environments in Canada that could relate to the data above
-
-# Result
-
-To dive deeper into the causations of this trend in outbreak reports, we show the frequency trend of top-3 causative agents respectively over the same time period (@fig-causations).
 ```{r}
@@
 
-## First discussion point {#sec-first-point}
+(brief summary of what was done in the paper)
+
+## First discussion point
 
@@
 # Appendix {-}
-INFO OF RAW DATA, CONDITIONS OF DATA COLLECTION, ETC
 
 # Additional data details
-NEED TO MODIFY
+
+## Other Similar Datasets
+
+## Raw Data Variables Overview
 
@@
 # References
-NEED TO ADD REFERENCES

### 2024-09-27 paper/paper.qmd
paper revision

@@
 date-format: long
-abstract: "(specify the general area of the paper and encourage the reader) (specify the dataset and methods at a general level) (specify headline result). (implications)."
+abstract: "Toronto is currently at the crisis of being swallowed up by the cost of dealing with pertussis, an infectious respiratory disease.
+This paper analyzes disease outbreak data in Toronto healthcare institutions to apprehend its trend and come up with countermeasures.
+The analysis indicates that respiratory illnesses and long term care home institutions need intensive management strategies.
+Furthermore, the paper discusses necessary action plans for both the policymakers and healthcare authorities of Toronto."
 format: pdf
@@
 library(lubridate)
+library(modelsummary)
+library(knitr)
+library(kableExtra)
 
@@
 
-(after clearly describing the dataset, use tables to display relevant cross-tabs and graphs to contrast groups)
-(thoroughly discuss the variables in the dataset we are using. graph the analysis data and table for summary stats)
-
-The 'Outbreaks in Toronto Healthcare Institutions' dataset [@toronto_outbreaks_dataset], obtained from City of Toronto Open Data Portal, was used for the trend analysis of outbreak occurrence.
+'Outbreaks in Toronto Healthcare Institutions' dataset [@toronto_outbreaks_dataset], obtained from City of Toronto Open Data Portal, was used for the trend analysis of outbreak occurrence.
 Published by Toronto Public Health under the requirement to monitor symptoms of infections, this dataset contains reports of suspected and confirmed outbreaks of gastroenteric (e.g., nausea, vomiting, diarrhea, fever) and respiratory (e.g., cough, runny nose, sore throat, fever) illness.
@@
 
-In this paper, R for statistical computing [@R] is used to handle data.
+In this paper, R for statistical computing [@R] was used to handle data.
 In particular, packages like `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `lubridate` [@lubridate] were utilized to analyze data from the City of Toronto [@opendatatoronto].
-Data from year 2016 to 2024 were combined, and several variables were removed in the filtering process for analysis data.
+Data from year 2016 to 2024 were combined, resulting in 5253 observations in the dataset.
+The selected variables for analysis are the following: 'outbreak_setting', 'type_of_outbreak', 'date_outbreak_began'.
+The main criteria considered was whether the information could be helpful for Toronto healthcare institutions to devise preemptive measures for outbreak occurrences.
+More specifically, the reason for selecting 'date_outbreak_began' is because it is considered more important to accurately determine the onset time of disease to immediately prevent the surge of infected population.
+In order to use long-term data, which is more effective in capturing the complexity and stochastic nature of disease outbreaks [@bmj_outbreak_prediction], the 'year' information was extracted as the variable for analysis.
+@tbl-part_of_dataset shows an extract of the dataset and @tbl-summary_stats communicates the summary statistics of the analysis data.
+For more detailed information about the dataset and each variable, refer to Appendix \ref{sec:appendix}.
+
+```{r}
+#| label: tbl-part_of_dataset
+#| tbl-cap: A dataset of outbreak occurrence in Toronto healthcare institutions
+#| echo: false
+#| tbl.pos: 'H'
+
+data |>
+  slice(1:10) |>
+  kable(
+    col.names = c("Outbreak Setting", "Type of Outbreak", "Date Outbreak Began", "Year"),
+    align = 'c'
+  ) |>
+  kable_styling(latex_options = "H")
+
+```
+
+```{r}
+#| label: tbl-summary_stats
+#| tbl-cap: Descriptive statistics for toronto healthcare dataset
+#| echo: false
+
+datasummary_balance(
+  formula = ~type_of_outbreak,
+  data = data |>
+    filter(type_of_outbreak %in% c("Respiratory", "Enteric", "Other")),
+  dinm = FALSE,
+  notes = "Data Source: Toronto Public Health"
+) |>
+  kable_styling(latex_options = c("scale_down", "H"))
+```
+
+\newpage
 
 ## Results
+
+What is of our primary interest is the occurrence patterns of outbreaks in Toronto.
+First, we are going to analyze it from a timely perspective.
+Through @fig-outbreak, which shows the number of outbreak reports by year, we can see the distribution of occurrence frequency over time.
+The number of outbreaks has increased sharply as of 2022, and since then, the number has been relatively high until recently.
 
@@
 #| echo: false
+#| fig.pos: 'H'
 
@@
   theme_minimal() +
-  labs(x = "Date Period", y = "Number of Outbreaks Reported",
-       title = "Total Outbreak Counts Reported in Toronto Healthcare Institutions") +
+  labs(x = "Date", y = "Number of Outbreaks",
+       title = "Total Outbreaks Count Per Year") +
   scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
@@
 
-```{r}
-#| label: fig-setting
-#| fig-cap: Outbreaks Count by Outbreak Settings
-#| echo: false
-
-library(dplyr)
-library(ggplot2)
-
-outbreak_counts <- data %>%
-  count(outbreak_setting)
-
-ggplot(outbreak_counts, aes(x = "", y = n, fill = outbreak_setting)) +
-  geom_bar(stat = "identity", width = 1) +
-  coord_polar("y", start = 0) + 
-  theme_minimal() +
-  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
-                               "Retirement Home" = "#B3FFC6", 
-                               "Hospital-Chronic Care" = "#BA55D3", 
-                               "Hospital-Acute Care" = "#DDA0DD",
-                               "Hospital-Psychiatric" = "grey",
-                               "Shelter" = "brown",
-                               "Transitional Care" = "black")) +
-  labs(x = NULL, y = NULL, fill = "Outbreak Setting", 
-       title = "Number of Outbreaks Reported by Setting") +
-  theme(axis.text.x = element_blank(),  
-        axis.ticks = element_blank(),  
-        panel.grid = element_blank())  
-```
+For detailed analysis, the outbreak was largely divided into respiratory and enteric as separated by Toronto public health, and the trend of occurrence over time was re-analyzed.
+@fig-outbreaks_count_by_types shows that the number of respiratory outbreak cases is always relatively high, and the increase rate is also greater.
+In particular, a significant rise in the number of respiratory outbreaks is seen in year 2022, which is expected to have contributed substantially to the surge in the total number of outbreak cases.
+Furthermore, it can be seen that the large increase in the respiratory outbreak does not show an easily decreasing trend.
 
 ```{r}
-
-```
-
-```{r}
-#| label: fig-outbreaks count by types
+#| label: fig-outbreaks_count_by_types
 #| fig-cap: Outbreaks Count in Toronto Healthcare Institutions, by Types of Outbreak
@@
 #| message: false
+#| fig.pos: 'H'
 
@@
   scale_color_manual(values = c("Respiratory" = "red", "Enteric" = "blue", "Other" = "black")) +
-  labs(title = "Trend of Outbreak Counts by Type (2016-2024)",
-       x = "Year", y = "Outbreak Count", color = "Type of Outbreak") +
+  labs(title = "Outbreaks Count Per Year, by Type",
+       x = "Date", y = "Number of Outbreaks", color = "Type of Outbreak") +
   theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate the x-axis labels
   scale_x_discrete(drop = FALSE) # Ensure all years are shown on the x-axis
@@
 
+Next, we are going to look at the outbreak occurrence trends from a spatial perspective.
+@fig-setting shows the occurrence locations of all outbreak reports from 2016 to September 2024 in relative percentages.
+What accounted for the highest percentage with a noticeable difference compared to the others was LCTH, which stands for Long-Term Care Home, a residential facility providing 24-hour care and support for individuals with complex medical needs, typically the elderly or those with chronic conditions (@ontario_ltc).
+
 ```{r}
-#| label: fig-respiratory outbreaks count by outbreak settings
+#| label: fig-setting
+#| fig-cap: Outbreaks Count by Outbreak Settings
+#| echo: false
+#| fig.pos: 'H'
+
+outbreak_counts <- data %>%
+  count(outbreak_setting) %>%
+  mutate(percentage = n / sum(n) * 100)
+
+ggplot(outbreak_counts, aes(x = "", y = n, fill = outbreak_setting)) +
+  geom_bar(stat = "identity", width = 1) +
+  coord_polar("y", start = 0) + 
+  theme_minimal() +
+  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
+                               "Retirement Home" = "#B3FFC6", 
+                               "Hospital-Chronic Care" = "#BA55D3", 
+                               "Hospital-Acute Care" = "#DDA0DD",
+                               "Hospital-Psychiatric" = "grey",
+                               "Shelter" = "brown",
+                               "Transitional Care" = "black")) +
+  geom_text(aes(label = paste0(round(percentage, 1), "%")),
+            position = position_stack(vjust = 0.5),
+            size = 3) +  # Add percentage labels
+  labs(x = NULL, y = NULL, fill = "Outbreak Setting", 
+       title = "Outbreaks Count by Setting") +
+  theme(axis.text.x = element_blank(),  
+        axis.ticks = element_blank(),  
+        panel.grid = element_blank())  
+```
+
+Finally, to determine whether there is a relationship between the outbreak occurrence pattern over time and the place of occurrence, the data were re-expressed in the following graph.
+
+```{r}
+#| label: fig-respiratory_outbreaks_count_by_settings
 #| fig-cap: Respiratory Outbreaks Count, by Outbreak Settings
@@
 #| message: false
-
-library(dplyr)
-library(ggplot2)
+#| fig.pos: 'H'
 
@@
   group_by(year, outbreak_setting) %>%
-  summarise(outbreak_count = n(), .groups = 'drop')
+  summarise(outbreak_count = n(), .groups = 'drop') %>%
+  group_by(year) %>%
+  mutate(proportion = outbreak_count / sum(outbreak_count) * 100)  # Normalize by year
 
@@
   geom_col(position = "stack") +  # Stack bars based on outbreak settings
+  geom_text(aes(label = ifelse(proportion > 5, paste0(round(proportion, 1), "%"), "")), 
+            position = position_stack(vjust = 0.5), 
+            size = 2) +  # Add percentage labels with smaller font size
   scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
@@
                                "Transitional Care" = "black")) +
-  labs(title = "Annual Respiratory Outbreak Counts by Setting",
-       x = "Year",
-       y = "Outbreak Count",
+  labs(title = "Annual Respiratory Outbreaks Count by Setting",
+       x = "Date",
+       y = "Number of Outbreaks",
        fill = "Outbreak Setting") +
   theme_minimal() +
-  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
+  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1)) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate the x-axis labels
 ```
 
+@fig-respiratory_outbreaks_count_by_settings shows that the rate of respiratory outbreak occurrence according to its setting remains relatively similar.
+Something that stands out even if it's a little bit is that with the surge in respiratory outbreak in 2022, the outbreak rate of LTCH did not increase proportionally.
+Rather the outbreak rate in other places, such as hospitals in the chronic care sector and retirement home, showed an increasing trend.
+In addition, it can be predicted that in 2024, the rate of LTCH respiratory outbreak occurrence will be significantly higher than the previous three years.
+
 ```{r}
-#| label: fig-enteric outbreaks count by outbreak settings
+#| label: fig-enteric_outbreaks_count_by_settings
 #| fig-cap: Enteric Outbreaks Count, by Outbreak Settings
@@
                                "Transitional Care" = "black")) +
-  labs(title = "Annual Enteric Outbreak Counts by Setting",
-       x = "Year",
-       y = "Outbreak Count",
+  labs(title = "Annual Enteric Outbreaks Count by Setting",
+       x = "Date",
+       y = "Number of Outbreaks",
        fill = "Outbreak Setting") +
   theme_minimal() +
-  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
+  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1)) + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate the x-axis labels
+
 ```
-Talk way more about it.
+
+On the other hand, @fig-enteric_outbreaks_count_by_settings shows that regarding enteric outbreaks, occurrence rate in hospitals rather decreased in recent years.
+In addition, in 2021 and 2022, when the number of enteric outbreak cases was markedly low, it can be seen that there were outbreak reports only in LTCH and retirement home.
 
@@
 
-(brief summary of what was done in the paper)
+Under the purpose of predicting the possibility of a disease outbreak more accurately and specifically, an outbreak pattern analysis was conducted based on the reports of Toronto healthcare institutions.
+The following items will discuss how the above analysis results can be applied and practically used in the real world.
 
-## First discussion point
+## Policy Measures to be Established by Toronto to Prevent and Cope with Disease Outbreaks
 
-analyze the total outbreak frequency trend over the time period given.
-(in relation to the social/political background in healthcare)
+The timely outbreak trend analysis in Toronto healthcare institutions (@fig-outbreak) implies that the number of outbreaks has surged as of the time of certain incidents in 2022.
+The factors that might have substantially influenced this kind of pattern are the emergence of new COVID-19 variants and relaxed public health measures [@cihi_impact_covid19_2022].
+At this point of time period, Toronto eased certain pandemic-related restrictions, such as the removal of mandatory masking for visitors and caregivers in long-term care homes.
+This is in line with the results shown in @fig-outbreaks_count_by_types in that it explains the increase in the incidence of respiratory diseases.
 
-## Second discussion point
+In the current situation, we are dealing with the crisis of pertussis infections, which is a typical respiratory disease.
+As we can see from @fig-respiratory_outbreaks_count_by_settings, a surge in respiratory outbreak occurrence affects every settings of public healthcare institutions.
+Accordingly, the first step to address this might be to strengthen vaccination campaigns.
+Pertussis outbreaks are preventable through proper vaccination coverage, and expanding booster doses to not only children but also adults and healthcare workers could substantially reduce the spread of disease [@canada_pertussis_2024].
+Also, since lack of PPE in Canada's public healthcare settings has been a key issue in exacerbating the spread of infections [@cfp_ltc_2020], the government should ensure adequate healthcare capacity by increasing funding and staffing, especially in LTCHs and hospitals.
 
-figure out what might be the main causative agent for the rapid increase in outbreak reports starting from 2022.
-acknowledging the causative agent that we should be aware of might help us prevent further outbreak increase from happening
--> know what is causing this rise in outbreak reports and research about that specific disease settings
+As @fig-outbreaks_count_by_types and @fig-enteric_outbreaks_count_by_settings implies, enteric outbreak occurrence rates are relatively low in the present time period.
+Therefore, it would be more efficient to focus on securing sufficient resources for preventing respiratory diesases rather than enteric.
+In addition, during periods when the incidence of enteric outbreak is low, the disease appears intensively only in LTCH and retirement homes.
+It would be a good idea to concentrate resources dealing with enteric diseases on those sites to achieve effective medical resource allocation.
 
-## Third discussion point
+## Countermeasures in LTCH for Mitigating Outbreak Occurrences
 
-from the results of the data analysis above, discuss a way we could prevent further increase of outbreak occurrence.
-in largely three dimensions: personal/ society/ government
+The spatial outbreak trend analysis (@fig-setting) implies that LTCH is where most of the outbreaks in Toronto take place.
+Considering the nature of long-term care homes, where some of the most vulnerable populations reside in a congregate living environment, the analysis above seems to be a natural result.
+Moreover, there were major challenges in LTCHs due to shortage of personal protective equipment(PPE) and inadequate staffing [@cihi_impact_covid19_2022].
+However, rather than calling it inevitable, we should make efforts to reduce the incidence of outbreak at the location, even a little.
+
+One suggestion is to create small, dedicated "cohorts" of staff assigned to specific groups of residents, rather than having staff move freely between floors [@cfp_ltc_2020].
+This would minimize staff interaction across larger groups and prevent outbreaks from sweeping through entire homes.
+Cohorting has already been proven effective in hospital settings, and is expected to also be beneficial in long-term care facilities.
 
@@
 
-Dividing a year into just two could influence the analysis result. (ex. weather affecting the outbreak occurrence)
-Many 'N/A', 'Unable to identify' data in Causative Agent.
+The study started with a grand intention of devising specific countermeasures for disease outbreaks by the analysis of outbreak trends sorted into time and place criteria.
+However, the conclusions derived from data analysis results seem to only put more emphasis on maintaining public health awareness at all times.
+Contrary to my original intention, it does not provide as much effective guidance for specific healthcare institutions in reallocating their resources.
+In addition, the dataset that I used had insufficient information for year 2024, since data from October 2024 and after was not possibly obtained.
+This could have affected the results of trend analysis because all the other data for year 2016 to 2023 had occurrence reports for the whole year.
 
-For further inspection, taking the place information in consideration might also have been meaningful.
+Further analysis using the seasonal information of this dataset might be intriguing.
+In particular, respiratory disease occur more severely during the change of seasons.
+Analyzing quarterly disease outbreak trends is expected to lead conclusions on how different quarterly preparations can be made throughout the year.
+Additionally, given more time and resources, it will be possible to develop a model for predicting the possibility of future infectious diseases.
 
@@
 
-# Appendix {-}
+# Appendix {#sec:appendix}
 
-# Additional data details
+## Additional data details
 
-## Other Similar Datasets
+### Other Similar Datasets Considered
 
-## Raw Data Variables Overview
+Another similar dataset that I had put in consideration when starting this research was the 'Monthly Communicable Disease Surveillance Data' [@toronto_communicable_disease_data_2024].
+Although this dataset contains observations of countless different types of diseases, it has only two types of variables, the number of cases and occurrence rates.
+So, it was determined that in-depth research in the field could be difficult by using this dataset.
+
+### Raw Data Variables Overview
+
+A detailed description of each variable in the original dataset is as follows [@toronto_outbreaks_dataset].
+
+_id: Unique row identifier for Open Data database.
+Institution Name: Name of the healthcare institution where the outbreak has been declared.
+Institution Address: Street address of the institution.
+Outbreak Setting: The type of facility, in terms of the type of healthcare provided.
+Type of Outbreak: The mode of transmission classification description of the outbreak.
+Causative Agent-1: The first aetiological agent(s) that have been identified from one or more outbreak-related cases, when an agent is confirmed.
+Causative Agent-2: The second aetiological agent(s) that have been identified from one or more outbreak-related cases, when an agent is confirmed.
+Date Outbreak Began: The date the institution reports the outbreak to TPH.
+Date Declared Over: In general, respiratory and enteric outbreaks are declared over if no new cases have been detected in the facility for a period of time equal to atleast one full incubation period plus one full period of communicability for the particular outbreak organism.
+Active: Institutional outbreaks remain active until they have been declared over and the public health investigation is closed.
+
+@tbl-variable_stats shows the summary statistics for all variables in the raw data.
+
+```{r}
+#| label: tbl-variable_stats
+#| tbl-cap: Raw Data Variables Summary Statistics
+#| echo: false
+#| warning: false
+#| message: false
+#| tbl.pos: 'H'
+
+# Function to generate refined summary statistics and display them in a long format
+generate_refined_summary_long_format <- function(data) {
+
+  # Numeric summary
+  numeric_summary <- data %>%
+    select(where(is.numeric)) %>%
+    summarise(across(everything(),
+                     list(
+                       count = ~sum(!is.na(.)),
+                       mean = ~mean(., na.rm = TRUE),
+                       median = ~median(., na.rm = TRUE),
+                       sd = ~sd(., na.rm = TRUE),
+                       min = ~min(., na.rm = TRUE),
+                       max = ~max(., na.rm = TRUE)
+                     ),
+                     .names = "{.col}_{.fn}")) %>%
+    pivot_longer(cols = everything(),
+                 names_to = c("variable", ".value"),
+                 names_sep = "_")  # Pivot to long format
+
+  # Categorical summary
+  categorical_summary <- data %>%
+    select(where(is.character)) %>%
+    summarise(across(everything(),
+                     list(
+                       count = ~sum(!is.na(.)),
+                       unique = ~n_distinct(.),
+                       top = ~names(sort(table(.), decreasing = TRUE)[1]),
+                       freq = ~max(table(.), na.rm = TRUE)
+                     ),
+                     .names = "{.col}_{.fn}")) %>%
+    pivot_longer(cols = everything(),
+                 names_to = c("variable", ".value"),
+                 names_sep = "_")  # Pivot to long format
+
+  # Date summary
+  date_summary <- data %>%
+    select(where(is.Date)) %>%
+    summarise(across(everything(),
+                     list(
+                       count = ~sum(!is.na(.)),
+                       min = ~min(., na.rm = TRUE),
+                       max = ~max(., na.rm = TRUE)
+                     ),
+                     .names = "{.col}_{.fn}")) %>%
+    pivot_longer(cols = everything(),
+                 names_to = c("variable", ".value"),
+                 names_sep = "_")  # Pivot to long format
+
+  # Combine all summaries into one long table
+  combined_summary <- bind_rows(numeric_summary, categorical_summary, date_summary) %>%
+    slice(-c(1:6))  # Remove the first 6 rows
+
+  # Return the pretty long table
+  combined_summary %>%
+    kable(caption = "Refined Summary Statistics (Long Format)") %>%
+    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
+                  full_width = F,
+                  font_size = 8)
+}
+
+# Generate the summary statistics table after removing the first 6 rows
+generate_refined_summary_long_format(raw_data) |>
+  kable_styling(latex_options = c("scale_down", "H"), font_size = 8)
+
+```
 

### 2024-10-03 state_respondents.qmd
Add files via upload

---
title: "2022 American Community Surveys Analysis"
author: "Group 25"
format: pdf
editor: visual
---
```{r}
#| include: false
#| warning: false
#| message: false

library(dplyr)
library(knitr)
library(tidyverse)
library(kableExtra)
library(ggplot2)
```

# Method of Obtaining the Data
To retrieve data from IPUMS USA, we first login to the IPUMS website and selected "IPUMS USA." From there, we clicked "Get Data" and chose the "2022 ACS" sample under the "SELECT SAMPLE" section. To obtain state-level data, we selected "HOUSEHOLD" > "GEOGRAPHIC" and added the "STATEICP" variable to the data cart. For individual-level data, we went to the "PERSON" section and included the "EDUC" variable. After reviewing the cart, we clicked "CREATE DATA EXTRACT" and set the "DATA FORMAT" to "csv." Finally, we submitted the extract and then downloaded and saved the file locally (e.g., "usa_00001.csv") for use in R.

```{r}
#| echo: false
# Load the IPUMS data
ipums_data <- read.csv("usa_00001.csv")

# Filter for doctoral degree holders
doctoral_data <- ipums_data %>% filter(EDUCD == 116)
```

Next, we count the respondents with doctoral degrees grouped by states.
```{r}
#| echo: false
#| tbl-cap: doctoral respondents
# Count respondents with doctoral degrees by state
doctoral_count <- doctoral_data %>%
  group_by(STATEICP) %>%
  summarise(doctoral = n())
kable(head(doctoral_count))  %>%
  kable_styling(full_width = FALSE, font_size = 10) %>%
  column_spec(1:2, width = "2cm")
```

# Brief Overview of the Ratio Estimators Approach

The ratio estimator is a statistical method used to estimate population totals or averages by utilizing known sample ratios. It operates by calculating the ratio of a specific characteristic (in this paper, the number of individuals holding doctoral degrees) relative to the total population in a known sample (the state of California). This ratio is then applied to other regions or groups, based on the assumption that the relationship between the characteristic and the population remains consistent across different areas. This method is particularly effective when complete population data is unavailable, but sample data provides a reliable basis for inferring proportional relationships that can be generalized.

# Estimates and the Actual Number of Respondents

When assuming that California has 391,171 respondents across all educational levels, we can calculate the ratio for California. The Laplace ratio method works well when the relationship between the characteristic of interest, such as the proportion of doctoral degree holders, and the population is consistent across all units. In the example above, the ratio for California is 0.01619752. However, if this ratio is not representative of other states due to unobserved factors, the estimates derived from the method can become biased. This bias occurs when regional differences disrupt the assumption of proportionality, leading to inaccurate estimates. Factors like varying educational systems, socioeconomic conditions, or migration patterns between states can cause these distortions.
```{r}
#| echo: false
california_total_respondents <- 391171
california_doctoral_respondents <- doctoral_count %>%
  filter(STATEICP == 71) %>%
  pull(doctoral)
california_ratio <- california_doctoral_respondents / california_total_respondents
print(paste("The California ratio that was found was:", california_ratio))

```

Using the ratio estimator approach might estimate the total number of respondents for each state as follows. Assuming the ratio holds for all states, we can estimate the total respondents for every state.

The table below presents an overview of the distribution of respondents with doctoral degrees across different states in the U.S., which is crucial for the understanding of regional educational attainment trends. The first column, STATEICP, refers to the unique state or territory codes provided by the IPUMS system, where each number corresponds to a specific state. The second column, doctoral_respondents, provides the number of respondents with doctoral degrees (filtered by EDUCD == 116) in each state. For example, state code 1 shows 600 respondents with doctoral degrees, while state code 3 shows a significantly higher number with 2014 respondents. Other state codes show varying counts of respondents, such as 165 for state code 2, 244 for state code 4, and so on.
```{r}
#| echo: false
#| tbl-cap: estimated total respondents
doctoral_count <- doctoral_count %>%
  mutate(estimated_total = doctoral / california_ratio)
kable(head(doctoral_count))  %>%
  kable_styling(full_width = FALSE, font_size = 10) %>%
  column_spec(1:3, width = "2cm")
```

The process of comparing it with the actual data of respondents is as follows.The table below provides a comparison between the estimated total respondents, derived using the ratio estimator method, and the actual total respondents for each state. The doctoral_respondents column represents the number of individuals with doctoral degrees in each state, while the estimated_total_respondents column shows the projected total population for each state, calculated by applying the California ratio (0.01619752) to the number of doctoral respondents in each state. The actual_total_respondents column reflects the true population count for each state, which was directly obtained from the data. This allows for a comparison between the estimated and actual figures. In state code 1, the actual number of respondents is 37,369, slightly higher than the estimate of 37,042. For state code 3, the actual total is 73,077, significantly lower than the estimate of 124,340. These differences illustrate the potential inaccuracies that can arise from using a ratio estimator when the relationship between the doctoral respondents and the total population is not consistent across states. This highlights the importance of accounting for regional variations in such analyses.
```{r}
#| echo: false
#| tbl-cap: comparison of the estimated and actual total respondents

actual_total_count <- ipums_data %>%
  group_by(STATEICP) %>%
  summarise(actual_total = n())

doctoral_count <- doctoral_count %>%
  left_join(actual_total_count, by = "STATEICP")

kable(head(doctoral_count))  %>%
  kable_styling(full_width = FALSE, font_size = 10) %>%
  column_spec(1:4, width = "2cm")
```
```{r}
#| warning: false
#| message: false
#| echo: false
#| label: fig-compare
#| fig-cap: comparison of estimated and total respondents
#| fig-height: 9

# Load library
library(ggplot2)
library(reshape2)

# Further process the data
data_melted <- melt(doctoral_count, id.vars = "STATEICP", measure.vars = c("estimated_total", "actual_total"))

# Generate chart
ggplot(data_melted, aes(x = factor(data_melted$STATEICP), y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "State Code", y = "Number of Respondents", fill = "Respondent Type",
       title = "Comparison of Estimated and Actual Total Respondents") +
  theme_minimal() +
  coord_flip()
```

# Explanation of Why They are Different

A simple explanation for why the estimated number of respondents and actual number of respondents differ is that the naive assumption that all states share the same doctorate-respondent to total respondent ratio is false. The ratio estimator assumes that the proportion of respondents with doctoral degrees in California is representative of the proportion in other states. However, educational attainment can vary significantly due to differences in demographics, economic opportunities, and educational infrastructure across states. This variance leads to discrepancies between the estimated and actual counts. When the data used for estimation is based on a sample rather than a complete population census, random sampling variability introduces uncertainty into the calculated ratio, which in turn affects the accuracy of the estimates. This variability can lead to deviations between the sample-based ratio and the true population ratio, impacting the reliability of the estimation.

### 2024-10-10 paper/paper.qmd
Add files via upload

--- 2024-09-26 paper/paper.qmd
@@
 ---
-title: "Outbreaks in Toronto Healthcare Institutions After COVID-19"
-subtitle: "Analysis of the Frequency Trends and the Causative Agent"
+title: "My title"
+subtitle: "My subtitle if needed"
 author: 
   - Yunkyung Ko
-thanks: "Code and data are available at: https://github.com/koyunkyung/toronto_outbreaks"
+  - Chenika Bukes
+thanks: "Code and data are available at: [https://github.com/koyunkyung/spotify-API](https://github.com/koyunkyung/spotify-API)."
 date: today
 date-format: long
-abstract: "In Canada, outbreak incidence has been slowly increasing from June 2024 up to the beginning of the accumulating data period on August 10, 2024.
-(ADD CITATION https://health-infobase.canada.ca/covid-19/) Using the 'Outbreaks in Toronto Healthcare Institutions' dataset, we show how the trends of outbreak occurrence has been over the 5 years after COVID-19.
-(specify headline result).
-(implications)."
+abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
 format: pdf
@@
 library(tidyverse)
-library(opendatatoronto)
-library(ggplot2)
-library(dplyr)
-library(readr)
-library(lubridate)
-
-data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+library(spotifyr)
 ```
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR, @opendatatoronto and @rohan.
+Overview paragraph
 
-The remainder of this paper is structured as follows. @sec-data....
+Estimand paragraph
+
+Results paragraph
+
+Why it matters paragraph
+
+Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
 
@@
 
-@fig-outbreak shows the total number of outbreaks reported in Toronto healthcare institutions over the time period of 2020 to 2024.
-We can see that the numbers seem to decrease by the end of 2021, but exponentially increase from the start of 2022.
-A trend of relatively high outbreak occurrence rate continues till 2024.
+## Overview
+
+We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+
+Overview text
+
+## Measurement
+
+Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+
+## Outcome variables
+
+Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+
+Some of our data is of penguins (@fig-bills), from @palmerpenguins.
 
 ```{r}
+#| label: fig-bills
+#| fig-cap: Bills of penguins
+#| echo: false
 
+ggplot(penguins, aes(x = island, fill = species)) +
+  geom_bar(alpha = 0.8) +
+  scale_fill_manual(values = c("darkorange","purple","cyan4"),
+                    guide = "none") +
+  theme_minimal() +
+  facet_wrap(~species, ncol = 1) +
+  coord_flip()
 ```
 
-```{r}
-#| label: fig-outbreak
-#| fig-cap: Total Outbreaks Count in Toronto Healthcare Institutions
-#| echo: false
+Talk more about it.
 
-data |>
-  ggplot(mapping = aes(x = year)) +
-  geom_bar() +
-  theme_minimal() +
-  labs(x = "Date Period", y = "Number of Outbreaks Reported",
-       title = "Total Number of Outbreaks Reported in Toronto Healthcare Institutions") +
-  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
-
-```
+And also planes (@fig-planes).
+(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
 ```{r}
-#| label: fig-setting
-#| fig-cap: Outbreaks Count by Outbreak Settings
-#| echo: false
-
-library(dplyr)
-library(ggplot2)
-
-outbreak_counts <- data %>%
-  count(outbreak_setting)
-
-ggplot(outbreak_counts, aes(x = "", y = n, fill = outbreak_setting)) +
-  geom_bar(stat = "identity", width = 1) +
-  coord_polar("y", start = 0) + 
-  theme_minimal() +
-  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
-                               "Retirement Home" = "#B3FFC6", 
-                               "Hospital-Chronic Care" = "#BA55D3", 
-                               "Hospital-Acute Care" = "#DDA0DD",
-                               "Hospital-Psychiatric" = "grey",
-                               "Shelter" = "brown",
-                               "Transitional Care" = "black")) +
-  labs(x = NULL, y = NULL, fill = "Outbreak Setting", 
-       title = "Number of Outbreaks Reported by Setting") +
-  theme(axis.text.x = element_blank(),  
-        axis.ticks = element_blank(),  
-        panel.grid = element_blank())  
-```
-
-Contrary to the expectation that the number of outbreaks would have increased the most rapidly immediately after COVID-19, the analysis with actual data shows that the number of outbreaks is on the rise as we approach 2022.
-Since there are various variables that can affect the result above, we would like to consider this one by one and proceed with an accurate analysis of the outbreak occurrence situation.
-### NEED TO MODIFY
-# check for factors such as comparing the number of outbreaks as a whole vs. only in healthcare institutions
-# would the _begin, _end sorting of timeline affect the shape of the graph above?
-# search for background information of healthcare environments in Canada that could relate to the data above
-
-# Result
-
-To dive deeper into the causations of this trend in outbreak reports, we show the frequency trend of top-3 causative agents respectively over the same time period (@fig-causations).
-```{r}
-
-```
-
-```{r}
-#| label: fig-outbreaks count by types
-#| fig-cap: Outbreaks Count in Toronto Healthcare Institutions, by Types of Outbreak
+#| label: fig-planes
+#| fig-cap: Relationship between wing length and width
 #| echo: false
@@
 
-# Summarizing the data to get the number of outbreaks per month and type of outbreak
-annual_data <- data %>%
-  group_by(year, type_of_outbreak) %>%
-  summarise(outbreak_count = n()) %>%
-  ungroup()
+analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
 
-# Convert the year column to a factor with levels in chronological order
-annual_data$year <- factor(annual_data$year, levels = unique(annual_data$year))
+analysis_data |> 
+  ggplot(aes(x = width, y = length)) +
+  geom_point(alpha = 0.8) +
+  theme_minimal() +
+  labs(x = "Wing width (mm)",
+       y = "Wing length (mm)")
+```
 
-# Create the line plot
-ggplot(annual_data, aes(x = year, y = outbreak_count, group = type_of_outbreak, color = type_of_outbreak)) +
-  geom_line(size = 1) +
-  geom_point(size = 2) +
-  scale_color_manual(values = c("Respiratory" = "red", "Enteric" = "blue", "Other" = "black")) +
-  labs(title = "Trend of Outbreak Counts by Type (2016-2024)",
-       x = "Year", y = "Outbreak Count", color = "Type of Outbreak") +
-  theme_minimal() +
-  scale_x_discrete(drop = FALSE) # Ensure all years are shown on the x-axis
+Talk way more about it.
 
+## Predictor variables
+
+Add graphs, tables and text.
+
+Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+
+# Model
+
+The goal of our modelling strategy is twofold. Firstly,...
+
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+
+## Model set-up
+
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_i + \gamma_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+### Model justification
+
+We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+
+We can use maths by including latex between dollar signs, for instance $\theta$.
+
+# Results
+
+Our results are summarized in @tbl-modelresults.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+library(rstanarm)
+
+first_model <-
+  readRDS(file = here::here("models/first_model.rds"))
 ```
@@
 ```{r}
-#| label: fig-respiratory outbreaks count by outbreak settings
-#| fig-cap: Respiratory Outbreaks Count, by Outbreak Settings
 #| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
-#| message: false
 
-library(dplyr)
-library(ggplot2)
-
-# Filter for 'Respiratory' outbreaks
-respiratory_data <- data %>%
-  filter(type_of_outbreak == "Respiratory")
-
-# Aggregate data by year and outbreak setting
-annual_counts <- respiratory_data %>%
-  group_by(year, outbreak_setting) %>%
-  summarise(outbreak_count = n(), .groups = 'drop')
-
-# Create a stacked bar plot
-ggplot(annual_counts, aes(x = year, y = outbreak_count, fill = outbreak_setting)) +
-  geom_col(position = "stack") +  # Stack bars based on outbreak settings
-  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
-                               "Retirement Home" = "#B3FFC6", 
-                               "Hospital-Chronic Care" = "#BA55D3", 
-                               "Hospital-Acute Care" = "#DDA0DD",
-                               "Hospital-Psychiatric" = "grey",
-                               "Shelter" = "brown",
-                               "Transitional Care" = "black")) +
-  labs(title = "Annual Respiratory Outbreak Counts by Setting",
-       x = "Year",
-       y = "Outbreak Count",
-       fill = "Outbreak Setting") +
-  theme_minimal() +
-  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
+modelsummary::modelsummary(
+  list(
+    "First model" = first_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
 ```
-
-```{r}
-#| label: fig-enteric outbreaks count by outbreak settings
-#| fig-cap: Enteric Outbreaks Count, by Outbreak Settings
-#| echo: false
-#| warning: false
-#| message: false
-
-library(dplyr)
-library(ggplot2)
-
-# Filter for 'Enteric' outbreaks
-enteric_data <- data %>%
-  filter(type_of_outbreak == "Enteric")
-
-# Aggregate data by year and outbreak setting
-annual_counts <- enteric_data %>%
-  group_by(year, outbreak_setting) %>%
-  summarise(outbreak_count = n(), .groups = 'drop')
-
-# Create a stacked bar plot
-ggplot(annual_counts, aes(x = year, y = outbreak_count, fill = outbreak_setting)) +
-  geom_col(position = "stack") +  # Stack bars based on outbreak settings
-  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
-                               "Retirement Home" = "#B3FFC6", 
-                               "Hospital-Chronic Care" = "#BA55D3", 
-                               "Hospital-Acute Care" = "#DDA0DD",
-                               "Hospital-Psychiatric" = "grey",
-                               "Shelter" = "brown",
-                               "Transitional Care" = "black")) +
-  labs(title = "Annual Enteric Outbreak Counts by Setting",
-       x = "Year",
-       y = "Outbreak Count",
-       fill = "Outbreak Setting") +
-  theme_minimal() +
-  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))
-```
-Talk way more about it.
 
@@
 
-analyze the total outbreak frequency trend over the time period given.
-(in relation to the social/political background in healthcare)
+If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
 
@@
 
-figure out what might be the main causative agent for the rapid increase in outbreak reports starting from 2022.
-acknowledging the causative agent that we should be aware of might help us prevent further outbreak increase from happening
--> know what is causing this rise in outbreak reports and research about that specific disease settings
+Please don't use these as sub-heading labels - change them to be what your point actually is.
 
@@
 
-from the results of the data analysis above, discuss a way we could prevent further increase of outbreak occurrence.
-in largely three dimensions: personal/ society/ government
-
 ## Weaknesses and next steps
 
-Dividing a year into just two could influence the analysis result. (ex. weather affecting the outbreak occurrence)
-Many 'N/A', 'Unable to identify' data in Causative Agent.
-
-For further inspection, taking the place information in consideration might also have been meaningful.
+Weaknesses and next steps should also be included.
 
@@
 # Appendix {-}
-INFO OF RAW DATA, CONDITIONS OF DATA COLLECTION, ETC
 
 # Additional data details
-NEED TO MODIFY
+
+# Model details {#sec-model-details}
+
+## Posterior predictive check
+
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheckandposteriorvsprior
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+
+pp_check(first_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+posterior_vs_prior(first_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+```
+
+## Diagnostics
+
+@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+
+@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-stanareyouokay
+#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| layout-ncol: 2
+
+plot(first_model, "trace")
+
+plot(first_model, "rhat")
+```
 

### 2024-10-10 scripts/02-download_data.R

@@ -1,0 +1,22 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from https://open.spotify.com/
+# Author: Yunkyung Ko
+# Date: 10 October 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(spotifyr)
+
+#### Download data ####
+radiohead <- get_artist_audio_features("radiohead")
+the_national <- get_artist_audio_features("the national")
+
+
+#### Save data ####
+saveRDS(radiohead, "data/01-raw_data/radiohead.rds")
+         

### 2024-10-10 paper/paper.qmd
Add files via upload

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Coldplay vs Radiohead Danceability over Time"
+subtitle: "Coldplay reamins more danceable over the years"
 author: 
-  - Yunkyung Ko
   - Chenika Bukes
-thanks: "Code and data are available at: [https://github.com/koyunkyung/spotify-API](https://github.com/koyunkyung/spotify-API)."
+  - Khushaal Nandwani
+  - Yongkyung Ko
+thanks: "Code and data are available at: https://github.com/chenikabukes/RadioheadColdplay"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study explores the danceability of Coldplay and Radiohead over time by using data extracted from Spotify’s API.
+The results show that Coldplay has consistently maintained a higher level of danceability compared to Radiohead.
+These findings are relevant in understanding the musical evolution of both bands and their appeal to different types of listeners."
 format: pdf
@@
 library(tidyverse)
-library(spotifyr)
+
+radiohead <- readRDS("../data/01-raw_data/radiohead.rds")
+coldplay <- readRDS("../data/01-raw_data/coldplay.rds")
 ```
@@
 # Introduction
+The estimand in this paper is the average danceability of tracks by both Coldplay and Radiohead over time.
+Danceability describes how suitable a track is for dancing based on a combination of tempo, rhythm stability, beat strength, and overall regularity.
+This paper shows how this metric changes for each band and discusses the implications.
 
-Overview paragraph
-
-Estimand paragraph
-
-Results paragraph
-
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+Understanding how danceability evolves over time can provide insight into the bands' stylistic changes and the broader trends in the music industry.
+All analysis in this paper was completed using R [@citeR].
+The remainder of this paper is structured as follows.
+Section @sec-data describes the data collection and preparation.
+Section @sec-results presents the findings using the tidyverse package [@tidyverse], and Section @sec-discussion discusses the implications of these results.
 
@@
 
-## Overview
+The data used in this study was retrieved using the Spotify Web API and processed using the R package spotifyr [@Spotify].
+The package allows bulk retrieval of an artist’s entire discography along with associated audio features, including danceability, tempo, energy, and others.
+We accessed both Coldplay's and Radiohead’s discographies and calculated the average danceability of their songs by year of release.
+The data consists of metadata about each track, including the release year, danceability score, and album information.
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+# Results {#sec-results}
 
-Overview text
-
-## Measurement
-
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
-
-## Outcome variables
-
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
+In @fig-coldplay, we observe Coldplay's tracks exhibit relatively stable danceability over time, with scores ranging between 0.4 and 0.6, with slight increases toward the latest releases.
+Coldplay’s more consistent and upward trend aligns with their pop-rock style, which tends to favor rhythmically engaging and accessible music, contributing to a higher and more consistent danceability score.
+Radiohead’s danceability exhibits much more fluctuation, ranging from around 0.2 to 0.6.
+There is an increase during the early 2000s, but it dips again afterward.
+The lower and more fluctuating danceability for Radiohead reflects their tendency to diverge from mainstream pop structures.
 
@@
 #| warning: false
+#| label: fig-coldplay
+#| tbl-cap: "Danceability of Coldplay's and Radiohead's tracks over time"
 #| message: false
+#| fig-cap: "Danceability of Coldplay's and Radiohead's tracks over time"
 
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+rbind(coldplay, radiohead) |>
+  select(artist_name, album_release_date, danceability) |>
+  mutate(album_release_date = ymd(album_release_date)) |>
+  ggplot(aes( x = album_release_date, y = danceability, color = artist_name)) +
+  geom_point(alpha = 0.3) +
+  geom_smooth() +
+  theme_minimal() +
+  facet_wrap(facets = vars(artist_name), dir = "v") +
+  labs(
+    x = "Album release date",
+    y = "Danceability",
+    color = "Artist"
+  ) +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
 ```
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+# Discussion {#sec-discussion}
+Both bands show some increase in danceability in their later albums, though Coldplay's increase is more pronounced, reflecting their shift towards more pop-centric music.
 
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
+# Appendix
+Out of interest we decided to explore the correlation between danceability and tempo. The below results do not show a significant relationship and thus were not included in the main findings. 
+```{R}
 #| echo: false
@@
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| label: fig-corrplot_r
+#| fig-cap: "Danceability vs Tempo for Radiohead Tracks"
+#| fig-width: 2.5
+#| fig-height: 2
 
-plot(first_model, "trace")
+library(ggplot2)
+library(dplyr)
+radiohead <- readRDS("../data/01-raw_data/radiohead.rds")
 
-plot(first_model, "rhat")
+#### Radiohead
+ggplot(radiohead, aes(x = tempo, y = danceability)) +
+  geom_point(color = "blue", alpha = 0.6) +
+  geom_smooth(method = "lm", color = "red", se = FALSE) +
+  labs(x = "Tempo (BPM)",
+       y = "Danceability") +
+theme_minimal()
+
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-corrplot_c
+#| fig-cap: "Danceability vs Tempo for Coldplay Tracks"
+#| fig-width: 2.5
+#| fig-height: 2
+
+library(ggplot2)
+library(dplyr)
+coldplay <- readRDS("../data/01-raw_data/coldplay.rds")
+
+#### Coldplay
+ggplot(coldplay, aes(x = tempo, y = danceability)) +
+  geom_point(color = "green", alpha = 0.6) +
+  geom_smooth(method = "lm", color = "red", se = FALSE) +
+  labs(x = "Tempo (BPM)",
+       y = "Danceability") +
+  theme_minimal()
 ```
-
 \newpage 
-
+\newpage
 # References

### 2024-10-10 scripts/02-download_data.R

@@ -1,0 +1,26 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from the Spotify API
+# Author: Group 20 
+# Date: 10 October 2024
+# Contact: chenika.bukes@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: You have saved your spotify API credentials to the environment file. 
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+library(spotifyr)
+
+#### Download data ####
+radiohead <- get_artist_audio_features("radiohead")
+coldplay <- get_artist_audio_features("coldplay")
+
+
+
+#### Save data ####
+saveRDS(radiohead, "data/01-raw_data/radiohead.rds")
+saveRDS(coldplay, "data/01-raw_data/coldplay.rds")
+
+         

### 2024-10-10 paper/paper.qmd
Update paper.qmd

@@
   - Khushaal Nandwani
-  - Yongkyung Ko
+  - Yunkyung Ko
 thanks: "Code and data are available at: https://github.com/chenikabukes/RadioheadColdplay"

### 2024-10-13 scripts/02-download_data.R
Add files via upload

@@ -1,0 +1,25 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from the FiveThirtyEight's 2024 US Presidential Election Polls (National)
+# Author: Yunkyung Ko
+# Date: 12 October 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+
+#### Download data ####
+# [...ADD CODE HERE TO DOWNLOAD...]
+
+
+
+#### Save data ####
+# [...UPDATE THIS...]
+# change the_raw_data to whatever name you assigned when you downloaded it.
+write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+
+         

### 2024-10-13 scripts/00-simulate_data.R
download raw dataset

@@ -1,52 +1,124 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates data
+# Author: Yunkyung Ko
+# Date: 22 September 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: None
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
+library(lubridate)
 
 
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+
+## Total Outbreak Counts Reported in Toronto Healthcare Institutions
+# Set seed for reproducibility
+set.seed(123)
+
+# Define the start and end dates
+start_date <- as.Date("2015-01-01")
+end_date <- as.Date("2025-12-31")
+
+# Generate a sequence of monthly dates between start_date and end_date
+monthly_dates <- seq(from = start_date, to = end_date, by = "month")
+
+# Format dates to "YYYY-MM" without including the day
+formatted_dates <- format(monthly_dates, "%Y-%m")
+
+# Generate random Poisson distributed numbers with lambda = 10 for each month
+lambda <- 15
+poisson_values <- rpois(length(formatted_dates), lambda)
+
+# Create a data frame to hold the results
+sim_date_total <- data.frame(Date = formatted_dates, Number_of_Outbreaks = poisson_values)
+
+### Write CSV
+write_csv(sim_date_total, file = "data/raw_data/simulated_data/simulated_date_total.csv")
+
+
+
+
+
+## Number of Outbreaks Reported by Setting
+# Set seed for reproducibility
+set.seed(123)
+
+# Simulated data with random numbers
+sim_setting_total <- data.frame(
[…]

### 2024-10-13 scripts/01-download_data.R

@@ -1,0 +1,36 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from Open Data Toronto
+# Author: Yunkyung Ko
+# Date: 22 September 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(opendatatoronto)
+
+#### Download data ####
+package <- show_package("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
+resources <- list_package_resources("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
+datastore_resources <- filter(resources, tolower(format) %in% c("csv", "geojson"))
+
+resources_list <- filter(datastore_resources, row_number() %in% 1:9)
+
+datasets <- list()
+for (i in seq_len(nrow(resources_list))) {
+  datasets[[i]] <- get_resource(resources_list[i, ])
+}
+
+#### Save data ####
+# change datasets to whatever name you assigned when you downloaded it.
+combined_data <- bind_rows(datasets)
+write_csv(combined_data, "data/raw_data/raw_data.csv")
+
+# Code linting and styling
+library(lintr)
+library(styler)
+style_file(path = "scripts/01-download_data.R")
+lint(filename = "scripts/01-download_data.R")

### 2024-10-13 scripts/02-data_cleaning.R

@@ -1,0 +1,31 @@
+#### Preamble ####
+# Purpose: Cleans the raw healthcare data into an analysis dataset
+# Author: Yunkyung Ko
+# Date: 22 September 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to have downloaded the data.
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read_csv("data/raw_data/raw_data.csv")
+
+cleaned_data <- raw_data |>
+  janitor::clean_names() |>
+  select(outbreak_setting, type_of_outbreak, date_outbreak_began) |>
+  mutate(
+    # Extract year
+    year = substr(date_outbreak_began, 1, 4),
+  )
+
+#### Save data ####
+write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")
+
+# Code linting and styling
+library(lintr)
+library(styler)
+style_file(path = "scripts/02-data_cleaning.R")
+lint(filename = "scripts/02-data_cleaning.R")

### 2024-10-13 scripts/00-simulate_data.R
add files via upload

@@ -1,124 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates data
-# Author: Yunkyung Ko
-# Date: 22 September 2024
-# Contact: yunkyung.ko@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(lubridate)
-
-
-#### Simulate data ####
-
-## Total Outbreak Counts Reported in Toronto Healthcare Institutions
-# Set seed for reproducibility
-set.seed(123)
-
-# Define the start and end dates
-start_date <- as.Date("2015-01-01")
-end_date <- as.Date("2025-12-31")
-
-# Generate a sequence of monthly dates between start_date and end_date
-monthly_dates <- seq(from = start_date, to = end_date, by = "month")
-
-# Format dates to "YYYY-MM" without including the day
-formatted_dates <- format(monthly_dates, "%Y-%m")
-
-# Generate random Poisson distributed numbers with lambda = 10 for each month
-lambda <- 15
-poisson_values <- rpois(length(formatted_dates), lambda)
-
-# Create a data frame to hold the results
-sim_date_total <- data.frame(Date = formatted_dates, Number_of_Outbreaks = poisson_values)
-
-### Write CSV
-write_csv(sim_date_total, file = "data/raw_data/simulated_data/simulated_date_total.csv")
-
-
-
-
-
-## Number of Outbreaks Reported by Setting
-# Set seed for reproducibility
-set.seed(123)
-
-# Simulated data with random numbers
-sim_setting_total <- data.frame(
-  Setting = c("Hospital-Acute Care", "Hospital-Chronic Care", "Hospital-Psychiatric", "LTCH", "Retirement Home", "Shelter", "Transitional Care"),
-  Count = round(runif(7, min = 1, max = 1000)) # Generate random numbers between 1 and 1000
-)
-### Write CSV
-write_csv(sim_setting_total, file = "data/raw_data/simulated_data/simulated_setting_total.csv")
-
-
-
-
-
-## Trend of Outbreak Counts by Type
-# Set seed for reproducibility
-set.seed(123)
-
-## Simulated data for outbreak counts by type and year
-years <- 2015:2025
-
-# Generate random outbreak counts for each type of outbreak
-respiratory <- round(runif(length(years), min = 200, max = 1000)) # Respiratory outbreaks
[…]

### 2024-10-13 scripts/01-download_data.R

@@ -1,36 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from Open Data Toronto
-# Author: Yunkyung Ko
-# Date: 22 September 2024
-# Contact: yunkyung.ko@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(opendatatoronto)
-
-#### Download data ####
-package <- show_package("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
-resources <- list_package_resources("80ce0bd7-adb2-4568-b9d7-712f6ba38e4e")
-datastore_resources <- filter(resources, tolower(format) %in% c("csv", "geojson"))
-
-resources_list <- filter(datastore_resources, row_number() %in% 1:9)
-
-datasets <- list()
-for (i in seq_len(nrow(resources_list))) {
-  datasets[[i]] <- get_resource(resources_list[i, ])
-}
-
-#### Save data ####
-# change datasets to whatever name you assigned when you downloaded it.
-combined_data <- bind_rows(datasets)
-write_csv(combined_data, "data/raw_data/raw_data.csv")
-
-# Code linting and styling
-library(lintr)
-library(styler)
-style_file(path = "scripts/01-download_data.R")
-lint(filename = "scripts/01-download_data.R")

### 2024-10-13 scripts/01-test_simulated_data.R

@@ -1,88 +1,0 @@
-#### Preamble ####
-# Purpose: Tests the structure and validity of the simulated US presidential election polls data
-# Author: Yunkyung Ko
-# Date: 12 October 2024
-# Contact: yunkyung.ko@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `us_election` rproj
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
-
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
-
-
-#### Test data ####
-
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 151 rows.")
-}
-
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
-} else {
-  stop("Test Failed: The dataset does not have 3 columns.")
-}
-
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
-} else {
-  stop("Test Failed: The 'division' column contains duplicate values.")
-}
-
-# Check if the 'state' column contains only valid Australian state names
-valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", 
-                  "Western Australia", "Tasmania", "Northern Territory", 
-                  "Australian Capital Territory")
-
-if (all(analysis_data$state %in% valid_states)) {
-  message("Test Passed: The 'state' column contains only valid Australian state names.")
-} else {
-  stop("Test Failed: The 'state' column contains invalid state names.")
-}
-
-# Check if the 'party' column contains only valid party names
-valid_parties <- c("Labor", "Liberal", "Greens", "National", "Other")
-
-if (all(analysis_data$party %in% valid_parties)) {
-  message("Test Passed: The 'party' column contains only valid party names.")
-} else {
-  stop("Test Failed: The 'party' column contains invalid party names.")
-}
-
[…]

### 2024-10-13 scripts/02-data_cleaning.R

@@ -1,31 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw healthcare data into an analysis dataset
-# Author: Yunkyung Ko
-# Date: 22 September 2024
-# Contact: yunkyung.ko@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Need to have downloaded the data.
-# Any other information needed? None
-
-#### Workspace setup ####
-library(tidyverse)
-
-#### Clean data ####
-raw_data <- read_csv("data/raw_data/raw_data.csv")
-
-cleaned_data <- raw_data |>
-  janitor::clean_names() |>
-  select(outbreak_setting, type_of_outbreak, date_outbreak_began) |>
-  mutate(
-    # Extract year
-    year = substr(date_outbreak_began, 1, 4),
-  )
-
-#### Save data ####
-write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")
-
-# Code linting and styling
-library(lintr)
-library(styler)
-style_file(path = "scripts/02-data_cleaning.R")
-lint(filename = "scripts/02-data_cleaning.R")

### 2024-10-13 scripts/02-download_data.R

@@ -1,25 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from the FiveThirtyEight's 2024 US Presidential Election Polls (National)
-# Author: Yunkyung Ko
-# Date: 12 October 2024
-# Contact: yunkyung.ko@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
-
-
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
-
-         

### 2024-10-13 scripts/03-test_data.R

@@ -1,140 +1,0 @@
-#### Preamble ####
-# Purpose: Sanity check of the data, Hypotheis verification of how the simulated data would unfold
-# Author: Yunkyung Ko
-# Date: 22 September 2024
-# Contact: yunkyung.ko@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Need to have simulated data
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-
-#### Test data ####
-date_total <- read_csv("data/raw_data/simulated_data/simulated_date_total.csv")
-# Test for negative numbers
-date_total$Number_of_Outbreaks |> min() >= 0
-# Test for NAs
-all(is.na(date_total$Number_of_Outbreaks |> min() >= 0))
-
-# Create the bar plot
-p <- ggplot(date_total, aes(x = factor(Date), y = Number_of_Outbreaks)) +
-  geom_bar(stat = "identity", fill = "black") +
-  labs(
-    title = "Total Outbreak Counts (Simulation)",
-    x = "Date",
-    y = "Number of Outbreaks Reported"
-  ) +
-  theme_minimal()
-
-# Save the plot as a PNG file
-ggsave("data/raw_data/simulated_results/simulated_total_outbreak_count.png", plot = p, width = 10, height = 8, dpi = 300)
-
-
-
-
-#### Test data ####
-setting_total <- read_csv("data/raw_data/simulated_data/simulated_setting_total.csv")
-# Test for negative numbers
-setting_total$Count |> min() >= 0
-# Test for NAs
-all(is.na(setting_total$Count |> min() >= 0))
-
-# Create a pie chart
-p <- ggplot(setting_total, aes(x = "", y = Count, fill = Setting)) +
-  geom_bar(stat = "identity", width = 1) +
-  coord_polar("y") +
-  labs(title = "Outbreak Counts by Setting (Simulation)", fill = "Outbreak Setting") +
-  theme_void() +
-  scale_fill_manual(values = c("#d3a9d3", "#a455a4", "#b3b3b3", "#a2d6f9", "#b6f5d3", "#963232", "#000000"))
-
-# Save the plot as a PNG file
-ggsave("data/raw_data/simulated_results/simulated_outbreaks_by_setting.png", plot = p, width = 10, height = 8, dpi = 300)
-
-
-
-
-
-
-#### Test data ####
[…]

### 2024-10-13 paper/paper.qmd
initial framework & download raw data

--- 2024-10-10 paper/paper.qmd
@@
   - Yunkyung Ko
-  - Chenika Bukes
-thanks: "Code and data are available at: [https://github.com/koyunkyung/spotify-API](https://github.com/koyunkyung/spotify-API)."
+thanks: "Code and data are available at: [https://github.com/koyunkyung/us_election_2024](https://github.com/koyunkyung/us_election_2024)."
 date: today
@@
 library(tidyverse)
-library(spotifyr)
 ```

### 2024-10-13 scripts/01-test_simulated_data.R

@@ -1,0 +1,88 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the simulated US presidential election polls data
+# Author: Yunkyung Ko
+# Date: 12 October 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+  # - The `tidyverse` package must be installed and loaded
+  # - 00-simulate_data.R must have been run
+# Any other information needed? Make sure you are in the `us_election` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+
+# Test if the data was successfully loaded
+if (exists("analysis_data")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
+
+
+#### Test data ####
+
+# Check if the dataset has 151 rows
+if (nrow(analysis_data) == 151) {
+  message("Test Passed: The dataset has 151 rows.")
+} else {
+  stop("Test Failed: The dataset does not have 151 rows.")
+}
+
+# Check if the dataset has 3 columns
+if (ncol(analysis_data) == 3) {
+  message("Test Passed: The dataset has 3 columns.")
+} else {
+  stop("Test Failed: The dataset does not have 3 columns.")
+}
+
+# Check if all values in the 'division' column are unique
+if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
+  message("Test Passed: All values in 'division' are unique.")
+} else {
+  stop("Test Failed: The 'division' column contains duplicate values.")
+}
+
+# Check if the 'state' column contains only valid Australian state names
+valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", 
+                  "Western Australia", "Tasmania", "Northern Territory", 
+                  "Australian Capital Territory")
+
+if (all(analysis_data$state %in% valid_states)) {
+  message("Test Passed: The 'state' column contains only valid Australian state names.")
+} else {
+  stop("Test Failed: The 'state' column contains invalid state names.")
+}
+
+# Check if the 'party' column contains only valid party names
+valid_parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+
+if (all(analysis_data$party %in% valid_parties)) {
+  message("Test Passed: The 'party' column contains only valid party names.")
+} else {
+  stop("Test Failed: The 'party' column contains invalid party names.")
+}
+
[…]

### 2024-10-13 scripts/02-download_data.R

@@ -1,0 +1,19 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from the FiveThirtyEight's 2024 US Presidential Election Polls (National)
+# Author: Yunkyung Ko
+# Date: 12 October 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Download data ####
+raw_elections_data <-read_csv(file ="https://projects.fivethirtyeight.com/polls/data/president_polls.csv")
+
+#### Save data ####
+write_csv(raw_elections_data, "data/01-raw_data/raw_elections_data.csv")
+
+         

### 2024-10-13 scripts/03-clean_data.R

@@ -1,0 +1,44 @@
+#### Preamble ####
+# Purpose: Cleans the raw elections data
+# Author: Yunkyung Ko
+# Date: 12 October 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to have downloaded the data.
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Clean data ####
+raw_data <- read_csv("data/01-raw_data/raw_elections_data.csv")
+
+cleaned_data <-
+  raw_data |>
+  janitor::clean_names() |>
+  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
+  filter(wing_width_mm != "caw") |>
+  mutate(
+    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
+                                   "1.35",
+                                   flying_time_sec_first_timer)
+  ) |>
+  mutate(wing_width_mm = if_else(wing_width_mm == "490",
+                                 "49",
+                                 wing_width_mm)) |>
+  mutate(wing_width_mm = if_else(wing_width_mm == "6",
+                                 "60",
+                                 wing_width_mm)) |>
+  mutate(
+    wing_width_mm = as.numeric(wing_width_mm),
+    wing_length_mm = as.numeric(wing_length_mm),
+    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
+  ) |>
+  rename(flying_time = flying_time_sec_first_timer,
+         width = wing_width_mm,
+         length = wing_length_mm
+         ) |> 
+  tidyr::drop_na()
+
+#### Save data ####
+write_csv(cleaned_data, "outputs/data/analysis_data.csv")

### 2024-10-14 scripts/03-clean_data.R
select relevant variables for the analysis data

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw elections data
+# Purpose: Cleans the raw elections data and selects relevant variables for the analysis
 # Author: Yunkyung Ko
 # Date: 12 October 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
@@ -11,34 +11,13 @@
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("data/01-raw_data/raw_elections_data.csv")
+raw_elections_data <- read_csv("data/01-raw_data/raw_elections_data.csv")
 
 cleaned_data <-
-  raw_data |>
+  raw_elections_data |>
   janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
+  select(pollster_id, pollster, methodology, transparency_score, state, race_id, sample_size, party, answer) |>
   tidyr::drop_na()
 
 #### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+write_csv(cleaned_data, "data/02-analysis_data/elections_data.csv")

### 2024-10-19 scripts/00-simulate_data.R
simulation of poll data for Harris

@@ -1,12 +1,11 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates a dataset of US election polling data
+# Author: Yunkyung Ko
+# Date: 19 October 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Any other information needed? Make sure you are in the `us_election` rproj
 
 
 #### Workspace setup ####
@@ -14,39 +13,26 @@ library(tidyverse)
 set.seed(853)
 
 
-#### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+#### Simulate polling data for Kamala Harris ####
+# Simulate poll data
+sim_poll_data <- tibble(
+  poll_id = 1:1000,
+  pollster = sample(c("YouGov", "RMG Research", "CBS News", "Napolitan News", "Ipsos", "SurveyMonkey", "Other"), size = 1000, replace = TRUE),
+  numeric_grade = runif(1000, 2.7, 5.0), # Pollster quality between 2.7 and 5.0
+  pollscore = runif(1000, -1.5, 2.5),
+  candidate_name = sample(c("Kamala Harris", "Donald Trump"), size = 1000, replace = TRUE, prob = c(0.6, 0.4)),
+  pct = jitter(50 + rnorm(1000, mean = 0, sd = 5), amount = 2),
+  state = sample(c("Florida", "Pennsylvania", "Michigan", "Wisconsin", "Arizona", "Texas", NA), size = 1000, replace = TRUE),
+  date = as.Date('2024-07-01') + sample(0:90, 1000, replace = TRUE)
 )
-
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
-
-# Create a dataset by randomly assigning states and parties to divisions
-analysis_data <- tibble(
-  division = paste("Division", 1:151),  # Add "Division" to make it a character
[…]

### 2024-10-19 scripts/01-test_simulated_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated US presidential election polls data
+# Purpose: Tests the structure and validity of the simulated US presidential election polls data. In addition, plot the data to check the overall expected outcomes.
 # Author: Yunkyung Ko
-# Date: 12 October 2024
+# Date: 19 October 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
@@ -12,77 +12,99 @@
 
 #### Workspace setup ####
 library(tidyverse)
-
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+library(ggplot2)
+sim_harris_data <- read_csv("data/00-simulated_data/simulated_data.csv")
 
 # Test if the data was successfully loaded
-if (exists("analysis_data")) {
+if (exists("sim_harris_data")) {
   message("Test Passed: The dataset was successfully loaded.")
 } else {
   stop("Test Failed: The dataset could not be loaded.")
 }
 
 
-#### Test data ####
-
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+#### Test the structure and validity of simulated dataset ####
+# Check if the dataset has 540 rows
+if (nrow(sim_harris_data) == 540) {
+  message("Test Passed: The dataset has 540 rows.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: The dataset does not have 540 rows.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+# Check if the dataset has 8 columns
+if (ncol(sim_harris_data) == 8) {
+  message("Test Passed: The dataset has 8 columns.")
 } else {
-  stop("Test Failed: The dataset does not have 3 columns.")
+  stop("Test Failed: The dataset does not have 8 columns.")
 }
 
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
+# Check if the 'national' column contains valid binary values
+valid_NA <- c(0,1)
+if (all(sim_harris_data$national %in% valid_NA)) {
+  message("Test Passed: The 'national' column contains only valid binary values.")
 } else {
[…]

### 2024-10-19 scripts/03-clean_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw elections data and selects relevant variables for the analysis
 # Author: Yunkyung Ko
-# Date: 12 October 2024
+# Date: 13 October 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Need to have downloaded the data.

### 2024-10-20 scripts/03-clean_data.R
clean and filter dataset (harris_data)

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Cleans the raw elections data and selects relevant variables for the analysis
+# Purpose: Cleans the raw elections data and filters high-quality polls data
 # Author: Yunkyung Ko
-# Date: 13 October 2024
+# Date: 19 October 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Need to have downloaded the data.
@@ -11,13 +11,49 @@
 library(tidyverse)
 
 #### Clean data ####
+# Read in the data and clean variable names
 raw_elections_data <- read_csv("data/01-raw_data/raw_elections_data.csv")
-
-cleaned_data <-
+elections_data <-
   raw_elections_data |>
   janitor::clean_names() |>
-  select(pollster_id, pollster, methodology, transparency_score, state, race_id, sample_size, party, answer) |>
-  tidyr::drop_na()
+  select(pollster, numeric_grade, pollscore, state, end_date, candidate_name, sample_size, pct) |>
+  tidyr::drop_na() |>
+  mutate(
+    state = if_else(is.na(state), "National", state) # Handle missing states as "National"
+  ) |>
+  group_by(state) |>
+  mutate(
+    state_count = n()  # Count the number of polls per state
+  ) |>
+  ungroup() |>
+  mutate(
+    state = if_else(state_count < 60, "Other", state)  # Assign "Other" for states with fewer than 60 polls
+  ) |>
+  select(-state_count)  # Remove the 'state_count' column
+
+# Filter data to Harris estimates based on high-quality polls after she declared
+harris_data <- 
+  elections_data |>
+  filter(
+    candidate_name == "Kamala Harris",
+    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix. 
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state),
+    end_date = mdy(end_date)
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  mutate(
+    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
+  ) |>
+  group_by(pollster) |>
+  filter(n() > 35) |> # Filter for pollsters with more than 35 polls
+  ungroup() |>
+  select(-pct)  # Remove the 'state_count' and 'pct' columns
+
 
 #### Save data ####
[…]

### 2024-10-20 scripts/04-test_analysis_data.R

@@ -1,69 +1,75 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests the structure and validity of the US presidential election polls data.
+# Author: Yunkyung Ko
+# Date: 19 October 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Need to have downloaded and cleaned the data.
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
-
 
 #### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
+# Test that the dataset has 8 columns
+test_that("dataset has 8 columns", {
+  harris_data <- read_csv("../data/02-analysis_data/harris_elections_data.csv")
+  expect_equal(ncol(harris_data), 8)
 })
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
-
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
-
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
+# Test that the 'pollster' column is character type
+test_that("'pollster' is character", {
+  harris_data <- read_csv("../data/02-analysis_data/harris_elections_data.csv")
+  expect_type(harris_data$pollster, "character")
 })
 
 # Test that the 'state' column is character type
 test_that("'state' is character", {
-  expect_type(analysis_data$state, "character")
+  harris_data <- read_csv("../data/02-analysis_data/harris_elections_data.csv")
+  expect_type(harris_data$state, "character")
+})
+
+# Test that the 'candidate_name' column is character type
+test_that("'candidate_name' is character", {
[…]

### 2024-10-20 scripts/00-simulate_data.R
code linting and styling

@@ -23,16 +23,15 @@ sim_poll_data <- tibble(
   candidate_name = sample(c("Kamala Harris", "Donald Trump"), size = 1000, replace = TRUE, prob = c(0.6, 0.4)),
   pct = jitter(50 + rnorm(1000, mean = 0, sd = 5), amount = 2),
   state = sample(c("Florida", "Pennsylvania", "Michigan", "Wisconsin", "Arizona", "Texas", NA), size = 1000, replace = TRUE),
-  date = as.Date('2024-07-01') + sample(0:90, 1000, replace = TRUE)
+  date = as.Date("2024-07-01") + sample(0:90, 1000, replace = TRUE)
 )
 # Add 'national' binary variable
 sim_poll_data <- sim_poll_data %>%
-  mutate(national = if_else(is.na(state), 1, 0))  # 1 if state is NA, 0 otherwise
+  mutate(national = if_else(is.na(state), 1, 0)) # 1 if state is NA, 0 otherwise
 # Filter Harris data and keep only high-quality pollsters
 sim_harris_data <- sim_poll_data %>%
   filter(candidate_name == "Kamala Harris", numeric_grade >= 3.0) %>%
-  select(-state)  # Remove 'state' column
+  select(-state) # Remove 'state' column
 
 #### Save data ####
 write_csv(sim_harris_data, "data/00-simulated_data/simulated_data.csv")
-

### 2024-10-20 scripts/01-test_simulated_data.R

@@ -4,9 +4,9 @@
 # Date: 19 October 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
+# Pre-requisites:
+# - The `tidyverse` package must be installed and loaded
+# - 00-simulate_data.R must have been run
 # Any other information needed? Make sure you are in the `us_election` rproj
 
 
@@ -39,7 +39,7 @@ if (ncol(sim_harris_data) == 8) {
 }
 
 # Check if the 'national' column contains valid binary values
-valid_NA <- c(0,1)
+valid_NA <- c(0, 1)
 if (all(sim_harris_data$national %in% valid_NA)) {
   message("Test Passed: The 'national' column contains only valid binary values.")
 } else {
@@ -62,7 +62,7 @@ if (all(!is.na(sim_harris_data))) {
 }
 
 # Check if there are no empty strings in 'pollster' and 'candidate_name' columns
-if (all(sim_harris_data$pollster != "" & sim_harris_data$candidate_name != "" )) {
+if (all(sim_harris_data$pollster != "" & sim_harris_data$candidate_name != "")) {
   message("Test Passed: There are no empty strings in 'pollster' or 'candidate_name'.")
 } else {
   stop("Test Failed: There are empty strings in one or more columns.")
@@ -85,8 +85,10 @@ sim_harris_pollster_a <- # Plot by pollster
   geom_smooth() +
   theme_classic() +
   labs(y = "Harris Percent", x = "Date", title = "The Vote that Harris Received in the Poll, by Pollster.") +
-  scale_color_manual(values = c("YouGov" = "blue", "RMG Research" = "red", "CBS News" = "green",
-                                "Napolitan News" = "purple", "Ipsos" = "brown", "SurveyMonkey" = "yellow", "Other" = "grey")) +
+  scale_color_manual(values = c(
+    "YouGov" = "blue", "RMG Research" = "red", "CBS News" = "green",
+    "Napolitan News" = "purple", "Ipsos" = "brown", "SurveyMonkey" = "yellow", "Other" = "grey"
+  )) +
   theme(legend.position = "bottom")
 ggsave("data/00-simulated_data/test_plots/sim_harris_pollster.png", sim_harris_pollster_a, width = 8, height = 6, dpi = 300)
[…]

### 2024-10-20 scripts/02-download_data.R

@@ -11,9 +11,7 @@
 library(tidyverse)
 
 #### Download data ####
-raw_elections_data <-read_csv(file ="https://projects.fivethirtyeight.com/polls/data/president_polls.csv")
+raw_elections_data <- read_csv(file = "https://projects.fivethirtyeight.com/polls/data/president_polls.csv")
 
 #### Save data ####
 write_csv(raw_elections_data, "data/01-raw_data/raw_elections_data.csv")
-
-         

### 2024-10-20 scripts/03-clean_data.R

@@ -23,20 +23,20 @@ elections_data <-
   ) |>
   group_by(state) |>
   mutate(
-    state_count = n()  # Count the number of polls per state
+    state_count = n() # Count the number of polls per state
   ) |>
   ungroup() |>
   mutate(
-    state = if_else(state_count < 60, "Other", state)  # Assign "Other" for states with fewer than 60 polls
+    state = if_else(state_count < 60, "Other", state) # Assign "Other" for states with fewer than 60 polls
   ) |>
-  select(-state_count)  # Remove the 'state_count' column
+  select(-state_count) # Remove the 'state_count' column
 
 # Filter data to Harris estimates based on high-quality polls after she declared
-harris_data <- 
+harris_data <-
   elections_data |>
   filter(
     candidate_name == "Kamala Harris",
-    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix. 
+    numeric_grade >= 2.7 # Need to investigate this choice - come back and fix.
   ) |>
   mutate(
     state = if_else(is.na(state), "National", state),
@@ -49,11 +49,9 @@ harris_data <-
   group_by(pollster) |>
   filter(n() > 35) |> # Filter for pollsters with more than 35 polls
   ungroup() |>
-  select(-pct)  # Remove the 'state_count' and 'pct' columns
+  select(-pct) # Remove the 'state_count' and 'pct' columns
 
 
 #### Save data ####
 write_csv(elections_data, "data/02-analysis_data/elections_data.csv")
 write_csv(harris_data, "data/02-analysis_data/harris_elections_data.csv")
-
-

### 2024-10-20 scripts/04-test_analysis_data.R

@@ -45,9 +45,11 @@ test_that("no missing values in dataset", {
 })
 
 # Test that 'state' contains only valid US state or territory names
-valid_states <- c("Arizona", "California", "Florida", "Georgia", "Iowa", "Maryland", "Massachusetts",
-                  "Michigan", "Minesota", "Montana", "Nevada", "New Hampshire", "New Mexico", "New York",
-                  "North Carolina", "Ohio", "Pennsylvania", "Tennessee", "Texas", "Virginia", "Wisconsin", "Other")
+valid_states <- c(
+  "Arizona", "California", "Florida", "Georgia", "Iowa", "Maryland", "Massachusetts",
+  "Michigan", "Minesota", "Montana", "Nevada", "New Hampshire", "New Mexico", "New York",
+  "North Carolina", "Ohio", "Pennsylvania", "Tennessee", "Texas", "Virginia", "Wisconsin", "Other"
+)
 test_that("'state' contains valid US state names", {
   harris_data <- read_csv("../data/02-analysis_data/harris_elections_data.csv")
   expect_true(all(harris_data$state %in% valid_states))

### 2024-10-20 scripts/03-clean_data.R
add pct column

@@ -48,8 +48,7 @@ harris_data <-
   ) |>
   group_by(pollster) |>
   filter(n() > 35) |> # Filter for pollsters with more than 35 polls
-  ungroup() |>
-  select(-pct) # Remove the 'state_count' and 'pct' columns
+  ungroup()
 
 
 #### Save data ####

### 2024-10-20 scripts/04-test_analysis_data.R

@@ -14,10 +14,10 @@ library(testthat)
 
 
 #### Test data ####
-# Test that the dataset has 8 columns
-test_that("dataset has 8 columns", {
+# Test that the dataset has 9 columns
+test_that("dataset has 9 columns", {
   harris_data <- read_csv("../data/02-analysis_data/harris_elections_data.csv")
-  expect_equal(ncol(harris_data), 8)
+  expect_equal(ncol(harris_data), 9)
 })
 
 # Test that the 'pollster' column is character type

### 2024-10-20 scripts/05-exploratory_data_analysis.R
model design for model_date, model_date_pollster

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Yunkyung Ko
+# Date: 20 October 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
@@ -13,19 +13,9 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+harris_data <- read_csv("data/02-analysis_data/harris_elections_data.csv")
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
 
 
 #### Save model ####

### 2024-10-20 scripts/06-model_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Yunkyung Ko
+# Date: 20 October 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
@@ -13,25 +13,40 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+harris_data <- read_csv("data/02-analysis_data/harris_elections_data.csv")
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# Model 1: pct as a function of date
+model_date <- stan_glm(
+  formula = pct ~ end_date,
+  data = harris_data,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 853
+)
+
+# Model 2: pct as a function of date and pollster
+model_date_pollster <- stan_glm(
+  formula = pct ~ end_date + pollster,
+  data = harris_data,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 853
+)
 
 
 #### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  model_date,
+  file = "models/model_date.rds"
 )
 
+saveRDS(
+  model_date_pollster,
+  file = "models/model_date_pollster.rds"
+)
 

### 2024-10-20 paper/paper.qmd
bayesian model initiatives

@@
 library(tidyverse)
+
 ```
@@
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
 ```
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
 ```
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
 ```
@@
 
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-10-20 scripts/06-model_data.R
model initiatives

@@ -38,6 +38,41 @@ model_date_pollster <- stan_glm(
   seed = 853
 )
 
+# Bayesian Model 1: random intercept for pollster
+# Change 'pollster' and 'state' to factor variables
+harris_data <- harris_data |>
+  mutate(
+    pollster = factor(pollster),
+    state = factor(state)
+  )
+# Specify priors
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+model_formula_1 <- cbind(num_harris, sample_size - num_harris) ~ (1 | pollster)
+bayesian_model_1 <- stan_glmer(
+  formula = model_formula_1,
+  data = harris_data,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+
+# Bayesian Model 2: random intercepts for pollster and state
+model_formula_2 <- cbind(num_harris, sample_size - num_harris) ~ (1 | pollster) + (1 | state)
+
+bayesian_model_2 <- stan_glmer(
+  formula = model_formula_2,
+  data = harris_data,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
 
 #### Save model ####
 saveRDS(
@@ -50,3 +85,12 @@ saveRDS(
   file = "models/model_date_pollster.rds"
 )
 
+saveRDS(
+  bayesian_model_1,
+  file = "models/bayesian_model_1.rds"
+)
+
+saveRDS(
+  bayesian_model_2,
+  file = "models/bayesian_model_2.rds"
+)

### 2024-10-20 scripts/06-model_data.R
update script preamble

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
+# Purpose: Regression Models for predicting the percentage of supportive polls
 # Author: Yunkyung Ko
 # Date: 20 October 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Need to have downloaded the data.
+# Any other information needed? None
 
 
 #### Workspace setup ####

### 2024-10-21 paper/paper.qmd
update introduction & data overview first draft

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Kamala Harris Rises as the 2024 Election Front Runner: A Poll-of-Polls Forecast"
+subtitle: "How Poll Aggregation Reveals US Presidental Race Trends[!!!UPDATE IT TO MAIN RESULTS!!!]"
 author: 
@@
 library(tidyverse)
+library(janitor)
+library(lubridate)
+library(broom)
+library(modelsummary)
+library(rstanarm)
+library(splines)
+library(ggplot2)
+library(knitr)
+library(kableExtra)
 
@@
 
-Overview paragraph
+The US presidential election is an event that receives a lot of international attention due to its far-reaching implications beyond the country's borders.
+The effects of US elections are not only related to economy and international relations around the world, but also link to social and environmental issues such as climate change [@euromonitor2024].
+In the anticipation of the 2024 US electoral competition, this paper is aimed at predicting possible outcomes of the election by analyzing the level of support that Kamala Harris will gain.
 
-Estimand paragraph
+We forecast the support of Kamala Harris based on the polling results at the national and state levels, and apply a linear regression and a Bayesian approach.
+The main parameter of interest is the proportion of vote or support that Harris received in surveys, which is traced over time.
+By considering the effect of changes in poll-making organizations and geographical distinctions, our objective is to correct for variation across different voter bases with the pooling the polls approach [@poolingpolls].
 
-Results paragraph
+Our initial linear model examining the support for Harris over time suggests that the rate of support remained relatively stable around 47%, with no significant increase or decrease.
+However large variability was detected, as seen in the spread of points around the fitted line.
+Consequently, pollster-specific effects and state-level random effects were added to the model but resulted in even higher variations.
+Some pollsters or states consistently reported higher or lower support for Harris compared to others, proposing that [!!!!!MEANINGFUL RESULT ADDED LATER!!!!!].
 
-Why it matters paragraph
+These prognostics provide much more than merely forecasting the election in question.
+Hinting the trajectory the US might take in matters of foreign policy, economics, and global politics, predictions enable stakeholders worldwide to take preemptive measures for the resulted changes of a newly elected government [@csis2024].
+As such, this study not only contributes to the domestic political discourse but also provides a valuable tool for global actors seeking to navigate the uncertainty surrounding the 2024 US presidential election[@csis2024].
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The paper is structured as follows.
+Initially, @sec-data and @sec-model explores the data and methodology used, including filtering and modeling techniques applied to the polling data.
+Following that, @sec-results presents the results from the linear and Bayesian models, while the next section @sec-discussion discusses the broader implications of these findings.
+Finally, the paper concludes with remarks on future directions for research and applications of these models [@sec-discussion].
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+We use the statistical programming language R [@citeR] to analyze US presidential polling data from FiveThirtyEight [@fivethirtyeight2024], focusing on support for Kamala Harris.
+The dataset includes 15831 poll results from various national and state-level polls, with key variables such as pollster, sample size, percentage of support for Harris, and end date of the poll.
+Following the guidance of @tellingstories, we compiled the results of each opinion poll over a period of time and compared them taking into account the methodological peculiarities of polling by pollsters and geographical scope of the conducted polls.
 
-Overview text
+To ensure data quality, we filtered the dataset to include only polls that measured Kamala Harris' support, with a numeric grade of pollster 2.7 or higher for reliability.
+We also limited the analysis to polls conducted after July 21, 2024, when Harris officially declared her candidacy, and excluded pollsters with fewer than 35 polls to focus on those with sufficient data for robust results.
+
+In performing the analysis, we utilized several R packages.
+@tidyverse was used for data manipulation and visualization and @rstanarm, @modelsummary was respectively used for Bayesian modeling and generating model summaries.
+For visualizing results, @ggplot2 was used and @kableExtra helped format tables for presentation.
+These packages provided a framework for efficient data processing, modeling, and reporting.
 
@@
 ## Outcome variables
+
+### pct: the percentage of the vote or support that the candidate received in the poll
 
@@
 
+### end_date: the date the poll ended
+
+### pollster: the polling organization that conducted the poll
+
+mention that it was filtered by numeric_grade
+
+### pollscore: the score or reliability of the pollster in question
+
 Add graphs, tables and text.
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
 ```
@@
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
 ```
 
-# Discussion
+# Discussion {#sec-discussion}
 

### 2024-10-22 paper/paper.qmd
paper - data section first draft complete

@@
 library(tidyverse)
+library(readr)
 library(janitor)
@@
 library(kableExtra)
+library(gridExtra)
+library(here)
 
+raw_elections_data <- read_csv(here::here("data/01-raw_data/raw_elections_data.csv"))
+harris_data <- read_csv(here::here("data/02-analysis_data/harris_elections_data.csv"))
 ```
@@
 We use the statistical programming language R [@citeR] to analyze US presidential polling data from FiveThirtyEight [@fivethirtyeight2024], focusing on support for Kamala Harris.
-The dataset includes 15831 poll results from various national and state-level polls, with key variables such as pollster, sample size, percentage of support for Harris, and end date of the poll.
+The dataset includes a wide range of poll results from various national and state-level polls, with key variables such as pollster, sample size, percentage of support for Harris, and end date of the poll.
 Following the guidance of @tellingstories, we compiled the results of each opinion poll over a period of time and compared them taking into account the methodological peculiarities of polling by pollsters and geographical scope of the conducted polls.
@@
 To ensure data quality, we filtered the dataset to include only polls that measured Kamala Harris' support, with a numeric grade of pollster 2.7 or higher for reliability.
-We also limited the analysis to polls conducted after July 21, 2024, when Harris officially declared her candidacy, and excluded pollsters with fewer than 35 polls to focus on those with sufficient data for robust results.
+We also limited the analysis to polls conducted after July 21, 2024, when Harris officially declared her candidacy, and excluded pollsters with fewer than 30 polls to focus on those with sufficient data for robust results.
 
@@
 
+(Data last updated on [Day] [Month] 2024.) - [!!! MODIFY AT FINAL !!!]
+
 ## Measurement
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+[!!! MODIFY OBSV. NUMBERS AFTER DATA UPDATES !!!] - ERASE AT FINAL
+
+The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results (15971 observations, based on dataset available at Oct 21, 22:00).
+The polls conducted by various polling organizations capture voter preferences by taking a representative sample of the electorate and asking for the voters' candidate of choice.
+Surveys were conducted at the state and national levels, providing the wide perspective on public feelings across the country.
+
+Each poll represents a predictor of an actual event, namely voter opinion at a particular moment.
+Nevertheless, like all survey results, the raw data is susceptible to many potential limitations including the following: sampling error, variation in polling methods, distortion because of inappropriate survey responses such as missing data or response from respondents who misunderstood the questions [@tellingstories].
+
+While applying several filters to the original dataset such as restricting to those with a numeric grade of 2.7 or higher or pollsters with more than 30 polls improves data reliability, certain limitations still exist.
+Selection bias and sampling error remains as a concern, since polls always represent only part of the population.
+Differences in the way different organizations conducted their polling might introduce more inconsistencies.
+Finally, by focusing our attention on post-declaration polls only, we exclude earlier trends that could add more insight into how Harris' support has evolved over time.
 
@@
 
-### pct: the percentage of the vote or support that the candidate received in the poll
+### pct
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+The main variable of interest that we aim to forecast is the 'pct' variable, which represents the proportion of vote or support that a candidate received in the poll.
+@tbl-pctraw1 and @fig-pct2 respectively shows the summary statistics and distribution of the 'pct' variable in the original dataset [@fivethirtyeight2024].
+@tbl-pctfilter1 and @fig-pctfilter2 shows the summary statistics and distribution of the same variable, but in the filtered dataset that only comprises of the supporting votes for Harris from relatively high-quality polling organizations.
+Comparing the summary statistics for the raw data (@tbl-pctraw1) and filtered data (@tbl-pctfilter1), higher numbers were derived from data filtered only by Harris supporters.
+Also, @fig-pctfilter2 illustrates that a significant number of polls indicate support levels ranging from 40% to 50%, which suggests a stable yet not substantial endorsement.
+This proposes that Harris possesses a reliable foundational support, although her capacity to obtain a majority remains ambiguous.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: tbl-pctraw1
+#| tbl-cap: Summary Statistics for the 'pct' Variable (Raw Data)
 #| echo: false
+
+# Calculate summary statistics for the 'pct' variable
+summary_stats <- raw_elections_data %>%
+  summarize(
+    mean = mean(pct, na.rm = TRUE),
+    median = median(pct, na.rm = TRUE),
+    min = min(pct, na.rm = TRUE),
+    max = max(pct, na.rm = TRUE),
+    sd = sd(pct, na.rm = TRUE),
+    n = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+```{r}
+#| label: fig-pct2
+#| fig-cap: Distribution of the 'pct' Variable (Raw Data)
+#| echo: false
+
+# Plot the distribution of 'pct' using a histogram
+ggplot(raw_elections_data, aes(x = pct)) +
+  geom_histogram(binwidth = 5, fill = "blue", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Percentage of Support (pct)",
+       y = "Frequency")
+
+```
+```{r}
+#| label: tbl-pctfilter1
+#| tbl-cap: Summary Statistics for the 'pct' Variable (Filtered Data)
+#| echo: false
+
+# Calculate summary statistics for the 'pct' variable
+summary_stats <- harris_data %>%
+  summarize(
+    mean = mean(pct, na.rm = TRUE),
+    median = median(pct, na.rm = TRUE),
+    min = min(pct, na.rm = TRUE),
+    max = max(pct, na.rm = TRUE),
+    sd = sd(pct, na.rm = TRUE),
+    n = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+```{r}
+#| label: fig-pctfilter2
+#| fig-cap: Distribution of the 'pct' Variable (Filtered Data)
+#| echo: false
+
+# Plot the distribution of 'pct' using a histogram
+ggplot(harris_data, aes(x = pct)) +
+  geom_histogram(binwidth = 5, fill = "blue", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Percentage of Support (pct)",
+       y = "Frequency")
 
@@
 
-Talk more about it.
+## Predictor variables
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+### end_date
+
+[!!! MODIFY DATE RANGE AFTER DATA UPDATES !!!] - ERASE AT FINAL
+
+The date variable representing the time the poll was concluded helps was put into account to keep track of how support for a candidate changes in time.
+The reported end dates in the original dataset [@fivethirtyeight2024] ranges from 1 January, 2023 to 9 September, 2024 (@tbl-date).
+@fig-date shows that the polling data is more concentrated on survey results conducted in the recent period.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: tbl-date
+#| tbl-cap: Summary Statistics for the 'end_date' Variable (Raw Data)
 #| echo: false
-#| warning: false
-#| message: false
+
+summary_stats <- raw_elections_data %>%
+  summarize(
+    min = min(end_date, na.rm = TRUE),
+    max = max(end_date, na.rm = TRUE)
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+```{r}
+#| label: fig-date
+#| fig-cap: Distribution of the 'end_date' Variable (Raw Data)
+#| echo: false
+
+# Convert 'end_date' to a proper date format
+raw_elections_dataa <- raw_elections_data %>%
+  mutate(end_date = mdy(end_date))
+filtered_data <- raw_elections_dataa %>%
+  filter(end_date >= as.Date("2023-01-01"))
+
+# Distribution of 'end_date'
+ggplot(filtered_data, aes(x = end_date)) +
+  geom_histogram(fill = "grey", alpha = 0.7, color = "black", binwidth = 30) +  # Set binwidth explicitly
+  theme_minimal() +
+  labs(x = "End Date",
+       y = "Count")
 
@@
 
-Talk way more about it.
+[!!! MODIFY DATE RANGE AFTER DATA UPDATES !!!] - ERASE AT FINAL
 
-## Predictor variables
+When filtering the data, not only pollster quality and candidate type but also the date variable was considered.
+The filtered data contains only the polling data after the declaration of Harris.
+So, the date variable for filtered data ranges from 23 July, 2024 to 14 October, 2024 (@tbl-date2).
+@fig-date2 shows that overall, polling is conducted regularly but intensifies around specific dates.
 
-### end_date: the date the poll ended
+```{r}
+#| label: tbl-date2
+#| tbl-cap: Summary Statistics for the 'end_date' Variable (Filtered Data)
+#| echo: false
 
-### pollster: the polling organization that conducted the poll
+summary_stats <- harris_data %>%
+  summarize(
+    min = min(end_date, na.rm = TRUE),
+    max = max(end_date, na.rm = TRUE)
+  )
 
-mention that it was filtered by numeric_grade
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
 
-### pollscore: the score or reliability of the pollster in question
+```
 
-Add graphs, tables and text.
+```{r}
+#| label: fig-date2
+#| fig-cap: Distribution of the 'end_date' Variable (Filtered Data)
+#| echo: false
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+# Convert 'end_date' to a proper date format
+harris_dataa <- harris_data %>%
+  mutate(end_date = ymd(end_date))  # Change mdy() to ymd() or dmy() depending on the format
+
+# Distribution of 'end_date'
+ggplot(harris_dataa, aes(x = end_date)) +
+  geom_histogram(fill = "grey", alpha = 0.7, color = "black", binwidth = 10) +  # Set binwidth explicitly
+  theme_minimal() +
+  labs(x = "End Date",
+       y = "Count")
+
+```
+
+### pollster, state
+
+The 'pollster' and 'state' variable were selected to consider the effect of changes in poll-making organizations and geographical distinctions.
+The two variables respectively represent the polling organization that conducted the poll and the US state where the poll was conducted or focused.
+
+@tbl-predictors shows that the original dataset [@fivethirtyeight2024] contains 222 distinct pollsters and 54 distinct states.
+After filtering for high-quality polls and assigning 'other' for states with fewer than 60 polls, the analysis data contains 3 distinct poll-making organizations and 19 geographical distinctions as shown in @tbl-predictors.
+
+```{r}
+#| label: tbl-predictors
+#| tbl-cap: Number of Distinct Values for the 'pollster', 'state' Variable (Raw Data)
+#| echo: false
+
+summary_stats <- raw_elections_data %>%
+  summarize(
+    pollsters = n_distinct(pollster),
+    states = n_distinct(state)
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+
+```{r}
+#| label: tbl-predictors
+#| tbl-cap: Number of Distinct Values for the 'pollster', 'state' Variable (Filtered Data)
+#| echo: false
+
+summary_stats <- harris_data %>%
+  summarize(
+    pollsters = n_distinct(pollster),
+    states = n_distinct(state)
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+
+The distribution of polling counts for different pollsters in @fig-pollster suggests that the analysis data is dominated by a few pollsters, particularly Siena/NYT.
+Depending on their polling methodology, the general results may have potential biases.
+A detailed analysis of the polling methodology and possible errors of the organization will be covered in @sec-data-details.
+
+```{r}
+#| label: fig-pollster
+#| fig-cap: Distribution of the 'pollster' Variable (Filtered Data for Harris)
+#| echo: false
+
+# Distribution of 'pollster' (bar plot since it's categorical)
+ggplot(harris_data, aes(x = pollster)) +
+  geom_bar(fill = "purple", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Pollster",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```
+
+@fig-state displays the distribution of polls across different states in the analysis data.
+Pennsylvania, Arizona, and Georgia are the top 3 states with high number of polls while states like Montana, New Mexico, and Maryland have much fewer polls.
+Note that a significant number of national or unspecified state-level polls are aggregated in this analysis data regarding the high count in 'Other' category.
+The concentration of polls in certain states further suggests a strategic focus on areas likely to impact the election outcome [@11alive2024].
+
+```{r}
+#| label: fig-state
+#| fig-cap: Distribution of the 'state' Variable (Filtered Data for Harris)
+#| echo: false
+
+# Distribution of 'state' (bar plot)
+ggplot(harris_data, aes(x = reorder(state, -table(state)[state]))) +
+  geom_bar(fill = "black", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "State",
+       y = "Count") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+```
+
+### pollscore
+
+This variable was another factor we had put into consideration to check whether the validity of the polling questions affects the polling results.
+The 'pollscore' variable represents the score or reliability of the pollster in question.
+The numeric values are the error and bias that can be attributed to the pollster, which means negative numbers are better.
+@tbl-pollscore and @fig-pollscore suggests that while the majority of the polls are moderately to highly qualitative in the original dataset, a fraction of the polls with low-quality or no scores could add noise or uncertainty to the analysis.
+
+```{r}
+#| label: tbl-pollscore
+#| tbl-cap: Summary statistics for the 'pollscore' Variable (Raw Data)
+#| echo: false
+
+summary_stats <- raw_elections_data %>%
+  summarize(
+    mean = mean(pollscore, na.rm = TRUE),
+    median = median(pollscore, na.rm = TRUE),
+    min = min(pollscore, na.rm = TRUE),
+    max = max(pollscore, na.rm = TRUE),
+    sd = sd(pollscore, na.rm = TRUE),
+    n = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+
+```{r}
+#| label: fig-pollscore
+#| fig-cap: Distribution of the 'pollscore' Variable (Raw Data)
+#| echo: false
+
+# Distribution of 'pollscore' (bar plot)
+ggplot(raw_elections_data, aes(x = factor(pollscore))) +
+  geom_bar(fill = "skyblue", alpha = 0.7, color = "black") +
+  theme_minimal() +
+  labs(x = "Pollscore",
+       y = "Count")
+
+```
+
+After the filtering to polling data of high-quality polling organizations, we can find that the overall value and standard deviation of pollscores went down in @tbl-pollscore2.
+This implies that the polling data narrowed down to the responses from more reliable survey questions.
+@fig-pollscore2 also indicates that the data cleaning process effectively excluded less reliable sources, which can enhance the robustness of subsequent analyses.
+
+```{r}
+#| label: tbl-pollscore2
+#| tbl-cap: Summary statistics for the 'pollscore' Variable (Filtered Data)
+#| echo: false
+
+summary_stats <- harris_data %>%
+  summarize(
+    mean = mean(pollscore, na.rm = TRUE),
+    median = median(pollscore, na.rm = TRUE),
+    min = min(pollscore, na.rm = TRUE),
+    max = max(pollscore, na.rm = TRUE),
+    sd = sd(pollscore, na.rm = TRUE),
+    n = n()
+  )
+
+# Display the summary statistics in a nicely formatted table
+summary_stats %>%
+  kable() %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+
+```
+
+```{r}
+#| label: fig-pollscore2
+#| fig-cap: Distribution of the 'pollscore' Variable (Filtered Data)
+#| echo: false
+
+# Distribution of 'pollscore' (bar plot)
+ggplot(harris_data, aes(x = factor(pollscore))) +
+  geom_bar(fill = "skyblue", alpha = 0.7, color = "black", width = 0.5) +
+  theme_minimal() +
+  labs(x = "Pollscore",
+       y = "Count")
+
+```
 
@@
 
-# Appendix {-}
+# Appendix
 
-# Additional data details
+# Additional data details {#sec-data-details}
+
+## Pollster Methodology Overview and Evaluation
+
+## Idealized Methodolgy
+
+## Idealized Survey
 

### 2024-10-22 scripts/03-clean_data.R

@@ -47,7 +47,7 @@ harris_data <-
     num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
   ) |>
   group_by(pollster) |>
-  filter(n() > 35) |> # Filter for pollsters with more than 35 polls
+  filter(n() > 30) |> # Filter for pollsters with more than 35 polls
   ungroup()
 
 

### 2024-10-22 paper/paper.qmd
paper - model, appendix section first draft

@@
 title: "Kamala Harris Rises as the 2024 Election Front Runner: A Poll-of-Polls Forecast"
-subtitle: "How Poll Aggregation Reveals US Presidental Race Trends[!!!UPDATE IT TO MAIN RESULTS!!!]"
+subtitle: "How Poll Aggregation Reveals US Presidental Race Trends"
 author: 
@@
   summarize(
-    min = min(end_date, na.rm = TRUE),
-    max = max(end_date, na.rm = TRUE)
+    Min = min(end_date, na.rm = TRUE),
+    Max = max(end_date, na.rm = TRUE)
   )
@@
 
-@tbl-predictors shows that the original dataset [@fivethirtyeight2024] contains 222 distinct pollsters and 54 distinct states.
-After filtering for high-quality polls and assigning 'other' for states with fewer than 60 polls, the analysis data contains 3 distinct poll-making organizations and 19 geographical distinctions as shown in @tbl-predictors.
+@tbl-predictors1 shows that the original dataset [@fivethirtyeight2024] contains 222 distinct pollsters and 54 distinct states.
+After filtering for high-quality polls and assigning 'other' for states with fewer than 60 polls, the analysis data contains 3 distinct poll-making organizations and 19 geographical distinctions as shown in @tbl-predictors2.
 
 ```{r}
-#| label: tbl-predictors
+#| label: tbl-predictors1
 #| tbl-cap: Number of Distinct Values for the 'pollster', 'state' Variable (Raw Data)
@@
   summarize(
-    pollsters = n_distinct(pollster),
-    states = n_distinct(state)
+    Pollster = n_distinct(pollster),
+    State = n_distinct(state)
   )
@@
 ```{r}
-#| label: tbl-predictors
+#| label: tbl-predictors2
 #| tbl-cap: Number of Distinct Values for the 'pollster', 'state' Variable (Filtered Data)
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The goal of our modelling strategy is to estimate Kamala Harris's support percentage in the 2024 US election polls, accounting for potential variations over time, as well as across different pollsters and states.
+The model balances complexity with interpretability, incorporating both linear and Bayesian frameworks to capture patterns in the data.
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Define $y_i$ as the percentage of support that Kamala Harris receives in poll $i$.
+We begin with two simple linear models and progress to more complex Bayesian models that account for hierarchical structures.
 
+The following models outline our approach:
+
+### Linear Model by Date
+
+```{=latex}
 \begin{align}  
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i &= \beta_0 + \beta_1 \cdot \text{end\_date}_i + \epsilon_i \\
+\epsilon_i &\sim \text{Normal}(0, \sigma^2)
 \end{align}
+```
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+where:
+- $y_i$ is the percentage of support for Harris in poll $i$,
+- $\beta_0$ is the intercept,
+- $\beta_1$ represents the effect of the poll's end date,
+- $\epsilon_i$ is the error term.
+
+### Linear Model by Date and Pollster
+
+```{=latex}
+\begin{align}  
+y_i &= \beta_0 + \beta_1 \cdot \text{end\_date}_i + \gamma_{p[i]} + \epsilon_i \\
+\epsilon_i &\sim \text{Normal}(0, \sigma^2)
+\end{align}
+```
+
+where:
+- $\gamma_{p[i]}$ is a fixed effect for pollster $p$ conducting poll $i$ (e.g., Siena/NYT).
+
+### Bayesian Model with Random Intercept for Pollster
+
+```{=latex}
+\begin{align}  
+y_i | \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
+\mu_i &= \beta_0 + \beta_1 \cdot \text{end\_date}_i + \gamma_{p[i]} \\
+\gamma_p &\sim \text{Normal}(0, \sigma_{\gamma})
+\end{align}
+```
+
+where:
+- $\gamma_p$ is a random effect for pollster $p$.
+
+### Bayesian Model with Random Intercept for Pollster and State
+
+```{=latex}
+\begin{align}  
+y_i | \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
+\mu_i &= \beta_0 + \beta_1 \cdot \text{end\_date}_i + \gamma_{p[i]} + \delta_{s[i]} \\
+\gamma_p &\sim \text{Normal}(0, \sigma_{\gamma}) \\
+\delta_s &\sim \text{Normal}(0, \sigma_{\delta})
+\end{align}
+```
+
+where:
+- $\delta_{s[i]}$ is a random effect for state $s$.
+
+The Bayesian models are fit using `rstanarm` in R. The priors used are weakly informative:
+- $\beta_0 \sim \text{Normal}(0, 10)$
+- $\sigma \sim \text{Exponential}(1)$
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+Different pollsters and states induce variations in polling results, as pollsters may have distinct methodologies and states represent diverse voter bases.
+Incorporating random effects for both pollsters and states allows us to improve the robustness of the model.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+These models are run through the rstanarm package [@rstanarm] in R  [@citeR], which makes Bayesian modeling available through the use of Stan's strong inference engine.
+To validate the models, RMSE and WAIC have been made use of to check the goodness of fit; Bayesian models with reduced RMSE and WAIC outperform linear models.
+We use weakly informative priors; for example, $\beta_0 \sim \text{Normal}(0, 10)$ and $\sigma \sim \text{Exponential}(1)$.
+This reflects our initial uncertainty but prevents overfitting.
+The priors were chosen conservatively to ensure that the model remains consistent.
+
+Model diagnostics, including posterior predictive checks and convergence diagnostics, were carried out to ensure the reliability of the results.
+The Bayesian models converged successfully, as indicated by $\hat{R} = 1$ for all parameters.
+
+The main assumption in these models is that the pollster and state effects can be treated as random.
+This assumes that the effects are normally distributed across pollsters and states, which may not always be accurate.
+Additionally, the model assumes that polling data is representative of the actual electorate, an assumption that can be violated if polls are biased or have non-random sampling issues.
+Despite these limitations, the hierarchical structure allows us to capture important variability, making the model suitable for predicting Harris's support.
+Future improvements could involve incorporating time-varying effects or exploring interactions between pollsters and states.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+Our results are summarized in @tbl-modelresults1 and @tbl-modelresults2.
+[!!! ADD MORE WORDS !!!]
 
@@
 
+model_date <-
+  readRDS(file = here::here("models/model_date.rds"))
+
+model_date_pollster <-
+  readRDS(file = here::here("models/model_date_pollster.rds"))
+
+bayesian_model_1 <-
+  readRDS(file = here::here("models/bayesian_model_1.rds"))
+
+bayesian_model_2 <-
+  readRDS(file = here::here("models/bayesian_model_2.rds"))
+
 ```
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: tbl-modelresults1
+#| tbl-cap: "Linear models of support percentages for Harris based on date and pollster"
 #| warning: false
+
+modelsummary(
+  list(
+    "Linear by Date" = model_date,
+    "Linear by Date, Pollster" = model_date_pollster
+  ),
+  statistic = "mad",
+  fmt = 2)
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults2
+#| tbl-cap: "Bayesian models of support percentages for Harris based on pollster and state"
+#| warning: false
+
+modelsummary(
+  list(
+    "Bayesian with Pollster" = bayesian_model_1,
+    "Bayesian with Pollster, State" = bayesian_model_2
+  ),
+  statistic = "mad",
+  fmt = 2)
 
@@
 
-## First discussion point {#sec-first-point}
+## Stability in Support Levels for Harris {#sec-first-point}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## Pollster-Specific Variability {#sec-second-point}
 
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
+## State-Level Differences {#sec-third-point}
 
@@
 
-Weaknesses and next steps should also be included.
+Future studies could focus on the incorporation of time-varying effects to capture dynamic shifts in public opinion, especially at times of large campaign events.
+Interaction terms between pollster and state can be included in the model, since some polling organizations may be more effective or influential in specific regions.
+This can give further explanation to how pollsters and regional dynamics affect overall support for a candidate.
+Moreover, extending the model to incorporate voter demographics, such as age, gender, and education, could show which segments of the population are driving changes in support.
+
+While our model provides a foundation to understand Harris' polling support, it will take further refinements are necessary to enhance its accuracy and applicability.
+The combination of hierarchical modeling with Bayesian methods is effective at accounting for the heterogeneity across pollsters and states, yet there remain considerable directions of inquiry to explore at the intersection of polling methodology, voter behavior, and regional electoral dynamics.
 
@@
 
+The New York Times/Siena College polling partnership, the polling organization that accounted for the majority of polls in our analysis (@fig-pollster), conducts polls tailored for specific elections, such as state or national races [@nyt_siena_2020_methodology].
+Their sample size typically includes 600 to 1000 likely voters per poll, with oversampling in battleground states to capture regional nuances [@nyt_siena_2020_methodology].
+The methodology uses random-digit dialing (RDD) for landlines and mobile phones to ensure representative sample coverage across demographics.
+In addition, online surveys are administered to complement phone-based responses, ensuring broader accessibility [@nyt_siena_2020_methodology].
+The stratified random sampling approach is employed, where the population is divided into strata (based on demographic variables like race, education, and geography), and a random sample is drawn from each stratum [@tellingstories].
+This allows for precision in reflecting the political leanings and key demographic shifts in specific regions.
+
+The organization intends to enhance transparency in how public opinion is assessed, ensuring that questions are carefully designed to represent contemporary political discussions, and that the terminology is polished through a process of iterative testing to achieve clarity.
+They devote extensive resources to cognitive testing to ensure question wording reflects what the public thinks [@siena_scri_2024_poll].
+Their polling methodology stands out in that its strategic focus on using representative samples reflect political leanings and demographics of a region for more contextual and precise polling.
+Siena/NYT has its reputation for accurately predicting key battleground state outcomes during previous elections, such as Florida in 2016[@siena_nyt_perfect_partnership_2024].
+
+The limitations of Siena/NYT's methodology are the challenge of polling itself.
+Since polling is a "snapshot in time",  the results can fluctuate based on recent political events or campaign dynamics.
+Additionally, while the effort to represent a broad demographic is laudable, there are still issues with nonresponse bias in polling-particularly among the hard-to-reach voter or voters suspicious of polling organizations themselves [@pew2023polling].
+
 ## Idealized Methodolgy
 
+The proposed methodology for forecasting the 2024 U.S.
+presidential election with a budget of $100,000 would be designed as follows.
+First, a stratified random sampling method will be employed that allows for the capture of the demographic elements such as age, gender, race, and education level.
+This could alleviate bias and make the sample more representative of the population [@pew2023polling].
+The data collection process will encompass both telephone surveys and internet polling, thereby effectively engaging a diverse range of voters [@pew2023polling].
+Survey respondents would include older populations via conventional methods and younger, technologically intellectual individuals through digital platforms.
+This multifaceted approach enhances the reliability and inclusiveness of the gathered data.
+
+In addition to robust sampling, the methodology incorporates weighting methods that account for groups that are underrepresented, ensuring that the outcomes are not skewed by sampling errors [@pew2023polling].
+The final model would use Bayesian hierarchical modeling, which allows for more flexible modeling of uncertainty and variation across states, pollsters, and other external factors.
+These models, along with out-of-sample testing and cross-validation, enable accurate prediction sensitive to the dynamics of real-world changes, including political events [@pew2023polling].
+The inclusion of external factors, such as major political events, debates, or sudden economic shifts, would help the model remain responsive to rapid changes in voter sentiment.
+
 ## Idealized Survey
+
+The proposed survey questionnarie design is provided in the following link:
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In the first posterior predictive check (@fig-ppcheck-1), we compare the observed data with replicated data generated from the posterior distribution.
+This shows that the model is able to replicate the overall distribution of the observed data, with the replicated curves (light blue) closely following the true data (dark blue line).
+This indicates that the model fits the data well in terms of capturing the main pattern or trend [@stan2023posteriorchecks].
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In the second plot (@fig-ppcheck-2), the replicated data which had both the pollster and state variable as random intercepts, shows relatively closer approximation to the true data distribution.
+The narrowing of uncertainty in the posterior relative to the prior indicates the impact of the data on refining the model's predictions.
+This reassures that the model fits the data reasonably well and that the prior information has been appropriately updated by the observed data [@stan2023posteriorchecks].
 
@@
 #| warning: false
-#| label: fig-ppcheckandposteriorvsprior
+#| label: fig-ppcheck
 #| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+#| fig-cap: "Examining how the Bayesian model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check"]
 
+pp_check(bayesian_model_1) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+pp_check(bayesian_model_2) +
+  theme_classic() +
+  theme(legend.position = "bottom")
 ```
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+@fig-stanareyouokay-1 is a trace plot.
+The sampled values for posterior distribution of intercept parameter across iterations of the MCMC algorithm shows good convergence [@stan_mcmc_traces].
+The lines for the parameter appear to be stable and fluctuating around a central value without any clear trends or patterns.
+This suggests that the MCMC algorithm has likely converged, and the posterior samples are representative of the target distribution.
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-stanareyouokay-2 is a Rhat plot.
+The Rhat value is approximately 1.0 for the intercept, which shows that the variance within and between multiple chains have converged.
+An Rhat value close to 1 indicates that the chains have mixed well and are drawing from the same distribution while values significantly greater than 1 would indicate that further iterations are needed [@stan_mcmc_diagnostics].This suggests that the Bayesian models for both "pollster" and "state" have likely converged, and the results derived from these models are reliable.
 
@@
 
+plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+
+plot(bayesian_model_2, pars = "(Intercept)", prob = 0.95)
+
 ```

### 2024-10-22 paper/paper.qmd
first draft complete

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "In the anticipation of the 2024 US presidential election, this paper intends to analyze and predict the level of support that Kamala Harris will gain.
+The support polls for Harris remained relatively constant.
+However, there were some variations detected when accounting for pollster and geographical differences.
+Our objective is to propose a reliable electoral prediction considering those variations help global stakeholders take reasonable preemptive measures."
 format: pdf
@@
 Consequently, pollster-specific effects and state-level random effects were added to the model but resulted in even higher variations.
-Some pollsters or states consistently reported higher or lower support for Harris compared to others, proposing that [!!!!!MEANINGFUL RESULT ADDED LATER!!!!!].
+Some pollsters or states consistently reported higher or lower support for Harris compared to others, proposing the significance of considering distinct pollster and state environment when interpreting electoral polling results.
 
@@
 
+[!!! NEED MODIFICATION !!!]
+
 ## Stability in Support Levels for Harris {#sec-first-point}
+
+Kamala Harris' polling numbers show stability, a common trend for well-known candidates whose electorate solidifies in its opinion over time.
+Articles from The New York Times [@nyt2024harristrump] note that "once voters form opinions about candidates, those views rarely fluctuate significantly unless a major event occurs, such as a scandal or a high-profile policy shift.
+In the case of Harris, her polling numbers have been consistent, only changing slightly since she announced in 2024 (@fig-date2).
+This suggests that her supporters are more or less consistent and unlikely to change much as the election draws closer.
 
@@
 
+Various polling entities, including Siena/NYT and Quinnipiac, use distinct methodologies, resulting in potential systematic discrepancies within their outcomes.
+According to Michael Wines, the inherent biases of each pollster result in different interpretations of the elections prediction [@nyt2024pollproblems].
+For instance, Siena/NYT's live-interview method often skews slightly older and more conservative in its responses compared to online polling methods used by other organizations [@siena_scri_2024_poll].
+This variability makes clear that when interpreting for polling results, the source of the poll must be included at minimum.
+
 ## State-Level Differences {#sec-third-point}
+
+Certain states, particularly battleground states like Florida, Pennsylvania, and Georgia, show significantly different polling results compared to national trends due to regional issues and voter priorities [@politico2024battleground].
+These reflect state-specific issues, such as local economies, immigration, and access to healthcare.
+Surveys carried out within these states often show wider variation in candidate support, which makes state-level analysis necessary for more accurate predictions.
+Incorporating state-specific factors into electoral models allows a better representation of the diverse political landscapes seen across the United States, hence improving the precision of prediction in these key areas.
 

### 2024-10-22 paper/paper.qmd
idealized survey update

@@
 
-The proposed survey questionnarie design is provided in the following link:
+The proposed survey questionnarie design is provided in the following link: https://forms.gle/zQ8iJyPk3HhJYrKd9.
+
+::: {.layout}
+## Image Row
+
+::: {.column width="1/5"}
+![Survey Intro](../other/idealized_survey/survey_intro.png)
+:::
+
+::: {.column width="1/5"}
+![Survey Questions 1](../other/idealized_survey/survey_q1.png)
+:::
+
+::: {.column width="1/5"}
+![Survey Questions 2](../other/idealized_survey/survey_q2.png)
+:::
+
+::: {.column width="1/5"}
+![Survey Questions 3](../other/idealized_survey/survey_q3.png)
+:::
+
+::: {.column width="1/5"}
+![Survey Final Thanks](../other/idealized_survey/survey_outro.png)
+:::
+:::
 

### 2024-10-22 paper/paper.qmd
polling data updates to a more recent one

@@
 
-(Data last updated on [Day] [Month] 2024.) - [!!! MODIFY AT FINAL !!!]
+(Data last updated on 22 Oct 2024.) - [!!! MODIFY AT FINAL !!!]
 
@@
 
-The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results (15971 observations, based on dataset available at Oct 21, 22:00).
+The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results (16012 observations, based on dataset available at Oct 22, 13:00).
 The polls conducted by various polling organizations capture voter preferences by taking a representative sample of the electorate and asking for the voters' candidate of choice.

### 2024-10-24 paper/paper.qmd
R package reference & graph/table caption updates

@@
 In performing the analysis, we utilized several R packages.
-@tidyverse was used for data manipulation and visualization and @rstanarm, @modelsummary was respectively used for Bayesian modeling and generating model summaries.
-For visualizing results, @ggplot2 was used and @kableExtra helped format tables for presentation.
+`tidyverse` [@tidyverse] was used for data manipulation and visualization and `rstanarm` [@rstanarm], `modelsummary` [@modelsummary] was respectively used for Bayesian modeling and generating model summaries.
+For visualizing results, `ggplot2` [@ggplot2] was used and `kableExtra` [@kableExtra] helped format tables for presentation.
 These packages provided a framework for efficient data processing, modeling, and reporting.
@@
 #| label: tbl-pctraw1
-#| tbl-cap: Summary Statistics for the 'pct' Variable (Raw Data)
+#| tbl-cap: Summary statistics for the proportion(%) of support that candidates received in the poll
 #| echo: false
@@
   summarize(
-    mean = mean(pct, na.rm = TRUE),
-    median = median(pct, na.rm = TRUE),
-    min = min(pct, na.rm = TRUE),
-    max = max(pct, na.rm = TRUE),
-    sd = sd(pct, na.rm = TRUE),
+    mean = round(mean(pct, na.rm = TRUE), 2),
+    median = round(median(pct, na.rm = TRUE),2),
+    min = round(min(pct, na.rm = TRUE), 2),
+    max = round(max(pct, na.rm = TRUE), 2),
+    sd = round(sd(pct, na.rm = TRUE), 2),
     n = n()
@@
 #| label: fig-pct2
-#| fig-cap: Distribution of the 'pct' Variable (Raw Data)
+#| fig-cap: Distribution of the proportion(%) of support that candidates received in the poll
 #| echo: false
@@
 #| label: tbl-pctfilter1
-#| tbl-cap: Summary Statistics for the 'pct' Variable (Filtered Data)
+#| tbl-cap: Summary statistics for the proportion(%) of support that Harris received in high-quality polls
 #| echo: false
@@
   summarize(
-    mean = mean(pct, na.rm = TRUE),
-    median = median(pct, na.rm = TRUE),
-    min = min(pct, na.rm = TRUE),
-    max = max(pct, na.rm = TRUE),
-    sd = sd(pct, na.rm = TRUE),
+    mean = round(mean(pct, na.rm = TRUE), 2),
+    median = round(median(pct, na.rm = TRUE), 2),
+    min = round(min(pct, na.rm = TRUE), 2),
+    max = round(max(pct, na.rm = TRUE), 2),
+    sd = round(sd(pct, na.rm = TRUE), 2),
     n = n()
@@
 #| label: fig-pctfilter2
-#| fig-cap: Distribution of the 'pct' Variable (Filtered Data)
+#| fig-cap: Distribution of the proportion(%) of support that Harris received in high-quality polls
 #| echo: false
@@
 #| label: tbl-date
-#| tbl-cap: Summary Statistics for the 'end_date' Variable (Raw Data)
+#| tbl-cap: Summary statistics for the date the poll was concluded
 #| echo: false
@@
 #| label: fig-date
-#| fig-cap: Distribution of the 'end_date' Variable (Raw Data)
+#| fig-cap: Distribution of the date the poll was concluded
 #| echo: false
@@
 #| label: tbl-date2
-#| tbl-cap: Summary Statistics for the 'end_date' Variable (Filtered Data)
+#| tbl-cap: Summary statistics for the date the high-quality polls for Harris was concluded
 #| echo: false
@@
 #| label: fig-date2
-#| fig-cap: Distribution of the 'end_date' Variable (Filtered Data)
+#| fig-cap: Distribution of the date the high-quality polls for Harris was concluded
 #| echo: false
@@
 #| label: tbl-predictors1
-#| tbl-cap: Number of Distinct Values for the 'pollster', 'state' Variable (Raw Data)
+#| tbl-cap: Number of distinct polling organizations and US states where the poll was conducted
 #| echo: false
@@
 #| label: tbl-predictors2
-#| tbl-cap: Number of Distinct Values for the 'pollster', 'state' Variable (Filtered Data)
+#| tbl-cap: Number of distinct high-quality polling organizations and US states where more than 60 polls for Harris were conducted
 #| echo: false
@@
 #| label: fig-pollster
-#| fig-cap: Distribution of the 'pollster' Variable (Filtered Data for Harris)
+#| fig-cap: Distribution of polling organizations where high-quality polls for Harris were conducted
 #| echo: false
@@
 #| label: fig-state
-#| fig-cap: Distribution of the 'state' Variable (Filtered Data for Harris)
+#| fig-cap: Distribution of US states where the more than 60 polls for Harris were conducted
 #| echo: false
@@
 #| label: tbl-pollscore
-#| tbl-cap: Summary statistics for the 'pollscore' Variable (Raw Data)
+#| tbl-cap: Summary statistics for the reliability scores of pollsters
 #| echo: false
@@
   summarize(
-    mean = mean(pollscore, na.rm = TRUE),
-    median = median(pollscore, na.rm = TRUE),
-    min = min(pollscore, na.rm = TRUE),
-    max = max(pollscore, na.rm = TRUE),
-    sd = sd(pollscore, na.rm = TRUE),
+    mean = round(mean(pollscore, na.rm = TRUE),2),
+    median = round(median(pollscore, na.rm = TRUE),2),
+    min = round(min(pollscore, na.rm = TRUE),2),
+    max = round(max(pollscore, na.rm = TRUE),2),
+    sd = round(sd(pollscore, na.rm = TRUE),2),
     n = n()
@@
 #| label: fig-pollscore
-#| fig-cap: Distribution of the 'pollscore' Variable (Raw Data)
+#| fig-cap: Distribution of the reliability scores of pollsters
 #| echo: false
 
-# Distribution of 'pollscore' (bar plot)
+# Distribution of 'pollscore' (bar plot) with simplified x-axis labels
 ggplot(raw_elections_data, aes(x = factor(pollscore))) +
@@
   theme_minimal() +
+  scale_x_discrete(breaks = c(-1.5, -1, -0.5, 0, 0.5, 1, 1.5)) +  # Simplified labels at 0.5 intervals
   labs(x = "Pollscore",
@@
 #| label: tbl-pollscore2
-#| tbl-cap: Summary statistics for the 'pollscore' Variable (Filtered Data)
+#| tbl-cap: Summary statistics for the reliability scores of high-quality pollsters used for analysis
 #| echo: false
@@
 #| label: fig-pollscore2
-#| fig-cap: Distribution of the 'pollscore' Variable (Filtered Data)
+#| fig-cap: Distribution of the reliability scores of high-quality pollsters used for analysis
 #| echo: false
@@
 
-The proposed survey questionnarie design is provided in the following link: https://forms.gle/zQ8iJyPk3HhJYrKd9.
+The proposed survey questionnarie design is provided in the following link:
+https://forms.gle/zQ8iJyPk3HhJYrKd9.
 
 ::: {.layout}
-## Image Row
+## Survey Demo
 

### 2024-10-24 paper/paper.qmd
recent polling data update

@@
 
-(Data last updated on 22 Oct 2024.) - [!!! MODIFY AT FINAL !!!]
+(Data last updated on 24 Oct 2024.) - [!!! MODIFY AT FINAL !!!]
 
@@
 
-The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results (16012 observations, based on dataset available at Oct 22, 13:00).
+The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results (16264 observations, based on dataset available at Oct 24, 13:00).
 The polls conducted by various polling organizations capture voter preferences by taking a representative sample of the electorate and asking for the voters' candidate of choice.

### 2024-10-24 paper/paper.qmd
abstract & title modifications

@@
 ---
-title: "Kamala Harris Rises as the 2024 Election Front Runner: A Poll-of-Polls Forecast"
-subtitle: "How Poll Aggregation Reveals US Presidental Race Trends"
+title: "A Poll-of-Polls Forecast for the 2024 U.S. Presidential Election: Kamala Harris Emerges as a Leading Candidate with Constant Support Rates Hovering Around 50%"
+subtitle: "Multiple Linear Regression and Bayesian Models Accounting for Pollster and Geographical Disparities in Support Trends"
 author: 
@@
 abstract: "In the anticipation of the 2024 US presidential election, this paper intends to analyze and predict the level of support that Kamala Harris will gain.
-The support polls for Harris remained relatively constant.
-However, there were some variations detected when accounting for pollster and geographical differences.
-Our objective is to propose a reliable electoral prediction considering those variations help global stakeholders take reasonable preemptive measures."
+The support polls for Harris remained relatively constant around 47%.
+However, there were some variations detected when accounting for pollster and geographical differences, with Siena/NYT and Texas standing out with its large fluctuation in support rates.
+Our objective is to propose a reliable electoral prediction considering those variations help global stakeholders take reasonable preemptive measures.
+[!!!
+NEED MODIFICATIONS FOR THE MAIN RESULTS PART !!!]"
 format: pdf

### 2024-10-24 scripts/03-clean_data.R

@@ -31,6 +31,10 @@ elections_data <-
   ) |>
   select(-state_count) # Remove the 'state_count' column
 
+# Remove outliers (pct values greater than 100 or less than 0)
+elections_data <- elections_data |>
+  filter(pct >= 0 & pct <= 100)
+
 # Filter data to Harris estimates based on high-quality polls after she declared
 harris_data <-
   elections_data |>
@@ -48,7 +52,8 @@ harris_data <-
   ) |>
   group_by(pollster) |>
   filter(n() > 30) |> # Filter for pollsters with more than 35 polls
-  ungroup()
+  ungroup()  |>
+  filter(pct >= 0 & pct <= 100)
 
 
 #### Save data ####

### 2024-10-24 scripts/03-clean_data.R
script for making parquet files for analysis data

@@ -9,6 +9,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
 #### Clean data ####
 # Read in the data and clean variable names
@@ -59,3 +60,6 @@ harris_data <-
 #### Save data ####
 write_csv(elections_data, "data/02-analysis_data/elections_data.csv")
 write_csv(harris_data, "data/02-analysis_data/harris_elections_data.csv")
+
+write_parquet(elections_data, "data/02-analysis_data/elections_data.parquet")
+write_parquet(harris_data, "data/02-analysis_data/harris_elections_data.parquet")

### 2024-10-25 paper/paper.qmd
update 'correlation between predictor variables'

@@
 library(here)
+library(usmap)  # For the map of US states
+library(viridis)  # For a color palette
 
@@
 
-## Outcome variables
+## Outcome variable
 
-### pct
+### The Proportion(%) of Support that a Candidate Received in the Poll
 
@@
 
-### end_date
+### The Date the Poll was Concluded
 
@@
 
-The date variable representing the time the poll was concluded helps was put into account to keep track of how support for a candidate changes in time.
+The 'end_date' variable representing the time the poll was concluded was put into account to keep track of how support for a candidate changes in time.
 The reported end dates in the original dataset [@fivethirtyeight2024] ranges from 1 January, 2023 to 9 September, 2024 (@tbl-date).
@@
 
-### pollster, state
+### Pollster and State
 
@@
 
-### pollscore
+### Pollscore Measuring the Validity of Polling Questions
 
@@
        y = "Count")
+
+```
+
+## Correlation Between Predictor Variables
+
+By examining the relationships between predictor variables in the analysis dataset, we aim to identify potential biases in interpretation and model instability.
+This is expected to ensure a transparent understanding of the model's robustness and interpretability.
+
+### End Date and State
+
+@fig-correlation1 shows notable polling concentrations on certain states such as Pennsylvania, Ohio, and Nevada in week 2024-37 (September 9, 2024) or 2024-38 (September 15, 2024).
+This suggests that these states are key battelgrounds or areas of strategic focus during the election period.
+Moreover, polling activities are not consistent across all weeks and there are clear peaks in polling activity in week 2024-38 (September 15, 2024), which may correspond to significant political events, debates, or media focuses.
+
+```{r}
+#| label: fig-correlation1
+#| fig-cap: "Polling concentration over time by state *Note:* Strategic focus of polling is directed in battleground states such as Pennsylvania and Ohio in September 15, 2024"
+#| echo: false
+
+# Aggregate the number of polls by state and by week (or month)
+poll_time_state_2 <- harris_data %>%
+  mutate(week = format(as.Date(end_date), "%Y-%U")) %>%  # Group by week of the year
+  group_by(state, week) %>%
+  summarise(total_polls = n(), .groups = "drop")  # Adding .groups = "drop"
+
+# Plot a heatmap of polling activity by state and week
+ggplot(poll_time_state_2, aes(x = week, y = state, fill = total_polls)) +
+  geom_tile(color = "white") +
+  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey90") +
+  labs(x = "Week",
+       y = "State",
+       fill = "Total Polls") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+### End Date and Pollster
+
+@fig-correlation2 shows that Siena/NYT, which made up the largest proportion of polls in our analysis dataset, has concentrated polling efforts in week 2024-37 (September 9, 2024) and week 2024-38 (September 15, 2024).
+Notably active periods, possibly around key election events, might introduce temporal bias and overemphasize methodologies of Siena/NYT in that specific time period.
+Emerson shows overall consistency of polling activities across several weeks indicating steady involvement, while other polling organizations show more sporadic or minimal involvement.
+
+```{r}
+#| label: fig-correlation2
+#| fig-cap: "Polling concentration over time by pollster *Note:* Concentrated polling efforts of Siena/NYT in the first half of September, 2024 show temporal bias and potential overemphasizing of their polling methodologies around key election events."
+#| echo: false
+
+# Aggregate the number of polls by state and by week (or month)
+poll_time_pollster <- harris_data %>%
+  mutate(week = format(as.Date(end_date), "%Y-%U")) %>%  # Group by week of the year
+  group_by(pollster, week) %>%
+  summarise(total_polls = n(), .groups = "drop")  # Adding .groups = "drop"
+
+# Plot a heatmap of polling activity by state and week
+ggplot(poll_time_pollster, aes(x = week, y = pollster, fill = total_polls)) +
+  geom_tile(color = "white") +
+  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey90") +
+  labs(x = "Week",
+       y = "Pollster",
+       fill = "Total Polls") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+### Pollster and State
+
+@fig-correlation3 shows that Siena/NYT has the strongest state-level presence in key battleground states like Michigan and Arizona.
+Others like AtlasIntel and Beacon/Shaw has minimal polling coverage in targeted states such as Nevada and Georgia.
+This vairation in coverage could influence the overall analysis in that some states might receive disproportionately more attention from certain pollsters.
+
+```{r}
+#| label: fig-correlation3
+#| fig-cap: "Polling activity by pollster and state *Note:* Siena/NYT has strongest state-level polling coverage in key battleground states"
+#| echo: false
+
+# Step 1: Filter out NAs from pollster column and summarize data by state and pollster
+pollster_state_data <- harris_data %>%
+  filter(!is.na(pollster)) %>%  # Remove NAs from pollster
+  group_by(state, pollster) %>%
+  summarise(total_polls = n(), .groups = "drop")  # Summarize number of polls
+
+# Step 2: Plot the faceted map using plot_usmap
+plot_usmap(data = pollster_state_data, values = "total_polls", regions = "states") +
+  facet_wrap(~ pollster, ncol = 2) +  # Create a map for each pollster
+  scale_fill_distiller(palette = "Blues", direction = 1, na.value = "gray90") +  # Use blue color palette
+  labs(fill = "Total Polls") +
+  theme_void() +  # Remove axes and gridlines
+  theme(strip.text = element_text(size = 12))  # Customize facet labels
 

### 2024-10-25 scripts/03-clean_data.R

@@ -52,7 +52,7 @@ harris_data <-
     num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
   ) |>
   group_by(pollster) |>
-  filter(n() > 30) |> # Filter for pollsters with more than 35 polls
+  filter(n() > 20) |> # Filter for pollsters with more than 20 polls
   ungroup()  |>
   filter(pct >= 0 & pct <= 100)
 

### 2024-10-29 paper/paper.qmd
results graph updates

@@
 
-The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results (16264 observations, based on dataset available at Oct 24, 13:00).
+The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results (16506 observations, based on dataset available at Oct 29).
 The polls conducted by various polling organizations capture voter preferences by taking a representative sample of the electorate and asking for the voters' candidate of choice.
@@
 
-Our results are summarized in @tbl-modelresults1 and @tbl-modelresults2.
+## Results from the relation between the prediction variable and outcome variable of the analysis data
+
+@fig-resultplot1 shows an initial increase in support following Harris' announcement, peaking around mid-September.
+However, this initial momentum stabilized with minor decrease in support as the election gets closer.
+Moreover, the scattered data points around the summary line show high variance between individual polls.
+
+```{r}
+#| label: fig-resultplot1
+#| fig-cap: "Polling votes for Harris over time *Note:* The date starts from after Harris officially announced her campaign for 2024 U.S. presidential election on July 21, 2024."
+#| echo: false
+base_plot <- ggplot(harris_data, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date")
+
+# Plots poll estimates and overall smoothing
+base_plot +
+  geom_point() +
+  geom_smooth()
+```
+
+When @fig-resultplot2
+
+```{r}
+#| label: fig-resultplot2
+#| fig-cap: "Polling votes for Harris over time by pollster *Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
+presidential election on July 21, 2024.
+Only filtered pollsters with high-quality polling questions are shown in the following."
+#| echo: false
+# Color by pollster
+base_plot +
+  geom_point(aes(color = pollster)) +
+  geom_smooth() + 
+  theme(legend.position = "bottom")
+```
+
+```{r}
+#| label: fig-resultplot3
+#| fig-cap: "Polling votes for Harris over time by pollster (facets) *Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
+presidential election on July 21, 2024.
+Only filtered pollsters with high-quality polling questions are shown in the following."
+#| echo: false
+# Facet by pollster
+base_plot +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x)) +  # Use GAM smoothing to handle complex data
+  facet_wrap(vars(pollster))
+```
+
+## Results from the prediction model
+
+Prediction results derived from our model frameworks are summarized in @tbl-modelresults1 and @tbl-modelresults2.
 [!!! ADD MORE WORDS !!!]
@@
 
-## Stability in Support Levels for Harris {#sec-first-point}
+## Why Harris could beat her polls {#sec-first-point}
 
@@
 
-## Pollster-Specific Variability {#sec-second-point}
+## Pollsters herding around false consensus {#sec-second-point}
+
+The polling averages for Harris and Trump don't seem to show a big difference.
 

### 2024-10-31 paper/paper.qmd
discussion and abstract refinement

@@
 ---
-title: "A Poll-of-Polls Forecast for the 2024 U.S. Presidential Election: Kamala Harris Emerges as a Leading Candidate with Constant Support Rates Hovering Around 50%"
-subtitle: "Multiple Linear Regression and Bayesian Models Accounting for Pollster and Geographical Disparities in Support Trends"
+title: "A Poll-of-Polls Forecast for the 2024 U.S. Presidential Election: Analysis of Support Trends for Harris Accounting for Pollster and State Disparities"
+subtitle: "Kamla Harris Emerges as a Leading Candidate with Constant Support Rates Hovering Around 50.03%"
 author: 
@@
 date-format: long
-abstract: "In the anticipation of the 2024 US presidential election, this paper intends to analyze and predict the level of support that Kamala Harris will gain.
-The support polls for Harris remained relatively constant around 47%.
-However, there were some variations detected when accounting for pollster and geographical differences, with Siena/NYT and Texas standing out with its large fluctuation in support rates.
-Our objective is to propose a reliable electoral prediction considering those variations help global stakeholders take reasonable preemptive measures.
-[!!!
-NEED MODIFICATIONS FOR THE MAIN RESULTS PART !!!]"
+abstract: "In the anticipation of the 2024 US presidential election, this paper intends to provide a trend analysis of the support rates for Harris accounting for pollster and state specific factors.
+Supportive polls for Harris remained relatively constant around 50.03%.
+However, some polling organizations and battleground states such as Siena/NYT and Michigan, showed variations from the average national results.
+Our objective is to propose a reliable electoral prediction ny not only analyzing the summary statistics and correlations of date, pollster, and state variables, but also by constructing prediction models using a linear and a Bayesian model approach."
 format: pdf
@@
 
-Our initial linear model examining the support for Harris over time suggests that the rate of support remained relatively stable around 47%, with no significant increase or decrease.
+Our initial linear model examining the support for Harris over time suggests that the rate of support remained relatively stable around 50.03%, with no significant increase or decrease.
 However large variability was detected, as seen in the spread of points around the fitted line.
@@
 
-We use the statistical programming language R [@citeR] to analyze US presidential polling data from FiveThirtyEight [@fivethirtyeight2024], focusing on support for Kamala Harris.
+We use the statistical programming language R [@citeR] to analyze US presidential polling data from FiveThirtyEight [@fivethirtyeight2024], focusing on support for Kamala Harris (Note that the data was last updated on 30 Oct.
+2024).
 The dataset includes a wide range of poll results from various national and state-level polls, with key variables such as pollster, sample size, percentage of support for Harris, and end date of the poll.
@@
 
-(Data last updated on 24 Oct 2024.) - [!!! MODIFY AT FINAL !!!]
+(Data last updated on 30 Oct 2024.) - [!!! MODIFY AT FINAL !!!]
 
@@
 
-The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results (16506 observations, based on dataset available at Oct 29).
+The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results (16817 observations, based on dataset available at Oct 30).
 The polls conducted by various polling organizations capture voter preferences by taking a representative sample of the electorate and asking for the voters' candidate of choice.
@@
 The 'end_date' variable representing the time the poll was concluded was put into account to keep track of how support for a candidate changes in time.
-The reported end dates in the original dataset [@fivethirtyeight2024] ranges from 1 January, 2023 to 9 September, 2024 (@tbl-date).
+The reported end dates in the original dataset [@fivethirtyeight2024] ranges from 27 April, 2021 to 29 October, 2024 (@tbl-date).
 @fig-date shows that the polling data is more concentrated on survey results conducted in the recent period.
@@
 
-summary_stats <- raw_elections_data %>%
+# Convert 'end_date' to a proper date format
+raw_elections_dataa <- raw_elections_data %>%
+  mutate(end_date = mdy(end_date))
+filtered_data <- raw_elections_dataa %>%
+  filter(end_date >= as.Date("2023-01-01"))
+
+summary_stats <- raw_elections_dataa %>%
   summarize(
@@
 
-# Convert 'end_date' to a proper date format
-raw_elections_dataa <- raw_elections_data %>%
-  mutate(end_date = mdy(end_date))
-filtered_data <- raw_elections_dataa %>%
-  filter(end_date >= as.Date("2023-01-01"))
-
 # Distribution of 'end_date'
@@
 The filtered data contains only the polling data after the declaration of Harris.
-So, the date variable for filtered data ranges from 23 July, 2024 to 14 October, 2024 (@tbl-date2).
+So, the date variable for filtered data ranges from 23 July, 2024 to 29 October, 2024 (@tbl-date2).
 @fig-date2 shows that overall, polling is conducted regularly but intensifies around specific dates.
@@
 
-@tbl-predictors1 shows that the original dataset [@fivethirtyeight2024] contains 222 distinct pollsters and 54 distinct states.
-After filtering for high-quality polls and assigning 'other' for states with fewer than 60 polls, the analysis data contains 3 distinct poll-making organizations and 19 geographical distinctions as shown in @tbl-predictors2.
+@tbl-predictors1 shows that the original dataset [@fivethirtyeight2024] contains 232 distinct pollsters and 56 distinct states.
+After filtering for high-quality polls and assigning 'other' for states with fewer than 60 polls, the analysis data contains 8 distinct poll-making organizations and 20 geographical distinctions as shown in @tbl-predictors2.
 
@@
 Others like AtlasIntel and Beacon/Shaw has minimal polling coverage in targeted states such as Nevada and Georgia.
-This vairation in coverage could influence the overall analysis in that some states might receive disproportionately more attention from certain pollsters.
+This variation in coverage could influence the overall analysis in that some states might receive disproportionately more attention from certain pollsters.
 
@@
 
-@fig-resultplot1 shows an initial increase in support following Harris' announcement, peaking around mid-September.
+@fig-resultplot1 shows an initial increase in support for Harris since her declaration to join the electoral race, peaking around mid-September.
 However, this initial momentum stabilized with minor decrease in support as the election gets closer.
@@
 #| echo: false
+#| warning: false
+
 base_plot <- ggplot(harris_data, aes(x = end_date, y = pct)) +
@@
 
-When @fig-resultplot2
+When accounting for the pollster difference, @fig-resultplot2 and @fig-resultplot3 show variations in Harris' support levels that are obscured in the aggregated view of @fig-resultplot1.
+Siena/NYT (pink) shows relatively stable polling results for Harris, with fewer extreme values and a slight downward trend in October.
+On the other hand, Emerson (greem) shows more fluctuation throughout the period with a wider spread of polling values.
 
@@
 #| echo: false
+#| warning: false
+
 # Color by pollster
@@
 #| echo: false
+#| warning: false
+
 # Facet by pollster
@@
 
+When accounting for the state difference, @fig-resultplot4 and @fig-resultplot5 show variations in Harris' support levels.
+Critical battleground states like Pennsylvania and Michigan show distinct trends from the national analysis in @fig-resultplot1.
+Pennsylvania shows relatively stable trends slightly above the national average while Michigan shows fluctuations over time.
+
+```{r}
+#| label: fig-resultplot4
+#| fig-cap: "Polling votes for Harris over time by state *Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
+presidential election on July 21, 2024.
+Only filtered pollsters with high-quality polling questions are shown in the following."
+#| echo: false
+# Color by pollster
+#| warning: false
+
+base_plot +
+  geom_point(aes(color = state)) +
+  geom_smooth() + 
+  theme(legend.position = "bottom")
+```
+
+```{r}
+#| label: fig-resultplot5
+#| fig-cap: "Polling votes for Harris over time by state (facets) *Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
+presidential election on July 21, 2024.
+Only filtered pollsters with high-quality polling questions are shown in the following."
+#| echo: false
+#| warning: false
+
+# Facet by state
+base_plot +
+  geom_point() +
+  geom_smooth(method = "gam", formula = y ~ s(x)) +  # Use GAM smoothing to handle complex data
+  facet_wrap(vars(state))
+```
+
 ## Results from the prediction model
 
-Prediction results derived from our model frameworks are summarized in @tbl-modelresults1 and @tbl-modelresults2.
-[!!! ADD MORE WORDS !!!]
+Prediction results derived from our model frameworks are summarized in @tbl-modelresults1 and @tbl-modelresults2. In the linear models (@tbl-modelresults1)
+, higher $R^2$ and lower WAIC, RMSE values in the second column show that the inclusion of pollster effects improves accuracy and reliability of the prediction model.
+In the Bayesian models (@tbl-modelresults2), higher Log-Likelihood and lower ELPD, WAIC values in the fourth column shows that including both pollster and state effects enhances model performance.
 
@@
 
-[!!! NEED MODIFICATION !!!]
-
 ## Why Harris could beat her polls {#sec-first-point}
 
-Kamala Harris' polling numbers show stability, a common trend for well-known candidates whose electorate solidifies in its opinion over time.
-Articles from The New York Times [@nyt2024harristrump] note that "once voters form opinions about candidates, those views rarely fluctuate significantly unless a major event occurs, such as a scandal or a high-profile policy shift.
-In the case of Harris, her polling numbers have been consistent, only changing slightly since she announced in 2024 (@fig-date2).
-This suggests that her supporters are more or less consistent and unlikely to change much as the election draws closer.
+@tbl-pctraw1 and @tbl-pctfilter1 shows that filtering data based on more reliable polling criteria makes the average rate of support for Harris increase.
+Given the current polling landscape where Harris and Trump have nearly equivalent levels of support [@factcheck2024], higher support rates for Harris in the filtered analysis serves as substantial evidence that Harris could win the election.
+
+Furthermore, our data analysis that incorporates state-specific factors (@fig-resultplot5) reveals that Harris holds relatively stable support in key battelground states, including Pennsylvania, Michigan, and Wisconsin.
+As it will be discussed in more detail in @sec-third-point, this suggests that Harris' support base is both resilient and potentially insulated from the unpredictable swings [@kff2024].
+Regarding the winner-take-all nature of the electoral college, Harris' stability in these key states not only strengthens her position but also shows high probability in her victory, as even slight leads in battleground states can result in a decisive electoral advantage [@kff2024].
+The combination of broad polling support and specific regional strength positions Harris as a leading presidential candidate in the 2024 election.
 
@@
 
-The polling averages for Harris and Trump don't seem to show a big difference.
+@fig-resultplot2 and @fig-resultplot3 shows notable differences in the supportive trends for Harris when considering the variations of polling organizations.
+In @tbl-modelresults1, the inclusion of pollster-specific variables had statistical significance.
+This proposes that pollster-specific factors, like methodology and sample composition, produce a range of outcomes even for the same candidate's support levels [@nytimes2024].
+As  pollsters often fear publishing outlier results that might undermine credibility, they may adjust the results with a perceived consensus [@silver2024].
+This is likely to reduce the accuracy of polling aggregates, misrepresenting the true differences in public sentiment.
 
-Various polling entities, including Siena/NYT and Quinnipiac, use distinct methodologies, resulting in potential systematic discrepancies within their outcomes.
-According to Michael Wines, the inherent biases of each pollster result in different interpretations of the elections prediction [@nyt2024pollproblems].
-For instance, Siena/NYT's live-interview method often skews slightly older and more conservative in its responses compared to online polling methods used by other organizations [@siena_scri_2024_poll].
-This variability makes clear that when interpreting for polling results, the source of the poll must be included at minimum.
+To be more specific, this adjustment based on previous consensus data would limit the visibility of demographic shifts that might favor one candidate unexpectedly [@aapor_herding].
+For instance, the polling discrepancies in the 2016 and 2020 U.S election were partly attributed to underestimated certain voter segments supportive of Trump [@silver2024].
+Assessing the potential limitations of aggregated data, recognizing that it may reflect methodological biases more than independent public opinion should be on the way for a reliable prediction of electoral support.
 
-## State-Level Differences {#sec-third-point}
+## The electoral college and the power of battleground states {#sec-third-point}
 
-Certain states, particularly battleground states like Florida, Pennsylvania, and Georgia, show significantly different polling results compared to national trends due to regional issues and voter priorities [@politico2024battleground].
-These reflect state-specific issues, such as local economies, immigration, and access to healthcare.
-Surveys carried out within these states often show wider variation in candidate support, which makes state-level analysis necessary for more accurate predictions.
-Incorporating state-specific factors into electoral models allows a better representation of the diverse political landscapes seen across the United States, hence improving the precision of prediction in these key areas.
+In the U.S.
+electoral college, the Constitution assigns each state a set number of electoral based on its Congressional representation and a candidate must secure more than 270 votes to win the presidency [@whyy2024].
+This often results in candidates focusing their campaigns primarily on battleground states, where voter preferences are more fluid and results are less predictable [@ashcenter2024].
+@fig-resultplot5 show that key battleground states' polling trends differ from that of the national average (@fig-resultplot1).
+
+The stability of polling trends in Pennsylvania suggests a consistent voter base despite of the battleground status, which often means that even slight shifts in support could be pivotal [@ashcenter2024].
+Harris' focus on healthcare reform, especially her commitment to strengthening the Affordable Care Act (ACA), is expected to have particular resonance in Pennsylvania, where a large population of low-to-middle income residents benefit from ACA subsidies [@factcheck2024].
+On the other hand, Michigan, where there are economically diverse population, national economic policy discussions could swing voter sentiment more sharply and result in relatively large fluctuations [@kff2024].
+Surveys carried out within battleground states show wider variation in candidate support, making the incorporation of state-specific factors into electoral models allow a better representation of the diverse political landscapes seen across the United States.
 
@@
 
-Future studies could focus on the incorporation of time-varying effects to capture dynamic shifts in public opinion, especially at times of large campaign events.
-Interaction terms between pollster and state can be included in the model, since some polling organizations may be more effective or influential in specific regions.
+While our model provides a foundation to understand Harris' polling support, further refinements are necessary to enhance its accuracy and applicability.
+For instance, future studies could focus on the incorporation of voter demographics, such as age, gender, and education.
+Showing which segments of the population are driving changes in support, this can give further explanation to how certain policies or campaign events of candidates affect overall support for a candidate.
+In addition, interaction terms between pollster and state can be included in the model, considering that some states might receive disproportionately more attention from certain pollsters as shown in @fig-correlation3.
 This can give further explanation to how pollsters and regional dynamics affect overall support for a candidate.
-Moreover, extending the model to incorporate voter demographics, such as age, gender, and education, could show which segments of the population are driving changes in support.
 
-While our model provides a foundation to understand Harris' polling support, it will take further refinements are necessary to enhance its accuracy and applicability.
-The combination of hierarchical modeling with Bayesian methods is effective at accounting for the heterogeneity across pollsters and states, yet there remain considerable directions of inquiry to explore at the intersection of polling methodology, voter behavior, and regional electoral dynamics.
+This study has its constraints in that sampling error and systematic biases inevitably occur in the process of measuring social phenomenon with data.
+In particular, due to the nature of polling surveys, non-responses or misunderstanding of the survey questions result in missing data or outliers.
+Considering the possibility of prediction errors, we should take caution in interpreting the analysis and prediction results.
+By conducting more in-depth research of the domain knowledge and constructing appropriate variables to put into account, it would help derive accurate conclusions rather than just acknowledging the numbers given itself.
 

### 2024-10-31 scripts/00-simulate_data.R
simulation data and tests update

@@ -12,7 +12,6 @@
 library(tidyverse)
 set.seed(853)
 
-
 #### Simulate polling data for Kamala Harris ####
 # Simulate poll data
 sim_poll_data <- tibble(
@@ -22,16 +21,13 @@ sim_poll_data <- tibble(
   pollscore = runif(1000, -1.5, 2.5),
   candidate_name = sample(c("Kamala Harris", "Donald Trump"), size = 1000, replace = TRUE, prob = c(0.6, 0.4)),
   pct = jitter(50 + rnorm(1000, mean = 0, sd = 5), amount = 2),
-  state = sample(c("Florida", "Pennsylvania", "Michigan", "Wisconsin", "Arizona", "Texas", NA), size = 1000, replace = TRUE),
+  state = sample(c("Florida", "Pennsylvania", "Michigan", "Wisconsin", "Arizona", "Texas"), size = 1000, replace = TRUE),
   date = as.Date("2024-07-01") + sample(0:90, 1000, replace = TRUE)
 )
-# Add 'national' binary variable
-sim_poll_data <- sim_poll_data %>%
-  mutate(national = if_else(is.na(state), 1, 0)) # 1 if state is NA, 0 otherwise
+
 # Filter Harris data and keep only high-quality pollsters
 sim_harris_data <- sim_poll_data %>%
-  filter(candidate_name == "Kamala Harris", numeric_grade >= 3.0) %>%
-  select(-state) # Remove 'state' column
+  filter(candidate_name == "Kamala Harris", numeric_grade >= 3.0)
 
 #### Save data ####
 write_csv(sim_harris_data, "data/00-simulated_data/simulated_data.csv")

### 2024-10-31 scripts/01-test_simulated_data.R

@@ -38,12 +38,12 @@ if (ncol(sim_harris_data) == 8) {
   stop("Test Failed: The dataset does not have 8 columns.")
 }
 
-# Check if the 'national' column contains valid binary values
-valid_NA <- c(0, 1)
-if (all(sim_harris_data$national %in% valid_NA)) {
-  message("Test Passed: The 'national' column contains only valid binary values.")
+# Check if the 'state' column contains valid state names
+valid_states <- c("Florida", "Pennsylvania", "Michigan", "Wisconsin", "Arizona", "Texas")
+if (all(sim_harris_data$state %in% valid_states)) {
+  message("Test Passed: The 'state' column contains only valid state names.")
 } else {
-  stop("Test Failed: The 'national' column contains invalid values.")
+  stop("Test Failed: The 'state' column contains invalid state names.")
 }
 
 # Check if the 'pollster' column contains valid pollster names
@@ -61,9 +61,9 @@ if (all(!is.na(sim_harris_data))) {
   stop("Test Failed: The dataset contains missing values.")
 }
 
-# Check if there are no empty strings in 'pollster' and 'candidate_name' columns
-if (all(sim_harris_data$pollster != "" & sim_harris_data$candidate_name != "")) {
-  message("Test Passed: There are no empty strings in 'pollster' or 'candidate_name'.")
+# Check if there are no empty strings in 'state', 'pollster', and 'candidate_name' columns
+if (all(sim_harris_data$state != "" & sim_harris_data$pollster != "" & sim_harris_data$candidate_name != "")) {
+  message("Test Passed: There are no empty strings in 'state', 'pollster', or 'candidate_name'.")
 } else {
   stop("Test Failed: There are empty strings in one or more columns.")
 }
@@ -97,7 +97,7 @@ sim_harris_pollster_b <- ggplot(sim_harris_data, aes(x = date, y = pct)) +
   geom_smooth() +
   facet_wrap(~pollster) +
   theme_classic() +
-  labs(y = "Harris Percent", x = "Date", title = "The Vote that Harris Received in the Poll, by Pollster.")
+  labs(y = "Harris Percent", x = "Date", title = "The Vote that Harris Received in the Poll, by Pollster (Facets)")
[…]

### 2024-11-01 paper/paper.qmd
fix for tables/figures spilling out of the page

@@
 Following that, @sec-results presents the results from the linear and Bayesian models, while the next section @sec-discussion discusses the broader implications of these findings.
-Finally, the paper concludes with remarks on future directions for research and applications of these models [@sec-discussion].
+Finally, @sec-appendix provides an in-depth overview of the pollster methodology, outline an idealized survey methodology, and detail the modeling techniques employed in our analysis.
 
@@
 
-We use the statistical programming language R [@citeR] to analyze US presidential polling data from FiveThirtyEight [@fivethirtyeight2024], focusing on support for Kamala Harris (Note that the data was last updated on 30 Oct.
-2024).
-The dataset includes a wide range of poll results from various national and state-level polls, with key variables such as pollster, sample size, percentage of support for Harris, and end date of the poll.
+We use the statistical programming language R [@citeR] to analyze US presidential polling data from FiveThirtyEight [@fivethirtyeight2024], focusing on support for Kamala Harris.
+The dataset, last updated on 30 October 2024, includes a wide range of poll results from various national and state-level polls, with key variables such as pollster, sample size, percentage of support for Harris, and end date of the poll.
 Following the guidance of @tellingstories, we compiled the results of each opinion poll over a period of time and compared them taking into account the methodological peculiarities of polling by pollsters and geographical scope of the conducted polls.
@@
 
-(Data last updated on 30 Oct 2024.) - [!!! MODIFY AT FINAL !!!]
-
 ## Measurement
 
-[!!! MODIFY OBSV. NUMBERS AFTER DATA UPDATES !!!] - ERASE AT FINAL
-
-The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results (16817 observations, based on dataset available at Oct 30).
+The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results, showing 16817 observations based on dataset available at October 30.
 The polls conducted by various polling organizations capture voter preferences by taking a representative sample of the electorate and asking for the voters' candidate of choice.
@@
 
+\newpage
+
 ```{r}
@@
 #| echo: false
+#| tbl-align: center
+#| tbl-width: 80%
 
@@
 #| echo: false
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
 
@@
   labs(x = "Percentage of Support (pct)",
-       y = "Frequency")
+       y = "Count")
 
@@
 ```
+
+\newpage
+
 ```{r}
@@
 
-[!!! MODIFY DATE RANGE AFTER DATA UPDATES !!!] - ERASE AT FINAL
-
 The 'end_date' variable representing the time the poll was concluded was put into account to keep track of how support for a candidate changes in time.
-The reported end dates in the original dataset [@fivethirtyeight2024] ranges from 27 April, 2021 to 29 October, 2024 (@tbl-date).
+The reported end dates in the original dataset [@fivethirtyeight2024] ranges from 7 April, 2021 to 29 October, 2024 (@tbl-date).
 @fig-date shows that the polling data is more concentrated on survey results conducted in the recent period.
@@
 #| label: tbl-date
-#| tbl-cap: Summary statistics for the date the poll was concluded
-#| echo: false
+#| tbl-cap: Summary statistics for the date that the poll was concluded
+#| echo: false
+#| tbl-align: center
+#| tbl-width: 80%
 
@@
 ```
+
+\newpage
+
 ```{r}
 #| label: fig-date
-#| fig-cap: Distribution of the date the poll was concluded
-#| echo: false
+#| fig-cap: Distribution of the date that the poll was concluded
+#| echo: false
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
 
@@
 
-[!!! MODIFY DATE RANGE AFTER DATA UPDATES !!!] - ERASE AT FINAL
-
 When filtering the data, not only pollster quality and candidate type but also the date variable was considered.
 The filtered data contains only the polling data after the declaration of Harris.
-So, the date variable for filtered data ranges from 23 July, 2024 to 29 October, 2024 (@tbl-date2).
+So, the date variable for filtered data in @tbl-date2 ranges from 23 July, 2024 to 29 October, 2024.
 @fig-date2 shows that overall, polling is conducted regularly but intensifies around specific dates.
@@
 #| label: tbl-date2
-#| tbl-cap: Summary statistics for the date the high-quality polls for Harris was concluded
-#| echo: false
+#| tbl-cap: Summary statistics for the date that high-quality polls for Harris was concluded
+#| echo: false
+#| tbl-align: center
+#| tbl-width: 80%
 
@@
 
+\newpage
+
 ```{r}
 #| label: fig-date2
-#| fig-cap: Distribution of the date the high-quality polls for Harris was concluded
-#| echo: false
+#| fig-cap: Distribution of the date that high-quality polls for Harris was concluded
+#| echo: false
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
 
@@
 Depending on their polling methodology, the general results may have potential biases.
-A detailed analysis of the polling methodology and possible errors of the organization will be covered in @sec-data-details.
+A detailed analysis of the polling methodology and possible errors of the organization will be covered in @sec-appendix.
 
@@
 #| label: tbl-pollscore2
-#| tbl-cap: Summary statistics for the reliability scores of high-quality pollsters used for analysis
-#| echo: false
+#| tbl-cap: Summary statistics for the reliability scores of high-quality pollsters in the analysis data
+#| echo: false
+#| tbl-align: center
+#| tbl-width: 80%
 
@@
 
+\newpage
+
 ```{r}
 #| label: fig-pollscore2
-#| fig-cap: Distribution of the reliability scores of high-quality pollsters used for analysis
-#| echo: false
+#| fig-cap: Distribution of the reliability scores of high-quality pollsters in the analysis data
+#| echo: false
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
 
@@
 
-## Correlation Between Predictor Variables
+## Correlation between predictor variables
 
@@
 
+\newpage
+
 ```{r}
 #| label: fig-correlation1
-#| fig-cap: "Polling concentration over time by state *Note:* Strategic focus of polling is directed in battleground states such as Pennsylvania and Ohio in September 15, 2024"
-#| echo: false
+#| fig-cap: "Polling concentration over time by state (*Note:* Strategic focus of polling is directed in battleground states such as Pennsylvania and Ohio in September 15, 2024)"
+#| echo: false
+#| fig-align: center
+#| page-break-after: true
 
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text = element_text(size = 8))
 
@@
 
+\newpage
+
 ```{r}
 #| label: fig-correlation2
-#| fig-cap: "Polling concentration over time by pollster *Note:* Concentrated polling efforts of Siena/NYT in the first half of September, 2024 show temporal bias and potential overemphasizing of their polling methodologies around key election events."
-#| echo: false
+#| fig-cap: "Polling concentration over time by pollster (*Note:* Concentrated polling efforts of Siena/NYT are shown in the first half of September, 2024.)"
+#| echo: false
+#| fig-align: center
+#| page-break-after: true
 
@@
 
+\newpage
+
 ```{r}
 #| label: fig-correlation3
-#| fig-cap: "Polling activity by pollster and state *Note:* Siena/NYT has strongest state-level polling coverage in key battleground states"
-#| echo: false
+#| fig-cap: "Polling activity by pollster and state (*Note:* Siena/NYT has strongest state-level polling coverage in key battleground states)"
+#| echo: false
+#| page-break-after: true
 
@@
 plot_usmap(data = pollster_state_data, values = "total_polls", regions = "states") +
-  facet_wrap(~ pollster, ncol = 2) +  # Create a map for each pollster
+  facet_wrap(~ pollster, ncol = 4) +  # Create a map for each pollster
   scale_fill_distiller(palette = "Blues", direction = 1, na.value = "gray90") +  # Use blue color palette
@@
   theme_void() +  # Remove axes and gridlines
-  theme(strip.text = element_text(size = 12))  # Customize facet labels
+  theme(strip.text = element_text(size = 9))  # Customize facet labels
 
@@
 
-The goal of our modelling strategy is to estimate Kamala Harris's support percentage in the 2024 US election polls, accounting for potential variations over time, as well as across different pollsters and states.
-The model balances complexity with interpretability, incorporating both linear and Bayesian frameworks to capture patterns in the data.
-Background details and diagnostics are included in [Appendix -@sec-model-details].
+Our modelling strategy estimates Kamala Harris's support percentage in the 2024 US election polls, accounting for potential variations over time including pollster and state differences.
+The model balances complexity with interpretability, using both linear and Bayesian frameworks to capture patterns in the data.
+We turned from the initial linear model to a Bayesian model for robust predictions considering the variables' uncertainties.
+The Bayesian approach, which incorporates both prior beliefs and observed data to make predictions, allowed us to capture the variances among pollsters and states more accurately and responsively.
+Further background details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 
-## Results from the relation between the prediction variable and outcome variable of the analysis data
+## Results from examining the analysis data
 
@@
 
+\newpage
+
 ```{r}
 #| label: fig-resultplot1
-#| fig-cap: "Polling votes for Harris over time *Note:* The date starts from after Harris officially announced her campaign for 2024 U.S. presidential election on July 21, 2024."
+#| fig-cap: "Polling votes for Harris over time (*Note:* The date starts from after Harris officially announced her campaign for 2024 U.S. presidential election on July 21, 2024.)"
 #| echo: false
 #| warning: false
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
 
@@
 Siena/NYT (pink) shows relatively stable polling results for Harris, with fewer extreme values and a slight downward trend in October.
-On the other hand, Emerson (greem) shows more fluctuation throughout the period with a wider spread of polling values.
+On the other hand, Emerson (green) shows more fluctuation throughout the period with a wider spread of polling values.
 
@@
 #| label: fig-resultplot2
-#| fig-cap: "Polling votes for Harris over time by pollster *Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
+#| fig-cap: "Polling votes for Harris over time by pollster (*Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
 presidential election on July 21, 2024.
-Only filtered pollsters with high-quality polling questions are shown in the following."
+Only filtered pollsters with high-quality polling questions are shown in the following.)"
 #| echo: false
 #| warning: false
+#| message: false
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
 
@@
 #| label: fig-resultplot3
-#| fig-cap: "Polling votes for Harris over time by pollster (facets) *Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
+#| fig-cap: "Polling votes for Harris over time by pollster (facets) (*Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
 presidential election on July 21, 2024.
-Only filtered pollsters with high-quality polling questions are shown in the following."
+Only filtered pollsters with high-quality polling questions are shown in the following.)"
 #| echo: false
 #| warning: false
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
 
@@
 
+\newpage
+
 When accounting for the state difference, @fig-resultplot4 and @fig-resultplot5 show variations in Harris' support levels.
@@
 #| label: fig-resultplot4
-#| fig-cap: "Polling votes for Harris over time by state *Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
+#| fig-cap: "Polling votes for Harris over time by state (*Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
 presidential election on July 21, 2024.
-Only filtered pollsters with high-quality polling questions are shown in the following."
-#| echo: false
+Only filtered pollsters with high-quality polling questions are shown in the following.)"
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
+
 # Color by pollster
-#| warning: false
-
 base_plot +
@@
 #| warning: false
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
 
@@
 
+\newpage
+
 ## Results from the prediction model
 
-Prediction results derived from our model frameworks are summarized in @tbl-modelresults1 and @tbl-modelresults2. In the linear models (@tbl-modelresults1)
-, higher $R^2$ and lower WAIC, RMSE values in the second column show that the inclusion of pollster effects improves accuracy and reliability of the prediction model.
+Prediction results derived from our model frameworks are summarized in @tbl-modelresults1 and @tbl-modelresults2.
+In the linear models (@tbl-modelresults1), higher $R^2$ and lower WAIC, RMSE values in the second column show that the inclusion of pollster effects improves accuracy and reliability of the prediction model.
 In the Bayesian models (@tbl-modelresults2), higher Log-Likelihood and lower ELPD, WAIC values in the fourth column shows that including both pollster and state effects enhances model performance.
+
+\
 
@@
 #| warning: false
+#| tbl-width: 80% 
+#| page-break-after: true
+#| page-break-before: true
 
@@
 #| warning: false
+#| tbl-align: center
+#| tbl-width: 80% 
+#| page-break-after: true
+#| page-break-before: true
 
@@
 For instance, future studies could focus on the incorporation of voter demographics, such as age, gender, and education.
-Showing which segments of the population are driving changes in support, this can give further explanation to how certain policies or campaign events of candidates affect overall support for a candidate.
+Showing which segments of the population are driving changes in support, this can give further explanation to how certain policies or campaign events of candidates affect overall support for a candidate [@brown2024candidate].
 In addition, interaction terms between pollster and state can be included in the model, considering that some states might receive disproportionately more attention from certain pollsters as shown in @fig-correlation3.
-This can give further explanation to how pollsters and regional dynamics affect overall support for a candidate.
+This can give further explanation to how pollsters and regional dynamics affect overall support for a candidate [@pew2020electionpolls].
 
 This study has its constraints in that sampling error and systematic biases inevitably occur in the process of measuring social phenomenon with data.
-In particular, due to the nature of polling surveys, non-responses or misunderstanding of the survey questions result in missing data or outliers.
+In particular, due to the nature of polling surveys, non-responses or misunderstanding of the survey questions result in missing data or outliers [@pew2020electionpolls].
 Considering the possibility of prediction errors, we should take caution in interpreting the analysis and prediction results.
-By conducting more in-depth research of the domain knowledge and constructing appropriate variables to put into account, it would help derive accurate conclusions rather than just acknowledging the numbers given itself.
+By conducting more in-depth research of the domain knowledge and constructing appropriate variables to put into account, it would help derive accurate conclusions rather than just acknowledging the numbers given itself [@geeksforgeeks2023domainknowledge].
 
@@
 
-# Appendix
-
-# Additional data details {#sec-data-details}
-
-## Pollster Methodology Overview and Evaluation
+# Appendix {#sec-appendix}
+
+## Pollster methodology overview and evaluation
 
@@
 
-## Idealized Methodolgy
+## Idealized methodolgy
 
@@
 
-## Idealized Survey
-
-The proposed survey questionnarie design is provided in the following link:
+### Idealized survey
+
+The proposed survey questionnarie design is in the following link:
+\
 https://forms.gle/zQ8iJyPk3HhJYrKd9.
@@
 ::: {.layout}
-## Survey Demo
+### Survey demo
 
@@
 
-# Model details {#sec-model-details}
-
-## Posterior predictive check
+## Model details {#sec-model-details}
+
+### Posterior predictive check
 
@@
 #| fig-subcap: ["Posterior prediction check"]
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
 
@@
 
-## Diagnostics
+### Diagnostics
 
@@
 #| layout-ncol: 2
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
 

### 2024-11-01 paper/paper.qmd
idealized methodology section refinement

@@
 Since polling is a "snapshot in time",  the results can fluctuate based on recent political events or campaign dynamics.
-Additionally, while the effort to represent a broad demographic is laudable, there are still issues with nonresponse bias in polling-particularly among the hard-to-reach voter or voters suspicious of polling organizations themselves [@pew2023polling].
+Additionally, there are still issues with non-response bias in polling, particularly among the hard-to-reach voter or voters suspicious of polling organizations themselves [@pew2023polling].
 
@@
 presidential election with a budget of $100,000 would be designed as follows.
-First, a stratified random sampling method will be employed that allows for the capture of the demographic elements such as age, gender, race, and education level.
-This could alleviate bias and make the sample more representative of the population [@pew2023polling].
-The data collection process will encompass both telephone surveys and internet polling, thereby effectively engaging a diverse range of voters [@pew2023polling].
-Survey respondents would include older populations via conventional methods and younger, technologically intellectual individuals through digital platforms.
-This multifaceted approach enhances the reliability and inclusiveness of the gathered data.
-
-In addition to robust sampling, the methodology incorporates weighting methods that account for groups that are underrepresented, ensuring that the outcomes are not skewed by sampling errors [@pew2023polling].
+First, a stratified random sampling method will be employed that allows for the capture of the demographic elements such as age, gender, race, and education level [@pew2023polling].
+We will focus on 10 battleground states including Florida, Pennsylvania, Michigan, and Arizona, each receiving a budget of $7,000.
+For each state, $5,000 would be allocated for telephone and online surveys, while $2,000 would be allocated for training local staff for telephone outreach and conducting data entry to maintain data quality and consistency.
+This targeted investment could alleviate bias and make the sample more representative of the population [@pew2023polling], enhancing the reliability and inclusiveness of the gathered data.
+
+In addition to robust sampling, the methodology incorporates high-quality questionnaire design and testing.
+We would invest $15,000 in this with $5,000 on initial drafting and expert review to ensure questions are clear and unbiased.
+$7,000 would be spend on conducting pre-tests and pilot studeies for refinement [@driveresearch_cognitive_testing] and $3,000 would be allocated to final review and translation of the survey questions.
+Cognitive testing and iterative feedback loops are expected to improve the validity and accuracy of the collected data to derive reliable prediction results [@presser2004pretesting].
+
+Then, $10,000 would be assigned for data processing and weighting.
+Weighting methods that account for groups that are underrepresented ensure that the outcomes are not skewed by sampling errors [@pew2023polling].
+Post-stratification weighting which ensures that the sample accurately represents the U.S voting population will take up $5,000 while data cleaning and consistency checks for preventing bias from erroneous entries or duplicates will take up $5,000.
+
 The final model would use Bayesian hierarchical modeling, which allows for more flexible modeling of uncertainty and variation across states, pollsters, and other external factors.
 These models, along with out-of-sample testing and cross-validation, enable accurate prediction sensitive to the dynamics of real-world changes, including political events [@pew2023polling].
-The inclusion of external factors, such as major political events, debates, or sudden economic shifts, would help the model remain responsive to rapid changes in voter sentiment.
+By investing $3,000 on model setup and calibration and $2,000 on cross-validation and testing, we expect to ensure the model remains robust and provides reliable predictions under various electoral scenarios.
 
@@
 ::: {.column width="1/5"}
-![Survey Intro](../other/idealized_survey/survey_intro.png)
+![Survey intro](../other/idealized_survey/survey_intro.png)
 :::
@@
 ::: {.column width="1/5"}
-![Survey Questions 1](../other/idealized_survey/survey_q1.png)
+![Survey questions asking about demographics](../other/idealized_survey/survey_q1.png)
 :::
@@
 ::: {.column width="1/5"}
-![Survey Questions 2](../other/idealized_survey/survey_q2.png)
+![Survey questions asking for voting intentions and candidate favorability](../other/idealized_survey/survey_q2.png)
 :::
@@
 ::: {.column width="1/5"}
-![Survey Questions 3](../other/idealized_survey/survey_q3.png)
+![Survey questions asking for issues of interest and likelihood to vote](../other/idealized_survey/survey_q3.png)
 :::
@@
 ::: {.column width="1/5"}
-![Survey Final Thanks](../other/idealized_survey/survey_outro.png)
+![Survey final thanks](../other/idealized_survey/survey_outro.png)
 :::

### 2024-11-03 scripts/05-exploratory_data_analysis.R
delete unnecessary files

@@ -1,27 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Yunkyung Ko
-# Date: 20 October 2024
-# Contact: yunkyung.ko@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-
-#### Read data ####
-harris_data <- read_csv("data/02-analysis_data/harris_elections_data.csv")
-
-### Model data ####
-
-
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
-
-

### 2024-11-03 paper/paper.qmd
title, abstract, introduction refinement

@@
 ---
-title: "A Poll-of-Polls Forecast for the 2024 U.S. Presidential Election: Analysis of Support Trends for Harris Accounting for Pollster and State Disparities"
-subtitle: "Kamla Harris Emerges as a Leading Candidate with Constant Support Rates Hovering Around 50.03%"
+title: "A Poll-of-Polls Forecast for the 2024 U.S. Presidential Election: Analyzing Support Trends for Harris by Pollster and State"
+subtitle: "Kamala Harris Leads with Stable Support Around 48%"
 author: 
@@
 date-format: long
-abstract: "In the anticipation of the 2024 US presidential election, this paper intends to provide a trend analysis of the support rates for Harris accounting for pollster and state specific factors.
-Supportive polls for Harris remained relatively constant around 50.03%.
-However, some polling organizations and battleground states such as Siena/NYT and Michigan, showed variations from the average national results.
-Our objective is to propose a reliable electoral prediction ny not only analyzing the summary statistics and correlations of date, pollster, and state variables, but also by constructing prediction models using a linear and a Bayesian model approach."
+abstract: "In the anticipation of the 2024 U.S.
+presidential election, this paper analyzes trends in Harris's support, focusing on pollster and state-specific factors.
+Harris’s support has remained steady around 48%, though key pollsters like Siena/NYT and battleground states such as Michigan show notable deviations.
+Our goal is to propose a reliable electoral prediction by examining correlations among date, pollster, and state variables and developing predictions through linear and Bayesian models."
 format: pdf
@@
 
-The US presidential election is an event that receives a lot of international attention due to its far-reaching implications beyond the country's borders.
-The effects of US elections are not only related to economy and international relations around the world, but also link to social and environmental issues such as climate change [@euromonitor2024].
-In the anticipation of the 2024 US electoral competition, this paper is aimed at predicting possible outcomes of the election by analyzing the level of support that Kamala Harris will gain.
-
-We forecast the support of Kamala Harris based on the polling results at the national and state levels, and apply a linear regression and a Bayesian approach.
-The main parameter of interest is the proportion of vote or support that Harris received in surveys, which is traced over time.
-By considering the effect of changes in poll-making organizations and geographical distinctions, our objective is to correct for variation across different voter bases with the pooling the polls approach [@poolingpolls].
-
-Our initial linear model examining the support for Harris over time suggests that the rate of support remained relatively stable around 50.03%, with no significant increase or decrease.
-However large variability was detected, as seen in the spread of points around the fitted line.
-Consequently, pollster-specific effects and state-level random effects were added to the model but resulted in even higher variations.
-Some pollsters or states consistently reported higher or lower support for Harris compared to others, proposing the significance of considering distinct pollster and state environment when interpreting electoral polling results.
-
-These prognostics provide much more than merely forecasting the election in question.
-Hinting the trajectory the US might take in matters of foreign policy, economics, and global politics, predictions enable stakeholders worldwide to take preemptive measures for the resulted changes of a newly elected government [@csis2024].
-As such, this study not only contributes to the domestic political discourse but also provides a valuable tool for global actors seeking to navigate the uncertainty surrounding the 2024 US presidential election[@csis2024].
-
-The paper is structured as follows.
-Initially, @sec-data and @sec-model explores the data and methodology used, including filtering and modeling techniques applied to the polling data.
-Following that, @sec-results presents the results from the linear and Bayesian models, while the next section @sec-discussion discusses the broader implications of these findings.
-Finally, @sec-appendix provides an in-depth overview of the pollster methodology, outline an idealized survey methodology, and detail the modeling techniques employed in our analysis.
+The 2024 U.S.
+presidential election receives a lot of international attention due to its far-reaching implications, affecting global economy, foreign policy, and social issues like climate change [@euromonitor2024].
+This paper aims to predict possible outcomes of the election by analyzing Kamala Harris's support in the polling data with a linear regression and a Bayesian approach.
+
+Our estimand is Harris's support rate in polls, tracked over time and adjusted for pollster and state specific variations.
+Using a "pooling the polls" approach, we aim to correct for variations across different voter bases [@poolingpolls].
+Initial linear models show Harris's support rate remains stable at around 48% over time, though with large variability as seen in the spread of points around the fitted line.
+Some pollsters or states such as Siena/NYT and Michigan consistently reported higher or lower support for Harris compared to others, proposing the importance of accounting for specific polling contexts.
+
+Beyond election forecasting, this analysis signals global stakeholders on potential shifts in the U.S.
+policy and strategy [@csis2024].
+As such, this study not only contributes to the domestic political discourse but also provides a valuable tool for global actors seeking to navigate the uncertainty surrounding the 2024 U.S.
+presidential election [@csis2024].
+The paper is organized as follows: @sec-data and @sec-model detail the data and methodology used, including filtering and modeling techniques applied to the polling data.
+@sec-results presents findings from the linear and Bayesian models, while @sec-discussion interprets broader implications.
+Finally, @sec-appendix covers pollster methodology and ideal survey practices.
 
@@
 
-### The Proportion(%) of Support that a Candidate Received in the Poll
+### The Proportion (%) of Support that a Candidate Received in the Poll
 
@@
 Comparing the summary statistics for the raw data (@tbl-pctraw1) and filtered data (@tbl-pctfilter1), higher numbers were derived from data filtered only by Harris supporters.
-Also, @fig-pctfilter2 illustrates that a significant number of polls indicate support levels ranging from 40% to 50%, which suggests a stable yet not substantial endorsement.
+Also, @fig-pctfilter2 illustrates that a significant number of polls indicate support levels ranging from 45% to 50%, which suggests a stable yet not substantial endorsement.
 This proposes that Harris possesses a reliable foundational support, although her capacity to obtain a majority remains ambiguous.
@@
 #| label: tbl-pctraw1
-#| tbl-cap: Summary statistics for the proportion(%) of support that candidates received in the poll
+#| tbl-cap: Summary statistics for the proportion (%) of support that candidates received in the poll
 #| echo: false
@@
 #| label: fig-pct2
-#| fig-cap: Distribution of the proportion(%) of support that candidates received in the poll
+#| fig-cap: Distribution of the proportion (%) of support that candidates received in the poll
 #| echo: false
@@
 ggplot(raw_elections_data, aes(x = pct)) +
-  geom_histogram(binwidth = 5, fill = "blue", alpha = 0.7, color = "black") +
+  geom_histogram(binwidth = 3, fill = "blue", alpha = 0.7, color = "black") +
   theme_minimal() +
@@
 #| label: tbl-pctfilter1
-#| tbl-cap: Summary statistics for the proportion(%) of support that Harris received in high-quality polls
+#| tbl-cap: Summary statistics for the proportion (%) of support that Harris received in high-quality polls
 #| echo: false
@@
 #| label: fig-pctfilter2
-#| fig-cap: Distribution of the proportion(%) of support that Harris received in high-quality polls
+#| fig-cap: Distribution of the proportion (%) of support that Harris received in high-quality polls
 #| echo: false
@@
 ggplot(harris_data, aes(x = pct)) +
-  geom_histogram(binwidth = 5, fill = "blue", alpha = 0.7, color = "black") +
+  geom_histogram(binwidth = 2, fill = "blue", alpha = 0.7, color = "black") +
   theme_minimal() +

### 2024-11-03 scripts/00-simulate_data.R
update preamble of the scripts

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulates a dataset of US election polling data
 # Author: Yunkyung Ko
-# Date: 19 October 2024
+# Date: 3 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed

### 2024-11-03 scripts/01-test_simulated_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests the structure and validity of the simulated US presidential election polls data. In addition, plot the data to check the overall expected outcomes.
 # Author: Yunkyung Ko
-# Date: 19 October 2024
+# Date: 3 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:

### 2024-11-03 scripts/02-download_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from the FiveThirtyEight's 2024 US Presidential Election Polls (National)
 # Author: Yunkyung Ko
-# Date: 12 October 2024
+# Date: 3 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None

### 2024-11-03 scripts/03-clean_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw elections data and filters high-quality polls data
 # Author: Yunkyung Ko
-# Date: 19 October 2024
+# Date: 3 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Need to have downloaded the data.

### 2024-11-03 scripts/04-test_analysis_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests the structure and validity of the US presidential election polls data.
 # Author: Yunkyung Ko
-# Date: 19 October 2024
+# Date: 3 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Need to have downloaded and cleaned the data.

### 2024-11-03 scripts/05-model_data.R

@@ -1,0 +1,96 @@
+#### Preamble ####
+# Purpose: Regression Models for predicting the percentage of supportive polls
+# Author: Yunkyung Ko
+# Date: 3 November 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Need to have downloaded the data.
+# Any other information needed? None
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+harris_data <- read_csv("data/02-analysis_data/harris_elections_data.csv")
+
+### Model data ####
+# Model 1: pct as a function of date
+model_date <- stan_glm(
+  formula = pct ~ end_date,
+  data = harris_data,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 853
+)
+
+# Model 2: pct as a function of date and pollster
+model_date_pollster <- stan_glm(
+  formula = pct ~ end_date + pollster,
+  data = harris_data,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 853
+)
+
+# Bayesian Model 1: random intercept for pollster
+# Change 'pollster' and 'state' to factor variables
+harris_data <- harris_data |>
+  mutate(
+    pollster = factor(pollster),
+    state = factor(state)
+  )
+# Specify priors
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+model_formula_1 <- cbind(num_harris, sample_size - num_harris) ~ (1 | pollster)
+bayesian_model_1 <- stan_glmer(
+  formula = model_formula_1,
+  data = harris_data,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+
+# Bayesian Model 2: random intercepts for pollster and state
[…]

### 2024-11-03 paper/paper.qmd
add graphs of model estimates in 'Result' section

@@
 library(broom)
+library(broom.mixed)
 library(modelsummary)
@@
 library(viridis)  # For a color palette
+library(cowplot)
 
@@
 #| warning: false
+#| message: false
 #| fig-align: center
@@
 base_plot +
-  geom_point() +
+  geom_point(alpha = 0.45) +
   geom_smooth()
@@
 base_plot +
-  geom_point(aes(color = pollster)) +
+  geom_point(aes(color = pollster), alpha = 0.45) +
   geom_smooth() + 
@@
 #| label: fig-resultplot5
-#| fig-cap: "Polling votes for Harris over time by state (facets) *Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
+#| fig-cap: "Polling votes for Harris over time by state (facets) (*Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
 presidential election on July 21, 2024.
-Only filtered pollsters with high-quality polling questions are shown in the following."
+Only filtered pollsters with high-quality polling questions are shown in the following.)"
 #| echo: false
@@
 
+@fig-modelgraph1 and @fig-modelgraph2 show the trends in the percentage of supportive polls for Harris predicted by linear models, respectively representing a time-series analysis and consideration of pollster differences.
+Moreover, @fig-modelgraph3 and @fig-modelgraph4 show the supportive trends predicted by Bayesian models accounting for pollster and state differences.
+Each model, from the simple linear regression to the more complex Bayesian models, seem to reflect a relatively consistent support pattern over time.
+This alignment of underlying support trends across model shows the credibility of the observed overall support trend.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modelgraph1
+#| fig-cap: "Predicted percentage of supportive polls for Harris in the linear model by date"
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
+
+# Generate predictions and plot for Model 1 (pct ~ end_date)
+harris_data$pred_date <- predict(model_date, newdata = harris_data)
+
+ggplot(harris_data, aes(x = end_date, y = pct)) +
+  geom_point(color = "black", alpha = 0.6) +
+  geom_line(aes(y = pred_date), color = "blue") +
+  labs(
+    x = "Date",
+    y = "Predicted percentage of supportive polls"
+  ) +
+  theme_minimal()
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modelgraph2
+#| fig-cap: "Predicted percentage of supportive polls for Harris in the linear model by date and pollster (*Note:* The blue summary line calculates the mean of the predictions across all dates and pollsters.)"
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
+
+# Generate predictions for Model 2, but ignore the individual pollster grouping in the plot
+# Calculate the mean predictions across all pollsters for each end_date
+harris_data$pred_date_pollster <- predict(model_date_pollster, newdata = harris_data)
+
+# Calculate the mean of the predictions across all dates and pollsters
+overall_mean_pred <- mean(harris_data$pred_date_pollster, na.rm = TRUE)
+
+# Plot the observed data points and a single straight summary line at the overall mean prediction
+ggplot(harris_data, aes(x = end_date, y = pct)) +
+  geom_point(aes(color = pollster), alpha = 0.6) +  # Observed data with colors by pollster
+  geom_hline(yintercept = overall_mean_pred, color = "blue") +
+  labs(
+    x = "Date",
+    y = "Predicted percentage of supportive polls"
+  ) +
+  theme_minimal()
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modelgraph3
+#| fig-cap: "Predicted percentage of supportive polls for Harris in the Bayesian model with random intercept for `pollster` variable (*Note:* In order to avoid the overfitting problem, overall mean of prediction was separately calculated in creating the summary line.)"
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
+
+# Generate predictions for Bayesian Model 1
+harris_data$bayesian_pred_1 <- posterior_predict(bayesian_model_1, newdata = harris_data) |> apply(2, mean)
+
+# Calculate the overall mean prediction
+overall_mean_pred <- mean(harris_data$bayesian_pred_1 / harris_data$sample_size * 100, na.rm = TRUE)
+
+# Plot the observed data points and a single straight line at the overall mean prediction
+ggplot(harris_data, aes(x = end_date, y = num_harris / sample_size * 100)) +
+  geom_point(color = "black", alpha = 0.6) +
+  geom_hline(yintercept = overall_mean_pred, color = "blue") +
+  labs(
+   x = "Date",
+   y = "Predicted percentage of supportive polls"
+  ) +
+  theme_minimal()
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-modelgraph4
+#| fig-cap: "Predicted percentage of supportive polls for Harris in the Bayesian model with random intercept for `pollster` and `state` variable (*Note:* In order to avoid the overfitting problem, overall mean of prediction was separately calculated in creating the summary line.)"
+#| fig-align: center
+#| fig-width: 5
+#| page-break-after: true
+
+# Generate predictions for Bayesian Model 2
+harris_data$bayesian_pred_2 <- posterior_predict(bayesian_model_2, newdata = harris_data) |> apply(2, mean)
+
+# Calculate the overall mean prediction
+overall_mean_pred_2 <- mean(harris_data$bayesian_pred_2 / harris_data$sample_size * 100, na.rm = TRUE)
+
+# Plot the observed data points and a single straight line at the overall mean prediction
+ggplot(harris_data, aes(x = end_date, y = num_harris / sample_size * 100)) +
+  geom_point(color = "black", alpha = 0.6) +
+  geom_hline(yintercept = overall_mean_pred_2, color = "blue") +
+  labs(
+   x = "Date",
+   y = "Predicted percentage of supportive polls"
+  ) +
+  theme_minimal()
+
+```
+
 # Discussion {#sec-discussion}

### 2024-11-04 paper/paper.qmd
refine the data section

@@
 ggplot(harris_data, aes(x = pct)) +
-  geom_histogram(binwidth = 2, fill = "blue", alpha = 0.7, color = "black") +
+  geom_histogram(binwidth = 1.5, fill = "blue", alpha = 0.7, color = "black") +
   theme_minimal() +
@@
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-```
-
-### Pollscore Measuring the Validity of Polling Questions
-
-This variable was another factor we had put into consideration to check whether the validity of the polling questions affects the polling results.
-The 'pollscore' variable represents the score or reliability of the pollster in question.
-The numeric values are the error and bias that can be attributed to the pollster, which means negative numbers are better.
-@tbl-pollscore and @fig-pollscore suggests that while the majority of the polls are moderately to highly qualitative in the original dataset, a fraction of the polls with low-quality or no scores could add noise or uncertainty to the analysis.
-
-```{r}
-#| label: tbl-pollscore
-#| tbl-cap: Summary statistics for the reliability scores of pollsters
-#| echo: false
-#| tbl-align: center
-#| tbl-width: 80%
-
-summary_stats <- raw_elections_data %>%
-  summarize(
-    mean = round(mean(pollscore, na.rm = TRUE),2),
-    median = round(median(pollscore, na.rm = TRUE),2),
-    min = round(min(pollscore, na.rm = TRUE),2),
-    max = round(max(pollscore, na.rm = TRUE),2),
-    sd = round(sd(pollscore, na.rm = TRUE),2),
-    n = n()
-  )
-
-# Display the summary statistics in a nicely formatted table
-summary_stats %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
-
-```
-
-\newpage
-
-```{r}
-#| label: fig-pollscore
-#| fig-cap: Distribution of the reliability scores of pollsters
-#| echo: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Distribution of 'pollscore' (bar plot) with simplified x-axis labels
-ggplot(raw_elections_data, aes(x = factor(pollscore))) +
-  geom_bar(fill = "skyblue", alpha = 0.7, color = "black") +
-  theme_minimal() +
-  scale_x_discrete(breaks = c(-1.5, -1, -0.5, 0, 0.5, 1, 1.5)) +  # Simplified labels at 0.5 intervals
-  labs(x = "Pollscore",
-       y = "Count")
-
-```
-
-After the filtering to polling data of high-quality polling organizations, we can find that the overall value and standard deviation of pollscores went down in @tbl-pollscore2.
-This implies that the polling data narrowed down to the responses from more reliable survey questions.
-@fig-pollscore2 also indicates that the data cleaning process effectively excluded less reliable sources, which can enhance the robustness of subsequent analyses.
-
-```{r}
-#| label: tbl-pollscore2
-#| tbl-cap: Summary statistics for the reliability scores of high-quality pollsters in the analysis data
-#| echo: false
-#| tbl-align: center
-#| tbl-width: 80%
-
-summary_stats <- harris_data %>%
-  summarize(
-    mean = round(mean(pollscore, na.rm = TRUE),2),
-    median = round(median(pollscore, na.rm = TRUE),2),
-    min = round(min(pollscore, na.rm = TRUE),2),
-    max = round(max(pollscore, na.rm = TRUE),2),
-    sd = round(sd(pollscore, na.rm = TRUE),2),
-    n = n()
-  )
-
-# Display the summary statistics in a nicely formatted table
-summary_stats %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
-
-```
-
-\newpage
-
-```{r}
-#| label: fig-pollscore2
-#| fig-cap: Distribution of the reliability scores of high-quality pollsters in the analysis data
-#| echo: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Distribution of 'pollscore' (bar plot)
-ggplot(harris_data, aes(x = factor(pollscore))) +
-  geom_bar(fill = "skyblue", alpha = 0.7, color = "black", width = 0.5) +
-  theme_minimal() +
-  labs(x = "Pollscore",
-       y = "Count")
 

### 2024-11-04 paper/paper.qmd
refinement of 'measurement' and 'references'

@@
 
-Each poll represents a predictor of an actual event, namely voter opinion at a particular moment.
-Nevertheless, like all survey results, the raw data is susceptible to many potential limitations including the following: sampling error, variation in polling methods, distortion because of inappropriate survey responses such as missing data or response from respondents who misunderstood the questions [@tellingstories].
-
-While applying several filters to the original dataset such as restricting to those with a numeric grade of 2.7 or higher or pollsters with more than 30 polls improves data reliability, certain limitations still exist.
-Selection bias and sampling error remains as a concern, since polls always represent only part of the population.
-Differences in the way different organizations conducted their polling might introduce more inconsistencies.
-Finally, by focusing our attention on post-declaration polls only, we exclude earlier trends that could add more insight into how Harris' support has evolved over time.
+In converting an individual's complex set of opinions into "plausibly measurable" [@tellingstories] numbers, each poll attempts to capture and simplify voters' choice in a straightforward way.
+Polls capture a momentary snapshot of opinion, but this process is inevitably influenced by factors such as question phrasing, the timing of the poll, and contextual events.
+A person may have mixed or conflicting feelings about candidates, and the process of capturing these preferences in a single response imposes much error [@tellingstories].
+
+The outcome variable, which is intended to reflect overall support levels, is constructed from respondents' stated candidate preferences.
+As part of the data refinement process to enhance the reliability of predictions, we restricted our analysis to polls with a numeric quality grade of 2.7 or higher and to pollsters with over 30 polls in the dataset.
+We set a minimum threshold of 2.7 for the numeric grade to focus on the top 30% of the dataset, ensuring a sufficient sample size of 5,612 observations.
+Additionally, we included only pollsters with at least 30 polls, reasoning that a higher number of polls from a given pollster would enhance the reliability and validity of the aggregated data.
+
+While efforts to filter and refine the dataset mitigate some limitations, they cannot entirely eliminate the challenges inherent in measuring multifaceted concept like voter preference.
+Limitations including sampling error, response bias, and inconsistencies in polling methods remain.
+Some responses may also be distorted due to misunderstandings or incomplete answers, as respondents might interpret questions differently depending on context [@tellingstories].
+This measurement process is essentially an estimation, translating complex human sentiments into quantifiable data, subject to various limitations.
 
@@
 The Bayesian approach, which incorporates both prior beliefs and observed data to make predictions, allowed us to capture the variances among pollsters and states more accurately and responsively.
-Further background details and diagnostics are included in [Appendix -@sec-model-details].
+Further background details and diagnostics are included in @sec-model-details.
 
@@
 
-These models are run through the rstanarm package [@rstanarm] in R  [@citeR], which makes Bayesian modeling available through the use of Stan's strong inference engine.
-To validate the models, RMSE and WAIC have been made use of to check the goodness of fit; Bayesian models with reduced RMSE and WAIC outperform linear models.
+These models are run through the rstanarm package [@rstanarm] in R [@citeR].
+RMSE and WAIC were used to check for the goodness of fit of the model; Bayesian models with reduced RMSE and WAIC outperform linear models.
 We use weakly informative priors; for example, $\beta_0 \sim \text{Normal}(0, 10)$ and $\sigma \sim \text{Exponential}(1)$.
@@
 The Bayesian models converged successfully, as indicated by $\hat{R} = 1$ for all parameters.
+Further details about model diagnostics are included in @sec-model-details.
 
@@
 Despite these limitations, the hierarchical structure allows us to capture important variability, making the model suitable for predicting Harris's support.
-Future improvements could involve incorporating time-varying effects or exploring interactions between pollsters and states.
+Future improvements could involve incorporating time-varying effects or interaction terms between pollsters and states.
 
@@
 
-\
-
 ```{r}
@@
 #| warning: false
-#| tbl-width: 80% 
+#| tbl-width: 60% 
+#| tbl-height: 55%
 #| page-break-after: true
@@
 ```
-
-\newpage
 
@@
 ```
+
+\newpage
 
@@
 
+\newpage
+
 # Discussion {#sec-discussion}
@@
 \
-https://forms.gle/zQ8iJyPk3HhJYrKd9.
-
-::: {.layout}
+https://forms.gle/vehDvpDNDGZ8WCn47
+
+::: {.container}
 ### Survey demo
 
-::: {.column width="1/5"}
-![Survey intro](../other/idealized_survey/survey_intro.png)
+::: {.column width="30%"}
+![Survey Intro](../other/idealized_survey/survey_intro.png)
 :::
 
-::: {.column width="1/5"}
-![Survey questions asking about demographics](../other/idealized_survey/survey_q1.png)
+::: {.column width="30%"}
+![Questions asking for vote eligibility - citizenship, age](../other/idealized_survey/eligibility_1.png)
 :::
 
-::: {.column width="1/5"}
-![Survey questions asking for voting intentions and candidate favorability](../other/idealized_survey/survey_q2.png)
+::: {.column width="30%"}
+![Questions asking for vote eligibility - voter status and registration state](../other/idealized_survey/eligibility_2.png)
 :::
 
-::: {.column width="1/5"}
-![Survey questions asking for issues of interest and likelihood to vote](../other/idealized_survey/survey_q3.png)
+::: {.column width="100%"}
+::: {.columns}
+::: {.column width="33%"}
+![Demographics Questions](../other/idealized_survey/survey_q1.png)
 :::
-
-::: {.column width="1/5"}
-![Survey final thanks](../other/idealized_survey/survey_outro.png)
+::: {.column width="33%"}
+![Voting Intentions and Candidate Favorability](../other/idealized_survey/survey_q2.png)
 :::
+::: {.column width="33%"}
+![Issues of Interest and Likelihood to Vote](../other/idealized_survey/survey_q3.png)
+:::
+:::
+:::
+
+::: {.column width="30%"}
+![Survey Final Thanks](../other/idealized_survey/survey_outro.png)
+:::
+
 :::

### 2024-11-04 paper/paper.qmd
supplementation to idealized methodology contents

@@
 
-The New York Times/Siena College polling partnership, the polling organization that accounted for the majority of polls in our analysis (@fig-pollster), conducts polls tailored for specific elections, such as state or national races [@nyt_siena_2020_methodology].
-Their sample size typically includes 600 to 1000 likely voters per poll, with oversampling in battleground states to capture regional nuances [@nyt_siena_2020_methodology].
-The methodology uses random-digit dialing (RDD) for landlines and mobile phones to ensure representative sample coverage across demographics.
-In addition, online surveys are administered to complement phone-based responses, ensuring broader accessibility [@nyt_siena_2020_methodology].
-The stratified random sampling approach is employed, where the population is divided into strata (based on demographic variables like race, education, and geography), and a random sample is drawn from each stratum [@tellingstories].
-This allows for precision in reflecting the political leanings and key demographic shifts in specific regions.
-
-The organization intends to enhance transparency in how public opinion is assessed, ensuring that questions are carefully designed to represent contemporary political discussions, and that the terminology is polished through a process of iterative testing to achieve clarity.
-They devote extensive resources to cognitive testing to ensure question wording reflects what the public thinks [@siena_scri_2024_poll].
-Their polling methodology stands out in that its strategic focus on using representative samples reflect political leanings and demographics of a region for more contextual and precise polling.
-Siena/NYT has its reputation for accurately predicting key battleground state outcomes during previous elections, such as Florida in 2016[@siena_nyt_perfect_partnership_2024].
-
-The limitations of Siena/NYT's methodology are the challenge of polling itself.
-Since polling is a "snapshot in time",  the results can fluctuate based on recent political events or campaign dynamics.
-Additionally, there are still issues with non-response bias in polling, particularly among the hard-to-reach voter or voters suspicious of polling organizations themselves [@pew2023polling].
+The New York Times/Siena College polling partnership, the polling organization that accounts for the majority of polls in our analysis (@fig-pollster), employs a methodology tailored to capture voter preferences for specific elections, whether at the state or national level [@nyt_siena_2020_methodology].
+Sample sizes typically range from 600 to 1000 likely voters, with an emphasis on oversampling in battleground states to better reflect regional political dynamics [@nyt_siena_2020_methodology].
+
+To enhance representativeness, the organization uses a mixed-mode approach that combines random-digit dialing (RDD) for landlines and mobile phones with online surveys, increasing accessibility and demographic coverage [@nyt_siena_2020_methodology].
+The polling process relies on stratified random sampling, a technique where the population is divided into mutually exclusive and collectively exhaustive subgroups, or strata [@tellingstories], based on key demographic variables such as race, education, and geographic location.
+Within each stratum, random sampling ensures proportional representation.
+This approach is especially effective in capturing variation within diverse populations [@lohr2019sampling], as it controls for demographic factors that could otherwise introduce bias.
+However, stratified sampling can also introduce complexities, such as difficulty in defining strata boundaries, which can affect overall precision if strata are not well-chosen [@lohr2019sampling].
+
+In addition, the Siena/NYT intends to enhance transparency and reliability through cognitive testing of survey questions.
+Cognitive testing allows the pollster to evaluate whether respondents interpret questions as intended, which is a necessary process in minimizing response bias and improving question clarity [@presser2004pretesting].
+This iterative approach ensures that question wording resonates with respondents' current understanding of political issues, thus reduces measurement error[@siena_scri_2024_poll].
+
+A distinctive strength of this organization's approach is its commitment to accurately reflecting the demographics and political leanings of specific regions.
+This kind of approach has been validated in high-profile cases, such as the accurate prediction of battleground state outcomes in the 2016 Florida election [@siena_nyt_perfect_partnership_2024].
+By using robust sampling methods and iterative question refinement, their polls are well-positioned to reflect regional voter behavior with a higher degree of accuracy.
+
+However, challenges remain in achieving a fully representative sample.
+Polling captures a "snapshot in time"[@tellingstories] and is vulnerable to shifts caused by recent political events, which can introduce volatility into results.
+Additionally, non-response bias remains an issue, as certain groups that are distrustful of polling organizations or those difficult to reach by phone are often underrepresented [@pew2023polling].
+Simulation studies in the literature show that non-response bias can skew results, especially in polarized contexts, unless corrective weighting or follow-up efforts are applied [@presser2004pretesting].
+
+By continuously refining their approach by using corrective weighting [@nyt_siena_2020_methodology], Siena/NYT intends to remain a credible source of insights into the evolving dynamics of U.S.
+elections.
+To be more specific about this weighting method, it assigns different weights to responses based on demographic factors such as age, race, gender, or geographic location [@nyt_siena_2020_methodology].
+For instance, if a poll has fewer responses from young adults compared to their actual proportion in the population, each response from a young adult might be given a higher weight to compensate.
+This adjustment helps balance out the representation of various groups, making the poll more reflective of the overall population [@siena_nyt_perfect_partnership_2024].
 
@@
 First, a stratified random sampling method will be employed that allows for the capture of the demographic elements such as age, gender, race, and education level [@pew2023polling].
-We will focus on 10 battleground states including Florida, Pennsylvania, Michigan, and Arizona, each receiving a budget of $7,000.
-For each state, $5,000 would be allocated for telephone and online surveys, while $2,000 would be allocated for training local staff for telephone outreach and conducting data entry to maintain data quality and consistency.
-This targeted investment could alleviate bias and make the sample more representative of the population [@pew2023polling], enhancing the reliability and inclusiveness of the gathered data.
-
-In addition to robust sampling, the methodology incorporates high-quality questionnaire design and testing.
-We would invest $15,000 in this with $5,000 on initial drafting and expert review to ensure questions are clear and unbiased.
-$7,000 would be spend on conducting pre-tests and pilot studeies for refinement [@driveresearch_cognitive_testing] and $3,000 would be allocated to final review and translation of the survey questions.
-Cognitive testing and iterative feedback loops are expected to improve the validity and accuracy of the collected data to derive reliable prediction results [@presser2004pretesting].
-
-Then, $10,000 would be assigned for data processing and weighting.
-Weighting methods that account for groups that are underrepresented ensure that the outcomes are not skewed by sampling errors [@pew2023polling].
-Post-stratification weighting which ensures that the sample accurately represents the U.S voting population will take up $5,000 while data cleaning and consistency checks for preventing bias from erroneous entries or duplicates will take up $5,000.
+Stratified sampling involves dividing the population into mutually exclusive subgroups, or strata, based on demographic factors, then randomly sampling from each stratum [@tellingstories].
+This approach allows us to control for variation within subgroups, which improves representation and reduces sampling error compared to simple random sampling.
+However, as noted by Lohr (2019), defining appropriate strata boundaries can be challenging and may impact the precision of estimates if not carefully managed [@lohr2019sampling].
+By focusing on 10 battleground states including Florida, Pennsylvania, Michigan, and Arizona, we can capture the diverse political landscapes critical to the election outcome.
+Each state will receive a $7,000 budget, with $5,000 allocated to conducting telephone and online surveys and $2,000 for training local staff for outreach and data entry.
+This budget allocation aims to address potential sources of bias, ensuring the sample is reflective of the population and enhancing data reliability.
+
+To further strengthen our methodology, we will invest $15,000 in questionnaire design and testing.
+High-quality questionnaire design is essential for minimizing response bias and ensuring that questions are clear and unbiased [@tellingstories].
+With $5,000 on initial drafting and expert review, $7,000 on conducting pre-tests and pilot studies [@driveresearch_cognitive_testing] and $3,000 on final review and translation of the survey questions, we expect to improve the validity and reliability of the collected data.
+In particular, pre-tests, rooted in cognitive testing, is expected to help assess whether respondents interpret questions as intended, a critical step for reducing measurement error and enhancing question validity [@presser2004pretesting].
+
+Then, $10,000 will be dedicated to data weighting and cleaning.
+Post-stratification weighting will adjust for demographic imbalances in the sample, helping to correct for potential biases from non-response and underrepresented groups [@valliant2019poststratification].
+By assigning $5,000 for these weighting procedures, we aim to ensure that the sample aligns as accurately as possible with the U.S.
+voting population.
+Additionally, $5,000 will be allocated to data cleaning, which includes consistency checks and removal of duplicates or erroneous entries, further improving data quality and reducing bias [@pew2023polling].
+
+Our final prediction model will employ Bayesian hierarchical modeling, which allows us to account for uncertainty and variability across states, pollsters, and external factors.
+Bayesian models are particularly effective in election forecasting, as they enable dynamic updates based on real-world changes, such as shifts in public opinion following political events.
+By allocating $3,000 for model setup and calibration and $2,000 for cross-validation and out-of-sample testing, we ensure the model’s robustness in capturing complex electoral dynamics [@pew2023polling].
+This approach not only provides flexibility in handling diverse data sources but also improves predictive accuracy through structured handling of variability and uncertainty.
 
 The final model would use Bayesian hierarchical modeling, which allows for more flexible modeling of uncertainty and variation across states, pollsters, and other external factors.
-These models, along with out-of-sample testing and cross-validation, enable accurate prediction sensitive to the dynamics of real-world changes, including political events [@pew2023polling].
+These models, along with out-of-sample testing and cross-validation, enable accurate prediction sensitive to the dynamics of real-world changes, including political events [@stata2024bayesian].
 By investing $3,000 on model setup and calibration and $2,000 on cross-validation and testing, we expect to ensure the model remains robust and provides reliable predictions under various electoral scenarios.
+
+In summary, this methodology balances rigorous sampling, robust question design, and advanced statistical modeling, all underpinned by established survey and sampling principles.
+These practices aim to deliver a reliable and comprehensive forecast for the 2024 U.S.
+presidential election.
 
@@
 https://forms.gle/vehDvpDNDGZ8WCn47
+\
+Following the guidance of @tellingstories, we aimed to have survey questions that are conversational and flow from one to the next, grouped within topics [@elson2018question].
+We refined the wording of questions to be based on what the respondent would be comfortable with, simultaneously checking for whether the survey questions could appropriately relate to measuring the estimand [@tellingstories].
 

### 2024-11-04 scripts/06-model_data.R
remove duplicate files

@@ -1,96 +1,0 @@
-#### Preamble ####
-# Purpose: Regression Models for predicting the percentage of supportive polls
-# Author: Yunkyung Ko
-# Date: 20 October 2024
-# Contact: yunkyung.ko@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: Need to have downloaded the data.
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-
-#### Read data ####
-harris_data <- read_csv("data/02-analysis_data/harris_elections_data.csv")
-
-### Model data ####
-# Model 1: pct as a function of date
-model_date <- stan_glm(
-  formula = pct ~ end_date,
-  data = harris_data,
-  family = gaussian(),
-  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-  prior_aux = exponential(rate = 1, autoscale = TRUE),
-  seed = 853
-)
-
-# Model 2: pct as a function of date and pollster
-model_date_pollster <- stan_glm(
-  formula = pct ~ end_date + pollster,
-  data = harris_data,
-  family = gaussian(),
-  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-  prior_aux = exponential(rate = 1, autoscale = TRUE),
-  seed = 853
-)
-
-# Bayesian Model 1: random intercept for pollster
-# Change 'pollster' and 'state' to factor variables
-harris_data <- harris_data |>
-  mutate(
-    pollster = factor(pollster),
-    state = factor(state)
-  )
-# Specify priors
-priors <- normal(0, 2.5, autoscale = TRUE)
-
-model_formula_1 <- cbind(num_harris, sample_size - num_harris) ~ (1 | pollster)
-bayesian_model_1 <- stan_glmer(
-  formula = model_formula_1,
-  data = harris_data,
-  family = binomial(link = "logit"),
-  prior = priors,
-  prior_intercept = priors,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95
-)
-
-# Bayesian Model 2: random intercepts for pollster and state
[…]

### 2024-11-14 paper/paper.qmd
fix reference errors
first draft of paper update

--- 2024-11-04 paper/paper.qmd
@@
 ---
-title: "A Poll-of-Polls Forecast for the 2024 U.S. Presidential Election: Analyzing Support Trends for Harris by Pollster and State"
-subtitle: "Kamala Harris Leads with Stable Support Around 48%"
+title: "My title"
+subtitle: "My subtitle if needed"
 author: 
-  - Yunkyung Ko
-thanks: "Code and data are available at: [https://github.com/koyunkyung/us_election_2024](https://github.com/koyunkyung/us_election_2024)."
+  - First author
+  - Another author
+thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
 date: today
 date-format: long
-abstract: "In the anticipation of the 2024 U.S.
-presidential election, this paper analyzes trends in Harris's support, focusing on pollster and state-specific factors.
-Harris’s support has remained steady around 48%, though key pollsters like Siena/NYT and battleground states such as Michigan show notable deviations.
-Our goal is to propose a reliable electoral prediction by examining correlations among date, pollster, and state variables and developing predictions through linear and Bayesian models."
+abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
 format: pdf
@@
 bibliography: references.bib
-toc: true
-toc-title: "Table of Contents"
-toc-depth: 2
-toc-location: left
 ---
@@
 library(tidyverse)
-library(readr)
-library(janitor)
-library(lubridate)
-library(broom)
-library(broom.mixed)
-library(modelsummary)
-library(rstanarm)
-library(splines)
-library(ggplot2)
-library(knitr)
-library(kableExtra)
-library(gridExtra)
-library(here)
-library(usmap)  # For the map of US states
-library(viridis)  # For a color palette
-library(cowplot)
-
-raw_elections_data <- read_csv(here::here("data/01-raw_data/raw_elections_data.csv"))
-harris_data <- read_csv(here::here("data/02-analysis_data/harris_elections_data.csv"))
+library(palmerpenguins)
 ```
@@
 
-The 2024 U.S.
-presidential election receives a lot of international attention due to its far-reaching implications, affecting global economy, foreign policy, and social issues like climate change [@euromonitor2024].
-This paper aims to predict possible outcomes of the election by analyzing Kamala Harris's support in the polling data with a linear regression and a Bayesian approach.
+Overview paragraph
 
-Our estimand is Harris's support rate in polls, tracked over time and adjusted for pollster and state specific variations.
-Using a "pooling the polls" approach, we aim to correct for variations across different voter bases [@poolingpolls].
-Initial linear models show Harris's support rate remains stable at around 48% over time, though with large variability as seen in the spread of points around the fitted line.
-Some pollsters or states such as Siena/NYT and Michigan consistently reported higher or lower support for Harris compared to others, proposing the importance of accounting for specific polling contexts.
+Estimand paragraph
 
-Beyond election forecasting, this analysis signals global stakeholders on potential shifts in the U.S.
-policy and strategy [@csis2024].
-As such, this study not only contributes to the domestic political discourse but also provides a valuable tool for global actors seeking to navigate the uncertainty surrounding the 2024 U.S.
-presidential election [@csis2024].
-The paper is organized as follows: @sec-data and @sec-model detail the data and methodology used, including filtering and modeling techniques applied to the polling data.
-@sec-results presents findings from the linear and Bayesian models, while @sec-discussion interprets broader implications.
-Finally, @sec-appendix covers pollster methodology and ideal survey practices.
+Results paragraph
+
+Why it matters paragraph
+
+Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
 
@@
 
-We use the statistical programming language R [@citeR] to analyze US presidential polling data from FiveThirtyEight [@fivethirtyeight2024], focusing on support for Kamala Harris.
-The dataset, last updated on 30 October 2024, includes a wide range of poll results from various national and state-level polls, with key variables such as pollster, sample size, percentage of support for Harris, and end date of the poll.
-Following the guidance of @tellingstories, we compiled the results of each opinion poll over a period of time and compared them taking into account the methodological peculiarities of polling by pollsters and geographical scope of the conducted polls.
+We use the statistical programming language SQL [@citeR]
 
-To ensure data quality, we filtered the dataset to include only polls that measured Kamala Harris' support, with a numeric grade of pollster 2.7 or higher for reliability.
-We also limited the analysis to polls conducted after July 21, 2024, when Harris officially declared her candidacy, and excluded pollsters with fewer than 30 polls to focus on those with sufficient data for robust results.
+.... Our data [@shelter].... Following @tellingstories, we consider...
 
-In performing the analysis, we utilized several R packages.
-`tidyverse` [@tidyverse] was used for data manipulation and visualization and `rstanarm` [@rstanarm], `modelsummary` [@modelsummary] was respectively used for Bayesian modeling and generating model summaries.
-For visualizing results, `ggplot2` [@ggplot2] was used and `kableExtra` [@kableExtra] helped format tables for presentation.
-These packages provided a framework for efficient data processing, modeling, and reporting.
+Overview text
 
@@
 
-The original dataset sourced from FiveThirtyEight [@fivethirtyeight2024] aggregates a wide range of poll results, showing 16817 observations based on dataset available at October 30.
-The polls conducted by various polling organizations capture voter preferences by taking a representative sample of the electorate and asking for the voters' candidate of choice.
-Surveys were conducted at the state and national levels, providing the wide perspective on public feelings across the country.
+Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
 
-Each poll represents a predictor of an actual event, namely voter opinion at a particular moment.
-Nevertheless, like all survey results, the raw data is susceptible to many potential limitations including the following: sampling error, variation in polling methods, distortion because of inappropriate survey responses such as missing data or response from respondents who misunderstood the questions [@tellingstories].
+## Outcome variables
 
-While applying several filters to the original dataset such as restricting to those with a numeric grade of 2.7 or higher or pollsters with more than 30 polls improves data reliability, certain limitations still exist.
-Selection bias and sampling error remains as a concern, since polls always represent only part of the population.
-Differences in the way different organizations conducted their polling might introduce more inconsistencies.
-Finally, by focusing our attention on post-declaration polls only, we exclude earlier trends that could add more insight into how Harris' support has evolved over time.
-
-## Outcome variable
-
-### The Proportion (%) of Support that a Candidate Received in the Poll
-
-The main variable of interest that we aim to forecast is the 'pct' variable, which represents the proportion of vote or support that a candidate received in the poll.
-@tbl-pctraw1 and @fig-pct2 respectively shows the summary statistics and distribution of the 'pct' variable in the original dataset [@fivethirtyeight2024].
-@tbl-pctfilter1 and @fig-pctfilter2 shows the summary statistics and distribution of the same variable, but in the filtered dataset that only comprises of the supporting votes for Harris from relatively high-quality polling organizations.
-Comparing the summary statistics for the raw data (@tbl-pctraw1) and filtered data (@tbl-pctfilter1), higher numbers were derived from data filtered only by Harris supporters.
-Also, @fig-pctfilter2 illustrates that a significant number of polls indicate support levels ranging from 45% to 50%, which suggests a stable yet not substantial endorsement.
-This proposes that Harris possesses a reliable foundational support, although her capacity to obtain a majority remains ambiguous.
-
-\newpage
+### Price Levels
 
 ```{r}
-#| label: tbl-pctraw1
-#| tbl-cap: Summary statistics for the proportion (%) of support that candidates received in the poll
+#| label: fig-price
+#| fig-cap: Price level distributions across all observations used in the analysis
 #| echo: false
-#| tbl-align: center
-#| tbl-width: 80%
+#| warning: false
+#| message: false
 
-# Calculate summary statistics for the 'pct' variable
-summary_stats <- raw_elections_data %>%
-  summarize(
-    mean = round(mean(pct, na.rm = TRUE), 2),
-    median = round(median(pct, na.rm = TRUE),2),
-    min = round(min(pct, na.rm = TRUE), 2),
-    max = round(max(pct, na.rm = TRUE), 2),
-    sd = round(sd(pct, na.rm = TRUE), 2),
-    n = n()
-  )
+library(ggplot2)
+apple <- read.csv("../data/analysis_data/avg_price_by_vendor_apple.csv")
+mango <- read.csv("../data/analysis_data/avg_price_by_vendor_mango.csv")
+pear <- read.csv("../data/analysis_data/avg_price_by_vendor_pear.csv")
+orange <- read.csv("../data/analysis_data/avg_price_by_vendor_orange.csv")
 
-# Display the summary statistics in a nicely formatted table
-summary_stats %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
+vendor_avg_apple <- aggregate(avg_price ~ vendor, data = apple, mean)
+vendor_avg_mango <- aggregate(avg_price ~ vendor, data = mango, mean)
+vendor_avg_pear <- aggregate(avg_price ~ vendor, data = pear, mean)
+vendor_avg_orange <- aggregate(avg_price ~ vendor, data = orange, mean)
 
-```
-```{r}
-#| label: fig-pct2
-#| fig-cap: Distribution of the proportion (%) of support that candidates received in the poll
-#| echo: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
+vendor_avg_apple$fruit <- "apple"
+vendor_avg_mango$fruit <- "mango"
+vendor_avg_pear$fruit <- "pear"
+vendor_avg_orange$fruit <- "orange"
 
-# Plot the distribution of 'pct' using a histogram
-ggplot(raw_elections_data, aes(x = pct)) +
-  geom_histogram(binwidth = 3, fill = "blue", alpha = 0.7, color = "black") +
-  theme_minimal() +
-  labs(x = "Percentage of Support (pct)",
-       y = "Count")
+dataset <- rbind(vendor_avg_apple, vendor_avg_mango, vendor_avg_pear, vendor_avg_orange)
+datao <- rbind(apple, mango, pear, orange)
 
-```
-```{r}
-#| label: tbl-pctfilter1
-#| tbl-cap: Summary statistics for the proportion (%) of support that Harris received in high-quality polls
-#| echo: false
-#| tbl-align: center
-#| tbl-width: 80%
-
-# Calculate summary statistics for the 'pct' variable
-summary_stats <- harris_data %>%
-  summarize(
-    mean = round(mean(pct, na.rm = TRUE), 2),
-    median = round(median(pct, na.rm = TRUE), 2),
-    min = round(min(pct, na.rm = TRUE), 2),
-    max = round(max(pct, na.rm = TRUE), 2),
-    sd = round(sd(pct, na.rm = TRUE), 2),
-    n = n()
-  )
-
-# Display the summary statistics in a nicely formatted table
-summary_stats %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
-
-```
-
-\newpage
-
-```{r}
-#| label: fig-pctfilter2
-#| fig-cap: Distribution of the proportion (%) of support that Harris received in high-quality polls
-#| echo: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Plot the distribution of 'pct' using a histogram
-ggplot(harris_data, aes(x = pct)) +
-  geom_histogram(binwidth = 1.5, fill = "blue", alpha = 0.7, color = "black") +
-  theme_minimal() +
-  labs(x = "Percentage of Support (pct)",
-       y = "Count")
+# Plot the distribution of prices with a histogram
+ggplot(datao, aes(x = avg_price)) +
+  geom_histogram(binwidth = 1, fill = "grey", color = "black") + # Set bin width and colors
+  labs(
+       x = "Average Price",
+       y = "Frequency") +
+  theme_minimal()
 
@@
 
-### The Date the Poll was Concluded
+### Type of Fruits: The Product that We Specified to Compare the Price Distribution
 
-The 'end_date' variable representing the time the poll was concluded was put into account to keep track of how support for a candidate changes in time.
-The reported end dates in the original dataset [@fivethirtyeight2024] ranges from 7 April, 2021 to 29 October, 2024 (@tbl-date).
-@fig-date shows that the polling data is more concentrated on survey results conducted in the recent period.
+@fig-fruits shows the frequency of observations for four fruit types: apples, mangoes, oranges, and pears.
+Apples have the highest count, with around 50 instances, possibly due to popularity or market demand.
+Pears follow with a moderate count of about 20.
+Mangoes and oranges have the lowest counts, around 10 each, possibly due to factors like seasonality or regional availability.
 
 ```{r}
-#| label: tbl-date
-#| tbl-cap: Summary statistics for the date that the poll was concluded
-#| echo: false
-#| tbl-align: center
-#| tbl-width: 80%
-
-# Convert 'end_date' to a proper date format
-raw_elections_dataa <- raw_elections_data %>%
-  mutate(end_date = mdy(end_date))
-filtered_data <- raw_elections_dataa %>%
-  filter(end_date >= as.Date("2023-01-01"))
-
-summary_stats <- raw_elections_dataa %>%
-  summarize(
-    Min = min(end_date, na.rm = TRUE),
-    Max = max(end_date, na.rm = TRUE)
-  )
-
-# Display the summary statistics in a nicely formatted table
-summary_stats %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
-
-```
-
-\newpage
-
-```{r}
-#| label: fig-date
-#| fig-cap: Distribution of the date that the poll was concluded
-#| echo: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Distribution of 'end_date'
-ggplot(filtered_data, aes(x = end_date)) +
-  geom_histogram(fill = "grey", alpha = 0.7, color = "black", binwidth = 30) +  # Set binwidth explicitly
-  theme_minimal() +
-  labs(x = "End Date",
-       y = "Count")
-
-```
-
-When filtering the data, not only pollster quality and candidate type but also the date variable was considered.
-The filtered data contains only the polling data after the declaration of Harris.
-So, the date variable for filtered data in @tbl-date2 ranges from 23 July, 2024 to 29 October, 2024.
-@fig-date2 shows that overall, polling is conducted regularly but intensifies around specific dates.
-
-```{r}
-#| label: tbl-date2
-#| tbl-cap: Summary statistics for the date that high-quality polls for Harris was concluded
-#| echo: false
-#| tbl-align: center
-#| tbl-width: 80%
-
-summary_stats <- harris_data %>%
-  summarize(
-    Min = min(end_date, na.rm = TRUE),
-    Max = max(end_date, na.rm = TRUE)
-  )
-
-# Display the summary statistics in a nicely formatted table
-summary_stats %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
-
-```
-
-\newpage
-
-```{r}
-#| label: fig-date2
-#| fig-cap: Distribution of the date that high-quality polls for Harris was concluded
-#| echo: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Convert 'end_date' to a proper date format
-harris_dataa <- harris_data %>%
-  mutate(end_date = ymd(end_date))  # Change mdy() to ymd() or dmy() depending on the format
-
-# Distribution of 'end_date'
-ggplot(harris_dataa, aes(x = end_date)) +
-  geom_histogram(fill = "grey", alpha = 0.7, color = "black", binwidth = 10) +  # Set binwidth explicitly
-  theme_minimal() +
-  labs(x = "End Date",
-       y = "Count")
-
-```
-
-### Pollster and State
-
-The 'pollster' and 'state' variable were selected to consider the effect of changes in poll-making organizations and geographical distinctions.
-The two variables respectively represent the polling organization that conducted the poll and the US state where the poll was conducted or focused.
-
-@tbl-predictors1 shows that the original dataset [@fivethirtyeight2024] contains 232 distinct pollsters and 56 distinct states.
-After filtering for high-quality polls and assigning 'other' for states with fewer than 60 polls, the analysis data contains 8 distinct poll-making organizations and 20 geographical distinctions as shown in @tbl-predictors2.
-
-```{r}
-#| label: tbl-predictors1
-#| tbl-cap: Number of distinct polling organizations and US states where the poll was conducted
-#| echo: false
-#| tbl-align: center
-#| tbl-width: 80%
-
-summary_stats <- raw_elections_data %>%
-  summarize(
-    Pollster = n_distinct(pollster),
-    State = n_distinct(state)
-  )
-
-# Display the summary statistics in a nicely formatted table
-summary_stats %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
-
-```
-
-```{r}
-#| label: tbl-predictors2
-#| tbl-cap: Number of distinct high-quality polling organizations and US states where more than 60 polls for Harris were conducted
-#| echo: false
-#| tbl-align: center
-#| tbl-width: 80%
-
-summary_stats <- harris_data %>%
-  summarize(
-    Pollster = n_distinct(pollster),
-    State = n_distinct(state)
-  )
-
-# Display the summary statistics in a nicely formatted table
-summary_stats %>%
-  kable() %>%
-  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
-
-```
-
-The distribution of polling counts for different pollsters in @fig-pollster suggests that the analysis data is dominated by a few pollsters, particularly Siena/NYT.
-Depending on their polling methodology, the general results may have potential biases.
-A detailed analysis of the polling methodology and possible errors of the organization will be covered in @sec-appendix.
-
-```{r}
-#| label: fig-pollster
-#| fig-cap: Distribution of polling organizations where high-quality polls for Harris were conducted
-#| echo: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Distribution of 'pollster' (bar plot since it's categorical)
-ggplot(harris_data, aes(x = pollster)) +
-  geom_bar(fill = "purple", alpha = 0.7, color = "black") +
-  theme_minimal() +
-  labs(x = "Pollster",
-       y = "Count") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-```
-
-@fig-state displays the distribution of polls across different states in the analysis data.
-Pennsylvania, Arizona, and Georgia are the top 3 states with high number of polls while states like Montana, New Mexico, and Maryland have much fewer polls.
-Note that a significant number of national or unspecified state-level polls are aggregated in this analysis data regarding the high count in 'Other' category.
-The concentration of polls in certain states further suggests a strategic focus on areas likely to impact the election outcome [@11alive2024].
-
-```{r}
-#| label: fig-state
-#| fig-cap: Distribution of US states where the more than 60 polls for Harris were conducted
-#| echo: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Distribution of 'state' (bar plot)
-ggplot(harris_data, aes(x = reorder(state, -table(state)[state]))) +
-  geom_bar(fill = "black", alpha = 0.7, color = "black") +
-  theme_minimal() +
-  labs(x = "State",
-       y = "Count") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-
-```
-
-## Correlation between predictor variables
-
-By examining the relationships between predictor variables in the analysis dataset, we aim to identify potential biases in interpretation and model instability.
-This is expected to ensure a transparent understanding of the model's robustness and interpretability.
-
-### End Date and State
-
-@fig-correlation1 shows notable polling concentrations on certain states such as Pennsylvania, Ohio, and Nevada in week 2024-37 (September 9, 2024) or 2024-38 (September 15, 2024).
-This suggests that these states are key battelgrounds or areas of strategic focus during the election period.
-Moreover, polling activities are not consistent across all weeks and there are clear peaks in polling activity in week 2024-38 (September 15, 2024), which may correspond to significant political events, debates, or media focuses.
-
-\newpage
-
-```{r}
-#| label: fig-correlation1
-#| fig-cap: "Polling concentration over time by state (*Note:* Strategic focus of polling is directed in battleground states such as Pennsylvania and Ohio in September 15, 2024)"
-#| echo: false
-#| fig-align: center
-#| page-break-after: true
-
-# Aggregate the number of polls by state and by week (or month)
-poll_time_state_2 <- harris_data %>%
-  mutate(week = format(as.Date(end_date), "%Y-%U")) %>%  # Group by week of the year
-  group_by(state, week) %>%
-  summarise(total_polls = n(), .groups = "drop")  # Adding .groups = "drop"
-
-# Plot a heatmap of polling activity by state and week
-ggplot(poll_time_state_2, aes(x = week, y = state, fill = total_polls)) +
-  geom_tile(color = "white") +
-  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey90") +
-  labs(x = "Week",
-       y = "State",
-       fill = "Total Polls") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text = element_text(size = 8))
-
-```
-
-### End Date and Pollster
-
-@fig-correlation2 shows that Siena/NYT, which made up the largest proportion of polls in our analysis dataset, has concentrated polling efforts in week 2024-37 (September 9, 2024) and week 2024-38 (September 15, 2024).
-Notably active periods, possibly around key election events, might introduce temporal bias and overemphasize methodologies of Siena/NYT in that specific time period.
-Emerson shows overall consistency of polling activities across several weeks indicating steady involvement, while other polling organizations show more sporadic or minimal involvement.
-
-\newpage
-
-```{r}
-#| label: fig-correlation2
-#| fig-cap: "Polling concentration over time by pollster (*Note:* Concentrated polling efforts of Siena/NYT are shown in the first half of September, 2024.)"
-#| echo: false
-#| fig-align: center
-#| page-break-after: true
-
-# Aggregate the number of polls by state and by week (or month)
-poll_time_pollster <- harris_data %>%
-  mutate(week = format(as.Date(end_date), "%Y-%U")) %>%  # Group by week of the year
-  group_by(pollster, week) %>%
-  summarise(total_polls = n(), .groups = "drop")  # Adding .groups = "drop"
-
-# Plot a heatmap of polling activity by state and week
-ggplot(poll_time_pollster, aes(x = week, y = pollster, fill = total_polls)) +
-  geom_tile(color = "white") +
-  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey90") +
-  labs(x = "Week",
-       y = "Pollster",
-       fill = "Total Polls") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text = element_text(size = 8))
-
-```
-
-### Pollster and State
-
-@fig-correlation3 shows that Siena/NYT has the strongest state-level presence in key battleground states like Michigan and Arizona.
-Others like AtlasIntel and Beacon/Shaw has minimal polling coverage in targeted states such as Nevada and Georgia.
-This variation in coverage could influence the overall analysis in that some states might receive disproportionately more attention from certain pollsters.
-
-\newpage
-
-```{r}
-#| label: fig-correlation3
-#| fig-cap: "Polling activity by pollster and state (*Note:* Siena/NYT has strongest state-level polling coverage in key battleground states)"
-#| echo: false
-#| page-break-after: true
-
-# Step 1: Filter out NAs from pollster column and summarize data by state and pollster
-pollster_state_data <- harris_data %>%
-  filter(!is.na(pollster)) %>%  # Remove NAs from pollster
-  group_by(state, pollster) %>%
-  summarise(total_polls = n(), .groups = "drop")  # Summarize number of polls
-
-# Step 2: Plot the faceted map using plot_usmap
-plot_usmap(data = pollster_state_data, values = "total_polls", regions = "states") +
-  facet_wrap(~ pollster, ncol = 4) +  # Create a map for each pollster
-  scale_fill_distiller(palette = "Blues", direction = 1, na.value = "gray90") +  # Use blue color palette
-  labs(fill = "Total Polls") +
-  theme_void() +  # Remove axes and gridlines
-  theme(strip.text = element_text(size = 9))  # Customize facet labels
-
-```
-
-# Model {#sec-model}
-
-Our modelling strategy estimates Kamala Harris's support percentage in the 2024 US election polls, accounting for potential variations over time including pollster and state differences.
-The model balances complexity with interpretability, using both linear and Bayesian frameworks to capture patterns in the data.
-We turned from the initial linear model to a Bayesian model for robust predictions considering the variables' uncertainties.
-The Bayesian approach, which incorporates both prior beliefs and observed data to make predictions, allowed us to capture the variances among pollsters and states more accurately and responsively.
-Further background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the percentage of support that Kamala Harris receives in poll $i$.
-We begin with two simple linear models and progress to more complex Bayesian models that account for hierarchical structures.
-
-The following models outline our approach:
-
-### Linear Model by Date
-
-```{=latex}
-\begin{align}  
-y_i &= \beta_0 + \beta_1 \cdot \text{end\_date}_i + \epsilon_i \\
-\epsilon_i &\sim \text{Normal}(0, \sigma^2)
-\end{align}
-```
-
-where:
-- $y_i$ is the percentage of support for Harris in poll $i$,
-- $\beta_0$ is the intercept,
-- $\beta_1$ represents the effect of the poll's end date,
-- $\epsilon_i$ is the error term.
-
-### Linear Model by Date and Pollster
-
-```{=latex}
-\begin{align}  
-y_i &= \beta_0 + \beta_1 \cdot \text{end\_date}_i + \gamma_{p[i]} + \epsilon_i \\
-\epsilon_i &\sim \text{Normal}(0, \sigma^2)
-\end{align}
-```
-
-where:
-- $\gamma_{p[i]}$ is a fixed effect for pollster $p$ conducting poll $i$ (e.g., Siena/NYT).
-
-### Bayesian Model with Random Intercept for Pollster
-
-```{=latex}
-\begin{align}  
-y_i | \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
-\mu_i &= \beta_0 + \beta_1 \cdot \text{end\_date}_i + \gamma_{p[i]} \\
-\gamma_p &\sim \text{Normal}(0, \sigma_{\gamma})
-\end{align}
-```
-
-where:
-- $\gamma_p$ is a random effect for pollster $p$.
-
-### Bayesian Model with Random Intercept for Pollster and State
-
-```{=latex}
-\begin{align}  
-y_i | \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
-\mu_i &= \beta_0 + \beta_1 \cdot \text{end\_date}_i + \gamma_{p[i]} + \delta_{s[i]} \\
-\gamma_p &\sim \text{Normal}(0, \sigma_{\gamma}) \\
-\delta_s &\sim \text{Normal}(0, \sigma_{\delta})
-\end{align}
-```
-
-where:
-- $\delta_{s[i]}$ is a random effect for state $s$.
-
-The Bayesian models are fit using `rstanarm` in R. The priors used are weakly informative:
-- $\beta_0 \sim \text{Normal}(0, 10)$
-- $\sigma \sim \text{Exponential}(1)$
-
-### Model justification
-
-Different pollsters and states induce variations in polling results, as pollsters may have distinct methodologies and states represent diverse voter bases.
-Incorporating random effects for both pollsters and states allows us to improve the robustness of the model.
-
-These models are run through the rstanarm package [@rstanarm] in R  [@citeR], which makes Bayesian modeling available through the use of Stan's strong inference engine.
-To validate the models, RMSE and WAIC have been made use of to check the goodness of fit; Bayesian models with reduced RMSE and WAIC outperform linear models.
-We use weakly informative priors; for example, $\beta_0 \sim \text{Normal}(0, 10)$ and $\sigma \sim \text{Exponential}(1)$.
-This reflects our initial uncertainty but prevents overfitting.
-The priors were chosen conservatively to ensure that the model remains consistent.
-
-Model diagnostics, including posterior predictive checks and convergence diagnostics, were carried out to ensure the reliability of the results.
-The Bayesian models converged successfully, as indicated by $\hat{R} = 1$ for all parameters.
-
-The main assumption in these models is that the pollster and state effects can be treated as random.
-This assumes that the effects are normally distributed across pollsters and states, which may not always be accurate.
-Additionally, the model assumes that polling data is representative of the actual electorate, an assumption that can be violated if polls are biased or have non-random sampling issues.
-Despite these limitations, the hierarchical structure allows us to capture important variability, making the model suitable for predicting Harris's support.
-Future improvements could involve incorporating time-varying effects or exploring interactions between pollsters and states.
-
-# Results {#sec-results}
-
-## Results from examining the analysis data
-
-@fig-resultplot1 shows an initial increase in support for Harris since her declaration to join the electoral race, peaking around mid-September.
-However, this initial momentum stabilized with minor decrease in support as the election gets closer.
-Moreover, the scattered data points around the summary line show high variance between individual polls.
-
-\newpage
-
-```{r}
-#| label: fig-resultplot1
-#| fig-cap: "Polling votes for Harris over time (*Note:* The date starts from after Harris officially announced her campaign for 2024 U.S. presidential election on July 21, 2024.)"
+#| label: fig-fruits
+#| fig-cap: The distrubution of different categories of fruits used in the analysis
 #| echo: false
@@
 #| message: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
 
-base_plot <- ggplot(harris_data, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date")
+# Count the number of observations in each dataset
+counts <- data.frame(
+  fruit = c("apple", "mango", "pear", "orange"),
+  observation_count = c(nrow(apple), nrow(mango), nrow(pear), nrow(orange))
+)
 
-# Plots poll estimates and overall smoothing
-base_plot +
-  geom_point(alpha = 0.45) +
-  geom_smooth()
-```
-
-When accounting for the pollster difference, @fig-resultplot2 and @fig-resultplot3 show variations in Harris' support levels that are obscured in the aggregated view of @fig-resultplot1.
-Siena/NYT (pink) shows relatively stable polling results for Harris, with fewer extreme values and a slight downward trend in October.
-On the other hand, Emerson (green) shows more fluctuation throughout the period with a wider spread of polling values.
-
-```{r}
-#| label: fig-resultplot2
-#| fig-cap: "Polling votes for Harris over time by pollster (*Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
-presidential election on July 21, 2024.
-Only filtered pollsters with high-quality polling questions are shown in the following.)"
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Color by pollster
-base_plot +
-  geom_point(aes(color = pollster), alpha = 0.45) +
-  geom_smooth() + 
-  theme(legend.position = "bottom")
-```
-
-```{r}
-#| label: fig-resultplot3
-#| fig-cap: "Polling votes for Harris over time by pollster (facets) (*Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
-presidential election on July 21, 2024.
-Only filtered pollsters with high-quality polling questions are shown in the following.)"
-#| echo: false
-#| warning: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Facet by pollster
-base_plot +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x)) +  # Use GAM smoothing to handle complex data
-  facet_wrap(vars(pollster))
-```
-
-\newpage
-
-When accounting for the state difference, @fig-resultplot4 and @fig-resultplot5 show variations in Harris' support levels.
-Critical battleground states like Pennsylvania and Michigan show distinct trends from the national analysis in @fig-resultplot1.
-Pennsylvania shows relatively stable trends slightly above the national average while Michigan shows fluctuations over time.
-
-```{r}
-#| label: fig-resultplot4
-#| fig-cap: "Polling votes for Harris over time by state (*Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
-presidential election on July 21, 2024.
-Only filtered pollsters with high-quality polling questions are shown in the following.)"
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Color by pollster
-base_plot +
-  geom_point(aes(color = state)) +
-  geom_smooth() + 
-  theme(legend.position = "bottom")
-```
-
-```{r}
-#| label: fig-resultplot5
-#| fig-cap: "Polling votes for Harris over time by state (facets) (*Note:* The date starts from after Harris officially announced her campaign for 2024 U.S.
-presidential election on July 21, 2024.
-Only filtered pollsters with high-quality polling questions are shown in the following.)"
-#| echo: false
-#| warning: false
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Facet by state
-base_plot +
-  geom_point() +
-  geom_smooth(method = "gam", formula = y ~ s(x)) +  # Use GAM smoothing to handle complex data
-  facet_wrap(vars(state))
-```
-
-\newpage
-
-## Results from the prediction model
-
-Prediction results derived from our model frameworks are summarized in @tbl-modelresults1 and @tbl-modelresults2.
-In the linear models (@tbl-modelresults1), higher $R^2$ and lower WAIC, RMSE values in the second column show that the inclusion of pollster effects improves accuracy and reliability of the prediction model.
-In the Bayesian models (@tbl-modelresults2), higher Log-Likelihood and lower ELPD, WAIC values in the fourth column shows that including both pollster and state effects enhances model performance.
-
-\
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-model_date <-
-  readRDS(file = here::here("models/model_date.rds"))
-
-model_date_pollster <-
-  readRDS(file = here::here("models/model_date_pollster.rds"))
-
-bayesian_model_1 <-
-  readRDS(file = here::here("models/bayesian_model_1.rds"))
-
-bayesian_model_2 <-
-  readRDS(file = here::here("models/bayesian_model_2.rds"))
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults1
-#| tbl-cap: "Linear models of support percentages for Harris based on date and pollster"
-#| warning: false
-#| tbl-width: 80% 
-#| page-break-after: true
-#| page-break-before: true
-
-modelsummary(
-  list(
-    "Linear by Date" = model_date,
-    "Linear by Date, Pollster" = model_date_pollster
-  ),
-  statistic = "mad",
-  fmt = 2)
-
-```
-
-\newpage
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults2
-#| tbl-cap: "Bayesian models of support percentages for Harris based on pollster and state"
-#| warning: false
-#| tbl-align: center
-#| tbl-width: 80% 
-#| page-break-after: true
-#| page-break-before: true
-
-modelsummary(
-  list(
-    "Bayesian with Pollster" = bayesian_model_1,
-    "Bayesian with Pollster, State" = bayesian_model_2
-  ),
-  statistic = "mad",
-  fmt = 2)
-
-```
-
-@fig-modelgraph1 and @fig-modelgraph2 show the trends in the percentage of supportive polls for Harris predicted by linear models, respectively representing a time-series analysis and consideration of pollster differences.
-Moreover, @fig-modelgraph3 and @fig-modelgraph4 show the supportive trends predicted by Bayesian models accounting for pollster and state differences.
-Each model, from the simple linear regression to the more complex Bayesian models, seem to reflect a relatively consistent support pattern over time.
-This alignment of underlying support trends across model shows the credibility of the observed overall support trend.
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-modelgraph1
-#| fig-cap: "Predicted percentage of supportive polls for Harris in the linear model by date"
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Generate predictions and plot for Model 1 (pct ~ end_date)
-harris_data$pred_date <- predict(model_date, newdata = harris_data)
-
-ggplot(harris_data, aes(x = end_date, y = pct)) +
-  geom_point(color = "black", alpha = 0.6) +
-  geom_line(aes(y = pred_date), color = "blue") +
+# Plot the observation count for each fruit
+ggplot(counts, aes(x = fruit, y = observation_count)) +
+  geom_bar(stat = "identity", fill = "skyblue") +
   labs(
-    x = "Date",
-    y = "Predicted percentage of supportive polls"
-  ) +
-  theme_minimal()
-```
-
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-modelgraph2
-#| fig-cap: "Predicted percentage of supportive polls for Harris in the linear model by date and pollster (*Note:* The blue summary line calculates the mean of the predictions across all dates and pollsters.)"
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-# Generate predictions for Model 2, but ignore the individual pollster grouping in the plot
-# Calculate the mean predictions across all pollsters for each end_date
-harris_data$pred_date_pollster <- predict(model_date_pollster, newdata = harris_data)
-
-# Calculate the mean of the predictions across all dates and pollsters
-overall_mean_pred <- mean(harris_data$pred_date_pollster, na.rm = TRUE)
-
-# Plot the observed data points and a single straight summary line at the overall mean prediction
-ggplot(harris_data, aes(x = end_date, y = pct)) +
-  geom_point(aes(color = pollster), alpha = 0.6) +  # Observed data with colors by pollster
-  geom_hline(yintercept = overall_mean_pred, color = "blue") +
-  labs(
-    x = "Date",
-    y = "Predicted percentage of supportive polls"
-  ) +
+       x = "Fruit Type",
+       y = "Observation Count") +
   theme_minimal()
@@
 
+### Type of Vendors
+
+@fig-vendors provides a visualization of the number of observations for each vendor in the dataset, specifically Loblaws, NoFrills, and Voila.
+The x-axis represents these three vendors, while the y-axis shows the count of observations, likely indicating the number of unique products or price points recorded for each vendor.
+Each bar is labeled with the exact count: Loblaws has 33 observations, NoFrills has 31, and Voila has the highest at 36.
+These counts are relatively similar across vendors, suggesting a fairly balanced representation in the data, although Voila has a slight edge in the number of entries.
+This balanced distribution is beneficial for comparative analysis, as it reduces the risk of one vendor disproportionately influencing results.
+The slight difference in counts could introduce a minor weight toward Voila in any statistical analysis, but overall, the dataset’s even distribution supports reliable comparisons.
+Thus, any averages or other metrics calculated by vendor are likely to be comparable across Loblaws, NoFrills, and Voila, minimizing selection bias due to unequal observation counts.
+
 ```{r}
+#| label: fig-vendors
+#| fig-cap: The distribution of vendors used in the analysis
+#| echo: false
+#| warning: false
+
+# Plot the distribution of observation count by vendor with adjusted bar width and labels
+ggplot(datao, aes(x = vendor)) +
+  geom_bar(fill = "purple", width = 0.4) + # Narrower bar width
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) + # Add count labels above bars
+  labs(
+       x = "Vendor",
+       y = "Observation Count") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability
+```
+
+# Results
+
+@fig-results illustrates the average prices of four fruits—apples, mangoes, oranges, and pears—across three different vendors: Loblaws, NoFrills, and Voila.
+Each fruit is represented by a distinct color, as indicated in the legend: red for apples, green for mangoes, blue for oranges, and purple for pears.
+The bar plot component of the chart displays the average price of each fruit at the respective vendors.
+Notably, apples are priced highest at Loblaws and lowest at Voila.
+Mangoes exhibit minimal price variation across the vendors.
+Oranges show a significant price difference, being most expensive at Loblaws and least expensive at Voila.
+Similarly, pears are priced lowest at Voila.
+Superimposed on the bar plot is a line plot that connects the average prices of each fruit across the vendors, highlighting the price trends.
+The lines for apples, oranges, and pears demonstrate a downward trend from Loblaws to Voila, indicating decreasing prices.
+In contrast, the line for mangoes remains relatively flat, suggesting consistent pricing among the vendors.
+In summary, the chart effectively compares the pricing strategies of different vendors for various fruits, revealing that Voila generally offers lower prices for most fruits, while Loblaws tends to have higher prices.
+These visual trends provide insights into the pricing correlations among the vendors and set the stage for further analysis of potential causal factors influencing these pricing differences.
+
+```{r}
+#| label: fig-results
+#| fig-cap: Price distribution of different kinds of fruits(apple, mango, orange, pear) by vendors
 #| echo: false
@@
 #| message: false
-#| label: fig-modelgraph3
-#| fig-cap: "Predicted percentage of supportive polls for Harris in the Bayesian model with random intercept for `pollster` variable (*Note:* In order to avoid the overfitting problem, overall mean of prediction was separately calculated in creating the summary line.)"
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
 
-# Generate predictions for Bayesian Model 1
-harris_data$bayesian_pred_1 <- posterior_predict(bayesian_model_1, newdata = harris_data) |> apply(2, mean)
-
-# Calculate the overall mean prediction
-overall_mean_pred <- mean(harris_data$bayesian_pred_1 / harris_data$sample_size * 100, na.rm = TRUE)
-
-# Plot the observed data points and a single straight line at the overall mean prediction
-ggplot(harris_data, aes(x = end_date, y = num_harris / sample_size * 100)) +
-  geom_point(color = "black", alpha = 0.6) +
-  geom_hline(yintercept = overall_mean_pred, color = "blue") +
-  labs(
-   x = "Date",
-   y = "Predicted percentage of supportive polls"
-  ) +
-  theme_minimal()
-
+ggplot(dataset, aes(x = vendor, y = avg_price, fill = fruit, group = fruit)) +
+  geom_bar(stat = "identity", position = "dodge", aes(color = fruit)) +
+  geom_line(aes(color = fruit)) +
+  geom_point(aes(color = fruit)) +
+  facet_wrap(~fruit, scales = "free_y") +
+  labs(x = "Render", y = "Price") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ```
 
-```{r}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: fig-modelgraph4
-#| fig-cap: "Predicted percentage of supportive polls for Harris in the Bayesian model with random intercept for `pollster` and `state` variable (*Note:* In order to avoid the overfitting problem, overall mean of prediction was separately calculated in creating the summary line.)"
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
+# Discussion
 
-# Generate predictions for Bayesian Model 2
-harris_data$bayesian_pred_2 <- posterior_predict(bayesian_model_2, newdata = harris_data) |> apply(2, mean)
+## Correlation vs. Causation
 
-# Calculate the overall mean prediction
-overall_mean_pred_2 <- mean(harris_data$bayesian_pred_2 / harris_data$sample_size * 100, na.rm = TRUE)
+When analyzing the pricing of the same fruit in different supermarkets, it is very important to distinguish Correlation and Causation for understanding the root cause of price difference.
+We have observed that the prices of Loblaws apples, pears, orange and mango are always higher than those of other supermarkets (such as NoFrills and Voila), and the price of each fruit in Voila is always lower than that in other stores, which shows that there is a correlation between supermarket and fruit pricing.
+This correlation may be related to the market positioning, operating costs and supply chain management of supermarkets.
+For example, Loblaws is positioned in the high-end market and can provide better products and services to middle-class customers, so its pricing is higher.
+Voila mainly serves groups with limited budgets, so the price is lower.
+In addition, factors such as geographical location, rent and employee salary may lead to differences in operating costs, which in turn may affect pricing strategies.
+For example, Loblaws is located in the city center, the rent is high, and the salary of employees is relatively high because of the high service level.
+In supply chain management, for example, the procurement channels of different supermarkets and the efficiency of supply chain may lead to cost differences and affect the final selling price.
+In addition, to determine whether these factors really lead to price differences, further data and analysis are needed, such as market positioning research, cost structure analysis and supply chain evaluation.
+Through these in-depth analysis, we can understand the causal relationship of price difference more accurately.
 
-# Plot the observed data points and a single straight line at the overall mean prediction
-ggplot(harris_data, aes(x = end_date, y = num_harris / sample_size * 100)) +
-  geom_point(color = "black", alpha = 0.6) +
-  geom_hline(yintercept = overall_mean_pred_2, color = "blue") +
-  labs(
-   x = "Date",
-   y = "Predicted percentage of supportive polls"
-  ) +
-  theme_minimal()
+## Missing Data
 
-```
+In this dataset, missing data occurs because different supermarkets offer varying selections of fruit varieties.
+For example, some supermarkets may carry a wide range of apple varieties, while others only stock a few.
+As a result, when analyzing data across all supermarkets, information on certain apple varieties may be incomplete, with specific types available only from certain vendors.
+Similar patterns may apply to other types of fruit, leading to gaps in the dataset.
+This missing data reflects differences in product offerings across vendors, and it's essential to be aware of and consider this incompleteness when conducting further analysis.
 
-# Discussion {#sec-discussion}
+## Sources of Bias
+In this dataset, we study the price differences and potential changes of 4 fruits between different supermarkets.
+The price differences are due to the different positioning of each supermarket, Loblaw, NoFrills, and Voila.
 
-## Why Harris could beat her polls {#sec-first-point}
+### Selection Bias
+As mentioned before, selection bias in this dataset mainly comes from the differences in customer groups of different supermarkets.
+Each supermarket (such as Loblaws, NoFrills, and Voila) has a different market positioning and customer group.
+Customers who choose to shop at Loblaws usually prefer high-end goods or organic foods, while NoFrills customers may pay more attention to low prices and discounts.
+This customer preference leads to bias in the price data, that is, the price differences between different supermarkets not only reflect the market price of the product itself, but also are affected by customer selection behavior.
+Therefore, Loblaws' higher prices do not necessarily mean that all its fruits are priced higher than other supermarkets, but rather reflect that its customers are willing to pay a premium for higher quality or brand.
 
-@tbl-pctraw1 and @tbl-pctfilter1 shows that filtering data based on more reliable polling criteria makes the average rate of support for Harris increase.
-Given the current polling landscape where Harris and Trump have nearly equivalent levels of support [@factcheck2024], higher support rates for Harris in the filtered analysis serves as substantial evidence that Harris could win the election.
+### Measurement Bias
+In this dataset, measurement bias may come from the different ways in which supermarkets record or define the average price (avg_price).
+For example, Loblaws may be promoting a specific brand or organic fruit, making the recorded average price lower than the actual price, while NoFrills may use a specific price for a single batch, ignoring other higher price ranges.
+These differences in measurement may cause distortions in price data, making our price comparisons between different supermarkets biased.
 
-Furthermore, our data analysis that incorporates state-specific factors (@fig-resultplot5) reveals that Harris holds relatively stable support in key battelground states, including Pennsylvania, Michigan, and Wisconsin.
-As it will be discussed in more detail in @sec-third-point, this suggests that Harris' support base is both resilient and potentially insulated from the unpredictable swings [@kff2024].
-Regarding the winner-take-all nature of the electoral college, Harris' stability in these key states not only strengthens her position but also shows high probability in her victory, as even slight leads in battleground states can result in a decisive electoral advantage [@kff2024].
-The combination of broad polling support and specific regional strength positions Harris as a leading presidential candidate in the 2024 election.
-
-## Pollsters herding around false consensus {#sec-second-point}
-
-@fig-resultplot2 and @fig-resultplot3 shows notable differences in the supportive trends for Harris when considering the variations of polling organizations.
-In @tbl-modelresults1, the inclusion of pollster-specific variables had statistical significance.
-This proposes that pollster-specific factors, like methodology and sample composition, produce a range of outcomes even for the same candidate's support levels [@nytimes2024].
-As  pollsters often fear publishing outlier results that might undermine credibility, they may adjust the results with a perceived consensus [@silver2024].
-This is likely to reduce the accuracy of polling aggregates, misrepresenting the true differences in public sentiment.
-
-To be more specific, this adjustment based on previous consensus data would limit the visibility of demographic shifts that might favor one candidate unexpectedly [@aapor_herding].
-For instance, the polling discrepancies in the 2016 and 2020 U.S election were partly attributed to underestimated certain voter segments supportive of Trump [@silver2024].
-Assessing the potential limitations of aggregated data, recognizing that it may reflect methodological biases more than independent public opinion should be on the way for a reliable prediction of electoral support.
-
-## The electoral college and the power of battleground states {#sec-third-point}
-
-In the U.S.
-electoral college, the Constitution assigns each state a set number of electoral based on its Congressional representation and a candidate must secure more than 270 votes to win the presidency [@whyy2024].
-This often results in candidates focusing their campaigns primarily on battleground states, where voter preferences are more fluid and results are less predictable [@ashcenter2024].
-@fig-resultplot5 show that key battleground states' polling trends differ from that of the national average (@fig-resultplot1).
-
-The stability of polling trends in Pennsylvania suggests a consistent voter base despite of the battleground status, which often means that even slight shifts in support could be pivotal [@ashcenter2024].
-Harris' focus on healthcare reform, especially her commitment to strengthening the Affordable Care Act (ACA), is expected to have particular resonance in Pennsylvania, where a large population of low-to-middle income residents benefit from ACA subsidies [@factcheck2024].
-On the other hand, Michigan, where there are economically diverse population, national economic policy discussions could swing voter sentiment more sharply and result in relatively large fluctuations [@kff2024].
-Surveys carried out within battleground states show wider variation in candidate support, making the incorporation of state-specific factors into electoral models allow a better representation of the diverse political landscapes seen across the United States.
-
-## Weaknesses and next steps
-
-While our model provides a foundation to understand Harris' polling support, further refinements are necessary to enhance its accuracy and applicability.
-For instance, future studies could focus on the incorporation of voter demographics, such as age, gender, and education.
-Showing which segments of the population are driving changes in support, this can give further explanation to how certain policies or campaign events of candidates affect overall support for a candidate [@brown2024candidate].
-In addition, interaction terms between pollster and state can be included in the model, considering that some states might receive disproportionately more attention from certain pollsters as shown in @fig-correlation3.
-This can give further explanation to how pollsters and regional dynamics affect overall support for a candidate [@pew2020electionpolls].
-
-This study has its constraints in that sampling error and systematic biases inevitably occur in the process of measuring social phenomenon with data.
-In particular, due to the nature of polling surveys, non-responses or misunderstanding of the survey questions result in missing data or outliers [@pew2020electionpolls].
-Considering the possibility of prediction errors, we should take caution in interpreting the analysis and prediction results.
-By conducting more in-depth research of the domain knowledge and constructing appropriate variables to put into account, it would help derive accurate conclusions rather than just acknowledging the numbers given itself [@geeksforgeeks2023domainknowledge].
-
-\newpage
-
-\appendix
-
-# Appendix {#sec-appendix}
-
-## Pollster methodology overview and evaluation
-
-The New York Times/Siena College polling partnership, the polling organization that accounted for the majority of polls in our analysis (@fig-pollster), conducts polls tailored for specific elections, such as state or national races [@nyt_siena_2020_methodology].
-Their sample size typically includes 600 to 1000 likely voters per poll, with oversampling in battleground states to capture regional nuances [@nyt_siena_2020_methodology].
-The methodology uses random-digit dialing (RDD) for landlines and mobile phones to ensure representative sample coverage across demographics.
-In addition, online surveys are administered to complement phone-based responses, ensuring broader accessibility [@nyt_siena_2020_methodology].
-The stratified random sampling approach is employed, where the population is divided into strata (based on demographic variables like race, education, and geography), and a random sample is drawn from each stratum [@tellingstories].
-This allows for precision in reflecting the political leanings and key demographic shifts in specific regions.
-
-The organization intends to enhance transparency in how public opinion is assessed, ensuring that questions are carefully designed to represent contemporary political discussions, and that the terminology is polished through a process of iterative testing to achieve clarity.
-They devote extensive resources to cognitive testing to ensure question wording reflects what the public thinks [@siena_scri_2024_poll].
-Their polling methodology stands out in that its strategic focus on using representative samples reflect political leanings and demographics of a region for more contextual and precise polling.
-Siena/NYT has its reputation for accurately predicting key battleground state outcomes during previous elections, such as Florida in 2016[@siena_nyt_perfect_partnership_2024].
-
-The limitations of Siena/NYT's methodology are the challenge of polling itself.
-Since polling is a "snapshot in time",  the results can fluctuate based on recent political events or campaign dynamics.
-Additionally, there are still issues with non-response bias in polling, particularly among the hard-to-reach voter or voters suspicious of polling organizations themselves [@pew2023polling].
-
-## Idealized methodolgy
-
-The proposed methodology for forecasting the 2024 U.S.
-presidential election with a budget of $100,000 would be designed as follows.
-First, a stratified random sampling method will be employed that allows for the capture of the demographic elements such as age, gender, race, and education level [@pew2023polling].
-We will focus on 10 battleground states including Florida, Pennsylvania, Michigan, and Arizona, each receiving a budget of $7,000.
-For each state, $5,000 would be allocated for telephone and online surveys, while $2,000 would be allocated for training local staff for telephone outreach and conducting data entry to maintain data quality and consistency.
-This targeted investment could alleviate bias and make the sample more representative of the population [@pew2023polling], enhancing the reliability and inclusiveness of the gathered data.
-
-In addition to robust sampling, the methodology incorporates high-quality questionnaire design and testing.
-We would invest $15,000 in this with $5,000 on initial drafting and expert review to ensure questions are clear and unbiased.
-$7,000 would be spend on conducting pre-tests and pilot studeies for refinement [@driveresearch_cognitive_testing] and $3,000 would be allocated to final review and translation of the survey questions.
-Cognitive testing and iterative feedback loops are expected to improve the validity and accuracy of the collected data to derive reliable prediction results [@presser2004pretesting].
-
-Then, $10,000 would be assigned for data processing and weighting.
-Weighting methods that account for groups that are underrepresented ensure that the outcomes are not skewed by sampling errors [@pew2023polling].
-Post-stratification weighting which ensures that the sample accurately represents the U.S voting population will take up $5,000 while data cleaning and consistency checks for preventing bias from erroneous entries or duplicates will take up $5,000.
-
-The final model would use Bayesian hierarchical modeling, which allows for more flexible modeling of uncertainty and variation across states, pollsters, and other external factors.
-These models, along with out-of-sample testing and cross-validation, enable accurate prediction sensitive to the dynamics of real-world changes, including political events [@pew2023polling].
-By investing $3,000 on model setup and calibration and $2,000 on cross-validation and testing, we expect to ensure the model remains robust and provides reliable predictions under various electoral scenarios.
-
-### Idealized survey
-
-The proposed survey questionnarie design is in the following link:
-\
-https://forms.gle/zQ8iJyPk3HhJYrKd9.
-
-::: {.layout}
-### Survey demo
-
-::: {.column width="1/5"}
-![Survey intro](../other/idealized_survey/survey_intro.png)
-:::
-
-::: {.column width="1/5"}
-![Survey questions asking about demographics](../other/idealized_survey/survey_q1.png)
-:::
-
-::: {.column width="1/5"}
-![Survey questions asking for voting intentions and candidate favorability](../other/idealized_survey/survey_q2.png)
-:::
-
-::: {.column width="1/5"}
-![Survey questions asking for issues of interest and likelihood to vote](../other/idealized_survey/survey_q3.png)
-:::
-
-::: {.column width="1/5"}
-![Survey final thanks](../other/idealized_survey/survey_outro.png)
-:::
-:::
-
-## Model details {#sec-model-details}
-
-### Posterior predictive check
-
-In the first posterior predictive check (@fig-ppcheck-1), we compare the observed data with replicated data generated from the posterior distribution.
-This shows that the model is able to replicate the overall distribution of the observed data, with the replicated curves (light blue) closely following the true data (dark blue line).
-This indicates that the model fits the data well in terms of capturing the main pattern or trend [@stan2023posteriorchecks].
-
-In the second plot (@fig-ppcheck-2), the replicated data which had both the pollster and state variable as random intercepts, shows relatively closer approximation to the true data distribution.
-The narrowing of uncertainty in the posterior relative to the prior indicates the impact of the data on refining the model's predictions.
-This reassures that the model fits the data reasonably well and that the prior information has been appropriately updated by the observed data [@stan2023posteriorchecks].
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheck
-#| layout-ncol: 2
-#| fig-cap: "Examining how the Bayesian model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check"]
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-pp_check(bayesian_model_1) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-pp_check(bayesian_model_2) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-```
-
-### Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot.
-The sampled values for posterior distribution of intercept parameter across iterations of the MCMC algorithm shows good convergence [@stan_mcmc_traces].
-The lines for the parameter appear to be stable and fluctuating around a central value without any clear trends or patterns.
-This suggests that the MCMC algorithm has likely converged, and the posterior samples are representative of the target distribution.
-
-@fig-stanareyouokay-2 is a Rhat plot.
-The Rhat value is approximately 1.0 for the intercept, which shows that the variance within and between multiple chains have converged.
-An Rhat value close to 1 indicates that the chains have mixed well and are drawing from the same distribution while values significantly greater than 1 would indicate that further iterations are needed [@stan_mcmc_diagnostics].This suggests that the Bayesian models for both "pollster" and "state" have likely converged, and the results derived from these models are reliable.
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-#| fig-align: center
-#| fig-width: 5
-#| page-break-after: true
-
-plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
-
-plot(bayesian_model_2, pars = "(Intercept)", prob = 0.95)
-
-```
+### Confounding Bias
+In this dataset, confounding bias may come from factors such as fruit type, quality, packaging type, seasonal changes, etc.
+For example, Loblaws may be more inclined to sell high-quality or organic apples, which are usually priced higher, and this tendency is not shown separately in the data.
+Therefore, the price difference between supermarkets may not be due to the supermarket itself, but to uncontrolled fruit quality or seasonal factors.
 

### 2024-11-14 paper/paper.qmd
update introduction and title

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Price Comparison of Common Fruits Across Major Vendors"
+subtitle: "An Analysis Revealing Loblaws as the Highest Priced Vendor Among Apples, Oranges, Mangoes, and Pears"
 author: 
@@
   - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+thanks: "Code and data are available at: [https://github.com/koyunkyung/canadian_grocery](https://github.com/koyunkyung/canadian_grocery)."
 date: today
@@
 
-```{r}
-#| include: false
-#| warning: false
-#| message: false
-
-library(tidyverse)
-library(palmerpenguins)
-```
-
 # Introduction
 
-Overview paragraph
-
-Estimand paragraph
-
-Results paragraph
-
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+This study investigates price differences for four common fruits—apples, oranges, mangoes, and pears—across three major vendors: Loblaws, NoFrills, and Voila.
+By analyzing the average prices for each fruit type at each vendor, we aim to identify price patterns and determine which vendor consistently has the highest prices.
+This analysis provides insight into each vendor’s pricing strategy and helps consumers make informed decisions based on their preferences and budget.
+Our findings indicate that Loblaws generally has the highest average prices for these fruits among the vendors studied.
+In this paper, we also address several important analytical considerations to strengthen our findings.
+First, we examine correlation vs.
+causation in @sec-discussion1 to clarify that observed price differences among vendors may correlate with various factors, but further analysis would be needed to establish causation.
+Second, in @sec-discussion2, we discuss the impact of missing data due to the fact that different vendors offer a varying selection of fruit varieties, which affects the completeness of the dataset.
+Finally, in @sec-discussion3 we identify potential sources of bias—such as selection bias, measurement bias, and confounding factors—that may influence our results and interpretations.
 
@@
 
-We use the statistical programming language SQL [@citeR]
-
-.... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+The dataset was analyzed using R [@citeR] and SQL [@sqlite], and use Tidyverse [@citeTidyverse], Knitr [@citeKnitr], ggplot2[@citeGgplot2] and Dplyr [@citeDplyr].
+Data are extract from Hammer Dataset [@hammerdataset].
+The dataset provides detailed information on the average prices of four types of fruits—apples, mangoes, pears, and oranges—across three vendors: Loblaws, NoFrills, and Voila.
+It includes columns for the vendor name, the average price, and the fruit type.
+Each fruit is represented by multiple entries from each vendor, allowing for comprehensive comparisons of price distributions.
+In addition to the average price, the dataset contains observation counts for each fruit and vendor, which indicate that apples are the most frequently recorded fruit, followed by pears, while mangoes and oranges have fewer observations, possibly reflecting factors like consumer demand or seasonality.
+Vendor representation is fairly balanced, with Voila having a slight edge in the number of observations compared to Loblaws and NoFrills, supporting reliable comparative analysis.
+The dataset structure thus enables an analysis of price levels and distributions across different fruits and vendors, with potential insights into market trends and vendor-specific pricing strategies.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+This dataset, while useful for analyzing price distribution, is a constructed representation of purchasing behavior and may not fully capture the complexities of real-world consumer actions.
+Since the data does not include actual transaction records or contextual factors influencing purchases (such as seasonality, brand loyalty, or regional demand), it may lack critical elements that drive genuine purchasing decisions.
+Consequently, insights derived from this dataset should be interpreted with caution, acknowledging that it represents an approximation rather than a comprehensive model of consumer behavior.
 
@@
 ### Price Levels
+
+@fig-price reveals that the majority of products are priced below $10, indicating a concentration of lower-priced items in this dataset.
+A smaller number of items fall in the higher price ranges, with few products priced above $20, suggesting that high-priced items are outliers rather than the norm.
+This distribution suggests a market trend towards affordability or a focus on budget-friendly options for the products included in the analysis.
 
@@
 
-## Correlation vs. Causation
+## Correlation vs. Causation {#sec-discussion1}
 
@@
 
-## Missing Data
+## Missing Data {#sec-discussion2}
 
@@
 
-## Sources of Bias
+## Sources of Bias {#sec-discussion3}
 In this dataset, we study the price differences and potential changes of 4 fruits between different supermarkets.

### 2024-11-14 paper/paper.qmd
update abstract and authors

@@
 author: 
-  - First author
-  - Another author
+  - Yunkyung Ko
+  - Yawen Tan
+  - Shanjie Jiao
+  - XIzi Sun
 thanks: "Code and data are available at: [https://github.com/koyunkyung/canadian_grocery](https://github.com/koyunkyung/canadian_grocery)."
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study examines the price differences of four common fruits—apples, oranges, mangoes, and pears—across three major Canadian grocery vendors: Loblaws, NoFrills, and Voila.
+By analyzing the average prices of these fruits, we aim to identify price patterns and assess which vendor tends to offer higher or lower prices.
+Our results show that Loblaws generally has the highest prices, while Voila offers the lowest.
+This suggests variations in pricing strategies tied to each vendor's market positioning.
+Additionally, the study explores analytical considerations, including the distinction between correlation and causation, the impact of missing data, and sources of potential bias such as selection, measurement, and confounding biases.
+These factors are critical in understanding the observed pricing differences and ensuring the robustness of the analysis.
+This research provides insights for consumers making budget-conscious shopping decisions and highlights methodological challenges in analyzing observational data."
 format: pdf

### 2024-11-16 scripts/01-test_simulated_data.R
Add files via upload

@@ -1,0 +1,89 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the simulated Australian 
+  #electoral divisions dataset.
+# Author: Yunkyung Ko
+# Date: 15 November 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+  # - The `tidyverse` package must be installed and loaded
+  # - 00-simulate_data.R must have been run
+# Any other information needed? Make sure you are in the `infant_health` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+
+# Test if the data was successfully loaded
+if (exists("analysis_data")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
+
+
+#### Test data ####
+
+# Check if the dataset has 151 rows
+if (nrow(analysis_data) == 151) {
+  message("Test Passed: The dataset has 151 rows.")
+} else {
+  stop("Test Failed: The dataset does not have 151 rows.")
+}
+
+# Check if the dataset has 3 columns
+if (ncol(analysis_data) == 3) {
+  message("Test Passed: The dataset has 3 columns.")
+} else {
+  stop("Test Failed: The dataset does not have 3 columns.")
+}
+
+# Check if all values in the 'division' column are unique
+if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
+  message("Test Passed: All values in 'division' are unique.")
+} else {
+  stop("Test Failed: The 'division' column contains duplicate values.")
+}
+
+# Check if the 'state' column contains only valid Australian state names
+valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", 
+                  "Western Australia", "Tasmania", "Northern Territory", 
+                  "Australian Capital Territory")
+
+if (all(analysis_data$state %in% valid_states)) {
+  message("Test Passed: The 'state' column contains only valid Australian state names.")
+} else {
+  stop("Test Failed: The 'state' column contains invalid state names.")
+}
+
+# Check if the 'party' column contains only valid party names
+valid_parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+
+if (all(analysis_data$party %in% valid_parties)) {
+  message("Test Passed: The 'party' column contains only valid party names.")
+} else {
[…]

### 2024-11-18 scripts/00-simulate_data.R
update analysis, simulation data scripts

@@ -2,7 +2,7 @@
 # Purpose: Simulates a dataset of Australian electoral divisions, including the 
   #state and party that won each division.
 # Author: Yunkyung Ko
-# Date: 17 November 2024
+# Date: 18 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
@@ -15,42 +15,36 @@ set.seed(853)
 
 
 #### Simulate data ####
-simulation_data <- tibble(
-  dem_momage = sample(18:45, 1000, replace = TRUE), # Mother Age: 18-45
-  dem_dadage = sample(18:60, 1000, replace = TRUE), # Father Age: 18-60
-  mom_bmi = round(runif(1000, 18, 40), 1), # BMI: 18-40
-  mom_notobaco = sample(0:1, 1000, replace = TRUE, prob = c(0.8, 0.2)), # 80% Non-Smoker, 20% Smoker
-  mom_noinfec = sample(c(0, 1, 9), 1000, replace = TRUE, prob = c(0.7, 0.2, 0.1)), # No Infection: 70%, True: 20%, Not Reported: 10%
-  med_previs = sample(c(1:30, 99), 1000, replace = TRUE, prob = c(rep(0.03, 30), 0.1)), # Prenatal Visits: 1~30 or 99
-  med_wic = sample(c("Yes", "No", "U"), 1000, replace = TRUE, prob = c(0.5, 0.4, 0.1)), # WIC: Yes(50%), No(40%), Unknown(10%)
-  inft_weight = sample(c(1000:4500, 9999), 1000, replace = TRUE, prob = c(rep(0.001, 3501), 0.01)), # Birth Weight: 1000~4500g, 9999 Not Stated
-  inft_gest = sample(c(22:42, 99), 1000, replace = TRUE, prob = c(rep(0.04, 21), 0.05)), # Gestation: 22~42 weeks or 99 unknown
-  no_abnorm = sample(c(0, 1, 9), 1000, replace = TRUE, prob = c(0.6, 0.3, 0.1)), # No Abnormal Conditions
-  no_congen = sample(c(0, 1, 9), 1000, replace = TRUE, prob = c(0.7, 0.2, 0.1))  # No Congenital Anomalies
+# Generate 1000 samples for each variable
+simulation_data <- data.frame(
[…]

### 2024-11-18 scripts/03-clean_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Selects relevant variables for analysis and removes the outliers(unknown or unreported).
 # Author: Yunkyung Ko
-# Date: 17 November 2024
+# Date: 18 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
@@ -22,40 +22,29 @@ infant_health_na <-
   janitor::clean_names() |>
   
   # select relevant variables for analysis
-  select(mager, fagecomb, bmi, f_tobaco, no_infec, previs, wic, dbwt, combgest, no_abnorm, no_congen) |>
+  select(ld_indl, ld_augm, ld_ster, ld_antb, ld_chor, ld_anes, apgar5) |>
   
   # handling outliers by converting 'unknown or not reported' values to NA
   mutate(
-    no_infec = ifelse(no_infec == 9, NA, no_infec),
-    previs = ifelse(previs == 99, NA, previs),
-    wic = ifelse(wic == "Y", "Yes", ifelse(wic == "N", "No", NA)),
-    dbwt = ifelse(dbwt == 9999, NA, dbwt),
-    combgest = ifelse(combgest == 99, NA, combgest),
-    no_abnorm = ifelse(no_abnorm == 9, NA, no_abnorm),
-    no_congen = ifelse(no_congen == 9, NA, no_congen)
+    apgar5 = ifelse(apgar5 == 99, NA, apgar5)
   ) |>
   
   mutate(
     # converting categorical text values into binary numerical values
-    wic = case_when(
-      wic == "Yes" ~ 1,
-      wic == "No" ~ 0,
-      TRUE ~ NA_real_),
-    
-    # reversing binary values in the variable so that healthier outcomes are represented by values closer to 1
-    f_tobaco = ifelse(f_tobaco == 1, 0, ifelse(f_tobaco == 0, 1, NA))
-  ) |>
+    across(c(ld_indl, ld_augm, ld_ster, ld_antb, ld_chor, ld_anes), ~ case_when(
+      . == "Y" ~ 1,
+      . == "N" ~ 0,
+      . == "U" ~ NA_real_,
+      TRUE ~ as.numeric(.)
+    ))) |>
   
   rename(
-    dem_momage = mager,
-    dem_dadage = fagecomb,
-    mom_bmi = bmi,
-    mom_notobaco = f_tobaco,
[…]

### 2024-11-18 scripts/04-test_analysis_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests the structure and validity of the analysis dataset.
 # Author: Yunkyung Ko
-# Date: 17 November 2024
+# Date: 18 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
@@ -16,14 +16,14 @@ library(testthat)
 infant_health <- read_csv("../data/02-analysis_data/infant_health.csv")
 
 #### Test dataset structure ####
-# Test that the dataset has 3479516 rows
-test_that("dataset has 3479516 rows", {
-  expect_equal(nrow(infant_health), 3479516)
+# Test that the dataset has 3588588 rows
+test_that("dataset has 3588588 rows", {
+  expect_equal(nrow(infant_health), 3588588)
 })
 
-# Test that the dataset has 11 columns
-test_that("dataset has 11 columns", {
-  expect_equal(ncol(infant_health), 11)
+# Test that the dataset has 7 columns
+test_that("dataset has 7 columns", {
+  expect_equal(ncol(infant_health), 7)
 })
 
 #### Test for missing values ####
@@ -33,31 +33,15 @@ test_that("no missing values in dataset", {
 })
 
 #### Test for variable types ####
-# Define expected types for each column - numeric types
-# Test variable types
-test_that("numeric variables have correct types", {
-  expected_types <- list(
-    dem_momage = "double",    
-    dem_dadage = "double",    
-    mom_bmi = "double",       
-    med_previs = "double",    
-    inft_weight = "double",   
-    inft_gest = "double"      
-  )
-  for (var in names(expected_types)) {
-    if (var %in% colnames(infant_health)) {
-      expect_type(infant_health[[var]], expected_types[[var]]
-      )
-    } else {
-      fail(paste("Variable", var, "is not in the dataset."))
-    }
-  }
+# Test numeric variables
+test_that("apgar5 is of double type", {
+  expect_type(infant_health$apgar5, "double")
 })
 
 
 # Test binary variables
 test_that("binary variables contain only 0 or 1", {
-  binary_variables <- c("mom_notobaco", "mom_noinfec", "med_wic", "no_abnorm", "no_congen")  # Add all binary variable names here
+  binary_variables <- c("indc", "augmt", "ster", "antb", "chor", "anes")  # Add all binary variable names here
   for (var in binary_variables) {
[…]

### 2024-11-19 scripts/01-test_simulated_data.R
update scripts for simulation tests

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests the structure and validity of the simulated infant health dataset. Shows expected results of the logistic regression model with simulated data.
 # Author: Yunkyung Ko
-# Date: 17 November 2024
+# Date: 18 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
@@ -12,7 +12,6 @@
 
 #### Workspace setup ####
 library(tidyverse)
-
 sim_data <- read_csv("data/00-simulated_data/simulated_data.csv")
 
 # Test if the data was successfully loaded
@@ -24,24 +23,22 @@ if (exists("sim_data")) {
 
 
 #### Test dataset structure ####
-
-# Check if the dataset has 584 rows
-if (nrow(sim_data) == 584) {
-  message("Test Passed: The dataset has 584 rows.")
+# Check if the dataset has 671 rows
+if (nrow(sim_data) == 671) {
+  message("Test Passed: The dataset has 671 rows.")
 } else {
-  stop("Test Failed: The dataset does not have 584 rows.")
+  stop("Test Failed: The dataset does not have 671 rows.")
 }
 
-# Check if the dataset has 11 columns
-if (ncol(sim_data) == 11) {
-  message("Test Passed: The dataset has 11 columns.")
+# Check if the dataset has 7 columns
+if (ncol(sim_data) == 7) {
+  message("Test Passed: The dataset has 7 columns.")
 } else {
-  stop("Test Failed: The dataset does not have 11 columns.")
+  stop("Test Failed: The dataset does not have 7 columns.")
 }
 
 
 #### Test for missing values ####
-
 # Check if there are any missing values in the dataset
 if (all(!is.na(sim_data))) {
   message("Test Passed: The dataset contains no missing values.")
@@ -51,69 +48,22 @@ if (all(!is.na(sim_data))) {
 
 
 #### Test for outliers (unknown or unreported values) ####
-
-# Define expected outlier codes for each variable
-outlier_codes <- list(
-  mom_noinfec = 9,
-  med_previs = 99,
-  inft_weight = 9999,
-  inft_gest = 99,
-  no_abnorm = 9,
-  no_congen = 9
-)
-
-# Check for outliers in each variable
-outlier_results <- lapply(names(outlier_codes), function(var) {
-  if (var %in% colnames(sim_data)) {
[…]

### 2024-11-20 scripts/06-model_data.R
update codes for modeling data

@@ -1,37 +1,52 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models for predicting the health status of the infant with data of drugs usage during labor and delivery
+# Author: Yunkyung Ko
+# Date: 18 November 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites:
+  # - The `tidyverse`, `rstanarm` package must be installed and loaded
+  # - 03-clean_data.R must have been run
+# Any other information needed? Make sure you are in the `infant_health` rproj
 
 
 #### Workspace setup ####
-library(tidyverse)
+library(readr)
+library(randomForest)
+library(caret)
 library(rstanarm)
+library(ranger)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+infant_health <- read_csv("data/02-analysis_data/infant_health.csv")
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+
+# Model 1: Random Forest for Predicting apgar5
+set.seed(853)  # Ensure reproducibility
+randfor <- ranger(
+  formula = apgar5 ~ indc + augmt + ster + antb + chor + anes,  # Formula for the model
+  data = infant_health,                                        # Dataset
+  num.trees = 500,                                             # Correct argument for number of trees
+  mtry = 3,                                                    # Number of variables tried at each split
+  importance = "impurity",                                     # Variable importance
+  num.threads = 4                                              # Number of cores for parallelism
+)
+
+
+# Model 2: Bayesian GLM for Predicting apgar5
+bayeslinear <- stan_glm(
+  formula = apgar5 ~ indc + augmt + ster + antb + chor + anes,  # Formula for the model
[…]

### 2024-11-21 paper/paper.qmd
update graphs and tables for data / result section

--- 2024-11-14 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "How Drug Usage During Labor and Delivery Affects Infant Health: A Predictive Analysis of APGAR5 Scores"
+subtitle: "!!! MAIN RESULTS !!!"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Yunkyung Ko
+thanks: "Code and data are available at: [https://github.com/koyunkyung/infant_health](https://github.com/koyunkyung/infant_health)."
 date: today
@@
 library(tidyverse)
-library(palmerpenguins)
+library(data.table)
+library(arrow)
+library(scales)
+library(dplyr)
+library(knitr)
+
+raw_data <- fread("../data/01-raw_data/infant_data.csv" , showProgress = FALSE)
+analysis_data <- read_csv("../data/02-analysis_data/infant_health.csv")
+balanced_data <- read_csv("../data/02-analysis_data/infant_balanced.csv")
+
 ```
@@
 
-We use the statistical programming language SQL [@citeR]
-
-.... Our data [@shelter].... Following @tellingstories, we consider...
+We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
 
@@
 
-### Price Levels
+### Apgar Score: a measure the infant's chance of surviving the first year of life
+
+The Apgar score is a measure of the need for resuscitation [@natality2023] to the infant, which is the act of bringing someone back to life or waking them [@cambridgeResuscitation].
+It is a test given to newborns soon after birth (5 minutes) to check 'Appearance(skin color)', 'Pulse(heart rate)', 'Grimace response(reflexes)', 'Activity(muscle tone)', 'Respiration(breathing rate and effort)' [@kidsHealthApgar].
+Each is rated on a scale of 0 to 2, with 2 being the best score [@kidsHealthApgar].
+Apgar scores range from 0 to 10, with a score of 7 or higher indicating that the neonate is in good to excellent physical condition [@natality2023].
+
+@fig-apgar1 shows that the majority of infants in the raw dataset achieve a high APGAR5 score, clustering around 9 and 10.
+Very few observations exist for lower scores, reflecting rare instances of significant distress at birth.
+Even after filtering the dataset by selecting the relevant variables for analysis and removing the NA values, @fig-apgar2 shows that the observations are overly clustered around high APGAR5 scores.
 
 ```{r}
-#| label: fig-price
-#| fig-cap: Price level distributions across all observations used in the analysis
+#| label: fig-apgar1
+#| fig-cap: "The distribution of APGAR5 scores across the entire observation in the original dataset **Note:** Unknown or unreported observations were excluded so that the distribution could be clearly visualized"
 #| echo: false
@@
 
-library(ggplot2)
-apple <- read.csv("../data/analysis_data/avg_price_by_vendor_apple.csv")
-mango <- read.csv("../data/analysis_data/avg_price_by_vendor_mango.csv")
-pear <- read.csv("../data/analysis_data/avg_price_by_vendor_pear.csv")
-orange <- read.csv("../data/analysis_data/avg_price_by_vendor_orange.csv")
+# Remove rows where 'apgar5' equals 99
+filtered_data <- raw_data %>%
+  filter(apgar5 != 99)
 
-vendor_avg_apple <- aggregate(avg_price ~ vendor, data = apple, mean)
-vendor_avg_mango <- aggregate(avg_price ~ vendor, data = mango, mean)
-vendor_avg_pear <- aggregate(avg_price ~ vendor, data = pear, mean)
-vendor_avg_orange <- aggregate(avg_price ~ vendor, data = orange, mean)
+# Histogram for 'apgar5'
+filtered_data |>
+  ggplot(aes(x = apgar5)) +
+  geom_histogram(aes(y = after_stat(count)), 
+    binwidth = 1, 
+    color = "grey", 
+    fill = "skyblue", 
+    alpha = 0.7) +
 
-vendor_avg_apple$fruit <- "apple"
-vendor_avg_mango$fruit <- "mango"
-vendor_avg_pear$fruit <- "pear"
-vendor_avg_orange$fruit <- "orange"
+  # Add labels
+  labs(x = "APGAR5 Score",
+       y = "Count of Observations") +
 
-dataset <- rbind(vendor_avg_apple, vendor_avg_mango, vendor_avg_pear, vendor_avg_orange)
-datao <- rbind(apple, mango, pear, orange)
+  # Format y-axis numbers with commas
+  scale_y_continuous(labels = comma) +
 
-# Plot the distribution of prices with a histogram
-ggplot(datao, aes(x = avg_price)) +
-  geom_histogram(binwidth = 1, fill = "grey", color = "black") + # Set bin width and colors
-  labs(
-       x = "Average Price",
-       y = "Frequency") +
+  # Apply a clean theme
   theme_minimal()
+
+```
+
+```{r}
+#| label: fig-apgar2
+#| fig-cap: The distribution of APGAR5 scores across the filtered observation in analysis dataset, which selected the relevant variables and removed the NA values
+#| echo: false
+#| warning: false
+#| message: false
+
+# Histogram for 'apgar5'
+analysis_data |>
+  ggplot(aes(x = apgar5)) +
+  geom_histogram(aes(y = after_stat(count)), 
+    binwidth = 1, 
+    color = "grey", 
+    fill = "skyblue", 
+    alpha = 0.7) + 
+  # Add labels
+  labs(x = "APGAR5 Score",
+       y = "Count of Observations") +
+  # Format y-axis numbers with commas
+  scale_y_continuous(labels = comma) +
+  # Apply a clean theme
+  theme_minimal()
+
+```
+
+To ensure a balanced analysis and accurately assess the effects of drug treatment during delivery and labor, the data was refined to achieve a more even distribution of observations across APGAR5 scores.
+Based on the lowest observation count of 2,065 in the original distribution, the number of observations for each APGAR5 score was set to 2,000 as shown in @fig-apgar3.
+Random sampling was used to ensure a fair distribution across all score levels.
+
+```{r}
+#| label: fig-apgar3
+#| fig-cap: The distribution of APGAR5 scores across the filtered observation in analysis dataset
+#| echo: false
+#| warning: false
+#| message: false
+
+# Generate a table for APGAR5 score distribution
+apgar_table <- balanced_data %>%
+  group_by(apgar5) %>%  # Group by APGAR5 score
+  summarise(
+    Count = n()  # Count of observations
+  ) %>%
+  arrange(apgar5)  # Ensure rows are ordered by APGAR5 score
+
+# Transpose the table for horizontal display
+apgar_table_transposed <- t(as.matrix(apgar_table$Count))  # Transpose only the counts
+colnames(apgar_table_transposed) <- apgar_table$apgar5  # Set APGAR5 scores as column names
+rownames(apgar_table_transposed) <- "Count"  # Set the row name
+
+# Convert to a data frame for display
+apgar_table_transposed_df <- as.data.frame(apgar_table_transposed)
+
+# Display the transposed table using kable
+knitr::kable(
+  apgar_table_transposed_df)
 
@@
 
-### Type of Fruits: The Product that We Specified to Compare the Price Distribution
+All of the predictor variables used in the analysis are classified into the same category, which is the 'characteristics of labor and delivery'.
+This item, which contains 6 separate checkboxes that the respondent can choose from, allows for the reporting of more than one characteristic and includes a choice of "None of the above"[@natality2023].
 
-@fig-fruits shows the frequency of observations for four fruit types: apples, mangoes, oranges, and pears.
-Apples have the highest count, with around 50 instances, possibly due to popularity or market demand.
-Pears follow with a moderate count of about 20.
-Mangoes and oranges have the lowest counts, around 10 each, possibly due to factors like seasonality or regional availability.
+### Number of Treatments Used During Delivery and Labor
+
+@fig-numtreat shows that the majority of observations involve 0 to 2 treatments during labor and delivery, with a steep decline in counts for 3 or more treatments.
+Most births reported in this dataset occur with minimal medical intervention, and higher number of treatments are relatively rare.
 
 ```{r}
-#| label: fig-fruits
-#| fig-cap: The distrubution of different categories of fruits used in the analysis
+#| label: fig-numtreat
+#| fig-cap: The distribution of observations across different number of treatments used during delivery and labor
 #| echo: false
@@
 
-# Count the number of observations in each dataset
-counts <- data.frame(
-  fruit = c("apple", "mango", "pear", "orange"),
-  observation_count = c(nrow(apple), nrow(mango), nrow(pear), nrow(orange))
+anal_data <- balanced_data %>%
+  mutate(
+    drug_count = rowSums(select(., c(indc, augmt, ster, antb, chor, anes)) == 1, na.rm = TRUE)
   )
 
-# Plot the observation count for each fruit
-ggplot(counts, aes(x = fruit, y = observation_count)) +
-  geom_bar(stat = "identity", fill = "skyblue") +
+# Calculate observation counts for each drug_count
+drug_count_summary <- anal_data %>%
+  group_by(drug_count) %>%
+  summarise(observation_count = n())  # Count observations for each drug_count
+
+# Plot the counts with all x-axis values displayed
+ggplot(drug_count_summary, aes(x = drug_count, y = observation_count)) +
+  geom_bar(stat = "identity", fill = "darkgreen", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(observation_count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
   labs(
-       x = "Fruit Type",
-       y = "Observation Count") +
-  theme_minimal()
+    x = "Number of Treatments Used During Labor and Delivery",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(labels = comma) +  # Format numbers on the y-axis
+  scale_x_continuous(breaks = seq(min(drug_count_summary$drug_count), max(drug_count_summary$drug_count), by = 1)) +  # Ensure all x-axis values are shown
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 12),
+    axis.title = element_text(size = 13)
+  )
 
@@
 
-### Type of Vendors
-
-@fig-vendors provides a visualization of the number of observations for each vendor in the dataset, specifically Loblaws, NoFrills, and Voila.
-The x-axis represents these three vendors, while the y-axis shows the count of observations, likely indicating the number of unique products or price points recorded for each vendor.
-Each bar is labeled with the exact count: Loblaws has 33 observations, NoFrills has 31, and Voila has the highest at 36.
-These counts are relatively similar across vendors, suggesting a fairly balanced representation in the data, although Voila has a slight edge in the number of entries.
-This balanced distribution is beneficial for comparative analysis, as it reduces the risk of one vendor disproportionately influencing results.
-The slight difference in counts could introduce a minor weight toward Voila in any statistical analysis, but overall, the dataset’s even distribution supports reliable comparisons.
-Thus, any averages or other metrics calculated by vendor are likely to be comparable across Loblaws, NoFrills, and Voila, minimizing selection bias due to unequal observation counts.
+### Type of Treatments Used During Delivery and Labor
 
 ```{r}
-#| label: fig-vendors
-#| fig-cap: The distribution of vendors used in the analysis
-#| echo: false
-#| warning: false
-
-# Plot the distribution of observation count by vendor with adjusted bar width and labels
-ggplot(datao, aes(x = vendor)) +
-  geom_bar(fill = "purple", width = 0.4) + # Narrower bar width
-  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) + # Add count labels above bars
-  labs(
-       x = "Vendor",
-       y = "Observation Count") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability
-```
-
-# Results
-
-@fig-results illustrates the average prices of four fruits—apples, mangoes, oranges, and pears—across three different vendors: Loblaws, NoFrills, and Voila.
-Each fruit is represented by a distinct color, as indicated in the legend: red for apples, green for mangoes, blue for oranges, and purple for pears.
-The bar plot component of the chart displays the average price of each fruit at the respective vendors.
-Notably, apples are priced highest at Loblaws and lowest at Voila.
-Mangoes exhibit minimal price variation across the vendors.
-Oranges show a significant price difference, being most expensive at Loblaws and least expensive at Voila.
-Similarly, pears are priced lowest at Voila.
-Superimposed on the bar plot is a line plot that connects the average prices of each fruit across the vendors, highlighting the price trends.
-The lines for apples, oranges, and pears demonstrate a downward trend from Loblaws to Voila, indicating decreasing prices.
-In contrast, the line for mangoes remains relatively flat, suggesting consistent pricing among the vendors.
-In summary, the chart effectively compares the pricing strategies of different vendors for various fruits, revealing that Voila generally offers lower prices for most fruits, while Loblaws tends to have higher prices.
-These visual trends provide insights into the pricing correlations among the vendors and set the stage for further analysis of potential causal factors influencing these pricing differences.
-
-```{r}
-#| label: fig-results
-#| fig-cap: Price distribution of different kinds of fruits(apple, mango, orange, pear) by vendors
+#| label: fig-typetreat
+#| fig-cap: The distribution of observations across different type of treatments used during delivery and labor
 #| echo: false
@@
 
-ggplot(dataset, aes(x = vendor, y = avg_price, fill = fruit, group = fruit)) +
-  geom_bar(stat = "identity", position = "dodge", aes(color = fruit)) +
-  geom_line(aes(color = fruit)) +
-  geom_point(aes(color = fruit)) +
-  facet_wrap(~fruit, scales = "free_y") +
-  labs(x = "Render", y = "Price") +
+# Define custom labels and order for Treatment_Type
+custom_labels <- c(
+  "indc" = "Induction",
+  "augmt" = "Augmentation",
+  "ster" = "Steroids",
+  "antb" = "Antibiotics",
+  "chor" = "Chorioamnionitis",
+  "anes" = "Anesthesia"
+)
+# Reshape data to long format for treatment types
+treatment_summary <- balanced_data %>%
+  select(indc, augmt, ster, antb, chor, anes) %>%  # Select treatment columns
+  pivot_longer(
+    cols = everything(),  # Reshape all columns to long format
+    names_to = "Treatment_Type",  # Column for treatment type
+    values_to = "Value"  # Column for values (0 or 1)
+  ) %>%
+  filter(Value == 1) %>%  # Keep rows where treatment is used
+  group_by(Treatment_Type) %>%
+  summarise(Count = n(), .groups = "drop") %>%  # Count observations for each treatment type
+  mutate(
+    Treatment_Type = factor(Treatment_Type, levels = names(custom_labels), labels = custom_labels)  # Apply custom labels and order
+  )
+
+# Plot the distribution of observations for each treatment
+ggplot(treatment_summary, aes(x = Treatment_Type, y = Count, fill = Treatment_Type)) +
+  geom_bar(stat = "identity", fill = "darkgreen", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Treatment Type",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 10),
+    legend.position = "none"  # Remove legend for cleaner look
+  )
 ```
+
+### Induction of labor
+
+```{r}
+#| label: fig-indc1
+#| fig-cap: The distribution of observations in the raw dataset based on whether the infant received induction of labor or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_indl_summary1 <- raw_data %>%
+  group_by(ld_indl) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_indl = ifelse(ld_indl == "Y", "Yes", ifelse(ld_indl == "N", "No", "Unknown")),  # Label groups
+    ld_indl = factor(ld_indl, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_indl_summary1, aes(x = ld_indl, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Induction of Labor",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_indl_summary1$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+```{r}
+#| label: fig-indc2
+#| fig-cap: The distribution of observations in the analysis dataset on whether the infant received induction of labor or not,
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_indl_summary2 <- balanced_data %>%
+  group_by(indc) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_indl = ifelse(indc == 1, "Yes", ifelse(indc == 0 , "No", "Unknown")),  # Label groups
+    ld_indl = factor(ld_indl, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_indl_summary2, aes(x = ld_indl, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Induction of Labor",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_indl_summary2$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+
+### Augmentation of labor
+
+```{r}
+#| label: fig-augmt1
+#| fig-cap: The distribution observations in the raw dataset based on whether the infant received augmentation of labor or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_augm_summary1 <- raw_data %>%
+  group_by(ld_augm) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_augm = ifelse(ld_augm == "Y", "Yes", ifelse(ld_augm == "N", "No", "Unknown")),  # Label groups
+    ld_augm = factor(ld_augm, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_augm_summary1, aes(x = ld_augm, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Augmentation of Labor",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_augm_summary1$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+```{r}
+#| label: fig-augmt2
+#| fig-cap: The distribution observations in the analysis datset on whether the infant received augmentation of labor or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_augm_summary2 <- balanced_data %>%
+  group_by(augmt) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_augm = ifelse(augmt == 1, "Yes", ifelse(augmt == 0 , "No", "Unknown")),  # Label groups
+    ld_augm = factor(ld_augm, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_augm_summary2, aes(x = ld_augm, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Augmentation of Labor",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_augm_summary2$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+
+### Steroids (glucocorticoids) for fetal lung maturation received by the mother before delivery
+
+```{r}
+#| label: fig-ster1
+#| fig-cap: The distribution observations in the raw dataset based on whether the infant received steroid treatments or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_ster_summary1 <- raw_data %>%
+  group_by(ld_ster) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_ster = ifelse(ld_ster == "Y", "Yes", ifelse(ld_ster == "N", "No", "Unknown")),  # Label groups
+    ld_ster = factor(ld_ster, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_ster_summary1, aes(x = ld_ster, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Steroids Usage",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_ster_summary1$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+```{r}
+#| label: fig-ster2
+#| fig-cap: The distribution observations in the analysis datset on whether the infant received steroid treatments or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_ster_summary2 <- balanced_data %>%
+  group_by(ster) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_ster = ifelse(ster == 1, "Yes", ifelse(ster == 0 , "No", "Unknown")),  # Label groups
+    ld_ster = factor(ld_ster, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_ster_summary2, aes(x = ld_ster, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Steroids Usage",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_ster_summary2$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+
+### Antibiotics received by the mother during delivery
+
+```{r}
+#| label: fig-antb1
+#| fig-cap: The distribution observations in the raw dataset based on whether the infant received antibiotic treatments or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_antb_summary1 <- raw_data %>%
+  group_by(ld_antb) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_antb = ifelse(ld_antb == "Y", "Yes", ifelse(ld_antb == "N", "No", "Unknown")),  # Label groups
+    ld_antb = factor(ld_antb, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_antb_summary1, aes(x = ld_antb, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Antibiotics Usage",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_antb_summary1$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+
+```{r}
+#| label: fig-antb2
+#| fig-cap: The distribution observations in the analysis datset on whether the infant received antibiotic treatments or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_antb_summary2 <- balanced_data %>%
+  group_by(antb) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_antb = ifelse(antb == 1, "Yes", ifelse(antb == 0 , "No", "Unknown")),  # Label groups
+    ld_antb = factor(ld_antb, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_antb_summary2, aes(x = ld_antb, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Antibiotics Usage",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_antb_summary2$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+
+### Clinical chorioamnionitis diagnosed during labor or maternal temperature over 38 degrees celcius (100.4 degrees fahrenheit)
+
+```{r}
+#| label: fig-chor1
+#| fig-cap: The distribution observations in the raw dataset based on whether the infant received chorioamnionitis treatments or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_chor_summary1 <- raw_data %>%
+  group_by(ld_chor) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_chor = ifelse(ld_chor == "Y", "Yes", ifelse(ld_chor == "N", "No", "Unknown")),  # Label groups
+    ld_chor = factor(ld_chor, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_chor_summary1, aes(x = ld_chor, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Chorioamnionitis Usage",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_chor_summary1$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+```{r}
+#| label: fig-chor2
+#| fig-cap: The distribution observations in the analysis datset on whether the infant received chorioamnionitis treatments or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_chor_summary2 <- balanced_data %>%
+  group_by(chor) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_chor = ifelse(chor == 1, "Yes", ifelse(chor == 0 , "No", "Unknown")),  # Label groups
+    ld_chor = factor(ld_chor, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_chor_summary2, aes(x = ld_chor, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Chorioamnionitis Usage",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_chor_summary2$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+
+### Epidural or spinal anesthesia during labor
+
+```{r}
+#| label: fig-anes1
+#| fig-cap: The distribution observations in the raw dataset based on whether the infant received anesthesia treatments or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_anes_summary1 <- raw_data %>%
+  group_by(ld_anes) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_anes = ifelse(ld_anes == "Y", "Yes", ifelse(ld_anes == "N", "No", "Unknown")),  # Label groups
+    ld_anes = factor(ld_anes, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_anes_summary1, aes(x = ld_anes, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Anesthesia Usage",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_anes_summary1$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+
+```{r}
+#| label: fig-anes2
+#| fig-cap: The distribution observations in the analysis datset on whether the infant received anesthesia treatments or not
+#| echo: false
+#| warning: false
+#| message: false
+
+# Prepare the data
+ld_anes_summary2 <- balanced_data %>%
+  group_by(anes) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  mutate(
+    ld_anes = ifelse(anes == 1, "Yes", ifelse(anes == 0 , "No", "Unknown")),  # Label groups
+    ld_anes = factor(ld_anes, levels = c("Yes", "No", "Unknown"))  # Specify factor levels for order
+  )
+
+# Plot the bar graph
+ggplot(ld_anes_summary2, aes(x = ld_anes, y = Count)) +
+  geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  labs(
+    x = "Anesthesia Usage",
+    y = "Observation Count"
+  ) +
+  scale_y_continuous(limits = c(0, max(ld_anes_summary2$Count) + 1000), expand = c(0, 0)) +
+  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 11),
+    axis.title = element_text(size = 13)
+  )
+
+```
+
+## Correlation between predictor variables
+
+### Induction of Labor and Augmentation of Labor
+
+```{r}
+#| label: fig-corr1
+#| fig-cap: Mosaic plot showing the correlation between induction and augmentation of labor
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load necessary library
+library(vcd)
+
+# Create a contingency table for "Induction of Labor" and "Augmentation of Labor"
+contingency_table <- xtabs(~ indc + augmt, data = balanced_data)
+
+# Perform chi-squared test to get expected frequencies
+chi_test <- chisq.test(contingency_table)
+expected <- chi_test$expected  # Expected frequencies
+residuals <- (contingency_table - expected) / sqrt(expected)  # Compute residuals
+
+# Define a custom shading function manually
+custom_shading <- function(residuals) {
+  breaks <- c(-Inf, -2, 0, 2, Inf)  # Breaks for shading categories
+  colors <- c("#E8F5E9", "#C8E6C9", "#A5D6A7", "#388E3C")  # Light to dark green
+  as.character(cut(residuals, breaks = breaks, labels = colors, include.lowest = TRUE))
+}
+
+# Assign shading to each cell in the contingency table
+color_map <- matrix(custom_shading(as.vector(residuals)), 
+                    nrow = nrow(contingency_table), 
+                    ncol = ncol(contingency_table))
+
+# Create the mosaic plot with custom colors
+mosaic(
+  contingency_table,
+  gp = gpar(fill = color_map),  # Apply the color map
+  labeling_args = list(
+    set_varnames = c(indc = "Induction of Labor", augmt = "Augmentation of Labor")  # Custom axis labels
+  ))
+
+```
+
+### Usage of Steroids and Antibiotics
+
+```{r}
+#| label: fig-corr2
+#| fig-cap: Mosaic plot showing the correlation between usage of steroids and antibiotics
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load necessary library
+library(vcd)
+
+# Create a contingency table for "Induction of Labor" and "Augmentation of Labor"
+contingency_table <- xtabs(~ ster + antb, data = balanced_data)
+
+# Perform chi-squared test to get expected frequencies
+chi_test <- chisq.test(contingency_table)
+expected <- chi_test$expected  # Expected frequencies
+residuals <- (contingency_table - expected) / sqrt(expected)  # Compute residuals
+
+# Define a custom shading function manually
+custom_shading <- function(residuals) {
+  breaks <- c(-Inf, -2, 0, 2, Inf)  # Breaks for shading categories
+  colors <- c("#E8F5E9", "#C8E6C9", "#A5D6A7", "#388E3C")  # Light to dark green
+  as.character(cut(residuals, breaks = breaks, labels = colors, include.lowest = TRUE))
+}
+
+# Assign shading to each cell in the contingency table
+color_map <- matrix(custom_shading(as.vector(residuals)), 
+                    nrow = nrow(contingency_table), 
+                    ncol = ncol(contingency_table))
+
+# Create the mosaic plot with custom colors
+mosaic(
+  contingency_table,
+  gp = gpar(fill = color_map),  # Apply the color map
+  labeling_args = list(
+    set_varnames = c(ster = "Steroids", antb = "Antibiotics")  # Custom axis labels
+  ))
+
+```
+
+### Usage of Chorioamnionitis and Antibiotics
+
+```{r}
+#| label: fig-corr3
+#| fig-cap: Mosaic plot showing the correlation between usage of chorioamnionitis and antibiotics
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load necessary library
+library(vcd)
+
+# Create a contingency table for "Induction of Labor" and "Augmentation of Labor"
+contingency_table <- xtabs(~ chor + antb, data = balanced_data)
+
+# Perform chi-squared test to get expected frequencies
+chi_test <- chisq.test(contingency_table)
+expected <- chi_test$expected  # Expected frequencies
+residuals <- (contingency_table - expected) / sqrt(expected)  # Compute residuals
+
+# Define a custom shading function manually
+custom_shading <- function(residuals) {
+  breaks <- c(-Inf, -2, 0, 2, Inf)  # Breaks for shading categories
+  colors <- c("#E8F5E9", "#C8E6C9", "#A5D6A7", "#388E3C")  # Light to dark green
+  as.character(cut(residuals, breaks = breaks, labels = colors, include.lowest = TRUE))
+}
+
+# Assign shading to each cell in the contingency table
+color_map <- matrix(custom_shading(as.vector(residuals)), 
+                    nrow = nrow(contingency_table), 
+                    ncol = ncol(contingency_table))
+
+# Create the mosaic plot with custom colors
+mosaic(
+  contingency_table,
+  gp = gpar(fill = color_map),  # Apply the color map
+  labeling_args = list(
+    set_varnames = c(chor = "Chorioamnionitis", antb = "Antibiotics")  # Custom axis labels
+  ))
+
+```
+
+# Model
+
+The goal of our modelling strategy is twofold. Firstly,...
+
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+
+## Model set-up
+
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_i + \gamma_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+### Model justification
+
+We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+
+We can use maths by including latex between dollar signs, for instance $\theta$.
+
+# Results
+
+## Results from examining the analysis dataset
+
+```{r}
+#| label: fig-result1
+#| fig-cap: The mean of APGAR5 scores by the number of treatments recieved during labor and delivery
+#| echo: false
+#| warning: false
+#| message: false
+
+anal_data <- balanced_data %>%
+  mutate(
+    drug_count = rowSums(select(., c(indc, augmt, ster, antb, chor, anes)) == 1, na.rm = TRUE)
+  )
+
+summary_data <- anal_data %>%
+  group_by(drug_count) %>%
+  summarise(
+    mean_apgar5 = mean(apgar5),
+    sd_apgar5 = sd(apgar5)
+  )
+
+ggplot(summary_data, aes(x = factor(drug_count), y = mean_apgar5)) +
+  geom_bar(
+    stat = "identity",
+    width = 0.5,  # Adjust width for thinner bars
+    fill = "skyblue",
+    color = "black",
+    alpha = 0.7
+  ) +
+  geom_errorbar(
+    aes(ymin = mean_apgar5 - sd_apgar5, ymax = mean_apgar5 + sd_apgar5),
+    width = 0.2,  # Width of error bars remains the same
+    color = "blue"
+  ) +
+  geom_text(
+    aes(label = round(mean_apgar5, 1)),
+    vjust = -0.5,
+    size = 4,
+    color = "black"
+  ) +
+  labs(
+    x = "Number of Treatments During Labor and Delivery",
+    y = "Mean of APGAR5 Score"
+  ) +
+  scale_y_continuous(
+    limits = c(0, max(summary_data$mean_apgar5 + summary_data$sd_apgar5) + 1),
+    breaks = seq(0, 10, by = 1)
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text = element_text(size = 12),
+    axis.title = element_text(size = 13)
+  )
+
+```
+
+```{r}
+#| label: fig-result2
+#| fig-cap: Heatmap of the overall correlation between treatments administered during delivery and APGAR5 scores
+#| echo: false
+#| warning: false
+#| message: false
+
+heatmap_data <- balanced_data %>%
+  pivot_longer(
+    cols = c(indc, augmt, ster, antb, chor, anes), 
+    names_to = "Drug_Usage", 
+    values_to = "Usage"
+  ) %>%
+  filter(Usage == 1) %>%  # Keep rows where drug usage is "Yes" (1)
+  group_by(Drug_Usage, apgar5) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  pivot_wider(names_from = apgar5, values_from = Count, values_fill = 0)
+
+# Convert the reshaped data back to long format for ggplot
+heatmap_long <- heatmap_data %>%
+  pivot_longer(-Drug_Usage, names_to = "APGAR5", values_to = "Count") %>%
+  mutate(APGAR5 = as.numeric(APGAR5))
+
+# Create heatmap
+# Create a mapping of custom labels for drug usage categories
+custom_labels <- c(
+  indc = "Induction",
+  augmt = "Augmentation",
+  ster = "Steroids",
+  antb = "Antibiotics",
+  chor = "Chorioamnionitis",
+  anes = "Anesthesia"
+)
+ggplot(heatmap_long, aes(x = APGAR5, y = Drug_Usage, fill = Count)) +
+  geom_tile(color = "white") +
+  scale_fill_gradient(
+    low = "skyblue", 
+    high = "blue", 
+    name = "Count of Yes Responses", 
+    guide = guide_colorbar(
+      name = "Yes Responses",
+      barwidth = 1,  # Adjust the width of the legend
+      barheight = 6  # Adjust the height of the legend
+    )
+  ) +
+  scale_x_continuous(breaks = 0:10) +  # Show all integers from 0 to 10
+  scale_y_discrete(labels = custom_labels) +  # Apply custom labels to y-axis (Drug Usage Categories)
+  labs(
+    x = "Five-Minute APGAR Score",
+    y = "Treatments During Labor and Delivery"
+  ) +
+  theme_minimal(base_size = 10) +
+  theme(
+    axis.text.x = element_text(hjust = 1, vjust = 1),  # Angle and align x-axis text
+    axis.text.y = element_text(size = 9),  # Increase y-axis text size
+    legend.text = element_text(size = 7),  # Reduce legend text size
+    legend.title = element_text(size = 7),  # Reduce legend title size
+  )
+
+```
+
+```{r}
+#| label: fig-indcscore
+#| fig-cap: APGAR5 score distribution for the usage of induction labor
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter data for Induction of Labor (indc == 1)
+indc_data <- balanced_data %>% filter(indc == 1)
+
+# Compute the counts for each APGAR5 score
+indc_data_summary <- indc_data %>%
+  group_by(apgar5) %>%
+  summarise(Count = n(), .groups = "drop")
+
+# Plot the bar chart with a trendline
+ggplot(indc_data_summary, aes(x = apgar5, y = Count)) +
+  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
+  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
+  geom_point(color = "red", size = 3) +                        # Highlight points
+  labs(
+    x = "Five-Minute APGAR Score",
+    y = "Observation Count"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = 0:10) +
+  scale_y_continuous(limits = c(0, max(indc_data_summary$Count) + 100), expand = c(0, 0)) +
+  theme(
+    axis.text.x = element_text(hjust = 1, vjust = 1),
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+
+```{r}
+#| label: fig-augmtscore
+#| fig-cap: APGAR5 score distribution for the usage of augmentation labor
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter data for Augmentation of Labor (augmt == 1)
+augmt_data <- balanced_data %>% filter(augmt == 1)
+
+# Compute the counts for each APGAR5 score
+augmt_data_summary <- augmt_data %>%
+  group_by(apgar5) %>%
+  summarise(Count = n(), .groups = "drop")
+
+# Plot the bar chart with a trendline
+ggplot(augmt_data_summary, aes(x = apgar5, y = Count)) +
+  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
+  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
+  geom_point(color = "red", size = 3) +                        # Highlight points
+  labs(
+    x = "Five-Minute APGAR Score",
+    y = "Observation Count"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = 0:10) +
+  scale_y_continuous(limits = c(0, max(augmt_data_summary$Count) + 100), expand = c(0, 0)) +
+  theme(
+    axis.text.x = element_text(hjust = 1, vjust = 1),
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+
+```{r}
+#| label: fig-sterscore
+#| fig-cap: APGAR5 score distribution for the usage of steroids
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter data for Steroids (ster == 1)
+ster_data <- balanced_data %>% filter(ster == 1)
+
+# Compute the counts for each APGAR5 score
+ster_data_summary <- ster_data %>%
+  group_by(apgar5) %>%
+  summarise(Count = n(), .groups = "drop")
+
+# Plot the bar chart with a trendline
+ggplot(ster_data_summary, aes(x = apgar5, y = Count)) +
+  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
+  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
+  geom_point(color = "red", size = 3) +                        # Highlight points
+  labs(
+    x = "Five-Minute APGAR Score",
+    y = "Observation Count"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = 0:10) +
+  scale_y_continuous(limits = c(0, max(ster_data_summary$Count) + 100), expand = c(0, 0)) +
+  theme(
+    axis.text.x = element_text(hjust = 1, vjust = 1),
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+
+```{r}
+#| label: fig-antbscore
+#| fig-cap: APGAR5 score distribution for the usage of antibiotics
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter data for Antibiotics (antb == 1)
+antb_data <- balanced_data %>% filter(antb == 1)
+
+# Compute the counts for each APGAR5 score
+antb_data_summary <- antb_data %>%
+  group_by(apgar5) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  arrange(apgar5)  # Ensure data is sorted by APGAR5
+
+# Plot the bar chart with a trendline
+ggplot(antb_data_summary, aes(x = apgar5, y = Count)) +
+  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
+  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
+  geom_point(color = "red", size = 3) +                        # Highlight points
+  labs(
+    x = "Five-Minute APGAR Score",
+    y = "Observation Count"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = 0:10) +
+  scale_y_continuous(limits = c(0, max(antb_data_summary$Count) + 100), expand = c(0, 0)) +
+  theme(
+    axis.text.x = element_text(hjust = 1, vjust = 1),
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+
+```{r}
+#| label: fig-chorscore
+#| fig-cap: APGAR5 score distribution for the usage of chorioamnionitis
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter data for Chorioamnionitis (chor == 1)
+chor_data <- balanced_data %>% filter(chor == 1)
+
+# Compute the counts for each APGAR5 score
+chor_data_summary <- chor_data %>%
+  group_by(apgar5) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  arrange(apgar5)  # Ensure data is sorted by APGAR5
+
+# Plot the bar chart with a trendline
+ggplot(chor_data_summary, aes(x = apgar5, y = Count)) +
+  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
+  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
+  geom_point(color = "red", size = 3) +                        # Highlight points
+  labs(
+    x = "Five-Minute APGAR Score",
+    y = "Observation Count"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = 0:10) +
+  scale_y_continuous(limits = c(0, max(chor_data_summary$Count) + 50), expand = c(0, 0)) +
+  theme(
+    axis.text.x = element_text(hjust = 1, vjust = 1),
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+
+```{r}
+#| label: fig-anesscore
+#| fig-cap: APGAR5 score distribution for the usage of anesthesia
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter data for Anesthesia (anes == 1)
+anes_data <- balanced_data %>% filter(anes == 1)
+
+# Compute the counts for each APGAR5 score
+anes_data_summary <- anes_data %>%
+  group_by(apgar5) %>%
+  summarise(Count = n(), .groups = "drop") %>%
+  arrange(apgar5)  # Ensure data is sorted by APGAR5
+
+# Plot the bar chart with a trendline
+ggplot(anes_data_summary, aes(x = apgar5, y = Count)) +
+  geom_col(color = "black", fill = "skyblue", alpha = 0.7) +  # Bar chart
+  geom_line(aes(y = Count), color = "red", size = 1.2) +       # Add trendline
+  geom_point(color = "red", size = 3) +                        # Highlight points
+  labs(
+    x = "Five-Minute APGAR Score",
+    y = "Observation Count"
+  ) +
+  theme_minimal() +
+  scale_x_continuous(breaks = 0:10) +
+  scale_y_continuous(limits = c(0, max(anes_data_summary$Count) + 200), expand = c(0, 0)) +
+  theme(
+    axis.text.x = element_text(hjust = 1, vjust = 1),
+    axis.text.y = element_text(size = 12)
+  )
+
+```
+
+## Results from the prediction model
+
+Our results are summarized in @tbl-modelresults.
 
@@
 
-## Correlation vs. Causation
+## First discussion point {#sec-first-point}
 
-When analyzing the pricing of the same fruit in different supermarkets, it is very important to distinguish Correlation and Causation for understanding the root cause of price difference.
-We have observed that the prices of Loblaws apples, pears, orange and mango are always higher than those of other supermarkets (such as NoFrills and Voila), and the price of each fruit in Voila is always lower than that in other stores, which shows that there is a correlation between supermarket and fruit pricing.
-This correlation may be related to the market positioning, operating costs and supply chain management of supermarkets.
-For example, Loblaws is positioned in the high-end market and can provide better products and services to middle-class customers, so its pricing is higher.
-Voila mainly serves groups with limited budgets, so the price is lower.
-In addition, factors such as geographical location, rent and employee salary may lead to differences in operating costs, which in turn may affect pricing strategies.
-For example, Loblaws is located in the city center, the rent is high, and the salary of employees is relatively high because of the high service level.
-In supply chain management, for example, the procurement channels of different supermarkets and the efficiency of supply chain may lead to cost differences and affect the final selling price.
-In addition, to determine whether these factors really lead to price differences, further data and analysis are needed, such as market positioning research, cost structure analysis and supply chain evaluation.
-Through these in-depth analysis, we can understand the causal relationship of price difference more accurately.
+If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
 
-## Missing Data
+## Second discussion point
 
-In this dataset, missing data occurs because different supermarkets offer varying selections of fruit varieties.
-For example, some supermarkets may carry a wide range of apple varieties, while others only stock a few.
-As a result, when analyzing data across all supermarkets, information on certain apple varieties may be incomplete, with specific types available only from certain vendors.
-Similar patterns may apply to other types of fruit, leading to gaps in the dataset.
-This missing data reflects differences in product offerings across vendors, and it's essential to be aware of and consider this incompleteness when conducting further analysis.
+Please don't use these as sub-heading labels - change them to be what your point actually is.
 
-## Sources of Bias
-In this dataset, we study the price differences and potential changes of 4 fruits between different supermarkets.
-The price differences are due to the different positioning of each supermarket, Loblaw, NoFrills, and Voila.
+## Third discussion point
 
-### Selection Bias
-As mentioned before, selection bias in this dataset mainly comes from the differences in customer groups of different supermarkets.
-Each supermarket (such as Loblaws, NoFrills, and Voila) has a different market positioning and customer group.
-Customers who choose to shop at Loblaws usually prefer high-end goods or organic foods, while NoFrills customers may pay more attention to low prices and discounts.
-This customer preference leads to bias in the price data, that is, the price differences between different supermarkets not only reflect the market price of the product itself, but also are affected by customer selection behavior.
-Therefore, Loblaws' higher prices do not necessarily mean that all its fruits are priced higher than other supermarkets, but rather reflect that its customers are willing to pay a premium for higher quality or brand.
+## Weaknesses and next steps
 
-### Measurement Bias
-In this dataset, measurement bias may come from the different ways in which supermarkets record or define the average price (avg_price).
-For example, Loblaws may be promoting a specific brand or organic fruit, making the recorded average price lower than the actual price, while NoFrills may use a specific price for a single batch, ignoring other higher price ranges.
-These differences in measurement may cause distortions in price data, making our price comparisons between different supermarkets biased.
+Weaknesses and next steps should also be included.
 
-### Confounding Bias
-In this dataset, confounding bias may come from factors such as fruit type, quality, packaging type, seasonal changes, etc.
-For example, Loblaws may be more inclined to sell high-quality or organic apples, which are usually priced higher, and this tendency is not shown separately in the data.
-Therefore, the price difference between supermarkets may not be due to the supermarket itself, but to uncontrolled fruit quality or seasonal factors.
+\newpage
+
+\appendix
+
+# Appendix {-}
+
+# Additional data details
+
+# Model details {#sec-model-details}
+
+## Posterior predictive check
+
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+
+## Diagnostics
+
+@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+
+@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
 

### 2024-11-21 scripts/03-clean_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Selects relevant variables for analysis and removes the outliers(unknown or unreported).
 # Author: Yunkyung Ko
-# Date: 18 November 2024
+# Date: 21 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
@@ -50,6 +50,28 @@ infant_health_na <-
 infant_health <- infant_health_na %>%
   drop_na()
 
+library(dplyr)
+
+# Define the desired sample size per `apgar5` score
+# You can use the minimum count across groups or specify a fixed number
+desired_sample_size <- 2000  # Adjust this based on your needs
+
+# Stratified random sampling
+balanced_data <- infant_health %>%
+  group_by(apgar5) %>%
+  sample_n(size = min(desired_sample_size, n()), replace = FALSE) %>%  # Ensure no oversampling
+  ungroup()
+
+# Check the distribution of the new dataset
+balanced_data %>%
+  group_by(apgar5) %>%
+  summarise(Count = n()) %>%
+  print()
+
+
 #### Save data ####
 write_csv(infant_health, "data/02-analysis_data/infant_health.csv")
 write_parquet(infant_health, "data/02-analysis_data/infant_health.parquet")
+
+write_csv(balanced_data, "data/02-analysis_data/infant_balanced.csv")
+write_parquet(balanced_data, "data/02-analysis_data/infant_balanced.parquet")

### 2024-11-21 scripts/05-exploratory_data_analysis.R

@@ -1,8 +1,8 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models to predict the health status of an infant with the data of treatments/drug usage during delivery and labor
+# Author: Yunkyung Ko
+# Date: 18 November 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]

### 2024-11-21 paper/state_respondents.qmd
Add files via upload

--- 2024-10-03 state_respondents.qmd
@@
 ---
-title: "2022 American Community Surveys Analysis"
-author: "Group 25"
+title: "Estimating the Number of Doctoral Degree Holders in the U.S.: A Comparative Analysis Using 2022 ACS Data and Laplace Ratio Estimators"
+subtitle: "Kamala Harris Leads with Stable Support Around 48%"
+author: 
+  - Yunkyung Ko
+  - YingKe He
+  - Richard Guo
+  - Ruikang Wang
+  - Ziheng Zhong
+thanks: "Code and data are available at: [https://github.com/koyunkyung/state_respondents](https://github.com/koyunkyung/state_respondents)."
+date: today
+date-format: long
+abstract: ""
 format: pdf
-editor: visual
+number-sections: true
+bibliography: references.bib
+toc: true
+toc-title: "Table of Contents"
+toc-depth: 2
+toc-location: left
 ---
+
 ```{r}
@@
 # Method of Obtaining the Data
+
 To retrieve data from IPUMS USA, we first login to the IPUMS website and selected "IPUMS USA." From there, we clicked "Get Data" and chose the "2022 ACS" sample under the "SELECT SAMPLE" section.
-To obtain state-level data, we selected "HOUSEHOLD" > "GEOGRAPHIC" and added the "STATEICP" variable to the data cart.
+To obtain state-level data, we selected "HOUSEHOLD" \> "GEOGRAPHIC" and added the "STATEICP" variable to the data cart.
 For individual-level data, we went to the "PERSON" section and included the "EDUC" variable.
@@
 #| echo: false
+#| include: false
+#| warning: false
+#| message: false
+
 # Load the IPUMS data
-ipums_data <- read.csv("usa_00001.csv")
-
+ipums_data <- read.csv("../usa_00001.csv")
 # Filter for doctoral degree holders
@@
 Next, we count the respondents with doctoral degrees grouped by states.
+
 ```{r}
+#| label: tbl-pctraw1
+#| tbl-cap: The distribution of respondents with doctoral degrees across states with different STATEICPs (**Note:** The table is being displayed with only the first part shown to optimize space.)
 #| echo: false
-#| tbl-cap: doctoral respondents
+
 # Count respondents with doctoral degrees by state
@@
   summarise(doctoral = n())
+
+# Modify column names before passing to kable
+colnames(doctoral_count) <- c("State", "Respondents with Doctoral Degree")
+
+# Display the table with custom column labels
 kable(head(doctoral_count)) %>%
@@
   column_spec(1:2, width = "2cm")
+
 ```
@@
 Factors like varying educational systems, socioeconomic conditions, or migration patterns between states can cause these distortions.
+
 ```{r}
 #| echo: false
+
+# Assuming State is correctly renamed, and column for doctoral degree count is "Respondents with Doctoral Degree"
 california_total_respondents <- 391171
 california_doctoral_respondents <- doctoral_count %>%
-  filter(STATEICP == 71) %>%
-  pull(doctoral)
-california_ratio <- california_doctoral_respondents / california_total_respondents
-print(paste("The California ratio that was found was:", california_ratio))
+  filter(State == 71) %>%
+  pull(`Respondents with Doctoral Degree`)
+
+california_doctoral_respondents
 
@@
 Other state codes show varying counts of respondents, such as 165 for state code 2, 244 for state code 4, and so on.
+
 ```{r}
+#| label: tbl-pctraw2
+#| tbl-cap: The distribution of respondents with doctoral degrees across states with different STATEICPs (**Note:** The table is being displayed with only the first part shown to optimize space.)
 #| echo: false
-#| tbl-cap: estimated total respondents
+
+# Assuming california_ratio is already defined and available in your environment
 doctoral_count <- doctoral_count %>%
-  mutate(estimated_total = doctoral / california_ratio)
-kable(head(doctoral_count))  %>%
-  kable_styling(full_width = FALSE, font_size = 10) %>%
-  column_spec(1:3, width = "2cm")
+  mutate(estimated_total = `Respondents with Doctoral Degree` / california_ratio)
+
+# Check the updated dataframe
+head(doctoral_count)
+
 ```
@@
 This highlights the importance of accounting for regional variations in such analyses.
+
 ```{r}
@@
 ```
+
 ```{r}

### 2024-11-21 paper/state_respondents.qmd
fix for caption and referencing errors

@@
 ---
-title: "Estimating the Number of Doctoral Degree Holders in the U.S.: A Comparative Analysis Using 2022 ACS Data and Laplace Ratio Estimators"
-subtitle: "Kamala Harris Leads with Stable Support Around 48%"
+title: "Estimating the Number of Doctoral Degree Holders in the U.S.: A Comparative Analysis Using Laplace Ratio Estimators"
+subtitle: "Discrepancy Between Estimated and Actual Totals Show Regional Variations in Educational Attainment"
 author: 
@@
 date-format: long
-abstract: ""
+abstract: "This paper uses the 2022 American Community Survey (ACS) data from IPUMS USA to estimate the total number of respondents with doctoral degrees across U.S.
+states using the Laplace ratio estimator.
+The estimated totals are compared to actual state-level respondent counts, highlighting discrepancies caused by regional variations in educational attainment.
+The results emphasize the limitations of applying a single ratio across diverse states and suggest that regional differences should be accounted for when using such estimation methods."
 format: pdf
@@
 
-To retrieve data from IPUMS USA, we first login to the IPUMS website and selected "IPUMS USA." From there, we clicked "Get Data" and chose the "2022 ACS" sample under the "SELECT SAMPLE" section.
+To retrieve data from IPUMS USA [@ipums2022], we first login to the IPUMS website and selected "IPUMS USA." From there, we clicked "Get Data" and chose the "2022 ACS" sample under the "SELECT SAMPLE" section.
 To obtain state-level data, we selected "HOUSEHOLD" \> "GEOGRAPHIC" and added the "STATEICP" variable to the data cart.
 For individual-level data, we went to the "PERSON" section and included the "EDUC" variable.
-After reviewing the cart, we clicked "CREATE DATA EXTRACT" and set the "DATA FORMAT" to "csv." Finally, we submitted the extract and then downloaded and saved the file locally (e.g., "usa_00001.csv") for use in R.
+After reviewing the cart, we clicked "CREATE DATA EXTRACT" and set the "DATA FORMAT" to "csv." Finally, we submitted the extract and then downloaded and saved the file locally (e.g., "usa_00001.csv") for use in R [@citeR].
+The `tidyverse` package [@tidyverse] was used to enhance the efficiency of data manipulation and data visualization
 
@@
 
-Next, we count the respondents with doctoral degrees grouped by states.
+Next, we count the respondents with doctoral degrees grouped by states as shown in.
+@tbl-pctraw1 presents an overview of the distribution of respondents with doctoral degrees across different states in the U.S., which is crucial for the understanding of regional educational attainment trends.
+The first column, 'State', refers to the unique state or territory codes provided by the IPUMS system, where each number corresponds to a specific state.
+The second column, 'Respondents with Doctoral Degree', provides the number of respondents with doctoral degrees (filtered by EDUCD == 116) in each state.
+For example, state code 1 shows 600 respondents with doctoral degrees, while state code 3 shows a significantly higher number with 2014 respondents.
+Other state codes show varying counts of respondents, such as 165 for state code 2, 244 for state code 4, and so on.
 
@@
   kable_styling(full_width = FALSE, font_size = 10) %>%
-  column_spec(1:2, width = "2cm")
+  column_spec(1:2, width = "5cm")
 
@@
 The Laplace ratio method works well when the relationship between the characteristic of interest, such as the proportion of doctoral degree holders, and the population is consistent across all units.
-In the example above, the ratio for California is 0.01619752.
-However, if this ratio is not representative of other states due to unobserved factors, the estimates derived from the method can become biased.
-This bias occurs when regional differences disrupt the assumption of proportionality, leading to inaccurate estimates.
-Factors like varying educational systems, socioeconomic conditions, or migration patterns between states can cause these distortions.
+In the example below, the ratio for California is 0.01619752.
 
@@
 
-california_doctoral_respondents
+california_ratio <- round(california_doctoral_respondents / california_total_respondents, 5)
+
+paste0("The California ratio that was found: ", california_ratio)
 
@@
 
-Using the ratio estimator approach might estimate the total number of respondents for each state as follows.
+Using the ratio estimator approach estimates the total number of respondents for each state as shown in @tbl-pctraw2.
 Assuming the ratio holds for all states, we can estimate the total respondents for every state.
-
-The table below presents an overview of the distribution of respondents with doctoral degrees across different states in the U.S., which is crucial for the understanding of regional educational attainment trends.
-The first column, STATEICP, refers to the unique state or territory codes provided by the IPUMS system, where each number corresponds to a specific state.
-The second column, doctoral_respondents, provides the number of respondents with doctoral degrees (filtered by EDUCD == 116) in each state.
-For example, state code 1 shows 600 respondents with doctoral degrees, while state code 3 shows a significantly higher number with 2014 respondents.
-Other state codes show varying counts of respondents, such as 165 for state code 2, 244 for state code 4, and so on.
 
@@
 #| label: tbl-pctraw2
-#| tbl-cap: The distribution of respondents with doctoral degrees across states with different STATEICPs (**Note:** The table is being displayed with only the first part shown to optimize space.)
+#| tbl-cap: Total number of respondents in each state, estimated by using the ratio estimators approach (**Note:** The table is being displayed with only the first part shown to optimize space.)
 #| echo: false
+#| warning: false
 
@@
 doctoral_count <- doctoral_count %>%
-  mutate(estimated_total = `Respondents with Doctoral Degree` / california_ratio)
+  mutate(estimated_total = round(`Respondents with Doctoral Degree` / california_ratio, 2))
 
-# Check the updated dataframe
-head(doctoral_count)
+# Display the table with custom column labels and rounded numbers
+kable(head(doctoral_count, 10), col.names = c("State", "Respondents with Doctoral Degree", "Estimated Total Number of Respondents")) %>%
+  kable_styling(full_width = FALSE, font_size = 10) %>%
+  column_spec(1:3, width = "5cm")
 
@@
 
-The process of comparing it with the actual data of respondents is as follows.The table below provides a comparison between the estimated total respondents, derived using the ratio estimator method, and the actual total respondents for each state.
-The doctoral_respondents column represents the number of individuals with doctoral degrees in each state, while the estimated_total_respondents column shows the projected total population for each state, calculated by applying the California ratio (0.01619752) to the number of doctoral respondents in each state.
-The actual_total_respondents column reflects the true population count for each state, which was directly obtained from the data.
+@tbl-pctraw3 provides a comparison between the estimated total respondents, derived using the ratio estimator method, and the actual total respondents for each state.
+The 'Respondents with Doctoral Degree' column represents the number of individuals with doctoral degrees in each state, while the 'Estimated Total Respondents' column shows the projected total population for each state, calculated by applying the California ratio (0.01619752) to the number of doctoral respondents in each state.
+The 'Actual Total Respondent' column reflects the true population count for each state, which was directly obtained from the data.
 This allows for a comparison between the estimated and actual figures.
@@
 
+\newpage
+
 ```{r}
+#| label: tbl-pctraw3
+#| tbl-cap: The comparison of actual and estimated total number of respondents in each state (**Note:** The table is being displayed with only the first part shown to optimize space.)
 #| echo: false
-#| tbl-cap: comparison of the estimated and actual total respondents
+#| warning: false
 
@@
 
+# Rename 'STATEICP' to 'State' in actual_total_count
+actual_total_count <- actual_total_count %>%
+  rename(State = STATEICP)
+
+# Perform the left join and clean up redundant columns
 doctoral_count <- doctoral_count %>%
-  left_join(actual_total_count, by = "STATEICP")
+  left_join(actual_total_count, by = "State") %>%
+  select(1:4) %>% 
+  rename("Estimated Total Respondents" = "estimated_total", "Actual Total Respondents" = "actual_total")
 
+# Display the table with the updated column names
 kable(head(doctoral_count)) %>%
   kable_styling(full_width = FALSE, font_size = 10) %>%
-  column_spec(1:4, width = "2cm")
+  column_spec(1:4, width = "3cm")
+
 ```
+@fig-compare uses a bar chart to show the differences between estimated and actual total respondents across every state of the U.S.
 
@@
 #| label: fig-compare
-#| fig-cap: comparison of estimated and total respondents
+#| fig-cap: Comparison of estimated and actual total respondents in each state of U.S.
 #| fig-height: 9
@@
 # Further process the data
-data_melted <- melt(doctoral_count, id.vars = "STATEICP", measure.vars = c("estimated_total", "actual_total"))
+data_melted <- melt(doctoral_count, id.vars = "State", measure.vars = c("Estimated Total Respondents", "Actual Total Respondents"))
 
-# Generate chart
-ggplot(data_melted, aes(x = factor(data_melted$STATEICP), y = value, fill = variable)) +
+# Generate the bar chart with corrected mappings
+ggplot(data_melted, aes(x = factor(State), y = value, fill = variable)) +
   geom_bar(stat = "identity", position = "dodge") +
-  labs(x = "State Code", y = "Number of Respondents", fill = "Respondent Type",
-       title = "Comparison of Estimated and Actual Total Respondents") +
+  labs(x = "State Code (STATEICP)", y = "Number of Respondents", fill = "Type") +
   theme_minimal() +
-  coord_flip()
+  coord_flip() +
+  scale_y_continuous(labels = scales::comma_format())  # Format y-axis with commas
+
 ```
@@
 This variability can lead to deviations between the sample-based ratio and the true population ratio, impacting the reliability of the estimation.
+
+\newpage
+
+# References

### 2024-11-21 paper/state_respondents.qmd
refinement of subtitle

@@
 title: "Estimating the Number of Doctoral Degree Holders in the U.S.: A Comparative Analysis Using Laplace Ratio Estimators"
-subtitle: "Discrepancy Between Estimated and Actual Totals Show Regional Variations in Educational Attainment"
+subtitle: "Discrepancy Between Estimated and Actual Total Respondents Show Regional Variations in Educational Attainment"
 author: 

### 2024-11-21 paper/state_respondents.qmd
fix typos

--- 2024-11-21 paper/state_respondents.qmd
@@
 ---
-title: "Estimating the Number of Doctoral Degree Holders in the U.S.: A Comparative Analysis Using Laplace Ratio Estimators"
-subtitle: "Discrepancy Between Estimated and Actual Totals Show Regional Variations in Educational Attainment"
+title: "Estimating the Number of Doctoral Degree Holders in the US: A Comparative Analysis Using Laplace Ratio Estimators"
+subtitle: "Discrepancy Between Estimated and Actual Total Respondents Show Regional Variations in Educational Attainment"
 author: 

### 2024-11-23 scripts/01-test_simulated_data.R
update simulation tests for the expected results

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests the structure and validity of the simulated infant health dataset. Shows expected results of the logistic regression model with simulated data.
 # Author: Yunkyung Ko
-# Date: 18 November 2024
+# Date: 23 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
@@ -106,22 +106,54 @@ sim_data <- sim_data %>%
     drug_count = rowSums(select(., c(indc, augmt, ster, antb, chor, anes)) == 1, na.rm = TRUE)
   )
 
-# Graph 1: Relationship between Number of Drugs Used and APGAR Score
-num_drugs <- ggplot(sim_data, aes(x = drug_count, y = apgar5)) +
-  geom_point(shape = 4, color = "blue", size = 2, alpha = 0.8) +  # Add blue X-shaped points
-  geom_smooth(method = "lm", color = "red", fill = "red", alpha = 0.2) + # Add trend line and CI
-  labs(
-    title = "Relationship between Number of Treatments During Labor/Delivery\nand the Infant's Health Status",
-    x = "Number of Treatments (Drug Usage) During Labor and Delivery",
-    y = "Five-Minute APGAR Score"
+# Graph 1: Relationship between Number of Treatments Used and APGAR Score
+simm_data <- sim_data %>%
+  mutate(
+    drug_count = rowSums(select(., c(indc, augmt, ster, antb, chor, anes)) == 1, na.rm = TRUE)
+  )
+
+summary_data_s <- simm_data %>%
+  group_by(drug_count) %>%
+  summarise(
+    mean_apgar5 = mean(apgar5),
+    sd_apgar5 = sd(apgar5)
+  )
+
+num_drugs <- ggplot(summary_data_s, aes(x = factor(drug_count), y = mean_apgar5)) +
+  geom_bar(
+    stat = "identity",
+    width = 0.5,  # Adjust width for thinner bars
+    fill = "skyblue",
+    color = "black",
+    alpha = 0.7
   ) +
-  scale_y_continuous(limits = c(0, 10), breaks = 0:10) +  # Ensure y-axis ranges from 0 to 10
+  geom_errorbar(
+    aes(ymin = mean_apgar5 - sd_apgar5, ymax = mean_apgar5 + sd_apgar5),
+    width = 0.2,  # Width of error bars remains the same
[…]

### 2024-11-23 scripts/03-clean_data.R
update data structure for prediction model

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Selects relevant variables for analysis and removes the outliers(unknown or unreported).
 # Author: Yunkyung Ko
-# Date: 21 November 2024
+# Date: 23 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
@@ -50,8 +50,6 @@ infant_health_na <-
 infant_health <- infant_health_na %>%
   drop_na()
 
-library(dplyr)
-
 # Define the desired sample size per `apgar5` score
 # You can use the minimum count across groups or specify a fixed number
 desired_sample_size <- 2000  # Adjust this based on your needs
@@ -62,12 +60,9 @@ balanced_data <- infant_health %>%
   sample_n(size = min(desired_sample_size, n()), replace = FALSE) %>%  # Ensure no oversampling
   ungroup()
 
-# Check the distribution of the new dataset
-balanced_data %>%
-  group_by(apgar5) %>%
-  summarise(Count = n()) %>%
-  print()
-
+# Convert the columns into factor so that they could be used for random forest prediction
+balanced_data <- balanced_data %>%
+  mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), ~ as.factor(.)))
 
 #### Save data ####
 write_csv(infant_health, "data/02-analysis_data/infant_health.csv")

### 2024-11-23 scripts/04-test_analysis_data.R

@@ -1,60 +1,75 @@
 #### Preamble ####
 # Purpose: Tests the structure and validity of the analysis dataset.
 # Author: Yunkyung Ko
-# Date: 18 November 2024
+# Date: 23 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
-  # - The `tidyverse`, `testthat` package must be installed and loaded
-  # - 03-clean_data.R must have been run 
+# - The `tidyverse`, `testthat` package must be installed and loaded
+# - 03-clean_data.R must have been run 
 # Any other information needed? Make sure you are in the `infant_health` rproj
-
 
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
-infant_health <- read_csv("../data/02-analysis_data/infant_health.csv")
 
-#### Test dataset structure ####
-# Test that the dataset has 3588588 rows
-test_that("dataset has 3588588 rows", {
-  expect_equal(nrow(infant_health), 3588588)
+# Load the dataset
+balanced_data <- read_csv("../data/02-analysis_data/infant_balanced.csv")
+balanced_data <- balanced_data %>%
+  mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), ~ as.factor(.)))
+
+#### Test 1: Dataset structure ####
+test_that("Dataset structure is correct", {
+  # Test that the dataset has 22000 rows
+  expect_equal(nrow(balanced_data), 22000)
+  
+  # Test that the dataset has 7 columns
+  expect_equal(ncol(balanced_data), 7)
+  
+  # Test column names
+  expected_columns <- c("indc", "augmt", "ster", "antb", "chor", "anes", "apgar5")
+  expect_true(all(expected_columns %in% colnames(balanced_data)),
+              "The dataset is missing one or more required columns.")
 })
 
-# Test that the dataset has 7 columns
-test_that("dataset has 7 columns", {
-  expect_equal(ncol(infant_health), 7)
-})
-
-#### Test for missing values ####
-# Test that there are no missing values in the dataset
-test_that("no missing values in dataset", {
-  expect_true(all(!is.na(infant_health)))
-})
-
-#### Test for variable types ####
-# Test numeric variables
-test_that("apgar5 is of double type", {
-  expect_type(infant_health$apgar5, "double")
-})
-
-
-# Test binary variables
-test_that("binary variables contain only 0 or 1", {
[…]

### 2024-11-23 scripts/06-model_data.R

@@ -34,19 +34,8 @@ randfor <- ranger(
 )
 
 
-# Model 2: Bayesian GLM for Predicting apgar5
-bayeslinear <- stan_glm(
-  formula = apgar5 ~ indc + augmt + ster + antb + chor + anes,  # Formula for the model
-  data = analysis_data,                                         # Dataset
-  family = gaussian(),                                          # Assumes normal errors for continuous outcome
-  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),  # Priors for coefficients
-  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE), # Prior for intercept
-  prior_aux = exponential(rate = 1, autoscale = TRUE),          # Prior for auxiliary parameters (e.g., sigma)
-  seed = 853                                                    # Seed for reproducibility
-)
-
 
 #### Save model ####
 saveRDS(randfor, file = "models/random_forest.rds")
-saveRDS(bayeslinear, file = "models/bayesian_glm.rds")
+
 

### 2024-11-23 paper/paper.qmd
tables and graphs for data/model results

@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The data transforms the phenomenon of infant health status into data, defining measurable variables that can capture the complex factors influencing the APGAR5 score.
+This transformation is achieved through the collection of information on various treatments and conditions experienced during labor and delivery, such as steroid use, chorioamnionitis, antibiotics, and anesthesia, among others.
+These variables, which are typically binary, are then recorded through surveys or medical records, using checkboxes or scales that allow for quantification.
+The APGAR5 score itself, is also a simplified numerical representation of an infant's health, which may overlook the complexities of their condition.
+
+By structuring these variables into numerical or categorical forms, the data can be analyzed and used to model the relationship between treatments and the APGAR5 score.
+However, this process involves challenges such as measurement error and biases, especially when data collection relies on self-reporting or clinician-recorded information.
 
@@
     high = "blue", 
-    name = "Count of Yes Responses", 
+    name = "Yes Responses", 
     guide = guide_colorbar(
@@
 
-Our results are summarized in @tbl-modelresults.
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+library(randomForest)
+library(caret)
+
+rf_model <-
+  readRDS(file = here::here("models/random_forest.rds"))
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-rfsummary1
+#| tbl-cap: Variable importance of treatments during labor and delivery on infant health, based on Random Forest model
+#| warning: false
+
+# Load necessary libraries
+library(tidyverse)
+library(caret)
+library(knitr)
+library(kableExtra)
+
+# Custom labels for the variables
+custom_labels <- c(
+  "indc1" = "Induction of Labor",
+  "augmt1" = "Augmentation of Labor",
+  "ster1" = "Steroids",
+  "antb1" = "Antibiotics",
+  "chor1" = "Chorioamnionitis",
+  "anes1" = "Anesthesia"
+)
+
+# Extract variable importance using caret
+importance_data <- varImp(rf_model, scale = TRUE)
+
+# Convert varImp.train object to a data frame
+importance_table <- as.data.frame(importance_data$importance)
+
+# Add variable names as a new column
+importance_table$Variable <- rownames(importance_table)
+
+# Transform the data frame
+importance_table <- importance_table %>%
+  mutate(
+    `Treatment Type` = recode(Variable, !!!custom_labels),  # Replace with custom labels
+    `Variable Importance` = Overall / sum(Overall) * 100  # Convert to percentages
+  ) %>%
+  select(`Treatment Type`, `Variable Importance`) %>%  # Select relevant columns
+  arrange(match(`Treatment Type`, custom_labels)) %>%  # Sort by custom_labels order
+  mutate(`Variable Importance` = round(`Variable Importance`, 2))  # Round to 2 decimal places
+
+# Display the importance table using kableExtra
+importance_table %>%
+  kable(
+    col.names = c("Treatment Type", "Variable Importance (%)")) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-rfsumbar
+#| fig-cap: Variable importance of treatments during labor and delivery on infant health, expressed in bar plots
+#| warning: false
+
+# Load necessary library
+library(ggplot2)
+
+# Horizontal bar plot
+ggplot(importance_table, aes(x = `Variable Importance`, y = `Treatment Type`)) +
+  geom_bar(stat = "identity", fill = "#DDA0DD", color = "grey") +
+  labs(
+    x = "Variable Importance (%)",
+    y = "Treatment Type"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.title.y = element_text(size = 12),
+    axis.text.x = element_text(size = 10),
+    axis.text.y = element_text(size = 10)
+  )
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-rfsummary2
+#| tbl-cap: Model accuracy metrics for Random Forest Model
+#| warning: false
+#| message: false
+
+# Load necessary libraries
+library(tidyverse)
+library(caret)
+library(knitr)
+library(kableExtra)
+
+# Predict on test data
+train_data <- read_csv("../data/02-analysis_data/train_test_data/train_data.csv")
+test_data <- read_csv("../data/02-analysis_data/train_test_data/test_data.csv")
+train_data <- train_data %>%
+  mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), as.factor))
+test_data <- test_data %>%
+  mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), as.factor))
+predictions <- predict(rf_model, newdata = test_data)
+
+# Calculate performance metrics
+actual <- test_data$apgar5
+rmse <- sqrt(mean((predictions - actual)^2))  # RMSE
+mae <- mean(abs(predictions - actual))       # MAE
+rss <- sum((predictions - actual)^2)         # Residual Sum of Squares
+tss <- sum((actual - mean(actual))^2)        # Total Sum of Squares
+r_squared <- 1 - (rss / tss)                 # R-Squared
+adjusted_r_squared <- 1 - ((1 - r_squared) * ((nrow(test_data) - 1) / (nrow(test_data) - length(rf_model$importance) - 1)))
+
+# Create a summary table
+accuracy_table <- tibble(
+  Metric = c("RMSE", "MAE", "R-Squared", "Adjusted R-Squared"),
+  Value = c(round(rmse, 2), round(mae, 2), round(r_squared, 2), round(adjusted_r_squared, 2))
+)
+
+# Display the table
+accuracy_table %>%
+  kable(
+    col.names = c("Metric", "Value")
+  ) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
 
@@
 
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
 \newpage

### 2024-11-23 scripts/03-clean_data.R
update the model for better accuracy

@@ -13,6 +13,7 @@
 library(data.table)
 library(tidyverse)
 library(arrow)
+library(caret)
 
 #### Clean data ####
 raw_data <- fread("data/01-raw_data/infant_data.csv")
@@ -64,9 +65,25 @@ balanced_data <- infant_health %>%
 balanced_data <- balanced_data %>%
   mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), ~ as.factor(.)))
 
+# Split data into training and testing sets to use it for prediction and accuracy check
+set.seed(853) # For reproducibility
+train_indices <- createDataPartition(balanced_data$apgar5, p = 0.7, list = FALSE)
+train_data <- balanced_data[train_indices, ]
+test_data <- balanced_data[-train_indices, ]
+
+
+
 #### Save data ####
 write_csv(infant_health, "data/02-analysis_data/infant_health.csv")
 write_parquet(infant_health, "data/02-analysis_data/infant_health.parquet")
 
 write_csv(balanced_data, "data/02-analysis_data/infant_balanced.csv")
 write_parquet(balanced_data, "data/02-analysis_data/infant_balanced.parquet")
+
+write_csv(train_data, "data/02-analysis_data/train_test_data/train_data.csv")
+write_parquet(train_data, "data/02-analysis_data/train_test_data/train_data.parquet")
+
+write_csv(test_data, "data/02-analysis_data/train_test_data/test_data.csv")
+write_parquet(test_data, "data/02-analysis_data/train_test_data/test_data.parquet")
+
+

### 2024-11-23 scripts/06-model_data.R

@@ -1,41 +1,51 @@
 #### Preamble ####
 # Purpose: Models for predicting the health status of the infant with data of drugs usage during labor and delivery
 # Author: Yunkyung Ko
-# Date: 18 November 2024
+# Date: 23 November 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
-  # - The `tidyverse`, `rstanarm` package must be installed and loaded
+  # - The `tidyverse`, `randomForest` package must be installed and loaded
   # - 03-clean_data.R must have been run
 # Any other information needed? Make sure you are in the `infant_health` rproj
 
 
 #### Workspace setup ####
-library(readr)
 library(randomForest)
 library(caret)
-library(rstanarm)
-library(ranger)
+library(tidyverse)
 
 #### Read data ####
-infant_health <- read_csv("data/02-analysis_data/infant_health.csv")
+infant_balanced <- read_csv("data/02-analysis_data/infant_balanced.csv")
+train_data <- read_csv("data/02-analysis_data/train_test_data/train_data.csv")
+test_data <- read_csv("data/02-analysis_data/train_test_data/test_data.csv")
 
-### Model data ####
+# Ensure binary variables are factors
+infant_balanced <- infant_balanced %>%
+  mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), as.factor))
+train_data <- train_data %>%
+  mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), as.factor))
+test_data <- test_data %>%
+  mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), as.factor))
 
-# Model 1: Random Forest for Predicting apgar5
-set.seed(853)  # Ensure reproducibility
-randfor <- ranger(
-  formula = apgar5 ~ indc + augmt + ster + antb + chor + anes,  # Formula for the model
-  data = infant_health,                                        # Dataset
-  num.trees = 500,                                             # Correct argument for number of trees
-  mtry = 3,                                                    # Number of variables tried at each split
-  importance = "impurity",                                     # Variable importance
-  num.threads = 4                                              # Number of cores for parallelism
+# Hyperparameter tuning for Random Forest
+set.seed(853)
+control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation
[…]

### 2024-11-26 scripts/05-exploratory_data_analysis.R
delete unnecessary files

@@ -1,37 +1,0 @@
-#### Preamble ####
-# Purpose: Models to predict the health status of an infant with the data of treatments/drug usage during delivery and labor
-# Author: Yunkyung Ko
-# Date: 18 November 2024
-# Contact: yunkyung.ko@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
-
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
-
-
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
-
-

### 2024-11-26 scripts/06-model_data.R
delete duplicate files

@@ -1,51 +1,0 @@
-#### Preamble ####
-# Purpose: Models for predicting the health status of the infant with data of drugs usage during labor and delivery
-# Author: Yunkyung Ko
-# Date: 23 November 2024
-# Contact: yunkyung.ko@mail.utoronto.ca
-# License: MIT
-# Pre-requisites:
-  # - The `tidyverse`, `randomForest` package must be installed and loaded
-  # - 03-clean_data.R must have been run
-# Any other information needed? Make sure you are in the `infant_health` rproj
-
-
-#### Workspace setup ####
-library(randomForest)
-library(caret)
-library(tidyverse)
-
-#### Read data ####
-infant_balanced <- read_csv("data/02-analysis_data/infant_balanced.csv")
-train_data <- read_csv("data/02-analysis_data/train_test_data/train_data.csv")
-test_data <- read_csv("data/02-analysis_data/train_test_data/test_data.csv")
-
-# Ensure binary variables are factors
-infant_balanced <- infant_balanced %>%
-  mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), as.factor))
-train_data <- train_data %>%
-  mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), as.factor))
-test_data <- test_data %>%
-  mutate(across(c("indc", "augmt", "ster", "antb", "chor", "anes"), as.factor))
-
-# Hyperparameter tuning for Random Forest
-set.seed(853)
-control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation
-grid <- expand.grid(mtry = seq(2, 6, by = 1))  # Search for mtry between 2 and 6
-
-rf_tuned <- train(
-  apgar5 ~ indc + augmt + ster + antb + chor + anes,
-  data = train_data,
-  method = "rf",
-  trControl = control,
-  tuneGrid = grid,
-  ntree = 500,  # Number of trees
-  importance = TRUE
-)
-
-
-#### Save model ####
-saveRDS(rf_tuned, file = "models/random_forest.rds")
-
-
-

### 2024-11-26 paper/paper.qmd
paper draft data and result section update

--- 2024-11-21 paper/paper.qmd
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+We use the '2023 Natality Data for the United States' to analyze the relationship between drug treatments administered to infants and their health outcomes.
+Provided by NBER and sourced from the National Center for Health Statistics [@rawdata], this dataset contains 3,605,081 live birth statistics across the U.S., including demographic details, health metrics, and geographical breakdowns by state and country [@natality2023].
+To represent treatments directly affecting the infant, variables under the category of 'Characteristics of Labor and Delivery' were selected, which capture medical interventions administered to the mother or infant during the critical period of childbirth [@cdc_facility_worksheet_2016].
+Six different types of treatments are under this category, which are labor induction, augmentation, the use of steroids, antibiotics, chorioamnionitis, and anesthesia [@natality2023].
+For the representation of infant health outcomes, the 'APGAR5' variable was selected, which shows the 5-minute Apgar score, a widely recognized indicator of an infant's immediate health status post-delivery [@medicalnewstoday_apgar_scores].
+
+To ensure a balanced distribution of observations across all Apgar5 score groups, we applied stratified random sampling and designated the sample size per group as 2,000.
+In addition, for data quality, we standardized the treatment measuring variables, containing respondents' answers of "Yes", "No", or "Unknown", into binary numerical variables.
+Then, the variables were converted into factors to prepare the data for predictive modeling[@analyticsvidhya2015], and split into training and testing sets to support machine learning workflows[@brownlee2019].
+
+In performing the analysis, we utilized the statistical programming language R [@citeR] and several R packages.
+`tidyverse` [@tidyverse] was used for data manipulation, and `ggplot2`[@ggplot2] was used for visualizing results in graphical methods.
+`randomForest` [@randomForest] and `caret` [@caret] were used for Random Forest modeling while `rstanarm` [@rstanarm] was used for Bayesian linear modeling in the process of generating predictions.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The data transforms the phenomenon of infant health status into data, defining quantifiable factors that can capture the complex impacts of maternal and neonatal pharmacological interventions [@rawdata].
+The transformation is achieved through the collection of information on various treatments and conditions experienced during labor and delivery, such as steroid use, chorioamnionitis, antibiotics, and anesthesia, among others [@natality2023].
+These variables, which were recorded through surveys or medical records, uses checkboxes or scales that allow for quantification [@cdc_facility_worksheet_2016].
+Errors such as inconsistencies in how data is recorded can arise during this kind of measurement process [@tellingstories].
+For instance, subjective interpretations of checkbox responses or variability in medical record documentation can lead to measurement error [@healthknowledge2024].
+These errors may result from differences in definitions of treatments across facilities or the accuracy of self-reported information in surveys [@healthknowledge2024].
+
+The 5-minute Apgar score itself, is also a simplified numerical representation of an infant's health, which may overlook the complexities of their condition.
+While useful for rapid clinical decision-making, the Apgar score may fail to capture nuanced aspects of an infant's condition, such as underlying metabolic imbalances, subtle neurological issues, or long-term impact of perinatal complications [@aap2006].
+
+By structuring these variables into categorical or numerical forms, the data can be analyzed and used to model the relationship between treatments and the APGAR5 score.
+Specifically, drug usage during labor and delivery was encoded as binary variables (@sec-predictor).
+However, this has a limitation that it disregards the quantitative intensity, dosage, or frequency of drug administration, leading to incomplete modeling of the treatment's impact [@geeksforgeeks_feature_encoding].
+This simplification may also amplify noise in the data by grouping dissimilar observations into the same category, reducing the model's ability to detect subtle relationships [@geeksforgeeks_feature_encoding].
 
@@
 
-The Apgar score is a measure of the need for resuscitation [@natality2023] to the infant, which is the act of bringing someone back to life or waking them [@cambridgeResuscitation].
+The Apgar score is a measure of the need for resuscitation [@natality2023] to the infant, which is "the act of bringing someone back to life or waking them" [@cambridgeResuscitation].
 It is a test given to newborns soon after birth (5 minutes) to check 'Appearance(skin color)', 'Pulse(heart rate)', 'Grimace response(reflexes)', 'Activity(muscle tone)', 'Respiration(breathing rate and effort)' [@kidsHealthApgar].
@@
 
-To ensure a balanced analysis and accurately assess the effects of drug treatment during delivery and labor, the data was refined to achieve a more even distribution of observations across APGAR5 scores.
+Therefore, to ensure a balanced analysis [@brownlee2020], the data was refined to achieve a more even distribution of observations across APGAR5 scores.
 Based on the lowest observation count of 2,065 in the original distribution, the number of observations for each APGAR5 score was set to 2,000 as shown in @fig-apgar3.
-Random sampling was used to ensure a fair distribution across all score levels.
+Random sampling was used for respective score groups to ensure a fair distribution across all score levels [@brownlee2020].
 
@@
 
-## Predictor variables
-
-All of the predictor variables used in the analysis are classified into the same category, which is the 'characteristics of labor and delivery'.
-This item, which contains 6 separate checkboxes that the respondent can choose from, allows for the reporting of more than one characteristic and includes a choice of "None of the above"[@natality2023].
+## Predictor variables {#sec-predictor}
+
+All of the predictor variables used in the analysis are classified under the same category, which is the 'Characteristics of labor and delivery' [@natality2023].
+This item, which contains 6 separate checkboxes that the respondent can choose from, allows for the reporting of more than one characteristic and includes a choice of "None of the above" [@cdcFacilityGuide].
 
@@
 ### Type of Treatments Used During Delivery and Labor
+
+@fig-typetreat shows that the distribution of treatment types varies significantly, with anesthesia being the most commonly recorded intervention, while chorioamnionitis is rare.
+Treatments like anesthesia and antibiotics, with higher observation counts, may have a greater impact on the observed outcomes, whereas less frequent interventions, such as steroids or chorioamnionitis, may require careful consideration to avoid biases due to smaller sample sizes [@selectstatistics2024].
 
@@
 
+@fig-indc1 shows that in the raw data, the "No" category dominates with over 2.4 million observations, while the "Yes" category trails with around 1.2 million.
+After filtering and refining the dataset, @fig-indc2 shows that the gap between the "Yes" and "No" category becomes narrow, with the "No" category with 15,752 and "Yes" category with 6,248.
+This kind of balancing process is expected to allow for a more robust evaluation of how labor induction might influence Apgar scores, reducing the risk of the prediction model underestimating or overestimating treatment effects [@isi2024].
+
 ```{r}
@@
 ```
+
 ```{r}
 #| label: fig-indc2
-#| fig-cap: The distribution of observations in the analysis dataset on whether the infant received induction of labor or not,
+#| fig-cap: The distribution of observations in the analysis dataset based on whether the infant received induction of labor or not
 #| echo: false
@@
 
+@fig-augmt1 shows that in the raw data, the "No" category overwhelmingly dominates with 2,850,569 observations compared to 752,302 in the "Yes" category.
+@fig-augmt2 narrow the gap with the "No" category reduced to 18,589 and "Yes" category adjusted to 3,411.
+
 ```{r}
@@
 ```
+
 ```{r}
 #| label: fig-augmt2
-#| fig-cap: The distribution observations in the analysis datset on whether the infant received augmentation of labor or not
+#| fig-cap: The distribution observations in the analysis datset based on whether the infant received augmentation of labor or not
 #| echo: false
@@
 
+@fig-ster1 shows that in the raw data, the "No" category heavily dominates with 3,464,273 observations compared to 138,598 in the "Yes" category, indicating severe imbalance.
+@fig-ster2 shows that the gap narrows after filtering and refining the dataset, with the "No" category to 19,662 and the "Yes" category to 2,338.
+
 ```{r}
@@
 #| label: fig-ster2
-#| fig-cap: The distribution observations in the analysis datset on whether the infant received steroid treatments or not
+#| fig-cap: The distribution observations in the analysis datset based on whether the infant received steroid treatments or not
 #| echo: false
@@
 
+@fig-antb1 shows that in the raw data, the "No" category dominates with 2,701,457 observations, while the "Yes" category has significantly fewer at 901,414.
+After balancing, @fig-antb2 shows that the "No" category is reduced to 15,828 and the "Yes" category is adjusted to 6,172, narrowing the gap.
+
 ```{r}
@@
 #| label: fig-antb2
-#| fig-cap: The distribution observations in the analysis datset on whether the infant received antibiotic treatments or not
+#| fig-cap: The distribution observations in the analysis datset based on whether the infant received antibiotic treatments or not
 #| echo: false
@@
 
+@fig-chor1 shows that the "No" category dominates with 3,539,254 observations, compared to 63,617 in the "Yes" category.
+@fig-chor2 shows that the extreme disparity between the two categories is addressed after balancing, with the "No" category having 21,213 and the "Yes" category having 787.
+
 ```{r}
@@
 #| label: fig-chor2
-#| fig-cap: The distribution observations in the analysis datset on whether the infant received chorioamnionitis treatments or not
+#| fig-cap: The distribution observations in the analysis datset based on whether the infant received chorioamnionitis treatments or not
 #| echo: false
@@
 
+@fig-anes1 shows that in the raw data, the "Yes" category dominates with 2,810,461 observations, while the "No" category has 792,410.
+The initial overrepresentation of the "Yes" category is addressed through the balancing process, with @fig-anes2 showing the "Yes" category of 15,387 and the "No" category of 6,613.
+
 ```{r}
@@
 #| label: fig-anes2
-#| fig-cap: The distribution observations in the analysis datset on whether the infant received anesthesia treatments or not
+#| fig-cap: The distribution observations in the analysis datset based on whether the infant received anesthesia treatments or not
 #| echo: false
@@
 
+@fig-corr1 shows that most observations fall into the category where neither induction nor augmentation is performed, indicating these interventions are less commonly used together.
+Moreover, smaller proportions are observed in combinations where one or both interventions are present, suggesting a potential correlation where the likelihood of augmentation increases when induction is performed [@friendly1992].
+
 ```{r}
 #| label: fig-corr1
-#| fig-cap: Mosaic plot showing the correlation between induction and augmentation of labor
+#| fig-cap: Mosaic plot showing the correlation between induction and augmentation of labor (**Note:** the shading intensity (from light green to dark green) represents the standardized residuals of the chi-squared test for the contingency table.)
 #| echo: false
@@
 
+In @fig-corr2, most cases fall into the "neither used" category (Steroids:0, Antibiotics:0), shown by the large, darkly shaded area.
+This means that it is very common for neither treatment to be used together during labor [@friendly1992].
+On the other hand, the smaller, lighter areas where either or both treatments are used show that these combinations happen much less often than expected [@friendly1992].
+This suggests that steroids and antibiotics are usually not given together, and their usage might depend on specific and separate medical needs rather than being commonly paired treatments.
+
 ```{r}
 #| label: fig-corr2
-#| fig-cap: Mosaic plot showing the correlation between usage of steroids and antibiotics
+#| fig-cap: Mosaic plot showing the correlation between usage of steroids and antibiotics (**Note:** the shading intensity (from light green to dark green) represents the standardized residuals of the chi-squared test for the contingency table.)
 #| echo: false
@@
 
+In @fig-corr3, the large, darkly shaded area for "Chorioamnionitis:0" and "Antibiotics:0" shows that cases where neither condition is present are more common than expected [@friendly1992].
+Conversely, the smaller and lighter areas, especially for "Chorioamnionitis:1" and "Antibiotics:1", suggest that when chorioamnionitis occurs, antibiotics are often used.
+The pattern reflects a likely positive correlation [@friendly1992] between chorioamnionitis and anitbiotics usage.
+
 ```{r}
 #| label: fig-corr3
-#| fig-cap: Mosaic plot showing the correlation between usage of chorioamnionitis and antibiotics
+#| fig-cap: Mosaic plot showing the correlation between usage of chorioamnionitis and antibiotics (**Note:** the shading intensity (from light green to dark green) represents the standardized residuals of the chi-squared test for the contingency table.)
 #| echo: false
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The goal of our modelling strategy is twofold.
+Firstly, a Random Forest model was utilized to identify significant predictors of infant health status, as measured by the 5-minute Apgar score.
+Secondly, a Bayesian Linear Model was applied to quantify the effects of medical interventions during labor and delivery on infant health outcomes, incorporating prior knowledge into the analysis framework.
+
+Here we briefly describe the Random Forest and Bayesian analysis models used to investigate these relationships. Background details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+Define $y_i$ as the APGAR5 score for the $i$th infant.
+Let the binary treatment variables $\text{indc}_i$, $\text{augmt}_i$, $\text{ster}_i$, $\text{antb}_i$, $\text{chor}_i$, and $\text{anes}_i$ represent the use of induction, augmentation, steroids, antibiotics, chorioamnionitis, and anesthesia, respectively, during labor and delivery.
+
+### Random Forest Model
+
+The Random Forest Model is a non-parametic method that predicts $y_i$ using an ensemble of decision trees [@mlnuggets2024]:
+
+$\hat{y}_i = \text{RandomForest}(X)$
+
+where $X$ represents the set of predictors: $(\text{indc}_i, \text{augmt}_i, \text{ster}_i, \text{antb}_i, \text{chor}_i, \text{anes}_i)$.
+
+We employ a grid search for hyperparameter tuning:
+  - $mtry$: Number of variables randomly sampled at each split (values: 2, 3, 4).
+  - $ntree$: Number of trees in the forest, fixed at 500.
+
+The `caret` package[@caret] in R[@citeR] is used to implement the Random Forest model, with cross-validation to minimize overfitting.
+
+### Bayesian Linear Model
+
+The Bayesian Linear Model assumes the following relationship for each infant $i$:
+
+$y_i | \mu_i, \sigma \sim \text{Normal}(\mu_i, \sigma)$
+
+$\mu_i = \alpha + \beta_1 \cdot \text{indc}_i + \beta_2 \cdot \text{augmt}_i + \beta_3 \cdot \text{ster}_i + \beta_4 \cdot \text{antb}_i + \beta_5 \cdot \text{chor}_i + \beta_6 \cdot \text{anes}_i$
+
+Priors are specified as:
+
+$\alpha \sim \text{Normal}(5, 2), \quad \beta_j \sim \text{Normal}(0, 2) \quad \text{for } j = 1, \ldots, 6, \quad \sigma \sim \text{Exponential}(1)$
+
+We fit the Bayesian model using the `stan_glm` function in the `rstanarm`[@rstanarm] package, which implements Markov Chian Monte Carlo (MCMC) sampling with 4 chaing and 2000 iterations per chain.
+We use the default priors from `rstanarm`[@rstanarm].
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+The Random Forest Model was chosen for its robustness in handling non-linear relationships and complex interactions among predictors [@geeksforgeeks_random_forest], which is essential when analyzing the combined effects of various medical interventions on infant health outcomes.
+This method also provides interpretable measures of variable importance [@geeksforgeeks_random_forest], aiding in understanding the relative impact of each treatment.
+
+The Bayesian Linear Model was employed to incorporate prior knowledge [@strimmerlab_priors] regarding the expected effects of treatments on the Apgar5 score.
+Utilizing weakly informative priors, such as Normal distributions centered around zero with a variance of 2, reflects a belief in moderate but uncertain associations between treatments and outcomes [@strimmerlab_priors].
+The prior for the intercept (Normal(5, 2)) aligns with the central tendency of Apgar5 scores reported in previous studies [@clevelandclinic_apgar].
+An exponential prior for the standard deviation parameter ensures positivity while discouraging extreme variability [@strimmerlab_priors].
+
+By integrating the Random Forest's predictive accuracy [@geeksforgeeks_random_forest] with the Bayesian Linear Model's capacity for inference and uncertainty quantification [@nguyen2018uncertainty], this dual-model approach offers a comprehensive understanding of the factors influencing infant health outcomes.
 
@@
 
+### Relationship between the number of treatments received during labor/delivery and the mean of Apgar5 scores
+
+@fig-result1 shows that the mean of Apgar5 scores remain relatively stable around 5, across different number of treatments during labor and delivery.
+The dark blue error bars, which are the variability or uncertainty around the mean scores, show the wide range of outcomes within each category.
+While a slight decline in mean scores is observed as treatments increase from 0 to 5, the score rebounds at 6 treatments, suggesting no clear linear relationship.
+This proposes that the number of treatments alone may not significantly affect Apgar5 scores.
+
 ```{r}
 #| label: fig-result1
-#| fig-cap: The mean of APGAR5 scores by the number of treatments recieved during labor and delivery
+#| fig-cap: The mean of Apgar5 scores by the number of treatments received during labor and delivery (**Note:** The dark blue lines represent error bars, indicating the variability or uncertainty around the mean Apgar5 scores for each number of treatments.)
 #| echo: false
@@
     width = 0.2,  # Width of error bars remains the same
-    color = "blue"
+    color = "darkblue"
   ) +
@@
     x = "Number of Treatments During Labor and Delivery",
-    y = "Mean of APGAR5 Score"
+    y = "Mean of Apgar5 Score"
   ) +
@@
 
+### Relationship between the type of treatments administered during labor/delivery and the Apgar5 scores
+
+@fig-result2 shows that most treatments during labor and delivery are associated with normal Apgar scores ranging from 7 to 10, as indicated by the darker blue shades in these columns.
+Treatments such as usage of anesthesia and antibiotics exhibit broader distributions across Apgar scores, suggesting their use in a wide range of delivery conditions.
+Lower Apgar scores ranging from 0 to 3 are relatively rare, shown by lighter shades across treatments.
+Moreover, drug usage of chorioamnionitis and steroids show fewer "Yes" responses overall, suggesting their more targeted use.
+
 ```{r}
 #| label: fig-result2
-#| fig-cap: Heatmap of the overall correlation between treatments administered during delivery and APGAR5 scores
+#| fig-cap: Heatmap of the overall correlation between treatments administered during labor/delivery and Apgar5 scores
 #| echo: false
@@
   labs(
-    x = "Five-Minute APGAR Score",
+    x = "Five-Minute Apgar Score",
     y = "Treatments During Labor and Delivery"
@@
 
+@fig-indcscore shows that observation counts increase steadily from scores 0 to 5, remain relatively stable between scores 5 and 9, and then drop sharply at score 10.
+This pattern suggests that most infants who experienced induction during labor tend to achieve mid-to-high Apgar5 scores, reflecting generally stable and favorable health outcomes.
+
 ```{r}
 #| label: fig-indcscore
-#| fig-cap: APGAR5 score distribution for the usage of induction labor
+#| fig-cap: Apgar5 score distribution for the usage of induction labor (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
 #| echo: false
@@
   labs(
-    x = "Five-Minute APGAR Score",
+    x = "Five-Minute Apgar Score",
     y = "Observation Count"
@@
 
+@fig-augmtscore shows that observation counts are lower at extreme scores of 0 and 10, steadily increase from scores 1 to 5, plateau between 5 and 8, and peak at score 9 before sharply declining at score 10.
+This pattern suggests that augmentation of labor is associated with mid-to-high Apgar5 scores, reflecting favorable infant health outcomes in most cases, while fewer cases exhibit very low or extremely high scores.
+
 ```{r}
 #| label: fig-augmtscore
-#| fig-cap: APGAR5 score distribution for the usage of augmentation labor
+#| fig-cap: Apgar5 score distribution for the usage of augmentation labor (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
 #| echo: false
@@
 
+@fig-sterscore shows that observation counts are relatively low at extreme scores of 0 and 10, increase steadily from scores 0 to 4, and peak at score 4 before gradually declining through scores 5 to 10.
+This pattern proposes that usage of steroids for fetal lung maturation received by the mother before delivery is associated with a concentration of Apgar5 scores in the mid-range, reflecting moderate infant health outcomes in most cases.
+
 ```{r}
 #| label: fig-sterscore
-#| fig-cap: APGAR5 score distribution for the usage of steroids
+#| fig-cap: Apgar5 score distribution for the usage of steroids (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
 #| echo: false
@@
 
+@fig-antbscore shows that observation counts rise gradually from scores 0 to 2, and maintain a steady plateau between scores 4 and 7, before declining through scores 8 to 10.
+This pattern suggests that infants exposed to antibiotics received by the mother during delivery tend to achieve moderate Apgar5 scores, with fewer observations at both low and high extremes.
+
 ```{r}
 #| label: fig-antbscore
-#| fig-cap: APGAR5 score distribution for the usage of antibiotics
+#| fig-cap: Apgar5 score distribution for the usage of antibiotics (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
 #| echo: false
@@
 
+@fig-chorscore shows that the observation count peaks sharply at a score of 1, then stabilizes between scores 2 and 5, followed by a gradual decline from scores 6 through 10.
+This distribution suggests that infants that were diagnosed clinical chorioamnionitis during labor or maternal temperature over 38 degrees celcius tend to achieve lower-to-mid range Apgar5 scores.
+
 ```{r}
 #| label: fig-chorscore
-#| fig-cap: APGAR5 score distribution for the usage of chorioamnionitis
+#| fig-cap: Apgar5 score distribution for the usage of chorioamnionitis (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
 #| echo: false
@@
 
+@fig-anesscore shows that observation counts steadily increase from scores 0 to 5.
+Beyond score 5, the counts plateau, maintaining a relatively stable distribution across scores 6 to 9, followed by a sharp decline at score 10.
+This suggests the widespread distribution of higher Apgar scores when epidural or spinal anesthesia is used during labor.
+
 ```{r}
 #| label: fig-anesscore
-#| fig-cap: APGAR5 score distribution for the usage of anesthesia
+#| fig-cap: Apgar5 score distribution for the usage of anesthesia (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
 #| echo: false
@@
 
-Our results are summarized in @tbl-modelresults.
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+library(randomForest)
+library(caret)
+library(rstanarm)
+
+rf_model <-
+  readRDS(file = here::here("models/random_forest.rds"))
+
+bayesian_model <-
+  readRDS(file = here::here("models/bayesian_lm.rds"))
+```
+
+### Random Forest Model Results
+
+@tbl-rfsummary1 and @fig-rfsumbar shows that usage of steroids (45.76%) is the most influential factor among the treatments during labor and delivery when predicting Apgar5 scores.
+It is followed by usage of chorioamnionitis (22.44%) and anesthesia (13.99%), which also contributes substantially.
+Meanwhile, the exposure to antibiotics (0.00%) shows no measurable impact in this context, and suggests its irrelevance to the model's predictions.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-rfsummary1
+#| tbl-cap: Variable importance of treatments during labor and delivery on infant health, based on Random Forest model
+#| warning: false
+
+# Load necessary libraries
+library(tidyverse)
+library(caret)
+library(knitr)
+library(kableExtra)
+
+# Custom labels for the variables
+custom_labels <- c(
+  "indc" = "Induction of Labor",
+  "augmt" = "Augmentation of Labor",
+  "ster" = "Steroids",
+  "antb" = "Antibiotics",
+  "chor" = "Chorioamnionitis",
+  "anes" = "Anesthesia"
+)
+
+# Extract variable importance using caret
+importance_data <- varImp(rf_model, scale = TRUE)
+
+# Convert varImp.train object to a data frame
+importance_table <- as.data.frame(importance_data$importance)
+
+# Add variable names as a new column
+importance_table$Variable <- rownames(importance_table)
+
+# Transform the data frame
+importance_table <- importance_table %>%
+  mutate(
+    `Treatment Type` = recode(Variable, !!!custom_labels),  # Correctly map to custom labels
+    `Variable Importance` = Overall / sum(Overall) * 100  # Convert to percentages
+  ) %>%
+  select(`Treatment Type`, `Variable Importance`) %>%  # Select relevant columns
+  mutate(`Variable Importance` = round(`Variable Importance`, 2)) %>%  # Round to 2 decimal places
+  arrange(match(`Treatment Type`, c("Induction of Labor", "Augmentation of Labor", 
+                                    "Steroids", "Antibiotics", 
+                                    "Chorioamnionitis", "Anesthesia")))  # Ensure correct order
+
+# Display the importance table using kableExtra
+importance_table %>%
+  kable(
+    col.names = c("Treatment Type", "Variable Importance (%)")) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-rfsumbar
+#| fig-cap: Variable importance of treatments during labor and delivery on infant health, expressed in bar plots
+#| warning: false
+
+# Load necessary library
+library(ggplot2)
+
+# Reorder the levels of 'Treatment Type' to match the custom labels
+importance_table$`Treatment Type` <- factor(
+  importance_table$`Treatment Type`,
+  levels = c("Induction of Labor", "Augmentation of Labor", "Steroids", "Antibiotics", "Chorioamnionitis", "Anesthesia")
+)
+
+# Create the horizontal bar plot with the specified order
+ggplot(importance_table, aes(x = `Variable Importance`, y = `Treatment Type`)) +
+  geom_bar(stat = "identity", fill = "#DDA0DD", color = "grey") +
+  labs(
+    x = "Variable Importance (%)",
+    y = "Treatment Type"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.title.x = element_text(size = 12),
+    axis.title.y = element_text(size = 12),
+    axis.text.x = element_text(size = 10),
+    axis.text.y = element_text(size = 10)
+  )
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-rfsummary2
+#| tbl-cap: Model accuracy metrics for Random Forest Model
+#| warning: false
+#| message: false
+
+# Load necessary libraries
+library(tidyverse)
+library(caret)
+library(knitr)
+library(kableExtra)
+
+# Predict on test data
+train_data <- read_csv("../data/02-analysis_data/train_test_data/train_data.csv")
+test_data <- read_csv("../data/02-analysis_data/train_test_data/test_data.csv")
+predictions <- predict(rf_model, newdata = test_data)
+
+# Calculate performance metrics
+actual <- test_data$apgar5
+rmse <- sqrt(mean((predictions - actual)^2))  # RMSE
+mae <- mean(abs(predictions - actual))       # MAE
+rss <- sum((predictions - actual)^2)         # Residual Sum of Squares
+tss <- sum((actual - mean(actual))^2)        # Total Sum of Squares
+r_squared <- 1 - (rss / tss)                 # R-Squared
+adjusted_r_squared <- 1 - ((1 - r_squared) * ((nrow(test_data) - 1) / (nrow(test_data) - length(rf_model$importance) - 1)))
+
+# Create a summary table
+accuracy_table <- tibble(
+  Metric = c("RMSE", "MAE", "R-Squared", "Adjusted R-Squared"),
+  Value = c(round(rmse, 2), round(mae, 2), round(r_squared, 2), round(adjusted_r_squared, 2))
+)
+
+# Display the table
+accuracy_table %>%
+  kable(
+    col.names = c("Metric", "Value")
+  ) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+### Bayesian Linear Model Results
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| warning: false
+
+modelsummary::modelsummary(
+  list(
+    "Bayesian Linear Model" = bayesian_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
 
@@
 
-## Second discussion point
+## Second discussion point {#sec-second-point}
 
@@
 
-## Third discussion point
+## Third discussion point {#sec-third-point}
 
@@
 
-# Appendix {-}
+# Appendix {#sec-appendix}
 
@@
 
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheckandposteriorvsprior
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+
+pp_check(bayesian_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+posterior_vs_prior(bayesian_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+```
+
 ## Diagnostics
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-stanareyouokay
+#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| layout-ncol: 2
+
+plot(bayesian_model, "trace")
+
+plot(bayesian_model, "rhat")
+```
+
 \newpage

### 2024-11-26 paper/paper.qmd
paper draft for results and discussions section

@@
 
-### Random Forest Model Results
+### Random Forest Model Results {#sec-randfor-result}
 
@@
 # Reorder the levels of 'Treatment Type' to match the custom labels
-importance_table$`Treatment Type` <- factor(
-  importance_table$`Treatment Type`,
+importance_table <- importance_table %>%
+  mutate(`Treatment Type` = factor(
+    `Treatment Type`,
     levels = c("Induction of Labor", "Augmentation of Labor", "Steroids", "Antibiotics", "Chorioamnionitis", "Anesthesia")
-)
+  ))
 
@@
 ```
+
+@tbl-rfsummary2 shows the Random Forest model, despite incorporating hyperparameter tuning, demonstrated suboptimal performance.
+Specifically, the Root Mean Squared Error (RMSE) of 3.12 and Mean Absolute Error (MAE) of 2.69 suggest that the model's predictions deviate significantly from the observed values on average [@vitalflux_metrics].
+Furthermore, the R-squared and Adjusted R-squared values of 0.02 suggest the model captures only 2% of the variance in the target variable [@vitalflux_metrics].
+These results propose that further refinement is necessary, potentially through reevaluating feature engineering, exploring additional hyperparameter configurations, or considering alternative modeling approaches better suited for the data [@AnalyticsVidhyaFeatureEngineering2021].
 
@@
 
+Our results from the Bayesian Linear Model are summarized in @tbl-modelresults.
+
 ```{r}
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| tbl-cap: Bayesian linear models quantifying the effects of medical interventions during labor and delivery on infant health outcomes
 #| warning: false
@@
 
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point {#sec-second-point}
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+## Pharmaceutical companies should focus on optimizing antenatal corticosteroid formulations {#sec-first-point}
+
+@sec-randfor-result reveals that maternal and neonatal treatments during labor, such as steroids and antibiotics, have differential impacts on infant health outcomes.
+Usage of steroids, which showed positive correlations with the Apgar5 score in @fig-sterscore, was actually affirmed their critical role in improving neonatal survival, especially in extremely preterm infants [@nih_prenatal_steroid_treatment].
+In particular, antenatal corticosteroids stand out as the core intervention for preterm births under the gestation period of 22 weeks, significantly reducing complications like respiratory distress syndrome by accelerating fetal lung development [@nih_prenatal_steroid_treatment].
+
+In accordance to these analyses results, pharmaceutical companies can focus on optimizing steroid formulations for safer and more effective administration during earlier gestations [@nih_prenatal_steroid_treatment].
+Current formulations of antenatal corticosteroids, like betamethasone, have shown strong efficacy in improving neonatal outcomes but may carry risks when administered too early or in repeated doses [@nichd_prenatal_steroid_treatment].
+Hence, adjusting dosages or delivery mechanisms could enhance safety while preserving their life-saving benefits for preterm infants.
+For instance, this could involve sustained-release versions or targeted delivery systems to minimize potential side effects on both the mother and fetus [@nichd_prenatal_steroid_treatment].
+
+## Pharmaceutical companies, policymakers, and healthcare providers should collaborate to develop targeted antibiotics and enforce evidence-based protocols {#sec-second-point}
+
+Contrary to the influence of antibiotics dosage, the use of antibiotics appears less impactful as shown in @fig-rfsumbar.
+While crucial for preventing infections, their overuse or inappropriate administration may disrupt the delicate balance of neonatal microbiota and contribute to antibiotic resistance [@nih_prenatal_steroid_treatment].
+To address this, pharmaceutical innovations could focus on creating narrower-spectrum antibiotics that are highly effective against labor-associated infections while preserving beneficial bacteria [@narrow_spectrum_antibiotics].
+In addition, exploring probiotics or adjunct therapies to counteract microbiota disruption could further improve outcomes for newborns exposed to antibiotics [@cdhf_probiotics_infants].
+These actions align with the broader goal of a precision medicine approach, tailoring treatments to minimize risks while maximizing benefits for both mothers and infants [@nih_prenatal_steroid_treatment].
+
+Policymakers and healthcare providers must also take active roles in this regard.
+Enforcing evidence-based protocols for these treatments is crucial to ensure their judicious application [@acog_clinical_guidelines].
+Guidelines should outline clear indications for use, supported by comprehensive training for obstetric care providers to minimize variability in practice [@who_operational_guidance].
+By prioritizing the safe and effective administration of treatments [@who_operational_guidance], stakeholders can achieve better neonatal outcomes and address the risks associated with inappropriate interventions.
 
@@
 
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+The variability in treatment efficacy as observed in @fig-result2 and @fig-rfsumbar furthermore suggests the significant influence of contextual factors such as healthcare access, infrastructure, and socio-economic disparities [@who_contextual_factors].
+For instance, in low-resource settings, the availability of prenatal steroids may be limited, leading to suboptimal outcomes for preterm infants [@who_contextual_factors].
+Additionally, disparities in healthcare provider expertise and institutional protocols can amplify these variations, making it difficult to ensure consistent care across different regions or populations [@ajog_healthcare_disparities].
+
+Governments and healthcare organizations have an essential role to play in mitigating these disparities.
+For example, investments in prenatal care infrastructure, ensuring the availability of essential medications and training obstetric care providers, would be recommended [@who_contextual_factors].
+Mobile health initiatives and telemedicine could also play a transformative role in reaching underserved populations, ensuring that even remote areas benefit from timely interventions [@ajog_healthcare_disparities].
+
+Moreover, pharmaceutical companies can contribute by designing drugs that are accessible and effective in diverse healthcare settings [@ajog_healthcare_disparities].
+For instance, creating heat-stables formulations or medications with extended shelf lives can help ensure their utility in regions with limited resources [@ajog_healthcare_disparities].
+Collaboration with local healthcare systems to distribute these innovations is equally important [@who_contextual_factors].
+Addressing the structural and systematic factors that impact treatment efficacy can enhance the accessibility and consistency of labor-related medical interventions.
+
+## Limitations of the study and directions for future research {#sec-weakeness}
+
+In order to accurately contextualize the findings of this analysis, several limitations must be acknowledged.
+First, the observational nature of the study may introduce confounding variables that were not accounted for, potentially biasing the results [@biostatistics_confounding].
+Unlike randomized controlled trials (RCTs), observational studies lack random assignment, making it challenging to establish causality [@pmc5073487].
+Confounding factors, such as socioeconomic status, pre-existing health conditions, or access to healthcare, could influence both the treatment received and the neonatal outcomes, thereby skewing the correlations observed [@biostatistics_confounding].
+
+Second, the study's findings may have limited applicability to the broader maternal population if the sample is not representative [@tellingstories].
+Factors such as geographic location, healthcare facility type, and demographic characteristics can influence treatment practices and outcomes [@bmjmedicine_observational_study].
+For instance, a study conducted in a tertiary care center may not reflect the experiences of patients in rural or under-resourced settings [@bmjmedicine_observational_study].
+
+Third, inconsistencies in data collection methods, such as subjective interpretations of medical records or variability in documentation practices across facilities, can lead to measurement errors [@healthknowledge2024].
+These discrepancies may affect the accuracy of recorded treatments and outcomes, introducing bias and affecting the reliability of the conclusions drawn [@cambridge_observational_studies].
+Standardized data collection protocols might be essential to minimize such errors [@cambridge_observational_studies].
+
+Addressing these limitations, future research should aim to incorporate more diverse populations to enhance external validity [@scribbr_external_validity].
+Conducting randomized controlled trials, where feasible, would provide more definitive evidence regarding the efficacy of specific treatments during labor [@pmc5073487].
+Moreover, integrating pharmacokinetic and pharmacodynamic analyses, which helps complete a picture of how drugs are processed in the body, could offer deeper insights into how much of the drug reaches the baby and how quickly it is absorbed [@pmc5446801].
+Efforts by pharmaceutical companies and policymakers could play a pivotal role in advancing the research agenda by supporting studies that explore the safety and effectiveness of treatments for pregnant populations [@bmj_clinical_trials_pregnancy].
+These investments would not only address a critical area of unmet medical need but also uphold an ethical commitment to improving the health and well-being of mothers and their children [@springer_drug_treatments_pregnancy].
 

### 2024-11-26 paper/paper.qmd
appendix model details update

@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check, with the dark line representing the observed data distribution and the lighter lines indicating the posterior predictive distribution.
+This shows the notable deviation between `y` and `y_rep`, particularly the oscillations in the observed data.
+It suggests that the model may not fully capture the underlying structure of the observed data and proposes the need for model refinement, such as including additional predictors or modifying priors [@tellingstories].
+
+@fig-ppcheckandposteriorvsprior-2 compares the prior and posterior distributions of the model parameters.
+Significant shifts, such as for 'Steroids' and 'Chorioamnionitis' suggest these variables have strong impacts on the outcome, while minimal shifts for 'Antibiotics' indicate limited evidence of their effect.
+We can also note that the credible intervals underscore uncertainty.
 
@@
 
+# Create a named vector for the full variable names
+variable_names <- c(
+  "ster" = "Steroids",
+  "sigma" = "Residual Std.",
+  "indc" = "Induction of Labor",
+  "chor" = "Chorioamnionitis",
+  "augmt" = "Augmentation of Labor",
+  "antb" = "Antibiotics",
+  "anes" = "Anesthesia",
+  "(Intercept)" = "(Intercept)"
+)
+
+# Update the plot with the full variable names
 posterior_vs_prior(bayesian_model) +
   theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
+  scale_color_manual(
+    values = RColorBrewer::brewer.pal(length(variable_names), "Set1"),  # Use the same color palette
+    labels = variable_names  # Apply full names to the legend
+  ) +
   theme(legend.position = "bottom") +
-  coord_flip()
+  coord_flip() +
+  scale_x_discrete(labels = variable_names)  # Apply the full names to y-axis
+
 ```
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-stanareyouokay-1, which is a trace plot, shows stable and consistent fluctuations around a central value across iterations for all four chains.
+This suggests that the Bayesian model's Markov Chain Monte Carlo (MCMC) sampling has converged appropriately, ensuring that the posterior distributions are valid for interpretation.
+
+@fig-stanareyouokay-2, which is a Rhat plot, shows that all Rhat values are very close to 1 and none exceed the threshold of 1.05, indicating that the chains have converged well.
+This suggests that the Markov Chain Monte Carlo (MCMC) sampling has achieved stability and the posterior distributions can be reliably interpreted.
 
@@
 
-plot(bayesian_model, "trace")
+# Create a named vector for full variable names
+variable_names <- c(
+  "(Intercept)" = "(Intercept)",
+  "indc" = "Induction of Labor",
+  "augmt" = "Augmentation of Labor",
+  "ster" = "Steroids",
+  "antb" = "Antibiotics",
+  "chor" = "Chorioamnionitis",
+  "anes" = "Anesthesia",
+  "sigma" = "Residual Standard Deviation"
+)
+
+# Extract and relabel trace plots
+bayesplot::mcmc_trace(
+  as.array(bayesian_model),  # Convert model to array for trace plotting
+  pars = names(variable_names),  # Use variable names
+  facet_args = list(labeller = ggplot2::labeller(.rows = variable_names))  # Apply custom labels
+) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
 

### 2024-11-26 paper/paper.qmd
formatting graphs/tables

@@
   labs(x = "APGAR5 Score",
-       y = "Count of Observations") +
+       y = "Observation Count") +
 
@@
   labs(x = "APGAR5 Score",
-       y = "Count of Observations") +
+       y = "Observation Count") +
   # Format y-axis numbers with commas
@@
 ggplot(drug_count_summary, aes(x = drug_count, y = observation_count)) +
-  geom_bar(stat = "identity", fill = "darkgreen", color = "grey") +  # Bar plot
-  geom_text(aes(label = comma(observation_count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  geom_bar(stat = "identity", fill = "darkgreen",alpha = 0.5, color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(observation_count)), vjust = -0.5, size = 3, color = "black") +  # Add counts above bars
   labs(
@@
   theme(
-    axis.text = element_text(size = 12),
-    axis.title = element_text(size = 13)
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 10)
   )
@@
 # Plot the distribution of observations for each treatment
-ggplot(treatment_summary, aes(x = Treatment_Type, y = Count, fill = Treatment_Type)) +
-  geom_bar(stat = "identity", fill = "darkgreen", color = "grey") +  # Bar plot
-  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
-  labs(
-    x = "Treatment Type",
-    y = "Observation Count"
-  ) +
-  scale_y_continuous(labels = comma) +  # Format y-axis labels with commas
+ggplot(treatment_summary, aes(x = Count, y = Treatment_Type, fill = Treatment_Type)) +
+  geom_bar(stat = "identity", fill = "darkgreen", alpha = 0.5, color = "grey") +  # Bar plot
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 3, color = "black") +  # Add counts above bars
+  labs(
+    x = "Observation Count",
+    y = "Treatment Type"
+  ) +
+  scale_x_continuous(labels = comma) +  # Format x-axis labels with commas
   theme_minimal() +
@@
   geom_bar(stat = "identity", fill = "#a9dfbf", color = "grey") +  # Bar plot
-  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 4, color = "black") +  # Add counts above bars
+  geom_text(aes(label = comma(Count)), vjust = -0.5, size = 3, color = "black") +  # Add counts above bars
   labs(
@@
   theme(
-    axis.text = element_text(size = 11),
-    axis.title = element_text(size = 13)
+    axis.text = element_text(size = 10),
+    axis.title = element_text(size = 10)
   )
@@
 
-# Discussion
-
-## Pharmaceutical companies should focus on optimizing antenatal corticosteroid formulations {#sec-first-point}
+# Discussion {#sec-discussion}
+
+## Optimization of anetenatal corticosteroid formulations {#sec-first-point}
 
@@
 
-## Pharmaceutical companies, policymakers, and healthcare providers should collaborate to develop targeted antibiotics and enforce evidence-based protocols {#sec-second-point}
+## Targeted antibiotics and enforcement of evidence-based protocols {#sec-second-point}
 
@@
 
-## Third discussion point {#sec-third-point}
+## Disparities in treatment efficacy influenced by contextual factors {#sec-third-point}
 

### 2024-11-26 paper/paper.qmd
introduction, abstract updates

@@
 ---
-title: "How Drug Usage During Labor and Delivery Affects Infant Health: A Predictive Analysis of APGAR5 Scores"
-subtitle: "!!! MAIN RESULTS !!!"
+title: "How Drug Usage During Labor and Delivery Affects Infant Health: A Predictive Analysis of Apgar5 Scores"
+subtitle: "Steroid Efficacy and Antibiotic Risks in Labor-Associated Care"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "How maternal and neonatal drug interventions during labor shape the health status of a newborn remains a key question in perinatal care.
+With the 2023 Natality Data for the United States, we examined the relationship between six labor-related treatments and Apgar5 scores, a standard measure of neonatal health.
+Using a Random Forest model for prediction and a Bayesian Linear Model for inference, we identified usage of steroids and chorioamnionitis as the most impactful, while antibiotics dosage showed minimal influence.
+The analysis shows significant disparities in treatment efficacy, and further suggests the need for optimized corticosteroid formulations or targeted antibiotic protocols."
 format: pdf
@@
 
-# Introduction
-
-Overview paragraph
-
-Estimand paragraph
-
-Results paragraph
-
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+# Introduction {#sec-introduction}
+
+The health of newborns is strongly influenced by the treatments administered during labor and delivery [@who_operational_guidance].
+These interventions, including steroids, antibiotics, and anesthesia, are intended to address complications and improve neonatal outcomes, but their effectiveness varies widely [@anesthesiology_obstetric_anesthesia].
+This paper uses the 2023 Natality Data for the United States [@rawdata] to analyze the impact of six specific treatments on neonatal health outcomes, measured using the Apgar5 score.
+
+The estimand of interest is the average treatment effect (ATE) of six labor-associated interventions on neonatal health outcomes, as measured by Apgar5 scores.
+Specifically, we aim to quantify how the administration of each treatment changes the Apgar score compared to scenarios where the treatment is not applied [@tellingstories].
+By applying predictive modeling via Random Forest and inferential analysis through Bayesian linear models, we assess not only the magnitude of these effects[@geeksforgeeks_random_forest] but also the underlying uncertainties[@strimmerlab_priors].
+
+The analysis shows that prenatal exposure to steroids significantly improve Apgar scores (@tbl-rfsummary1), underscoring their role as a cornerstone of perinatal care.
+Conversely, antibiotics dosage doesn't show much contribution to the score distribution (@tbl-rfsummary1) and raises concerns about potential risks, including microbiota disruption [@narrow_spectrum_antibiotics].
+Moreover, the analysis proposes significant disparities in the consistent application and effectiveness of these treatments (@fig-rfsumbar).
+
+This study has its significance in that it states the need to improve health outcomes for mothers and their newborns globally.
+As recognized by the World Health Organization, a healthy start in life has significant repercussions for a person's health and well-being during infancy, childhood, and adulthood [@who_maternal_health].
+Ensuring quality care in maternal and newborn health is integral to the right to health, equity, and the preservation of dignity for women and their babies [@who_maternal_health].
+
+The remainder of this paper is structured as follows: @sec-data describes the dataset and methodology and @sec-model exhibits the use of predictive and inferential models.
+@sec-results presents the results of the analysis, detailing the observed relationships between treatments and neonatal health outcomes.
+@sec-discussion discusses the implications of these findings, focusing on healthcare policy and clinical practice improvements.
+Finally, @sec-weakeness concludes with a discussion of the study's limitations and potential areas for future research.
+@sec-appendix provides additional data details and model diagnostics.
 
@@
     vjust = -0.5,
-    size = 4,
+    size = 3,
     color = "black"
@@
     axis.text.x = element_text(hjust = 1, vjust = 1),
-    axis.text.y = element_text(size = 12)
+    axis.text.y = element_text(size = 10)
   )
@@
   theme(
-    axis.title.x = element_text(size = 12),
-    axis.title.y = element_text(size = 12),
-    axis.text.x = element_text(size = 10),
-    axis.text.y = element_text(size = 10)
+    axis.title.x = element_text(size = 10),
+    axis.title.y = element_text(size = 10),
+    axis.text.x = element_text(size = 9),
+    axis.text.y = element_text(size = 9)
   )
@@
 
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
+## Additional data details
+
+### Survey design and sampling techniques
+
+The 2023 Natality Dataset provides detailed records of maternal and neonatal health outcomes across diverse populations in the United States.
+This obervational dataset aggregates data from hospital records, birth certificates, and surveys completed by healthcare providers.
+Its comprehensive scope includes variables like maternal age, gestational age, socio-economic status, and the application of specific labor-related treatments.
+
+--- The dataset represents a census of live births in the U.S., negating the need for probabilistic sampling.
+However, for this study, we employed a stratified sampling approach to create a balanced subset.
+Treatments like steroids and antibiotics often show skewed distributions, with underrepresentation in certain socio-economic or geographic groups.
+Stratified sampling ensures adequate representation of minority groups, enabling a more equitable evaluation of treatment effects.
+Simulations using bootstrapped resampling were conducted to validate that the subset retained key characteristics of the full dataset.---
+
+### Observational data considerations
+
+!!![NEED TO WORK MORE] !!!
+
+## Model details {#sec-model-details}
+
+### Posterior predictive check
 
@@
   "anes" = "Anesthesia",
-  "sigma" = "Residual Standard Deviation"
+  "sigma" = "Residual Std."
 )

### 2024-11-26 paper/paper.qmd
update paper first draft

@@
 
-The 2023 Natality Dataset provides detailed records of maternal and neonatal health outcomes across diverse populations in the United States.
-This obervational dataset aggregates data from hospital records, birth certificates, and surveys completed by healthcare providers.
-Its comprehensive scope includes variables like maternal age, gestational age, socio-economic status, and the application of specific labor-related treatments.
-
---- The dataset represents a census of live births in the U.S., negating the need for probabilistic sampling.
-However, for this study, we employed a stratified sampling approach to create a balanced subset.
-Treatments like steroids and antibiotics often show skewed distributions, with underrepresentation in certain socio-economic or geographic groups.
-Stratified sampling ensures adequate representation of minority groups, enabling a more equitable evaluation of treatment effects.
-Simulations using bootstrapped resampling were conducted to validate that the subset retained key characteristics of the full dataset.---
+The 2023 Natality Dataset [@rawdata] provides detailed records of maternal and neonatal health outcomes across diverse populations in the United States.
+This obervational dataset aggregates data from hospital records, birth certificates, and surveys completed by healthcare providers [@natality2023].
+Its comprehensive scope includes variables like maternal age, gestational age, socio-economic status, and the application of specific labor-related treatments [@cdc_facility_worksheet_2016].
+
+The natality data involves addressing potential biases inherent in observational datasets.
+For example, selection bias may arise if certain populations, such as undocumented individuals or those from rural areas, are underrepresented in the data [@tellingstories].
+Additionally, measurement errors, such as inconsistencies in self-reported variables like smoking status or prenatal care visits, can affect data reliability [@healthknowledge2024].
+To mitigate these challenges, researchers employ weighting techniques to align samples with population distributions and utilize statistical models like propensity score matching to adjust for confounding factors [@Stuart2021].
+
+Integrating survey methodologies with the 2023 US Natality Data enables researchers to enrich findings by combining administrative records with additional information collected through structured questionnaires.
+This hybrid approach allows for the exploration of causal relationships, such as the effects of prenatal care access on birth outcomes.
+By adhering to rigorous survey design principles, such as pilot testing and the Total Survey Error framework, and employing robust sampling strategies, researchers can ensure their findings are both internally valid and generalizable.
+These methods ultimately enhance the utility of natality data for informing public health policies and interventions aimed at improving maternal and child health outcomes.
 
@@
 
-!!![NEED TO WORK MORE] !!!
+Observational data serve as the main database in biomedical research, particularly when randomized controlled trials (RCTs) are impractical due to ethical or logistical constraints [@healthknowledge2024].
+These datasets, often derived from electronic health records, patient surveys, or longitudinal studies [@healthknowledge2024], provide valuable insights into real-world healthcare outcomes.
+However, their inherent limitations necessitate careful consideration to ensure robust and credible findings.
+Among the key challenges are confounding variables, selection bias, and measurement errors [@tellingstories].
+For instance, in analyzing the effects of prenatal steroids on neonatal outcomes, unmeasured factors such as maternal health or socioeconomic status can introduce confounding [@natality202].
+Advanced statistical techniques like propensity score matching or instrumental variables are essential to isolate causal effects from observational data [@Olier2023ObservationalData].
+
+Another critical issue is selection bias, where certain subgroups (e.g., uninsured individuals or marginalized populations) may be systematically underrepresented in the dataset, thus limiting the generalizability of findings [@tellingstories].
+Addressing this requires techniques such as weighting adjustments or sensitivity analyses and, where possible, integration with survey data to fill informational gaps.
+Additionally, measurement errors and missing data—common in healthcare records—can distort results [@healthknowledge2024].
+Strategies such as multiple imputation and rigorous data cleaning are indispensable for mitigating these challenges [@healthknowledge2024].
+Coupled with robust ethical safeguards, including informed consent and adherence to privacy standards, these measures ensure that observational studies maintain both scientific and ethical rigor [@healthknowledge2024].
+
+Causal inference remains a central challenge in using observational data, as the same individual cannot simultaneously experience treatment and non-treatment conditions [@tellingstories].
+To approximate counterfactuals, researchers often rely on the Neyman-Rubin potential outcomes framework or marginal structural models [@Groves2010].
+Complementing this with simulation studies can test the robustness of findings under varying assumptions.
+By adhering to established frameworks like the Total Survey Error Model and integrating observational datasets with structured surveys, researchers can produce more reliable and actionable insights [@CausalConversations2024].
+This approach not only strengthens the internal and external validity [@tellingstories] of the results but also highlights the potential of observational data to inform equitable and effective healthcare interventions.
 

### 2024-11-28 other/datasheet/datasheet_natality2023.qmd
upload datasheet

---
title: "Datasheet for '2023 Natality Data for the United States'"
subtitle: "Analysis of US Birth Records for Public Health Insights"
author: 
  - Yunkyung Ko
thanks: "Code and data are available at: [https://github.com/koyunkyung/infant_health](https://github.com/koyunkyung/infant_health)."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "The 2023 Natality Dataset provides birth record data for the United States, covering demographic, medical, and geographic variables. The dataset enables research on public health, maternal outcomes, and disparities in birth trends. The datasheet documents the creation, composition, and potential uses of the dataset while addressing its limitations and ethical considerations."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @cdc_facility_worksheet_2016.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*

The dataset was created to serve as a standardized source for analyzing trends in maternal and infant health across the United States [@natality2023]. Its primary purposes include:
  - Public health researching and reporting
  By capturing demographic, medical, and geographic variables, it supports studies on key public health indicators. The target of studies include rates of preterm births and low birth weight, and disparities in maternal health outcomes by race, ethnicity, or socioeconomic status [@cdc2023births].
  - Policy development and evaluation
  Policymakers make use of this dataset to evaluate the effectiveness of health interventions and programs, such as those under the health people of age 20 to 30 [@cdc2023evaluation]. It identifies geographic regions or sub-populations that require targeted healthcare intiatives or funding, and assess trends over time to guide future health policies and resource allocation [@cdc2023evaluation].
  - Addressing data gaps in maternal and infant health
  The dataset solves for the inconsistencies in reporting across states by providing a nationwide coverage of natality data [@bigdataframework2023]. It standardizes information collected from birth certificates, ensuring compatibility for cross-state comparisons and longitudinal studies [@cdc2023nvss].

2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*

The dataset was created by the National Center for Health Statistics (NCHS), a division of the Centers for Disease Control and Prevention (CDC) [@natality2023]. The NCHS, which is responsible for collecting, standardizing, and disseminating natality data, supports statistical reporting at both national and state levels [@cdc2023nvss]. This effort was part of the Vital Statistics Cooperative Program, which monitors key health trends and provides data for annual reports on births in the United States [@cdc2023wonder]. It furthermore supports a broad range of research tasks including the understanding of long-term health implications of birth conditions by linking with other datasets [@cdc2023wonder].

3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*

The creation of the 2023 Natality Dataset was funded by the U.S. Department of Health and Human Services (HHS) through the Vital Statistics Cooperative Program (VSCP) [@cdc2023vscp]. This program is administered by the Centers for Disease Control and Prevention (CDC) and financially supports state and local offices to collect, process, and transmit natality data to the National Center for Health Statistics (NCHS) [@cdc2023vscp]. This funding aims to ensure the nationwide consistency and quality of vital statistics reporting [@cdc2023nvss].

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*

The instances in the dataset represent individual live birth records from the United States, as documented on standardized birth certificates [@natality2023]. Each record contains detailed information about the birth, including demographic data (e.g., maternal age, race, education), medical details (e.g., prenatal care, delivery method, complications), and geographic identifiers (e.g., state and country of birth) [@natality2023]. These records are uniform in structure and do not include multiple types of instances, ensuring consistency across all data points [@cdc2023birthguidelines].

2. *How many instances are there in total (of each type, if appropriate)?*

The dataset contains a total of 3,605,081 instances, each representing a single live birth recorded in the United States during the year 2023 [@natality2023]. These instances include both U.S. resident births (3,596,017) and a smaller number of births to non-residents (9,064), showing a wide coverage of all registered live births nationwide [@natality2023].

3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*

The 2023 Natality Dataset is a census of all registered live births in the U.S, ensuring relatively complete nationwide coverage with low sampling error [@natality2023]. Data is collected uniformly through mandatory birth certificates and validated by the Vital Statistics Cooperative Program [@cdc2023birthguidelines]. While rare unregistered births may introduce minimal undercoverage, systematic quality checks ensure data consistency and relatively accurate representation across geographic, demographic, and medical dimensions [@cdc2023birthguidelines].

4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*

Each instance in the dataset consists of structured data derived from standardized birth certificates [@natality2023]. The data includes both raw fields (e.g., mother's age, infant's birth weight) and processed features (e.g., gestational age in weeks, calculated Apgar scores) [@natality2023]. These fields capture demographic, medical, and geographic information, such as:
  - *Demographics: 
    maternal age, race, education, marital status, residency
  - *Medical details:
    birth weight, gestational age, prenatal care visits, complications during pregnancy or delivery, method of delivery (e.g., cesarean)
  - *Geographic information:
    state and country of birth

5. *Is there a label or target associated with each instance? If so, please provide a description.*

With no explicit label or target variable, several fields can serve as target variables for analysis or modeling, depending on the research question. For instance, outcomes such as `low birth weight (<2500 grams)`, `preterm birth (<37 weeks gestation)`, or `Apgar score below 7` can be used as target variables in studies predicting adverse birth outcomes [@natality2023].

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*

Commonly missing fields in this dataset include details about the father (e.g., age, education) or specific prenatal care data [@rawdata]. It may often be due to incomplete reporting by healthcare providers or individuals [@bmjmedicine_observational_study]. Moreover, certain medical details, such as complications during pregnancy, may be underreported becuase of inconsistencies in documentation practices across states or hospitals [@bmjmedicine_observational_study]. These gaps typically arise from data collection challenges rather than intentional omissions and can be addressed through data validation and impuation methods where feasible [@pmc5073487].

7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- TBD
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-11-29 scripts/06-model_api.R
upload scripts for model api, shiny app

@@ -1,0 +1,80 @@
+#### Preamble ####
+# Purpose: API for serving predictions from Random Forest and Bayesian Linear Models for health status of infants
+# Author: Yunkyung Ko
+# Date: 28 November 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+# - The `plumber` package must be installed and loaded
+# - 05-model_data.R must have been run
+# Any other information needed? Make sure you are in the `infant_health` rproj
+
+# Load necessary library
+library(plumber)
+
+# Load models from RDS files
+random_forest_model <- readRDS("../models/random_forest.rds")
+bayesian_model <- readRDS("../models/bayesian_lm.rds")
+
+#* @apiTitle Infant Health Prediction API
+#* @apiDescription API for predicting infant health status based on drugs used during labor and delivery
+
+#* Predict using a numeric vector of features (JSON input)
+#* @param features A numeric vector of features (e.g., maternal age, drug usage indicators)
+#* @post /predict
+function(features) {
+  # Validate input
+  if (is.null(features) || !is.numeric(unlist(features))) {
+    stop("Invalid input: 'features' must be a numeric vector")
+  }
+  
+  # Convert input to numeric vector
+  features <- as.numeric(unlist(features))
+  
+  # Ensure features are in the correct format for prediction
+  features_df <- as.data.frame(t(features))  # Convert to data frame for compatibility
+  
+  # Predict using Random Forest model
+  rf_prediction <- predict(random_forest_model, newdata = features_df, type = "response")
+  
+  # Predict using Bayesian Linear model
+  bayesian_prediction <- predict(bayesian_model, newdata = features_df, type = "response")
+  
+  # Return predictions as a list
+  list(
+    random_forest_prediction = rf_prediction,
+    bayesian_linear_model_prediction = bayesian_prediction
+  )
+}
+
+#* Predict using a local CSV file path
+#* @post /predict_csv_local
+function() {
+  # Specify the correct relative path to the test_data.csv
+  csv_path <- "../data/02-analysis_data/train_test_data/test_data.csv"
+  
+  # Load the CSV file
+  features_df <- read.csv(csv_path)
+  
+  # Validate that all columns in the CSV are numeric
+  if (!all(sapply(features_df, is.numeric))) {
+    stop("Invalid input: All columns in the CSV file must be numeric")
+  }
+  
+  # Predict using Random Forest model
[…]

### 2024-11-29 scripts/07-shiny_app.R

@@ -1,0 +1,110 @@
+#### Preamble ####
+# Purpose: Shiny web application for visualizing and analyzing predictions from Random Forest and Bayesian Linear Models for infant health status
+# Author: Yunkyung Ko
+# Date: 29 November 2024
+# Contact: yunkyung.ko@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+# - The `shint`, `DT` package must be installed and loaded
+# - 05-model_data.R must have been run
+# - The user must provide a dataset with numeric features (e.g., `test_data.csv`)
+# Any other information needed? Make sure you are in the `infant_health` rproj
+
+
+library(shiny)
+library(randomForest)  # For prediction
+library(readr)         # For reading CSV files
+library(DT)            # For interactive data tables
+library(ggplot2)       # For visualization
+
+# Load trained models
+random_forest_model <- readRDS("models/random_forest.rds")
+bayesian_model <- readRDS("models/bayesian_lm.rds")
+
+# Define UI
+ui <- fluidPage(
+  titlePanel("Infant Health Prediction App"),
+  
+  sidebarLayout(
+    sidebarPanel(
+      fileInput("file", "Upload Infant Data (CSV)", accept = c(".csv")),
+      actionButton("predict", "Run Predictions"),
+      hr(),
+      h4("Options:"),
+      sliderInput("bins", "Number of Bins for Histogram:", min = 5, max = 50, value = 20)
+    ),
+    
+    mainPanel(
+      tabsetPanel(
+        tabPanel("Data Preview", DTOutput("data_table")),
+        tabPanel("Prediction Results", DTOutput("predictions_table")),
+        tabPanel("Visualizations",
+                 plotOutput("rf_hist"),
+                 plotOutput("bayesian_hist"))
+      )
+    )
+  )
+)
+
+# Define server logic
+server <- function(input, output, session) {
+  # Reactive value to store uploaded data
+  reactive_data <- reactiveVal(NULL)
+  
+  # Load and display uploaded dataset
+  observeEvent(input$file, {
+    req(input$file)
+    data <- read_csv(input$file$datapath)
+    reactive_data(data)
+  })
+  
+  output$data_table <- renderDT({
+    req(reactive_data())
+    datatable(reactive_data(), options = list(pageLength = 5), caption = "Uploaded Dataset Preview")
+  })
+  
+  # Run predictions when button is clicked
+  predictions <- reactive({
[…]

### 2024-11-29 paper/paper.qmd
'enhancement' parts upload

@@
 #| label: fig-apgar1
-#| fig-cap: "The distribution of APGAR5 scores across the entire observation in the original dataset **Note:** Unknown or unreported observations were excluded so that the distribution could be clearly visualized"
+#| fig-cap: "The distribution of Apgar5 scores across the entire observation in the original dataset (**Note:** Unknown or unreported observations were excluded so that the distribution could be clearly visualized)"
 #| echo: false
@@
 #| label: fig-apgar2
-#| fig-cap: The distribution of APGAR5 scores across the filtered observation in analysis dataset, which selected the relevant variables and removed the NA values
+#| fig-cap: "The distribution of Apgar5 scores across the filtered observation in analysis dataset, which selected the relevant variables and removed the NA values"
 #| echo: false
@@
 #| label: fig-apgar3
-#| fig-cap: The distribution of APGAR5 scores across the filtered observation in analysis dataset
+#| fig-cap: "The distribution of Apgar5 scores across the filtered observation in analysis dataset"
 #| echo: false
@@
 #| label: fig-numtreat
-#| fig-cap: The distribution of observations across different number of treatments used during delivery and labor
+#| fig-cap: "The distribution of observations across different number of treatments used during delivery and labor"
 #| echo: false
@@
 #| message: false
+#| fig-width: 5
+#| fig-height: 4
 
@@
 
+\newpage
+
 ### Augmentation of labor
@@
 
+\newpage
+
 ### Steroids (glucocorticoids) for fetal lung maturation received by the mother before delivery
@@
 
+\newpage
+
 ### Antibiotics received by the mother during delivery
@@
 
+\newpage
+
 ### Clinical chorioamnionitis diagnosed during labor or maternal temperature over 38 degrees celcius (100.4 degrees fahrenheit)
@@
 
+\newpage
+
 ### Epidural or spinal anesthesia during labor
@@
 
+\newpage
+
 ## Correlation between predictor variables
@@
 
+\newpage
+
 @fig-indcscore shows that observation counts increase steadily from scores 0 to 5, remain relatively stable between scores 5 and 9, and then drop sharply at score 10.
@@
 
+\newpage
+
 @fig-augmtscore shows that observation counts are lower at extreme scores of 0 and 10, steadily increase from scores 1 to 5, plateau between 5 and 8, and peak at score 9 before sharply declining at score 10.
@@
 
+\newpage
+
 @fig-sterscore shows that observation counts are relatively low at extreme scores of 0 and 10, increase steadily from scores 0 to 4, and peak at score 4 before gradually declining through scores 5 to 10.
@@
 
+\newpage
+
 @fig-antbscore shows that observation counts rise gradually from scores 0 to 2, and maintain a steady plateau between scores 4 and 7, before declining through scores 8 to 10.
@@
 
+\newpage
+
 @fig-chorscore shows that the observation count peaks sharply at a score of 1, then stabilizes between scores 2 and 5, followed by a gradual decline from scores 6 through 10.
@@
 
+\newpage
+
 @fig-anesscore shows that observation counts steadily increase from scores 0 to 5.
@@
 
+\newpage
+
 @tbl-rfsummary2 shows the Random Forest model, despite incorporating hyperparameter tuning, demonstrated suboptimal performance.
@@
 Among the key challenges are confounding variables, selection bias, and measurement errors [@tellingstories].
-For instance, in analyzing the effects of prenatal steroids on neonatal outcomes, unmeasured factors such as maternal health or socioeconomic status can introduce confounding [@natality202].
+For instance, in analyzing the effects of prenatal steroids on neonatal outcomes, unmeasured factors such as maternal health or socioeconomic status can introduce confounding [@natality2023].
 Advanced statistical techniques like propensity score matching or instrumental variables are essential to isolate causal effects from observational data [@Olier2023ObservationalData].
@@
 
+### Model API Predictions
+
+#### API execution details
+
+The 'Infant Health Prediction API' provides predictions for infant health status based on maternal and delivery-related factors.
+By accepting structured datasets in CSV format, it offers researchers a powerful tool to generate real-time predictions using advanced maching learning models.
+
+- Endpoint URL: `http://127.0.0.1:3505/predict_csv_local`
+
+- Description: This endpoint processes a local CSV file containing test data and provides predictions from both the Random Forest and Bayesian Linear models.
+
+  - Random Forest Model:
+
+    Handles complex interactions between variables to generate robust predictions.
+
+  - Bayesian Linear Model:
+
+    Offers interpretable results with uncertainty quantification.
+
+#### Input requirements
+
+- The input file must be a CSV containing only numeric columns.
+
+- Required features include drug usage indicators and delivery-related factors.
+
+- Missing or non-numeric data will result in validation errors.
+
+#### Output requirements
+
+The API returns predictions in JSON format. Each entry corresponds to a row in the input dataset, with predictions for models:
+
+    "Random_Forest_Predictions": 5.0648,
+
+    "Bayesian_Linear_Model_Predictions": 5.1659
+
+The CURL command used to trigger the API:
+
+```bash
+
+curl -X POST 'http://127.0.0.1:3505/predict_csv_local' -H 'accept: */*' -d ''
+
+```
+
+### Shiny web applications for visualizing and analyzing predictions
+
+The 'Infant Health Prediction App' is a Shiny web application that enables users to:
+- Upload a dataset containing infant-related data in CSV format.
+- Run predictions for infant health status using both a Random Forest model and a Bayesian Linear model.
+- View prediction results in an interactive table.
+- Explore visualizations of the predictions through histograms.
+
+#### Features
+1. Data Upload and Preview: Users can upload datasets and preview them in an interactive table.
+2. Model Predictions: Predictions from both models are displayed in a table for easy comparison.
+3. Visualizations: Histograms provide insights into the distribution of predictions from each model.
+
+#### How to Run the App
+To launch the app locally, use the following R command:
+
+```r
+
+library(shiny)
+
+runApp("../scripts/07-shiny_app.R")
+
+```
+
+\newpage
+
+#### App Interface
+Below is the interface of the Infant Health Prediction App:
+
+![App Interface](../models/results/interface_screenshot.png)
+
+\newpage
+
+#### Prediction Table
+The predictions are displayed in an interactive table:
+
+![Prediction Table from the Shiny Web Application](../models/results/prediction_table_screenshot.png)
+
+\newpage
+
+#### Visualizations
+The histograms show the distribution of model predictions:
+
+![Random Forest Histogram from the Shiny Web Application](../models/results/random_forest_histogram.png)
+
+![Bayesian Linear Model Histogram from the Shiny Web Application](../models/results/bayesian_linear_histogram.png)
+
 \newpage

### 2024-11-29 other/datasheet/datasheet_natality2023.qmd
update datasheet

@@
 thanks: "Code and data are available at: [https://github.com/koyunkyung/infant_health](https://github.com/koyunkyung/infant_health)."
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
+date: today
+date-format: long
 abstract: "The 2023 Natality Dataset provides birth record data for the United States, covering demographic, medical, and geographic variables.
@@
 
-Extract of the questions from @cdc_facility_worksheet_2016.
+Extract of the questions from @gebru2021datasheetsdatasets.
 
@@
 The dataset was created to serve as a standardized source for analyzing trends in maternal and infant health across the United States [@natality2023]. Its primary purposes include:
-  - Public health researching and reporting
+
+  - Public health researching and reporting:
+
   By capturing demographic, medical, and geographic variables, it supports studies on key public health indicators.
 The target of studies include rates of preterm births and low birth weight, and disparities in maternal health outcomes by race, ethnicity, or socioeconomic status [@cdc2023births].
-  - Policy development and evaluation
+
+  - Policy development and evaluation:
+
   Policymakers make use of this dataset to evaluate the effectiveness of health interventions and programs, such as those under the health people of age 20 to 30 [@cdc2023evaluation].
 It identifies geographic regions or sub-populations that require targeted healthcare intiatives or funding, and assess trends over time to guide future health policies and resource allocation [@cdc2023evaluation].
-  - Addressing data gaps in maternal and infant health
+
+  - Addressing data gaps in maternal and infant health:
+
   The dataset solves for the inconsistencies in reporting across states by providing a nationwide coverage of natality data [@bigdataframework2023].
@@
 These fields capture demographic, medical, and geographic information, such as:
-  - *Demographics: 
+
+  - Demographics:
+
     maternal age, race, education, marital status, residency
-  - *Medical details:
+
+  - Medical details:
+
     birth weight, gestational age, prenatal care visits, complications during pregnancy or delivery, method of delivery (e.g., cesarean)
-  - *Geographic information:
+
+  - Geographic information:
+
     state and country of birth
@@
 7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
-	- TBD
+
+Relationships between individual instances are not explicitly defined in this dataset and each record represents a standalone live birth [@cdc2023birthguidelines].
+Every individual instance is not linked to other births, even in cases of familial connections (e.g., siblings born in different years) [@cdc2023birthguidelines].
+However, relationships can be inferred indirectly through shared geographic or demographic attributes, such as births occurring within the same country or among mothers of similar age and race [@natality2023].
+These inferred connections are not explicitly encoded, but can be analyzed statistically for patterns or trends.
+
 8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
-	- TBD
+
+The Centers for Disease Control and Prevention (CDC) does not prescribe specific data splits for this dataset [@cdc2023birthguidelines], but researchers are encouraged to determine appropriate data partitioning strategies.
+For instance, they can divide data into training, validation, and testing sets based on their specific analytical objectives and methodological requirements [@sharpsight2023].
+This approach may ensure the data to be utilized effectively in addressing diverse research questions.
+
 9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
-	- TBD
+
+The 2023 Natality Dataset contains relatively high-quality data due to its collection and standardization by the National Center for Health Statistics (NCHS).
+The NCHS employs rigorous data cleaning protocols including checks for internal consistency, cross-referencing with prior datasets for anomalies, and imputing missing values where appropriate [@cdc2023natality].
+However, there are potential errors, noise, and redundancies that arise from the complexities of data collection across states and facilities:
+
+  - Incomplete Data:
+
+    Certain fields, such as prenatal care visits, father’s information (e.g., age, education), and specific medical details, may be missing [@rawdata].
+These gaps often result from incomplete reporting by healthcare providers or omitted information from parents [@cdc2023vscp].
+
+  - Reporting Variability:
+
+    Differences in how states and hospitals document medical complications, delivery methods, or maternal health conditions can lead to inconsistencies [@pmc5073487].
+For example, complications during labor may be underreported in some jurisdictions due to differences in medical coding practices or provider discretion [@pmc5073487].
+
+  - Data Noise:
+
+    Some data points may include noise due to typographical errors, such as invalid maternal ages (e.g., extreme outliers) or implausible gestational ages [@bmjmedicine_observational_study].
+These errors, though infrequent, can affect statistical analyses if not addressed.
+
+  - Geographical and Temporal Variations:
+
+    Geographic disparities in healthcare access and reporting capabilities can lead to systematic biases, such as underrepresentation of certain rural areas or delays in data submission [@cdc2023vscp].
+
 10.
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- TBD
+
+The '2023 Natality Data for the United States' is largely self-contained, as it contains extensive birth records collected through standardized birth certificates across all United States and territories [@natality2023].
+The dataset does not rely on external resources like websites, tweets, or other datasets for its core content.
+However, researchers can enhance analyses by linking this dataset to other publicly available resources, such as socioeconomic data from the American Community Survey(ACS) or environmental data from the EPA [@cdc2023wonder].
+
+For access to more specific geographic data, such as specific countries or smaller regions, researchers may need to submit a formal data request to the National Center for Health Statistics (NCHS) [@cdc2023wonder].
+These requests are subject to review and approval to ensure compliance with privacy and ethical standards [@cdc2023vscp].
+Such restrictions are in place to prevent the misuse of sensitive data and to protect individual privacy [@cdc2023vscp].
+
+  **a) Guarantees of Continuity**
+
+  The datsaet itself is archived and maintained by the National Center for Health Statistics (NCHS), ensuring its availability over time [@cdc2023vscp].
+External resources for linkage, like ACS data, are also federally maintained, providing similar guarantees of continuity [@census2023acs].
+
+  **b) Archival Versions**
+
+  The NCHS maintains archival versions of the natality dataset, preserving the data as it existed at the time of release [@rawdata]. These archives include documentation to support reproducibilty.
+
+  **c) Restrictions**
+
+  While the dataset is free of charge and unrestricted for general use, requests for detailed geographic data may require additional permissions.
+Users must also follow ethical guidelines when linking with external datasets to prevent privacy violations or misuse [@cdc2023vscp].
+
 11.
@@
 If so, please provide a description.*
-	- TBD
+
+The dataset does not contain data that is protected by legal privilege or doctor-patient confidentiality, as it is publicly available and anaonymized to safeguard individual privacy [@rawdata].
+Personally identifiable information (PII), such as names, Social Security numbers, or exact residential addresses, is excluded to comply with federal privacy laws and regulations, including the Health Insurance Portability and Accountability Act (HIPAA) [@natality2023].
+
+However, the dataset includes sensitive demographic and health-related information, such as maternal age, race, education, and detailed medical data (e.g., complications during pregnancy, prenatal care visits).
+While this information is anonymized, it is still considered sensitive due to its potential to identify individuals when combined with external datasets.
+To mitigate risks, the National Center for Health Statistics (NCHS) applies strict data handling policies, and access to specific geographic data, such as detailed county-level information, requires additional permissions and adherence to ethical guidelines [@natality2023].
+These measures ensure that the dataset is suitable for public health research while minimizing risks associated with confidentiality breaches.
+
 12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
-	- TBD
+
+The dataset does not contain data that is inherently offensive, insulting, or threatening.
+However, certain fields may include sensitive demographic or medical information, such as maternal age, race, ethnicity, and health outcomes (e.g., complications during pregnancy, infant mortality).
+When misinterpreted or presented without proper context, this data could inadvertently reinforce stereotypes or cause anxiety, particularly in discussions around health disparities or adverse outcomes [@cdc2023vscp].
+
+For instance, data highlighting disparities in preterm birth rates among different racial or ethnic groups could be misused or misrepresented in ways that perpetuate stigma or bias [@bmjmedicine_observational_study].
+To address this, researchers and policymakers using the dataset are encouraged to approach analyses with sensitivity, ensuring that findings are contextualized and presented in a manner that respects the dignity and privacy of all individuals represented [@natality2023].
+
 13.
@@
 If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
-	- TBD
+
+Sub-populations identified by the dataset based on demographic and medical attributes [@natality2023] include:
+
+  - Age:
+
+  The maternal age variable allows for sub-populations such as adolescent mothers (<20 years old), mothers of advanced maternal age (35+ years), and age-specific birth trends.
+
+  - Race and Ethnicity:
+
+  Maternal race and ethnicity are recorded, enabling analyses of sub-populations such as non-Hispanic, White, Black, Hispanic, and Asian mothers.
+This enables studies on health disparities, such as differences in preterm birth or low birth weight rates among racial and ethnic groups.
+
+  - Geographic Location:
+
+  Births are categorized by state and country, allowing for sub-populations based on urban or rural settings and regional disparities in healthcare access or outcomes.
+
+  - Medical Factors:
+
+  The dataset identifies sub-populations based on medical characteristics, such as mothers with complications during pregnancy, preterm births, or low birth weight infants.
+
 14.
@@
 If so, please describe how.*
-	- TBD
+
+It is not possible to directly identify individuals from the dataset, as it is anonymized and excludes personally identifiable information (PII) such as names, addresses, and Social Security numbers [@natality2023].
+However, indirect identification could be theoretically possible if the dataset is combined with external information.
+For instance, a record with unique combinations of attributes—such as maternal age, race, specific birth complications, and detailed geographic location—could potentially be matched to external data sources like local news reports or publicly available registries [@cdc2023vscp].
+To mitigate such risks, the dataset limits access to granular geographic details (e.g., county-level data) and enforces strict ethical guidelines for use  [@cdc2023vscp].
+These measures significantly reduce the likelihood of re-identification while preserving the dataset’s utility for research.
+
 15.
@@
 If so, please provide a description.*
-	- TBD
-16. *Any other comments?*
-	- TBD
+
+The dataset contains sensitive data, primarily related to health and demographics. Sensitive fields include [@cdc2023vscp]:
+
+  - Race and Ethnic Origins:
+
+    Records of maternal race and ethnicity are included in the dataset.
+While this information is critical for analyzing disparities in birth outcomes, these fields must be interpreted carefully to avoid reinforcing stereotypes, as disparities often reflect systemic inequalities rather than intrinsic differences.
+
+  - Health Data:
+
+    - Maternal Health: Data on complications during pregnancy, delivery methods (e.g., cesarean sections), prenatal care visits
+
+    - Infant Outcomes: Metrics like Apgar scores, birth weight, gestational age, whether the birth was preterm
+
+    These fields are sensitive as they relate to medical conditions that could reveal vulnerabilities or personal health histories if misused.
+
+  - Geographic Location:
+
+    Even though the dataset provides state-level and limited geographic information, such as country or metropolitan area, access to detailed geographic data is possible after special permission.
+Geographic information, when combined with other datasets, could lead to indirect identification or stigmatization of specific communities with poor health outcomes.
+
+In order to minimize the risk of misuse or harm to individuals and communities, researchers must handle these fields responsible, contextualizing analyses to avoid reinforcing biases or stigmatizing groups [@cdc2023vscp].
+Access to more specific geographic details is restricted to approved uses, ensuring compliance with federal privacy regulations, including HIPAA and ethical guidelines from the National Center for Health Statistics (NCHS) [@cdc2023evaluation].
 
@@
 If so, please describe how.*
-	- TBD
+
+The data associated with each instance in the dataset was acquired from standardized birth certificates across the United States.
+These certificates are filled out at the time of birth by healthcare providers, including physicians, midwives, and hospital staff, based on direct observations during the delivery process and information reported by the mother or other reliable sources (e.g., medical records) [@cdc2023vscp].
+
+- Directly Observable Data:
+
+Certain data points, such as the infant's weight at birth, gestational age, and Apgar scores, are measured or observed by healthcare professionals.
+
+- Reported by Subjects:
+
+Information such as the mother's age, education level, race/ethnicity, and prenatal care visits is often provided by the mother or documented from her medical history.
+
+- Indirectly Inferred Data:
+
+Some fields, such as gestational age or complications during delivery, may be derived from medical evaluations, clinical records, or healthcare provider notes.
+
+The following methods were used to ensure data quality and consistency [@cdc2023birthguidelines]:
+
+- Standardization:
+
+The dataset adheres to the National Center for Health Statistics (NCHS) guidelines, ensuring that data collection and categorization are uniform across all states.
+
+- Validation:
+
+Healthcare professionals validate critical fields such as birth weight and gestational age by cross-referencing clinical records.
+Data entry undergoes quality checks by state vital statistics offices before submission to the NCHS.
+
+- Verification:
+
+Regular training for data collectors and feedback loops between hospitals and state agencies help minimize errors and improve reliability.
+
 2.
@@
 How were these mechanisms or procedures validated?*
-	- TBD
+
+The data in the dataset was collected through a combination of manual and automated mechanisms, standardized by the National Vital Statistics System (NVSS) under the National Center for Health Statistics (NCHS).
+The key mechanisms and procedures include [@cdc2023nvss]:
+
+  1. Manual Data Entry:
+
+Birth certificates are completed by healthcare providers (e,g,, physicians, nurses, or midwives) and administrative staff in hospitals or birthing centers.
+These individuals manually document details such as maternal demographics, birth outcomes, and medical procedures based on observations, patient interviews, and clinical records.
+
+  2. Standardized Electronic Systems:
+
+Many states utilize Electronic Birth Registration Systems (EBRS) to digitize the process [@ebrs2024].
+These systems allow hospital staff to enter birth certificate data electronically, reducing manual errors [@ebrs2024].
+The EBRS software enforces validation rules, such as checking for logical consistency (e.g., gestational age aligning with birth weight) and completeness (e.g., ensuring no mandatory fields are left blank) [@ebrs_dashboard2024].
+
+  3. Data Transmission:
+
+Once entered, data is transmitted from state vital statistics offices to the NCHS through secure channels [@cdc2023nvss].
+This ensures uniform data collection across all states and territories.
+The NCHS applies additional validation algorithms to identify and flag inconsistencies, such as outliers in birth weight or maternal age [@cdc2023vscp].
+
+  4. Validation Procedures:
+
+Training: Healthcare providers and staff responsible for data collection are trained periodically to ensure accuracy and compliance with NCHS guidelines.
+
+Audits: Routine quality assurance audits are conducted to compare entered data with source documents (e.g., medical records) and identify discrepancies.
+
+Feedback Loops: States receive feedback reports from the NCHS to address anomalies and improve future data collection processes.
+
+By leveraging a combination of manual human curation, electronic systems, and rigorous validation protocols, the dataset maintains a relatively high standard of accuracy and reliability, making it a trusted resource for public health research and policy development.
+
 3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-	- TBD
+
+As part of the National Vital Statistics System (NVSS), the dataset includes all reported live births based on data collected through birth certificates, ensuring extensive coverage of the target population [@cdc2023nvss].
+While the dataset itself is exhaustive and not a sample, researchers often create subsets or samples for specific analyses.
+These subsets can be generated using the following strategies:
+
+  - Deterministic Sampling:
+
+    - Researchers may filter data based on predetermined criteria [@deterministic_sampling2024], such as geographic region (e.g., focusing on births in a specific state) or maternal age groups (e.g., teenage mothers or mothers over 40).
+
+    - This approach ensures that the subset aligns with the research objectives but may not represent the entire population.
+
+  - Probabilistic Sampling:
+
+    - For statistical analyses, probabilistic methods such as stratified random sampling can be employed [@tellingstories].
+For example, researchers might stratify the data by race/ethnicity or socioeconomic status to ensure representation across key demographic groups.
+
+    - Sampling weights, often provided in supplementary datasets, can adjust for any oversampling or undersampling to maintain population-level estimates [@tellingstories].
+
+  - Exclusions Due to Missing or Invalid Data:
+
+    - In practice, records with missing, incomplete, or invalid fields may be excluded from specific analyses [@pmc_sampling2013].
+For instance, births with unrecorded gestational ages or birth weights might be omitted in studies focusing on preterm births.
+
 4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- TBD
+
+The data collection process involved healthcare providers, administrative staff, and state-level vital statistics offices [@cdc2023vscp].
+Healthcare providers, including physicians, nurses, and midwives, recorded information during and after delivery, while hospital administrative staff assisted in completing birth certificates.
+State-level vital statistics offices ensured data entry and validation before transmitting the records to the National Center for Health Statistics (NCHS) [@cdc2023vscp].
+These individuals were compensated as part of their professional roles within healthcare and administrative systems, and no additional payments were made specifically for data collection [@cdc2023birthguidelines].
+The NCHS also provided training and feedback to ensure standardized data quality across all contributors [@cdc2023vscp].
+
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- TBD
+
+The data was collected continuously throughout the calendar year of 2023, with records submitted by hospitals, birthing centers, and state vital statistics offices shortly after each birth [@cdc2023vscp].
+The data collection timeframe aligns with the creation timeframe of the instances, as birth certificates are typically completed and validated within days or weeks of the birth event [@cdc2023birthguidelines].
+This real-time collection ensures that the dataset accurately reflects births occurring in 2023 without significant delays or retrospective adjustments.
+
 6.
@@
 If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+
+The National Vital Statistics System (NVSS), where strict privacy and data protection regulations are operated, is where the birth certificate data was collected through [@natality2023].
+While the dataset itself does not require an Institutional Review Board (IRB) review due to its aggregated and de-identified nature, the underlying processes for collecting and managing birth certificate data adhere to ethical standards outlined by the National Center for Health Statistics (NCHS) and state-level public health agencies [@cdc_privacy2024].
+These processes ensure compliance with federal and state laws, such as the Public Health Service Act, to protect individual confidentiality and data integrity [@cdc_privacy2024].
+Additional ethical safeguards include restricted access to identifiable data and routine audits to maintain data security [@cdc_privacy2024].
+
 7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- TBD
+
+The data was not collected directly from individuals but obtained via third parties, specifically through state vital statistics offices.
+These offices collect data from birth certificates completed by healthcare providers and administrative staff at hospitals and birthing centers [@cdc2023vscp].
+The birth certificate information includes both observed data and data reported by the mother, ensuring comprehensive coverage of birth-related variables [@natality2023].
+
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- TBD
+
+Individuals are indirectly notified about the data collection as part of the birth registration process, which is legally mandated in the United States [@cdc_privacy2024].
+Hospitals and healthcare providers inform mothers that birth-related data is recorded on official birth certificates.
+Notices regarding the use of this data for statistical purposes, including anonymization and public health research, are typically included in state guidelines or hospital documentation [@cdc2023vscp].
+
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- TBD
+
+Consent for data collection and use is implied as part of the legal requirement for birth registration in the United States [@cdc_privacy2024].
+Parents or guardians provide the necessary information to healthcare providers, who then complete the birth certificate [@cdc2023vscp].
+Explicit consent for the statistical use of de-identified data is not required, as the data is anonymized and used solely for public health purposes under federal law [@cdc_privacy2024].
+
 10.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- TBD
+
+Because the data is de-identified and collected as part of a legal process, individuals do not have a mechanism to revoke their consent for data collection [@cdc_privacy2024].
+However, stringent privacy protections ensure that the data is used only for statistical and public health purposes [@statcan_trustlaws2024].
+
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
-12. *Any other comments?*
-	- TBD
+
+The National Center for Health Statistics (NCHS) conducts routine data protection impact analyses to assess the risks and benefits of using the dataset [@cdc_privacy2024].
+These analyses ensure that the de-identification process effectively protects individuals' privacy while maintaining the dataset's utility for public health research [@gdpr_dpia2024].
+Outcomes of these reviews have led to stringent privacy policies and restricted access to personally identifiable information [@cdc_privacy2024].
 
@@
 If not, you may skip the remaining questions in this section.*
-	- TBD
+
+Preprocessing and cleaning were performed on the raw birth certificate data before compiling the 2023 Natality Dataset. The following are the steps included [@cdc2023birthguidelines]:
+
+  - Validation and Cleaning:
+
+    - Removal of records with missing or invalid values for critical fields such as birth weight, gestational age, or maternal age.
+
+    - Logical consistency checks to ensure values align with realistic ranges (e.g., birth weights matching gestational age ranges).
+
+  - Standardization:
+
+    - Data fields were standardized based on the guidelines provided by the National Center for Health Statistics (NCHS) to ensure uniformity across states.
+
+    - Categorical variables (e.g., race, method of delivery) were encoded into predefined formats to facilitate consistency in analysis.
+
+  - Derived Variables:
+
+    - Additional fields, such as age groups or preterm birth indicators, were derived from the raw data for easier stratification in research.
+
 2.
@@
 If so, please provide a link or other access point to the "raw" data.*
-	- TBD
+
+The raw data is retained alongside the processed dataset to ensure transparency and allow for future analysis [@natality2023].
+Researchers and authorized personnel can request access to the raw, unprocessed data through the NCHS Vital Statistics Restricted Data Program.
+Details on accessing raw data can be found here: https://www.cdc.gov/nchs/data/data_access_and_resources_booklet_web.pdf
+
 3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- TBD
-4. *Any other comments?*
-	- TBD
+
+The preprocessing and cleaning processes were carried out using state-level Electronic Birth Registration Systems (EBRS) and federal validation tools provided by the NCHS [@natality2023].
+While the exact software used at the state level varies, the guidelines and validation tools employed by the NCHS are publicly documented [@cdc2023vscp].
+More information on these tools can be accessed here: [https://www.cdc.gov/nchs/nvss/births.htm](https://www.cdc.gov/nchs/nvss/births.htm)
 
@@
 1. *Has the dataset been used for any tasks already? If so, please provide a description.*
-	- TBD
+
+The datasets has already been widely used for public health research and reporting.
+For instance, it has been used to analyze trends in preterm births, low birth weight rates, and disparities in maternal health outcomes across demographic groups [@cdc2023evaluation].
+Policymakers and public health agencies have utilized the dataset to monitor key health indicators and evaluate the effectiveness of maternal and infant health programs [@cdc2023evaluation].
+
 2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
-	- TBD
+
+The dataset is commonly referenced in public health reports published by the Centers for Disease Control and Prevention (CDC) and other research institutions.
+A repository of related publications can be accessed through the CDC WONDER system: [https://wonder.cdc.gov/natality.html](https://wonder.cdc.gov/natality.html).
+Additional scholarly articles citing the dataset are available via platforms like PubMed [https://pubmed.ncbi.nlm.nih.gov/](https://pubmed.ncbi.nlm.nih.gov/) and Google Scholar [https://scholar.google.com/](https://scholar.google.com/).
+
 3. *What (other) tasks could the dataset be used for?*
-	- TBD
+
+- Epidemiology: Studying correlations between maternal risk factors (e.g., obesity, smoking) and infant health outcomes.
+
+- Policy Evaluation: Assessing the impact of interventions like prenatal care programs or smoking cessation campaigns.
+
+- Healthcare Resource Allocation: Identifying geographic areas with higher rates of adverse outcomes to guide resource distribution.
+
+- Machine Learning Models: Developing predictive models for maternal and infant health risks.
+
 4.
@@
 Is there anything a dataset consumer could do to mitigate these risks or harms?*
-	- TBD
-5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
-6. *Any other comments?*
-	- TBD
+
+There are some considerations that dataset consumers of 2023 Natality Data should consider [@bmjmedicine_observational_study]:
+
+- Sampling Bias: While the dataset includes all registered births, populations underrepresented in healthcare systems may have incomplete or inaccurate data.
+
+- De-Identification: Data anonymization limits the ability to link records to external datasets, potentially reducing its utility for longitudinal studies.
+
+- Data Cleaning: Removal of records with missing values might unintentionally exclude marginalized populations, introducing bias.
+To mitigate these risks, researchers should document preprocessing steps transparently and consider applying statistical adjustments to address biases.
 
@@
 If so, please provide a description.*
-	- TBD
+
+The dataset is distributed to third parties, including researchers, public health organizations, and policymakers, through the Centers for Disease Control and Prevention (CDC) and its associated systems, such as the CDC WONDER platform [https://wonder.cdc.gov/](https://wonder.cdc.gov/).
+This distribution ensures that the dataset is accessible for public health analysis and academic research.
+
 2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
-	- TBD
+
+The dataset is primarily distributed through the CDC WONDER platform as downloadable files or via an interactive query system.
+Some datasets may also be made available through restricted-use agreements for more detailed analysis.
+While the dataset does not have a specific Digital Object Identifier (DOI), it is assigned unique identifiers through the NVSS [@cdc_data_access2024].
+Access it here: [https://wonder.cdc.gov/](https://wonder.cdc.gov/)
+
 3. *When will the dataset be distributed?*
-	- TBD
+
+The dataset is distributed on an annual basis, typically several months after the close of the calendar year to allow for validation and quality assurance [@cdc_nvss_release2024].
+The 2023 Natality Dataset was made available in mid-2024 [@natality2023], following the standard release schedule.
+
 4.
@@
 If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
-	- TBD
+
+The dataset is distributed under public use terms as defined by the CDC [@cdc_data_access2024].
+Users must adhere to the terms outlined in the CDC's Data Use Agreement, which restricts the misuse of data for reidentification or commercial exploitation [@cdc_data_access2024].
+Detailed terms are available here: [/https://www.cdc.gov/nchs/data/data_access_and_resources_booklet_web.pdf](/https://www.cdc.gov/nchs/data/data_access_and_resources_booklet_web.pdf)
+
 5.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
-	- TBD
+
+No third-party IP-based restrictions apply to the dataset.
+However, access to restricted-use versions of the dataset, containing additional variables or identifiers, requires approval and is subject to stringent use agreements to ensure data confidentiality [@cdc_data_access2024].
+
 6.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
-	- TBD
-7. *Any other comments?*
-	- TBD
+
+No specific export controls apply to the public-use version of the dataset.
+However, restricted-use data may be subject to additional controls to comply with U.S.
+privacy laws, such as the Health Insurance Portability and Accountability Act (HIPAA) [@natality2023].
+These controls ensure the dataset remains compliant with national data privacy regulations [@cdc_data_access2024].
 
@@
 1. *Who will be supporting/hosting/maintaining the dataset?*
-	- TBD
+
+The 2023 Natality Dataset is hosted and maintained by the National Center for Health Statistics (NCHS) as part of the National Vital Statistics System (NVSS) [@natality2023].
+The NCHS ensures the dataset is updated, accurate, and accessible to researchers and public health professionals [@cdc2023natality].
+
 2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
-	- TBD
+
+The NCHS can be contacted through its official support channels.
+For inquiries about the dataset, email: nchsquery@cdc.gov, or visit their website: [https://www.cdc.gov/nchs/nhis/contact.htm](https://www.cdc.gov/nchs/nhis/contact.htm)
+
 3. *Is there an erratum? If so, please provide a link or other access point.*
-	- TBD
+
+Errata are issued as needed to correct errors in previously released datasets.
+Updates and corrections are communicated through the NCHS website: [https://www.cdc.gov/nchs/products/errata.htm](https://www.cdc.gov/nchs/products/errata.htm)
+
 4.
@@
 If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
-	- TBD
+
+The dataset is updated annually to include new data, correct errors, and improve data quality.
+Updates are performed by state vital statistics offices and the NCHS [@cdc2023natality].
+Changes are communicated to users through public notices on the NCHS website [https://www.cdc.gov/nchs/index.html](https://www.cdc.gov/nchs/index.html) and platforms like CDC WONDER [https://wonder.cdc.gov/](https://wonder.cdc.gov/).
+
 5.
@@
 If so, please describe these limits and explain how they will be enforced.*
-	- TBD
+
+As the dataset is anonymized and used for statistical purposes, there are no strict limits on its retention.
+However, personally identifiable information (PII) is removed to ensure compliance with privacy laws [@cdc_privacy2024], such as the Public Health Service Act, and original PII-containing records are securely stored by state authorities [@public_health_service_act].
+
 6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
-	- TBD
+
+Older versions of the dataset remain publicly accessible for historical analysis but are not actively updated.
+Researchers are encouraged to use the latest version for current analyses.
+Notices about dataset updates and versioning are provided on the NCHS website: [https://www.cdc.gov/nchs/nvss/index.htm](https://www.cdc.gov/nchs/nvss/index.htm)
+
 7.
@@
 If so, please provide a description.*
-	- TBD
-8. *Any other comments?*
-	- TBD
+
+There is no formal mechanism for external contributions or augmentation of the dataset, as it is based on standardized birth certificate data collected through the NVSS.
+Researchers interested in collaborative efforts can propose supplemental analyses or request access to restricted-use datasets  [@cdc_privacy2024] by applying through the NCHS Restricted Data Access Program: [https://www.cdc.gov/nchs/dqs/index.html](https://www.cdc.gov/nchs/dqs/index.html)
 

### 2024-12-01 paper/paper.qmd
revise abstract and introduction

@@
 abstract: "How maternal and neonatal drug interventions during labor shape the health status of a newborn remains a key question in perinatal care.
-With the 2023 Natality Data for the United States, we examined the relationship between six labor-related treatments and Apgar5 scores, a standard measure of neonatal health.
+With the 2023 Natality Data for the United States, the relationship between six labor-related treatments and Apgar5 scores, a standard measure of neonatal health, was examined.
 Using a Random Forest model for prediction and a Bayesian Linear Model for inference, we identified usage of steroids and chorioamnionitis as the most impactful, while antibiotics dosage showed minimal influence.
-The analysis shows significant disparities in treatment efficacy, and further suggests the need for optimized corticosteroid formulations or targeted antibiotic protocols."
+The analysis shows disparities in treatment efficacy, and further suggests the need for optimized corticosteroid formulations or targeted antibiotic protocols."
 format: pdf
@@
 
-The health of newborns is strongly influenced by the treatments administered during labor and delivery [@who_operational_guidance].
-These interventions, including steroids, antibiotics, and anesthesia, are intended to address complications and improve neonatal outcomes, but their effectiveness varies widely [@anesthesiology_obstetric_anesthesia].
-This paper uses the 2023 Natality Data for the United States [@rawdata] to analyze the impact of six specific treatments on neonatal health outcomes, measured using the Apgar5 score.
-
-The estimand of interest is the average treatment effect (ATE) of six labor-associated interventions on neonatal health outcomes, as measured by Apgar5 scores.
-Specifically, we aim to quantify how the administration of each treatment changes the Apgar score compared to scenarios where the treatment is not applied [@tellingstories].
-By applying predictive modeling via Random Forest and inferential analysis through Bayesian linear models, we assess not only the magnitude of these effects[@geeksforgeeks_random_forest] but also the underlying uncertainties[@strimmerlab_priors].
-
-The analysis shows that prenatal exposure to steroids significantly improve Apgar scores (@tbl-rfsummary1), underscoring their role as a cornerstone of perinatal care.
-Conversely, antibiotics dosage doesn't show much contribution to the score distribution (@tbl-rfsummary1) and raises concerns about potential risks, including microbiota disruption [@narrow_spectrum_antibiotics].
-Moreover, the analysis proposes significant disparities in the consistent application and effectiveness of these treatments (@fig-rfsumbar).
-
-This study has its significance in that it states the need to improve health outcomes for mothers and their newborns globally.
+Labor and birth is a time of excitement and anticipation, along with uncertainty and anxiety [@canada2023maternity].
+This emotional complexity is often heightened by decisions surrounding medical interventions that directly impact maternal and newborn health [@reproductive2019labor].
+Women frequently report anxiety in medication use especially during labor, driven by concerns about potential side effects [@reproductive2019labor] and insufficient information [@canada2023maternity].
+To address these concerns, this paper analyzes the 2023 Natality Data for the United States [@rawdata], specifically evaluating the relationship between six labor-related treatments and neonatal health outcomes, as measured by the Apgar5 score [@natality2023].
+
+The estimand is the average treatment effect (ATE) of six labor-associated interventions on neonatal health outcomes, quantified by Apgar5 scores [@rawdata].
+Specifically, we aim to estimate how the administration of each treatment changes the Apgar score compared to scenarios where the treatment is not applied [@tellingstories].
+By applying predictive modeling through a Random Forest model and inferential analysis through a Bayesian linear model, we assess not only the magnitude of these effects[@geeksforgeeks_random_forest] but also the underlying uncertainties[@strimmerlab_priors].
+
+The analysis shows that prenatal exposure to steroids significantly improve Apgar scores (@tbl-rfsummary1).
+Conversely, antibiotics dosage doesn't show as much contribution to the score distribution (@tbl-rfsummary1) and rather raises concerns about potential risks, including microbiota disruption [@narrow_spectrum_antibiotics].
+The paper therefore proposes significant disparities in the consistent application and effectiveness of these treatments (@fig-rfsumbar).
+
+This study has its significance in that it states the need to improve health outcomes for mothers and their newborns.
 As recognized by the World Health Organization, a healthy start in life has significant repercussions for a person's health and well-being during infancy, childhood, and adulthood [@who_maternal_health].

### 2024-12-03 paper/paper.qmd
supplement explanations for survey, sampling tech

@@
 ---
-title: "How Drug Usage During Labor and Delivery Affects Infant Health: A Predictive Analysis of Apgar5 Scores"
-subtitle: "Steroid Efficacy and Antibiotic Risks in Labor-Associated Care"
+title: "The Effects of Labor-Associated Drug Interventions on Neonatal Health Outcomes: A Predictive Analysis of Apgar5 Scores"
+subtitle: "Steroids Enhance Neonatal Health, While Antibiotics Show Limited Effectiveness"
 author: 
@@
 
+More detailed information of data measurement including the survey and sampling methods are shown in @sec-datadetails.
+
 ## Outcome variables
@@
 #| label: fig-apgar2
-#| fig-cap: "The distribution of Apgar5 scores across the filtered observation in analysis dataset, which selected the relevant variables and removed the NA values"
+#| fig-cap: "The distribution of Apgar5 scores across the filtered observation in analysis dataset"
 #| echo: false
@@
 
+\newpage
+
 ### Type of Treatments Used During Delivery and Labor
@@
 
+\newpage
+
 ### Induction of labor
 
-@fig-indc1 shows that in the raw data, the "No" category dominates with over 2.4 million observations, while the "Yes" category trails with around 1.2 million.
-After filtering and refining the dataset, @fig-indc2 shows that the gap between the "Yes" and "No" category becomes narrow, with the "No" category with 15,752 and "Yes" category with 6,248.
+@fig-indc-1 shows that in the raw data, the "No" category dominates with over 2.4 million observations, while the "Yes" category trails with around 1.2 million.
+After filtering and refining the dataset, @fig-indc-2 shows that the gap between the "Yes" and "No" category becomes narrow, with the "No" category with 15,752 and "Yes" category with 6,248.
 This kind of balancing process is expected to allow for a more robust evaluation of how labor induction might influence Apgar scores, reducing the risk of the prediction model underestimating or overestimating treatment effects [@isi2024].
 
-\newpage
-
-```{r}
-#| label: fig-indc1
-#| fig-cap: The distribution of observations in the raw dataset based on whether the infant received induction of labor or not
-#| echo: false
-#| warning: false
-#| message: false
+```{r}
+#| label: fig-indc
+#| fig-cap: The distribution of observations based on whether the infant received induction of labor or not
+#| fig-subcap: ["raw data", "analysis data"]
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
 #| fig-width: 5
@@
 
-```
-
-\newpage
-
-```{r}
-#| label: fig-indc2
-#| fig-cap: The distribution of observations in the analysis dataset based on whether the infant received induction of labor or not
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-width: 5
-#| fig-height: 4
-
 # Prepare the data
@@
 
-@fig-augmt1 shows that in the raw data, the "No" category overwhelmingly dominates with 2,850,569 observations compared to 752,302 in the "Yes" category.
-@fig-augmt2 narrow the gap with the "No" category reduced to 18,589 and "Yes" category adjusted to 3,411.
-
-```{r}
-#| label: fig-augmt1
-#| fig-cap: The distribution observations in the raw dataset based on whether the infant received augmentation of labor or not
+@fig-augmt-1 shows that in the raw data, the "No" category overwhelmingly dominates with 2,850,569 observations compared to 752,302 in the "Yes" category.
+@fig-augmt-2 narrow the gap with the "No" category reduced to 18,589 and "Yes" category adjusted to 3,411.
+
+```{r}
+#| label: fig-augmt
+#| fig-cap: The distribution of observations based on whether the infant received augmentation of labor or not
+#| fig-subcap: ["raw data", "analysis data"]
+#| layout-ncol: 2
 #| echo: false
@@
 
-```
-
-\newpage
-
-```{r}
-#| label: fig-augmt2
-#| fig-cap: The distribution observations in the analysis datset based on whether the infant received augmentation of labor or not
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-width: 5
-#| fig-height: 4
-
 # Prepare the data
@@
 
-@fig-ster1 shows that in the raw data, the "No" category heavily dominates with 3,464,273 observations compared to 138,598 in the "Yes" category, indicating severe imbalance.
-@fig-ster2 shows that the gap narrows after filtering and refining the dataset, with the "No" category to 19,662 and the "Yes" category to 2,338.
-
-```{r}
-#| label: fig-ster1
-#| fig-cap: The distribution observations in the raw dataset based on whether the infant received steroid treatments or not
+@fig-ster-1 shows that in the raw data, the "No" category heavily dominates with 3,464,273 observations compared to 138,598 in the "Yes" category, indicating severe imbalance.
+@fig-ster-2 shows that the gap narrows after filtering and refining the dataset, with the "No" category to 19,662 and the "Yes" category to 2,338.
+
+```{r}
+#| label: fig-ster
+#| fig-cap: The distribution of observations based on whether the infant received steroid treatments or not
+#| fig-subcap: ["raw data", "analysis data"]
+#| layout-ncol: 2
 #| echo: false
@@
 
-```
-
-\newpage
-
-```{r}
-#| label: fig-ster2
-#| fig-cap: The distribution observations in the analysis datset based on whether the infant received steroid treatments or not
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-width: 5
-#| fig-height: 4
-
 # Prepare the data
@@
 
-@fig-antb1 shows that in the raw data, the "No" category dominates with 2,701,457 observations, while the "Yes" category has significantly fewer at 901,414.
-After balancing, @fig-antb2 shows that the "No" category is reduced to 15,828 and the "Yes" category is adjusted to 6,172, narrowing the gap.
-
-```{r}
-#| label: fig-antb1
-#| fig-cap: The distribution observations in the raw dataset based on whether the infant received antibiotic treatments or not
+@fig-antb-1 shows that in the raw data, the "No" category dominates with 2,701,457 observations, while the "Yes" category has significantly fewer at 901,414.
+After balancing, @fig-antb-2 shows that the "No" category is reduced to 15,828 and the "Yes" category is adjusted to 6,172, narrowing the gap.
+
+```{r}
+#| label: fig-antb
+#| fig-cap: The distribution of observations based on whether the infant received antibiotic treatments or not
+#| fig-subcap: ["raw data", "analysis data"]
+#| layout-ncol: 2
 #| echo: false
@@
 
-```
-
-\newpage
-
-```{r}
-#| label: fig-antb2
-#| fig-cap: The distribution observations in the analysis datset based on whether the infant received antibiotic treatments or not
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-width: 5
-#| fig-height: 4
-
 # Prepare the data
@@
 
-@fig-chor1 shows that the "No" category dominates with 3,539,254 observations, compared to 63,617 in the "Yes" category.
-@fig-chor2 shows that the extreme disparity between the two categories is addressed after balancing, with the "No" category having 21,213 and the "Yes" category having 787.
-
-```{r}
-#| label: fig-chor1
-#| fig-cap: The distribution observations in the raw dataset based on whether the infant received chorioamnionitis treatments or not
+@fig-chor-1 shows that the "No" category dominates with 3,539,254 observations, compared to 63,617 in the "Yes" category.
+@fig-chor-2 shows that the extreme disparity between the two categories is addressed after balancing, with the "No" category having 21,213 and the "Yes" category having 787.
+
+```{r}
+#| label: fig-chor
+#| fig-cap: The distribution of observations based on whether the infant received chorioamnionitis treatments or not
+#| fig-subcap: ["raw data", "analysis data"]
+#| layout-ncol: 2
 #| echo: false
@@
 
-```
-
-\newpage
-
-```{r}
-#| label: fig-chor2
-#| fig-cap: The distribution observations in the analysis datset based on whether the infant received chorioamnionitis treatments or not
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-width: 5
-#| fig-height: 4
-
 # Prepare the data
@@
 
-@fig-anes1 shows that in the raw data, the "Yes" category dominates with 2,810,461 observations, while the "No" category has 792,410.
-The initial overrepresentation of the "Yes" category is addressed through the balancing process, with @fig-anes2 showing the "Yes" category of 15,387 and the "No" category of 6,613.
-
-```{r}
-#| label: fig-anes1
-#| fig-cap: The distribution observations in the raw dataset based on whether the infant received anesthesia treatments or not
+@fig-anes-1 shows that in the raw data, the "Yes" category dominates with 2,810,461 observations, while the "No" category has 792,410.
+The initial overrepresentation of the "Yes" category is addressed through the balancing process, with @fig-anes-2 showing the "Yes" category of 15,387 and the "No" category of 6,613.
+
+```{r}
+#| label: fig-anes
+#| fig-cap: The distribution of observations based on whether the infant received anesthesia treatments or not
+#| fig-subcap: ["raw data", "analysis data"]
+#| layout-ncol: 2
 #| echo: false
@@
 
-```
-
-\newpage
-
-```{r}
-#| label: fig-anes2
-#| fig-cap: The distribution observations in the analysis datset based on whether the infant received anesthesia treatments or not
-#| echo: false
-#| warning: false
-#| message: false
-#| fig-width: 5
-#| fig-height: 4
-
 # Prepare the data
@@
 
+\newpage
+
 ### Usage of Steroids and Antibiotics
@@
 
+\newpage
+
 ### Usage of Chorioamnionitis and Antibiotics
@@
 
+\newpage
+
 # Model {#sec-model}
 
-The goal of our modelling strategy is twofold.
-Firstly, a Random Forest model was utilized to identify significant predictors of infant health status, as measured by the 5-minute Apgar score.
-Secondly, a Bayesian Linear Model was applied to quantify the effects of medical interventions during labor and delivery on infant health outcomes, incorporating prior knowledge into the analysis framework.
-
-Here we briefly describe the Random Forest and Bayesian analysis models used to investigate these relationships. Background details and diagnostics are included in [Appendix -@sec-model-details].
+The modeling strategy in this study serves two primary goals: identifying significant predictors of neonatal health outcomes and quantifying the effects of labor-related medical interventions on these outcomes, measured by the 5-minute Apgar score.
+This dual approach employs a Random Forest model for prediction and a Bayesian Linear Model for inference.
+Details of the model implementation and diagnostics are provided in [Appendix -@sec-model-details].
 
@@
 
-Define $y_i$ as the APGAR5 score for the $i$th infant.
-Let the binary treatment variables $\text{indc}_i$, $\text{augmt}_i$, $\text{ster}_i$, $\text{antb}_i$, $\text{chor}_i$, and $\text{anes}_i$ represent the use of induction, augmentation, steroids, antibiotics, chorioamnionitis, and anesthesia, respectively, during labor and delivery.
-
-### Random Forest Model
-
-The Random Forest Model is a non-parametic method that predicts $y_i$ using an ensemble of decision trees [@mlnuggets2024]:
+Define $y_i$ as the 5-minute Apgar score for the $i$th infant, a standard measure of neonatal health.
+Let the binary variables $\text{indc}_i$, $\text{augmt}_i$, $\text{ster}_i$, $\text{antb}_i$, $\text{chor}_i$, and $\text{anes}_i$ represent the use of induction, augmentation, steroids, antibiotics, chorioamnionitis, and anesthesia, respectively, during labor and delivery.
+
+### Model Specification
+
+#### Random Forest Model
+
+The Random Forest model is a non-parametic method that predicts $y_i$ using a collection of decision trees [@mlnuggets2024]. The predicted 5-minute Apgar score is modeled as:
 
@@
 
-We employ a grid search for hyperparameter tuning:
+Hyperparameters were selected via grid search:
+
 - $mtry$: Number of variables randomly sampled at each split (values: 2, 3, 4).
-  - $ntree$: Number of trees in the forest, fixed at 500.
-
-The `caret` package[@caret] in R[@citeR] is used to implement the Random Forest model, with cross-validation to minimize overfitting.
-
-### Bayesian Linear Model
+- $ntree$: Number of trees in the forest, fixed at 500 for computational efficiency and robust performance.
+
+Inherent assumptions to this model are the following [@appliedai_random_forest]:
+
+- Independence of Observations: Observations in the dataset are assumed to be independent of each other. 
+- Non-Linear Relationships: Random Forest inherently assumes the presence of non-linear relationships among predictors, making it suitable for complex interactions.
+- Equal Contribution of Predictors: Each predictor is initially treated with equal importance during sampling.
+
+The Random Forest model was implemented using the `caret` package [@caret] in R[@citeR]. To minimize overfitting, 10-fold cross-validation was employed.
+
+#### Bayesian Linear Model
 
@@
 
+where:
+
 $\mu_i = \alpha + \beta_1 \cdot \text{indc}_i + \beta_2 \cdot \text{augmt}_i + \beta_3 \cdot \text{ster}_i + \beta_4 \cdot \text{antb}_i + \beta_5 \cdot \text{chor}_i + \beta_6 \cdot \text{anes}_i$
 
-Priors are specified as:
-
-$\alpha \sim \text{Normal}(5, 2), \quad \beta_j \sim \text{Normal}(0, 2) \quad \text{for } j = 1, \ldots, 6, \quad \sigma \sim \text{Exponential}(1)$
-
-We fit the Bayesian model using the `stan_glm` function in the `rstanarm`[@rstanarm] package, which implements Markov Chian Monte Carlo (MCMC) sampling with 4 chaing and 2000 iterations per chain.
-We use the default priors from `rstanarm`[@rstanarm].
+The priors reflect weakly informative beliefs based on existing literature:
+
+- Intercept: $\alpha \sim \text{Normal}(5, 2)$, aligned with typical Apgar5 scores [@healthline_apgar_score].
+
+- Coefficients: $\beta_j \sim \text{Normal}(0, 2) \quad \text{for } j = 1, \ldots, 6$, reflecting moderate but uncertain effects.
+
+- Standard deviation: $\sigma \sim \text{Exponential}(1)$, ensuring positivity and discouraging extreme variability.
+
+Inherent assumptions to this model are the following [@geeksforgeeks_bayesian_regression]:
+
+- Independence of Observations: Each infant's outcome is assumed to be independent of others.
+
+- Normality of Residuals: The residuals $(y_i - \mu_i)$ are assumed to follow a normal distribution.
+
+- Homoskedasticity: The variance $(\sigma)$ is assumed to be constant across all levels of predictors.
+
+- Linear Relationships: The effects of predictors are assumed to be linear, which might oversimplify complex interactions.
+
+- Priors Representing Beliefs: The priors are assumed to reasonably reflect prior knowledge, ensuring that they do not dominate the posterior estimates.
+
+The model is implemented using the `stan_glm` function in the `rstanarm` package [@rstanarm].
+Markov Chian Monte Carlo (MCMC) sampling was conducted with 4 chains and 2000 iterations per chain.
+Convergence diagnostics were verified, and posterior predictive checks were conducted to assess model fit, as shown in @sec-model-details.
 
@@
 By integrating the Random Forest's predictive accuracy [@geeksforgeeks_random_forest] with the Bayesian Linear Model's capacity for inference and uncertainty quantification [@nguyen2018uncertainty], this dual-model approach offers a comprehensive understanding of the factors influencing infant health outcomes.
+However, generalizability remains a concern for both models, as the results are derived from a specific population in the U.S.
+and may not extend to broader or differing contexts due to data limitations.
+
+### Alternative Models Considered
+
+#### Logistic Regression
+
+Logistic regression, known for its simplicity and interpretability, is suited for binary outcomes [@aws_logistic_regression].
+Its application is limited in this study due to the continuous nature of the Apgar5 score, which cannot be effectively modeled using logistic regression.
+Additionally, logistic regression struggles to capture non-linear relationships and complex interactions among predictors [@aws_logistic_regression], which are critical features of this dataset.
+Consequently, it was excluded in favor of models better equipped to handle continuous outcomes and non-linear effects, such as Random Forest and Bayesian Linear Models.
+
+#### Gradient Boosting Machines (GBM)
+
+Gradient Boosting Machines (GBM) were also considered due to their ability to handle imbalanced datasets through techniques like weighting and boosting [@towardsdatascience_gbm].
+However, GBMs are computationally intensive and require careful hyperparameter tuning to avoid overfitting [@towardsdatascience_gbm].
+In addition, while GBMs excel at prediction, their variable importance measures are less interpretable compared to Random Forest [@geeksforgeeks_gbm_rf], which aligns better with the study's goal of understanding the relative impact of predictors.
+Given these considerations, Random Forest was preferred over GBM due to its computational efficiency, robustness in handling complex interactions, and the ease of interpreting variable importance.
 
@@
 
-## Additional data details
+## Additional data details {#sec-datadetails}
 
@@
 
-Integrating survey methodologies with the 2023 US Natality Data enables researchers to enrich findings by combining administrative records with additional information collected through structured questionnaires.
+Integrating survey methodologies with the 2023 US Natality Data enables researchers to enrich findings by combining administrative records with additional information collected through structured questionnaires [@groves2010total_survey_error].
 This hybrid approach allows for the exploration of causal relationships, such as the effects of prenatal care access on birth outcomes.
-By adhering to rigorous survey design principles, such as pilot testing and the Total Survey Error framework, and employing robust sampling strategies, researchers can ensure their findings are both internally valid and generalizable.
+By adhering to rigorous survey design principles, such as pilot testing and the Total Survey Error framework, and employing robust sampling strategies, researchers can ensure their findings are both internally valid and generalizable [@groves2010total_survey_error].
 These methods ultimately enhance the utility of natality data for informing public health policies and interventions aimed at improving maternal and child health outcomes.
@@
 This approach not only strengthens the internal and external validity [@tellingstories] of the results but also highlights the potential of observational data to inform equitable and effective healthcare interventions.
+
+### Idealized data collection framework
+
+In order to better answer the research question of how medical interventions during labor affect neonatal health, the data collection approach would integrate detailed clinical records, structured surveys, and longitudinal tracking.
+Clinical records should include precise data on medical interventions such as dosage, timing, and indications for treatment, allowing for a detailed analysis of dose-response relationships and temporal effects [@FrontiersPharmacology2020].
+Longitudinal data would extend beyond immediate neonatal outcomes to track developmental milestones, maternal recovery, and long-term child health [@MCHData2009].
+Moreover, structured surveys would gather information on maternal perceptions, socio-economic status, and access to prenatal care, designed with clarity and flow to minimize cognitive burden [@chetty2023annual].
+
+Enhancements to the sampling methodologies would also be made to improve the validity of the collected data.
+Stratified random sampling would ensure representativeness across demographic and geographic groups [@tellingstories], while pilot testing and adherence to the Total Survey Error framework would mitigate biases in question design and sampling [@groves2010total_survey_error].
+Statistical adjustments, such as propensity score matching and weighting, would address selection biases and confounding variables inherent in observational data [@pmc5378668].
+These techniques, coupled with frameworks like the Neyman-Rubin potential outcomes model, would support robust causal inference, even in the absence of randomized controlled trials [@causalconversations2024].
+
+By integrating diverse data sources, employing advanced statistical methods, and maintaining ethical considerations such as informed consent and participant privacy, researchers can capture complex interactions and contextual factors that influence neonatal health outcomes.
+Ultimately, this integrative approach not only addresses existing limitations in natality data but also generates actionable insights to guide clinical practices and shape public health policies aimed at improving maternal and neonatal health outcomes [@do2020quality].
 
@@
 library(shiny)
-
 runApp("../scripts/07-shiny_app.R")

### 2024-12-03 paper/paper.qmd
fix citation errors

@@
 
+\newpage
+
 ## Predictor variables {#sec-predictor}
@@
 
-## Model set-up
+## Model Specification
 
@@
 
-### Model Specification
-
-#### Random Forest Model
+### Random Forest Model
 
@@
 
-#### Bayesian Linear Model
+### Bayesian Linear Model
 
@@
 
-### Model justification
+## Model Justification
 
@@
 #| label: fig-result1
-#| fig-cap: The mean of Apgar5 scores by the number of treatments received during labor and delivery (**Note:** The dark blue lines represent error bars, indicating the variability or uncertainty around the mean Apgar5 scores for each number of treatments.)
+#| fig-cap: Average Apgar5 scores by the number of treatments received during labor and delivery (**Note:** The dark blue lines represent error bars, indicating the variability or uncertainty around the mean Apgar5 scores for each number of treatments.)
 #| echo: false
@@
 #| label: fig-indcscore
-#| fig-cap: Apgar5 score distribution for the usage of induction labor (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
+#| fig-cap: Apgar5 score distribution for infants who received induction of labor
 #| echo: false
@@
 #| label: fig-augmtscore
-#| fig-cap: Apgar5 score distribution for the usage of augmentation labor (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
+#| fig-cap: Apgar5 score distribution for infants who received augmentation of labor
 #| echo: false
@@
 #| label: fig-sterscore
-#| fig-cap: Apgar5 score distribution for the usage of steroids (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
+#| fig-cap: Apgar5 score distribution for infants who received steroids dosage
 #| echo: false
@@
 #| label: fig-antbscore
-#| fig-cap: Apgar5 score distribution for the usage of antibiotics (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
+#| fig-cap: Apgar5 score distribution for infants who received antibiotics dosage
 #| echo: false
@@
 #| label: fig-chorscore
-#| fig-cap: Apgar5 score distribution for the usage of chorioamnionitis (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
+#| fig-cap: Apgar5 score distribution for infants who received chorioamnionitis dosage
 #| echo: false
@@
 #| label: fig-anesscore
-#| fig-cap: Apgar5 score distribution for the usage of anesthesia (**Note:** The red line represents a visual summary of how observation counts change along the score range.)
+#| fig-cap: Apgar5 score distribution for infants who received anesthesia dosage
 #| echo: false
@@
 #| label: fig-rfsumbar
-#| fig-cap: Variable importance of treatments during labor and delivery on infant health, expressed in bar plots
+#| fig-cap: Horizontal bar plot showing the variable importance of treatments during labor and delivery on infant health
 #| warning: false
@@
 #| label: tbl-modelresults
-#| tbl-cap: Bayesian linear models quantifying the effects of medical interventions during labor and delivery on infant health outcomes
+#| tbl-cap: Summary of a bayesian linear model quantifying the effects of medical interventions during labor and delivery on infant health outcomes
 #| warning: false

### 2024-12-03 paper/paper.qmd
supplement details for idealized data collection

@@
 In order to better answer the research question of how medical interventions during labor affect neonatal health, the data collection approach would integrate detailed clinical records, structured surveys, and longitudinal tracking.
-Clinical records should include precise data on medical interventions such as dosage, timing, and indications for treatment, allowing for a detailed analysis of dose-response relationships and temporal effects [@FrontiersPharmacology2020].
-Longitudinal data would extend beyond immediate neonatal outcomes to track developmental milestones, maternal recovery, and long-term child health [@MCHData2009].
-Moreover, structured surveys would gather information on maternal perceptions, socio-economic status, and access to prenatal care, designed with clarity and flow to minimize cognitive burden [@chetty2023annual].
-
-Enhancements to the sampling methodologies would also be made to improve the validity of the collected data.
-Stratified random sampling would ensure representativeness across demographic and geographic groups [@tellingstories], while pilot testing and adherence to the Total Survey Error framework would mitigate biases in question design and sampling [@groves2010total_survey_error].
-Statistical adjustments, such as propensity score matching and weighting, would address selection biases and confounding variables inherent in observational data [@pmc5378668].
-These techniques, coupled with frameworks like the Neyman-Rubin potential outcomes model, would support robust causal inference, even in the absence of randomized controlled trials [@causalconversations2024].
-
-By integrating diverse data sources, employing advanced statistical methods, and maintaining ethical considerations such as informed consent and participant privacy, researchers can capture complex interactions and contextual factors that influence neonatal health outcomes.
-Ultimately, this integrative approach not only addresses existing limitations in natality data but also generates actionable insights to guide clinical practices and shape public health policies aimed at improving maternal and neonatal health outcomes [@do2020quality].
+Assuming a total budget of $1M, resources would be allocated to maximize data quality and ensure robust causal inference.
+
+Clinical records would require $300,000 to include precise details on medical interventions, such as dosage, timing, and indications for treatment, enabling a detailed analysis of dose-response relationships and temporal effects [@FrontiersPharmacology2020].
+This allocation would cover the development of standardized electronic medical records, data cleaning processes, and staff training to ensure consistent and accurate data entry [@ashtoncollegeEMR].
+
+Longitudinal tracking, supported by a $200,000 allocation, would extend beyond immediate neonatal outcomes to monitor developmental milestones, maternal recovery, and long-term child health [@MCHData2009].
+These funds would be utilized for setting up robust follow-up systems, providing incentives to maintain participant engagement, and integrating longitudinal data with clinical records for comprehensive analysis.
+
+In addition, structured surveys, allocated $250,000, would gather information on maternal perceptions, socio-economic status, and access to prenatal care.
+These surveys would be carefully designed with pilot testing and adherence to the Total Survey Error framework to mitigate biases and cognitive burden [@groves2010total_survey_error], ensuring clarity and representativeness across diverse demographic and geographic groups [@chetty2023annual].
+
+$150,000 would be dedicated to advanced analytical techniques, including statistical adjustments like propensity score matching and weighting, to address selection biases and confounding variables inherent in observational data [@pmc5378668].
+Frameworks such as the Neyman-Rubin potential outcomes model would further support robust causal inference, even in the absence of randomized controlled trials [@causalconversations2024].
+
+The remaining $100,000 would ensure ethical compliance through informed consent processes, data anonymization, and administrative oversight [@tellingstories].
+By integrating diverse data sources, employing advanced statistical methods, and maintaining rigorous ethical standards, this approach not only addresses existing limitations in natality data but also generates actionable insights to inform clinical practices and public health policies aimed at improving maternal and neonatal health outcomes [@do2020quality].
 
@@
 We can also note that the credible intervals underscore uncertainty.
+
+\newpage
 
@@
 
+\newpage
+
 ### Model API Predictions
@@
 #### Features
-1. Data Upload and Preview: Users can upload datasets and preview them in an interactive table.
-2. Model Predictions: Predictions from both models are displayed in a table for easy comparison.
-3. Visualizations: Histograms provide insights into the distribution of predictions from each model.
+- Data Upload and Preview: Users can upload datasets and preview them in an interactive table.
+- Model Predictions: Predictions from both models are displayed in a table for easy comparison.
+- Visualizations: Histograms provide insights into the distribution of predictions from each model.
 

### 2024-12-03 paper/paper.qmd
fix citation errors

@@
 
-We use the '2023 Natality Data for the United States' to analyze the relationship between drug treatments administered to infants and their health outcomes.
-Provided by NBER and sourced from the National Center for Health Statistics [@rawdata], this dataset contains 3,605,081 live birth statistics across the U.S., including demographic details, health metrics, and geographical breakdowns by state and country [@natality2023].
+We use the the statistical programming language R [@citeR] to analyze the relationship between maternal drug treatments and neonatal health outcomes in the '2023 Natality Data for the United States' [@rawdata].
+Provided by NBER and sourced from the National Center for Health Statistics, this dataset contains 3,605,081 live birth statistics across the U.S., including demographic details, health metrics, and geographical breakdowns by state and country [@natality2023].
 To represent treatments directly affecting the infant, variables under the category of 'Characteristics of Labor and Delivery' were selected, which capture medical interventions administered to the mother or infant during the critical period of childbirth [@cdc_facility_worksheet_2016].
@@
 
-In performing the analysis, we utilized the statistical programming language R [@citeR] and several R packages.
+In performing the analysis, we utilized several R packages.
 `tidyverse` [@tidyverse] was used for data manipulation, and `ggplot2`[@ggplot2] was used for visualizing results in graphical methods.

### 2024-12-03 paper/paper.qmd
fix typos

@@
 In performing the analysis, we utilized several R packages.
-`tidyverse` [@tidyverse] was used for data manipulation, and `ggplot2`[@ggplot2] was used for visualizing results in graphical methods.
+`tidyverse` [@tidyverse] was used for data manipulation, and `ggplot2` [@ggplot2] was used for visualizing results in graphical methods.
 `randomForest` [@randomForest] and `caret` [@caret] were used for Random Forest modeling while `rstanarm` [@rstanarm] was used for Bayesian linear modeling in the process of generating predictions.
@@
 
-- Endpoint URL: `http://127.0.0.1:3505/predict_csv_local`
+- Endpoint URL: [http://127.0.0.1:3505/predict_csv_local](http://127.0.0.1:3505/predict_csv_local)
 

### 2024-12-03 scripts/07-shiny_app.R
fix typos/ match dates in the preamble of scripts

@@ -1,13 +1,13 @@
 #### Preamble ####
 # Purpose: Shiny web application for visualizing and analyzing predictions from Random Forest and Bayesian Linear Models for infant health status
 # Author: Yunkyung Ko
-# Date: 29 November 2024
+# Date: 3 December 2024
 # Contact: yunkyung.ko@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
-# - The `shint`, `DT` package must be installed and loaded
+# - The `shiny`, `DT` package must be installed and loaded
 # - 05-model_data.R must have been run
-# - The user must provide a dataset with numeric features (e.g., `test_data.csv`)
+# - The user must provide a dataset with numeric features
 # Any other information needed? Make sure you are in the `infant_health` rproj
 
 

### 2025-02-17 toronto_outbreaks/paper/paper.qmd

---
title: "How Toronto Should Manage the Crisis in the Air"
subtitle: "Analysis of Disease Outbreak Patterns Based on Its Type and Settings"
author: 
  - Yunkyung Ko
thanks: "Code and data are available at: https://github.com/koyunkyung/toronto_outbreaks"
date: today
date-format: long
abstract: "Toronto is currently at the crisis of being swallowed up by the cost of dealing with pertussis, an infectious respiratory disease. This paper analyzes disease outbreak data in Toronto healthcare institutions to apprehend its trend and come up with countermeasures. The analysis indicates that respiratory illnesses and long term care home institutions need intensive management strategies. Furthermore, the paper discusses necessary action plans for both the policymakers and healthcare authorities of Toronto."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
toc-title: "Table of Contents"
toc-depth: 2
toc-location: left
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)
library(modelsummary)
library(knitr)
library(kableExtra)

raw_data <- read_csv(here::here("data/raw_data/raw_data.csv"))
data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction

As of September 9 2024, 1,016 infections of pertussis, a contagious respiratory disease also known as whopping cough, were reported in Ontario [@nowtoronto_whooping_cough]. In all of Toronto, the figure is more than triple the five-year, pre-pandemic average of 31 cases for the same period [@publichealth_ontario].

Unexpected disease outbreaks significantly disrupt society by inducing widespread fear, anxiety, and economic losses [@frontiers_outbreaks]. Especially in Canada, where people continue to face some of the longest medical wait times for treatment in the developed world [@torontosun_wait_times], costs are likely to increase exponentially in an instant. Before Toronto overspends the cost of medical care, this paper is intending to analyze the trends of outbreaks reported in Toronto healthcare institutions. The analysis focuses on figuring out how the frequency trends of outbreak differ by its type and occurrence settings.

The paper starts with showing the overall pattern of outbreak occurrence frequency over the time period of 2016 to 2024. Then, it sorts the outbreaks into two different types, respiratory and enteric, to reiterate the analysis of outbreak frequency patterns. Diving deeper into the analysis, outbreak settings would also come into consideration. Finally, discussions about how specific healthcare institutions should reallocate their resources for particular outbreak types would proceed.

# Data

## Data Overview

'Outbreaks in Toronto Healthcare Institutions' dataset [@toronto_outbreaks_dataset], obtained from City of Toronto Open Data Portal, was used for the trend analysis of outbreak occurrence. Published by Toronto Public Health under the requirement to monitor symptoms of infections, this dataset contains reports of suspected and confirmed outbreaks of gastroenteric (e.g., nausea, vomiting, diarrhea, fever) and respiratory (e.g., cough, runny nose, sore throat, fever) illness. An outbreak is defined as a localized increase in the rate of infection or illness, above that which is expected. Please note that this dataset was last updated on September 24, 2024, and contains only the information up to then.

In this paper, R for statistical computing [@R] was used to handle data. In particular, packages like `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `lubridate` [@lubridate] were utilized to analyze data from the City of Toronto [@opendatatoronto]. Data from year 2016 to 2024 were combined, resulting in 5253 observations in the dataset. The selected variables for analysis are the following: 'outbreak_setting', 'type_of_outbreak', 'date_outbreak_began'. The main criteria considered was whether the information could be helpful for Toronto healthcare institutions to devise preemptive measures for outbreak occurrences. More specifically, the reason for selecting 'date_outbreak_began' is because it is considered more important to accurately determine the onset time of disease to immediately prevent the surge of infected population. In order to use long-term data, which is more effective in capturing the complexity and stochastic nature of disease outbreaks [@bmj_outbreak_prediction], the 'year' information was extracted as the variable for analysis. @tbl-part_of_dataset shows an extract of the dataset and @tbl-summary_stats communicates the summary statistics of the analysis data. For more detailed information about the dataset and each variable, refer to Appendix \ref{sec:appendix}.

```{r}
#| label: tbl-part_of_dataset
#| tbl-cap: A dataset of outbreak occurrence in Toronto healthcare institutions
#| echo: false
#| tbl.pos: 'H'

data |>
  slice(1:10) |>
  kable(
    col.names = c("Outbreak Setting", "Type of Outbreak", "Date Outbreak Began", "Year"),
    align = 'c'
  ) |>
  kable_styling(latex_options = "H")

```

\

```{r}
#| label: tbl-summary_stats
#| tbl-cap: Descriptive statistics for toronto healthcare dataset
#| echo: false
#| tbl.pos: 'H'

datasummary_balance(
  formula = ~type_of_outbreak,
  data = data |>
    filter(type_of_outbreak %in% c("Respiratory", "Enteric", "Other")),
  dinm = FALSE,
  notes = "Data Source: Toronto Public Health"
) |>
  kable_styling(latex_options = c("scale_down", "H")) 
```

\newpage

## Results

What is of our primary interest is the occurrence patterns of outbreaks in Toronto. First, we are going to analyze it from a timely perspective. Through @fig-outbreak, which shows the number of outbreak reports by year, we can see the distribution of occurrence frequency over time. The number of outbreaks has increased sharply as of 2022, and since then, the number has been relatively high until recently.

```{r}
#| label: fig-outbreak
#| fig-cap: Total Outbreaks Count in Toronto Healthcare Institutions
#| echo: false
#| fig.pos: 'H'

data |>
  ggplot(mapping = aes(x = year)) +
  geom_bar() +
  theme_minimal() +
  labs(x = "Date", y = "Number of Outbreaks",
       title = "Total Outbreaks Count Per Year") +
  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1))

```

For detailed analysis, the outbreak was largely divided into respiratory and enteric as separated by Toronto public health, and the trend of occurrence over time was re-analyzed. @fig-outbreaks_count_by_types shows that the number of respiratory outbreak cases is always relatively high, and the increase rate is also greater. In particular, a significant rise in the number of respiratory outbreaks is seen in year 2022, which is expected to have contributed substantially to the surge in the total number of outbreak cases. Furthermore, it can be seen that the large increase in the respiratory outbreak does not show an easily decreasing trend.

```{r}
#| label: fig-outbreaks_count_by_types
#| fig-cap: Outbreaks Count in Toronto Healthcare Institutions, by Types of Outbreak
#| echo: false
#| warning: false
#| message: false
#| fig.pos: 'H'

# Summarizing the data to get the number of outbreaks per month and type of outbreak
annual_data <- data %>%
  group_by(year, type_of_outbreak) %>%
  summarise(outbreak_count = n()) %>%
  ungroup()

# Convert the year column to a factor with levels in chronological order
annual_data$year <- factor(annual_data$year, levels = unique(annual_data$year))

# Create the line plot
ggplot(annual_data, aes(x = year, y = outbreak_count, group = type_of_outbreak, color = type_of_outbreak)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c("Respiratory" = "red", "Enteric" = "blue", "Other" = "black")) +
  labs(title = "Outbreaks Count Per Year, by Type",
       x = "Date", y = "Number of Outbreaks", color = "Type of Outbreak") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate the x-axis labels
  scale_x_discrete(drop = FALSE) # Ensure all years are shown on the x-axis

```

Next, we are going to look at the outbreak occurrence trends from a spatial perspective. @fig-setting shows the occurrence locations of all outbreak reports from 2016 to September 2024 in relative percentages. What accounted for the highest percentage with a noticeable difference compared to the others was LCTH, which stands for Long-Term Care Home, a residential facility providing 24-hour care and support for individuals with complex medical needs, typically the elderly or those with chronic conditions (@ontario_ltc).

```{r}
#| label: fig-setting
#| fig-cap: Outbreaks Count by Outbreak Settings
#| echo: false
#| fig.pos: 'H'

outbreak_counts <- data %>%
  count(outbreak_setting) %>%
  mutate(percentage = n / sum(n) * 100)

ggplot(outbreak_counts, aes(x = "", y = n, fill = outbreak_setting)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) + 
  theme_minimal() +
  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
                               "Retirement Home" = "#B3FFC6", 
                               "Hospital-Chronic Care" = "#BA55D3", 
                               "Hospital-Acute Care" = "#DDA0DD",
                               "Hospital-Psychiatric" = "grey",
                               "Shelter" = "brown",
                               "Transitional Care" = "black")) +
  geom_text(aes(label = ifelse(percentage > 2, paste0(round(percentage, 1), "%"), "")),
            position = position_stack(vjust = 0.5),
            size = 3) +  # Adjust the size of the label text
  labs(x = NULL, y = NULL, fill = "Outbreak Setting", 
       title = "Outbreaks Count by Setting") +
  theme(axis.text.x = element_blank(),  
        axis.ticks = element_blank(),  
        panel.grid = element_blank(),
        legend.text = element_text(size = 8),  # Make legend text smaller
        legend.title = element_text(size = 10) # Optionally make the legend title smaller
  )  
```

Finally, to determine whether there is a relationship between the outbreak occurrence pattern over time and the place of occurrence, the data were re-expressed in the following graph.

```{r}
#| label: fig-respiratory_outbreaks_count_by_settings
#| fig-cap: Respiratory Outbreaks Count, by Outbreak Settings
#| echo: false
#| warning: false
#| message: false
#| fig.pos: 'H'

# Filter for 'Respiratory' outbreaks
respiratory_data <- data %>%
  filter(type_of_outbreak == "Respiratory")

# Aggregate data by year and outbreak setting
annual_counts <- respiratory_data %>%
  group_by(year, outbreak_setting) %>%
  summarise(outbreak_count = n(), .groups = 'drop') %>%
  group_by(year) %>%
  mutate(proportion = outbreak_count / sum(outbreak_count) * 100)  # Normalize by year

# Create a stacked bar plot
ggplot(annual_counts, aes(x = year, y = outbreak_count, fill = outbreak_setting)) +
  geom_col(position = "stack") +  # Stack bars based on outbreak settings
  geom_text(aes(label = ifelse(proportion > 5, paste0(round(proportion, 1), "%"), "")), 
            position = position_stack(vjust = 0.5), 
            size = 2) +  # Add percentage labels with smaller font size
  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
                               "Retirement Home" = "#B3FFC6", 
                               "Hospital-Chronic Care" = "#BA55D3", 
                               "Hospital-Acute Care" = "#DDA0DD",
                               "Hospital-Psychiatric" = "grey",
                               "Shelter" = "brown",
                               "Transitional Care" = "black")) +
  labs(title = "Annual Respiratory Outbreaks Count by Setting",
       x = "Date",
       y = "Number of Outbreaks",
       fill = "Outbreak Setting") +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate the x-axis labels
```

@fig-respiratory_outbreaks_count_by_settings shows that the rate of respiratory outbreak occurrence according to its setting remains relatively similar. Something that stands out even if it's a little bit is that with the surge in respiratory outbreak in 2022, the outbreak rate of LTCH did not increase proportionally. Rather the outbreak rate in other places, such as hospitals in the chronic care sector and retirement home, showed an increasing trend. In addition, it can be predicted that in 2024, the rate of LTCH respiratory outbreak occurrence will be significantly higher than the previous three years.

```{r}
#| label: fig-enteric_outbreaks_count_by_settings
#| fig-cap: Enteric Outbreaks Count, by Outbreak Settings
#| echo: false
#| warning: false
#| message: false
#| fig.pos: 'H'

# Filter for 'Enteric' outbreaks
enteric_data <- data %>%
  filter(type_of_outbreak == "Enteric")

# Aggregate data by year and outbreak setting
annual_counts <- enteric_data %>%
  group_by(year, outbreak_setting) %>%
  summarise(outbreak_count = n(), .groups = 'drop') %>%
  group_by(year) %>%
  mutate(proportion = outbreak_count / sum(outbreak_count) * 100)  # Normalize by year

# Create a stacked bar plot
ggplot(annual_counts, aes(x = year, y = outbreak_count, fill = outbreak_setting)) +
  geom_col(position = "stack") +  # Stack bars based on outbreak settings
  geom_text(aes(label = ifelse(proportion > 5, paste0(round(proportion, 1), "%"), "")), 
            position = position_stack(vjust = 0.5), 
            size = 2) +  # Add percentage labels with smaller font size
  scale_fill_manual(values = c("LTCH" = "#B4E1FF", 
                               "Retirement Home" = "#B3FFC6", 
                               "Hospital-Chronic Care" = "#BA55D3", 
                               "Hospital-Acute Care" = "#DDA0DD",
                               "Hospital-Psychiatric" = "grey",
                               "Shelter" = "brown",
                               "Transitional Care" = "black")) +
  labs(title = "Annual Enteric Outbreaks Count by Setting",
       x = "Date",
       y = "Number of Outbreaks",
       fill = "Outbreak Setting") +
  theme_minimal() +
  scale_x_continuous(breaks = seq(min(data$year), max(data$year), by = 1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate the x-axis labels

```

On the other hand, @fig-enteric_outbreaks_count_by_settings shows that regarding enteric outbreaks, occurrence rate in hospitals rather decreased in recent years. In addition, in 2021 and 2022, when the number of enteric outbreak cases was markedly low, it can be seen that there were outbreak reports only in LTCH and retirement home.

# Discussion

Under the purpose of predicting the possibility of a disease outbreak more accurately and specifically, an outbreak pattern analysis was conducted based on the reports of Toronto healthcare institutions. The following items will discuss how the above analysis results can be applied and practically used in the real world.

## Policy Measures to be Established by Toronto to Prevent and Cope with Disease Outbreaks

The timely outbreak trend analysis in Toronto healthcare institutions (@fig-outbreak) implies that the number of outbreaks has surged as of the time of certain incidents in 2022. The factors that might have substantially influenced this kind of pattern are the emergence of new COVID-19 variants and relaxed public health measures [@cihi_impact_covid19_2022]. At this point of time period, Toronto eased certain pandemic-related restrictions, such as the removal of mandatory masking for visitors and caregivers in long-term care homes. This is in line with the results shown in @fig-outbreaks_count_by_types in that it explains the increase in the incidence of respiratory diseases.

In the current situation, we are dealing with the crisis of pertussis infections, which is a typical respiratory disease. As we can see from @fig-respiratory_outbreaks_count_by_settings, a surge in respiratory outbreak occurrence affects every settings of public healthcare institutions. Accordingly, the first step to address this might be to strengthen vaccination campaigns. Pertussis outbreaks are preventable through proper vaccination coverage, and expanding booster doses to not only children but also adults and healthcare workers could substantially reduce the spread of disease [@canada_pertussis_2024]. Also, since lack of PPE in Canada's public healthcare settings has been a key issue in exacerbating the spread of infections [@cfp_ltc_2020], the government should ensure adequate healthcare capacity by increasing funding and staffing, especially in LTCHs and hospitals.

As @fig-outbreaks_count_by_types and @fig-enteric_outbreaks_count_by_settings implies, enteric outbreak occurrence rates are relatively low in the present time period. Therefore, it would be more efficient to focus on securing sufficient resources for preventing respiratory diesases rather than enteric. In addition, during periods when the incidence of enteric outbreak is low, the disease appears intensively only in LTCH and retirement homes. It would be a good idea to concentrate resources dealing with enteric diseases on those sites to achieve effective medical resource allocation.

## Countermeasures in LTCH for Mitigating Outbreak Occurrences

The spatial outbreak trend analysis (@fig-setting) implies that LTCH is where most of the outbreaks in Toronto take place. Considering the nature of long-term care homes, where some of the most vulnerable populations reside in a congregate living environment, the analysis above seems to be a natural result. Moreover, there were major challenges in LTCHs due to shortage of personal protective equipment(PPE) and inadequate staffing [@cihi_impact_covid19_2022]. However, rather than calling it inevitable, we should make efforts to reduce the incidence of outbreak at the location, even a little.

One suggestion is to create small, dedicated "cohorts" of staff assigned to specific groups of residents, rather than having staff move freely between floors [@cfp_ltc_2020]. This would minimize staff interaction across larger groups and prevent outbreaks from sweeping through entire homes. Cohorting has already been proven effective in hospital settings, and is expected to also be beneficial in long-term care facilities.

## Weaknesses and next steps

The study started with a grand intention of devising specific countermeasures for disease outbreaks by the analysis of outbreak trends sorted into time and place criteria. However, the conclusions derived from data analysis results seem to only put more emphasis on maintaining public health awareness at all times. Contrary to my original intention, it does not provide as much effective guidance for specific healthcare institutions in reallocating their resources. In addition, the dataset that I used had insufficient information for year 2024, since data from October 2024 and after was not possibly obtained. This could have affected the results of trend analysis because all the other data for year 2016 to 2023 had occurrence reports for the whole year.

Further analysis using the seasonal information of this dataset might be intriguing. In particular, respiratory disease occur more severely during the change of seasons. Analyzing quarterly disease outbreak trends is expected to lead conclusions on how different quarterly preparations can be made throughout the year. Additionally, given more time and resources, it will be possible to develop a model for predicting the possibility of future infectious diseases.

\newpage

\appendix

# Appendix {#sec:appendix}

## Additional data details

### Other Similar Datasets Considered

Another similar dataset that I had put in consideration when starting this research was the 'Monthly Communicable Disease Surveillance Data' [@toronto_communicable_disease_data_2024]. Although this dataset contains observations of countless different types of diseases, it has only two types of variables, the number of cases and occurrence rates. So, it was determined that in-depth research in the field could be difficult by using this dataset.

### Raw Data Variables Overview

A detailed description of each variable in the original dataset is as follows [@toronto_outbreaks_dataset].

_id: Unique row identifier for Open Data database.
Institution Name: Name of the healthcare institution where the outbreak has been declared.
Institution Address: Street address of the institution.
Outbreak Setting: The type of facility, in terms of the type of healthcare provided.
Type of Outbreak: The mode of transmission classification description of the outbreak.
Causative Agent-1: The first aetiological agent(s) that have been identified from one or more outbreak-related cases, when an agent is confirmed.
Causative Agent-2: The second aetiological agent(s) that have been identified from one or more outbreak-related cases, when an agent is confirmed.
Date Outbreak Began: The date the institution reports the outbreak to TPH.
Date Declared Over: In general, respiratory and enteric outbreaks are declared over if no new cases have been detected in the facility for a period of time equal to atleast one full incubation period plus one full period of communicability for the particular outbreak organism.
Active: Institutional outbreaks remain active until they have been declared over and the public health investigation is closed.

@tbl-variable_stats shows the summary statistics for all variables in the raw data.

```{r}
#| label: tbl-variable_stats
#| tbl-cap: Raw Data Variables Summary Statistics
#| echo: false
#| warning: false
#| message: false
#| tbl.pos: 'H'

# Function to generate refined summary statistics and display them in a long format
generate_refined_summary_long_format <- function(data) {

  # Numeric summary
  numeric_summary <- data %>%
    select(where(is.numeric)) %>%
    summarise(across(everything(),
                     list(
                       count = ~sum(!is.na(.)),
                       mean = ~mean(., na.rm = TRUE),
                       median = ~median(., na.rm = TRUE),
                       sd = ~sd(., na.rm = TRUE),
                       min = ~min(., na.rm = TRUE),
                       max = ~max(., na.rm = TRUE)
                     ),
                     .names = "{.col}_{.fn}")) %>%
    pivot_longer(cols = everything(),
                 names_to = c("variable", ".value"),
                 names_sep = "_")  # Pivot to long format

  # Categorical summary
  categorical_summary <- data %>%
    select(where(is.character)) %>%
    summarise(across(everything(),
                     list(
                       count = ~sum(!is.na(.)),
                       unique = ~n_distinct(.),
                       top = ~names(sort(table(.), decreasing = TRUE)[1]),
                       freq = ~max(table(.), na.rm = TRUE)
                     ),
                     .names = "{.col}_{.fn}")) %>%
    pivot_longer(cols = everything(),
                 names_to = c("variable", ".value"),
                 names_sep = "_")  # Pivot to long format

  # Date summary
  date_summary <- data %>%
    select(where(is.Date)) %>%
    summarise(across(everything(),
                     list(
                       count = ~sum(!is.na(.)),
                       min = ~min(., na.rm = TRUE),
                       max = ~max(., na.rm = TRUE)
                     ),
                     .names = "{.col}_{.fn}")) %>%
    pivot_longer(cols = everything(),
                 names_to = c("variable", ".value"),
                 names_sep = "_")  # Pivot to long format

  # Combine all summaries into one long table
  combined_summary <- bind_rows(numeric_summary, categorical_summary, date_summary) %>%
    slice(-c(1:6))  # Remove the first 6 rows

  # Return the pretty long table
  combined_summary %>%
    kable(caption = "Refined Summary Statistics (Long Format)") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                  full_width = F,
                  font_size = 8)
}

# Generate the summary statistics table after removing the first 6 rows
generate_refined_summary_long_format(raw_data) |>
  kable_styling(latex_options = c("scale_down", "H"), font_size = 8)

```

\newpage

# References

### 2025-02-17 canadian_grocery/paper/paper.qmd

---
title: "Price Comparison of Common Fruits Across Major Vendors"
subtitle: "An Analysis Revealing Loblaws as the Highest Priced Vendor Among Apples, Oranges, Mangoes, and Pears"
author: 
  - Yunkyung Ko
  - Yawen Tan
  - Shanjie Jiao
  - Xizi Sun
thanks: "Code and data are available at: [https://github.com/koyunkyung/canadian_grocery](https://github.com/koyunkyung/canadian_grocery)."
date: today
date-format: long
abstract: "This study examines the price differences of four common fruits—apples, oranges, mangoes, and pears—across three major Canadian grocery vendors: Loblaws, NoFrills, and Voila. By analyzing the average prices of these fruits, we aim to identify price patterns and assess which vendor tends to offer higher or lower prices. Our results show that Loblaws generally has the highest prices, while Voila offers the lowest. This suggests variations in pricing strategies tied to each vendor's market positioning. Additionally, the study explores analytical considerations, including the distinction between correlation and causation, the impact of missing data, and sources of potential bias such as selection, measurement, and confounding biases. These factors are critical in understanding the observed pricing differences and ensuring the robustness of the analysis. This research provides insights for consumers making budget-conscious shopping decisions and highlights methodological challenges in analyzing observational data."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

This study investigates price differences for four common fruits—apples, oranges, mangoes, and pears—across three major vendors: Loblaws, NoFrills, and Voila. By analyzing the average prices for each fruit type at each vendor, we aim to identify price patterns and determine which vendor consistently has the highest prices. This analysis provides insight into each vendor’s pricing strategy and helps consumers make informed decisions based on their preferences and budget. Our findings indicate that Loblaws generally has the highest average prices for these fruits among the vendors studied.
In this paper, we also address several important analytical considerations to strengthen our findings. First, we examine correlation vs. causation in @sec-discussion1 to clarify that observed price differences among vendors may correlate with various factors, but further analysis would be needed to establish causation. Second, in @sec-discussion2, we discuss the impact of missing data due to the fact that different vendors offer a varying selection of fruit varieties, which affects the completeness of the dataset. Finally, in @sec-discussion3 we identify potential sources of bias—such as selection bias, measurement bias, and confounding factors—that may influence our results and interpretations.

# Data {#sec-data}

## Overview

The dataset was analyzed using R [@citeR] and SQL [@sqlite], and use Tidyverse [@citeTidyverse], Knitr [@citeKnitr], ggplot2[@citeGgplot2] and Dplyr [@citeDplyr]. Data are extract from Hammer Dataset [@hammerdataset]. The dataset provides detailed information on the average prices of four types of fruits—apples, mangoes, pears, and oranges—across three vendors: Loblaws, NoFrills, and Voila. It includes columns for the vendor name, the average price, and the fruit type. Each fruit is represented by multiple entries from each vendor, allowing for comprehensive comparisons of price distributions. In addition to the average price, the dataset contains observation counts for each fruit and vendor, which indicate that apples are the most frequently recorded fruit, followed by pears, while mangoes and oranges have fewer observations, possibly reflecting factors like consumer demand or seasonality. Vendor representation is fairly balanced, with Voila having a slight edge in the number of observations compared to Loblaws and NoFrills, supporting reliable comparative analysis. The dataset structure thus enables an analysis of price levels and distributions across different fruits and vendors, with potential insights into market trends and vendor-specific pricing strategies.

```{r}
#| label: tbl-data
#| tbl-cap: First 6 entries of Analysis Dataset
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
apple <- read.csv("../data/analysis_data/avg_price_by_vendor_apple.csv")
mango <- read.csv("../data/analysis_data/avg_price_by_vendor_mango.csv")
pear <- read.csv("../data/analysis_data/avg_price_by_vendor_pear.csv")
orange <- read.csv("../data/analysis_data/avg_price_by_vendor_orange.csv")

vendor_avg_apple <- aggregate(avg_price ~ vendor, data = apple, mean)
vendor_avg_mango <- aggregate(avg_price ~ vendor, data = mango, mean)
vendor_avg_pear <- aggregate(avg_price ~ vendor, data = pear, mean)
vendor_avg_orange <- aggregate(avg_price ~ vendor, data = orange, mean)

vendor_avg_apple$fruit <- "apple"
vendor_avg_mango$fruit <- "mango"
vendor_avg_pear$fruit <- "pear"
vendor_avg_orange$fruit <- "orange"

dataset <- rbind(vendor_avg_apple, vendor_avg_mango, vendor_avg_pear, vendor_avg_orange)
datao <- rbind(apple, mango, pear, orange)

# Create table
# Round the average price to two decimal places
dataset$avg_price <- round(dataset$avg_price, 2)
head_data <- head(dataset,6)
knitr::kable(head_data, format = "simple", col.names = c('Vendor', 'Average Price', 'Fruit'))

```

## Measurement

This dataset, while useful for analyzing price distribution, is a constructed representation of purchasing behavior and may not fully capture the complexities of real-world consumer actions. Since the data does not include actual transaction records or contextual factors influencing purchases (such as seasonality, brand loyalty, or regional demand), it may lack critical elements that drive genuine purchasing decisions. Consequently, insights derived from this dataset should be interpreted with caution, acknowledging that it represents an approximation rather than a comprehensive model of consumer behavior.

## Outcome variables

### Price Levels

@fig-price reveals that the majority of products are priced below $10, indicating a concentration of lower-priced items in this dataset. A smaller number of items fall in the higher price ranges, with few products priced above $20, suggesting that high-priced items are outliers rather than the norm. This distribution suggests a market trend towards affordability or a focus on budget-friendly options for the products included in the analysis.

```{r}
#| label: fig-price
#| fig-cap: Price level distributions across all observations used in the analysis
#| echo: false
#| warning: false
#| message: false

# Plot the distribution of prices with a histogram
ggplot(datao, aes(x = avg_price)) +
  geom_histogram(binwidth = 1, fill = "grey", color = "black") + # Set bin width and colors
  labs(
       x = "Average Price",
       y = "Frequency") +
  theme_minimal()

```

## Predictor variables

### Type of Fruits: The Product that We Specified to Compare the Price Distribution

@fig-fruits shows the frequency of observations for four fruit types: apples, mangoes, oranges, and pears. Apples have the highest count, with around 50 instances, possibly due to popularity or market demand. Pears follow with a moderate count of about 20. Mangoes and oranges have the lowest counts, around 10 each, possibly due to factors like seasonality or regional availability.

```{r}
#| label: fig-fruits
#| fig-cap: The distrubution of different categories of fruits used in the analysis
#| echo: false
#| warning: false
#| message: false

# Count the number of observations in each dataset
counts <- data.frame(
  fruit = c("apple", "mango", "pear", "orange"),
  observation_count = c(nrow(apple), nrow(mango), nrow(pear), nrow(orange))
)

# Plot the observation count for each fruit
ggplot(counts, aes(x = fruit, y = observation_count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(
       x = "Fruit Type",
       y = "Observation Count") +
  theme_minimal()

```

### Type of Vendors

@fig-vendors provides a visualization of the number of observations for each vendor in the dataset, specifically Loblaws, NoFrills, and Voila. The x-axis represents these three vendors, while the y-axis shows the count of observations, likely indicating the number of unique products or price points recorded for each vendor. Each bar is labeled with the exact count: Loblaws has 33 observations, NoFrills has 31, and Voila has the highest at 36. These counts are relatively similar across vendors, suggesting a fairly balanced representation in the data, although Voila has a slight edge in the number of entries. This balanced distribution is beneficial for comparative analysis, as it reduces the risk of one vendor disproportionately influencing results. The slight difference in counts could introduce a minor weight toward Voila in any statistical analysis, but overall, the dataset’s even distribution supports reliable comparisons. Thus, any averages or other metrics calculated by vendor are likely to be comparable across Loblaws, NoFrills, and Voila, minimizing selection bias due to unequal observation counts.

```{r}
#| label: fig-vendors
#| fig-cap: The distribution of vendors used in the analysis
#| echo: false
#| warning: false

# Plot the distribution of observation count by vendor with adjusted bar width and labels
ggplot(datao, aes(x = vendor)) +
  geom_bar(fill = "purple", width = 0.4) + # Narrower bar width
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) + # Add count labels above bars
  labs(
       x = "Vendor",
       y = "Observation Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability
```

# Results

@fig-results illustrates the average prices of four fruits—apples, mangoes, oranges, and pears—across three different vendors: Loblaws, NoFrills, and Voila. Each fruit is represented by a distinct color, as indicated in the legend: red for apples, green for mangoes, blue for oranges, and purple for pears. The bar plot component of the chart displays the average price of each fruit at the respective vendors. Notably, apples are priced highest at Loblaws and lowest at Voila. Mangoes exhibit minimal price variation across the vendors. Oranges show a significant price difference, being most expensive at Loblaws and least expensive at Voila. Similarly, pears are priced lowest at Voila. Superimposed on the bar plot is a line plot that connects the average prices of each fruit across the vendors, highlighting the price trends. The lines for apples, oranges, and pears demonstrate a downward trend from Loblaws to Voila, indicating decreasing prices. In contrast, the line for mangoes remains relatively flat, suggesting consistent pricing among the vendors. In summary, the chart effectively compares the pricing strategies of different vendors for various fruits, revealing that Voila generally offers lower prices for most fruits, while Loblaws tends to have higher prices. These visual trends provide insights into the pricing correlations among the vendors and set the stage for further analysis of potential causal factors influencing these pricing differences.

```{r}
#| label: fig-results
#| fig-cap: Price distribution of different kinds of fruits(apple, mango, orange, pear) by vendors
#| echo: false
#| warning: false
#| message: false

ggplot(dataset, aes(x = vendor, y = avg_price, fill = fruit, group = fruit)) +
  geom_bar(stat = "identity", position = "dodge", aes(color = fruit)) +
  geom_line(aes(color = fruit)) +
  geom_point(aes(color = fruit)) +
  facet_wrap(~fruit, scales = "free_y") +
  labs(x = "Render", y = "Price") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Discussion

## Correlation vs. Causation {#sec-discussion1}

When analyzing the pricing of the same fruit in different supermarkets, it is very important to distinguish Correlation and Causation for understanding the root cause of price difference. We have observed that the prices of Loblaws apples, pears, orange and mango are always higher than those of other supermarkets (such as NoFrills and Voila), and the price of each fruit in Voila is always lower than that in other stores, which shows that there is a correlation between supermarket and fruit pricing. This correlation may be related to the market positioning, operating costs and supply chain management of supermarkets. For example, Loblaws is positioned in the high-end market and can provide better products and services to middle-class customers, so its pricing is higher. Voila mainly serves groups with limited budgets, so the price is lower. In addition, factors such as geographical location, rent and employee salary may lead to differences in operating costs, which in turn may affect pricing strategies. For example, Loblaws is located in the city center, the rent is high, and the salary of employees is relatively high because of the high service level. In supply chain management, for example, the procurement channels of different supermarkets and the efficiency of supply chain may lead to cost differences and affect the final selling price. In addition, to determine whether these factors really lead to price differences, further data and analysis are needed, such as market positioning research, cost structure analysis and supply chain evaluation. Through these in-depth analysis, we can understand the causal relationship of price difference more accurately.

## Missing Data {#sec-discussion2}

In this dataset, missing data occurs because different supermarkets offer varying selections of fruit varieties. For example, some supermarkets may carry a wide range of apple varieties, while others only stock a few. As a result, when analyzing data across all supermarkets, information on certain apple varieties may be incomplete, with specific types available only from certain vendors. Similar patterns may apply to other types of fruit, leading to gaps in the dataset. This missing data reflects differences in product offerings across vendors, and it's essential to be aware of and consider this incompleteness when conducting further analysis.

## Sources of Bias {#sec-discussion3}
In this dataset, we study the price differences and potential changes of 4 fruits between different supermarkets. The price differences are due to the different positioning of each supermarket, Loblaw, NoFrills, and Voila.

### Selection Bias
As mentioned before, selection bias in this dataset mainly comes from the differences in customer groups of different supermarkets. Each supermarket (such as Loblaws, NoFrills, and Voila) has a different market positioning and customer group. Customers who choose to shop at Loblaws usually prefer high-end goods or organic foods, while NoFrills customers may pay more attention to low prices and discounts. This customer preference leads to bias in the price data, that is, the price differences between different supermarkets not only reflect the market price of the product itself, but also are affected by customer selection behavior. Therefore, Loblaws' higher prices do not necessarily mean that all its fruits are priced higher than other supermarkets, but rather reflect that its customers are willing to pay a premium for higher quality or brand.

### Measurement Bias
In this dataset, measurement bias may come from the different ways in which supermarkets record or define the average price (avg_price). For example, Loblaws may be promoting a specific brand or organic fruit, making the recorded average price lower than the actual price, while NoFrills may use a specific price for a single batch, ignoring other higher price ranges. These differences in measurement may cause distortions in price data, making our price comparisons between different supermarkets biased.

### Confounding Bias
In this dataset, confounding bias may come from factors such as fruit type, quality, packaging type, seasonal changes, etc. For example, Loblaws may be more inclined to sell high-quality or organic apples, which are usually priced higher, and this tendency is not shown separately in the data. Therefore, the price difference between supermarkets may not be due to the supermarket itself, but to uncontrolled fruit quality or seasonal factors.

\newpage

# References

### 2025-02-17 infant_health/paper/paper.qmd

--- 2024-12-03 paper/paper.qmd
@@
 title: "The Effects of Labor-Associated Drug Interventions on Neonatal Health Outcomes: A Predictive Analysis of Apgar5 Scores"
-subtitle: "Steroids Enhance Neonatal Health, While Antibiotics Show Limited Effectiveness"
+subtitle: "Steroids enhance neonatal health, while antibiotics show limited effectiveness"
 author: 
@@
 
-By structuring these variables into categorical or numerical forms, the data can be analyzed and used to model the relationship between treatments and the APGAR5 score.
+By structuring these variables into categorical or numerical forms, the data can be analyzed and used to model the relationship between treatments and the Apgar5 score.
 Specifically, drug usage during labor and delivery was encoded as binary variables (@sec-predictor).
@@
 
-@fig-apgar1 shows that the majority of infants in the raw dataset achieve a high APGAR5 score, clustering around 9 and 10.
+@fig-apgar1 shows that many infants in the raw dataset achieve a high Apgar5 score, clustering around 9 and 10.
 Very few observations exist for lower scores, reflecting rare instances of significant distress at birth.
-Even after filtering the dataset by selecting the relevant variables for analysis and removing the NA values, @fig-apgar2 shows that the observations are overly clustered around high APGAR5 scores.
+Even after filtering the dataset by selecting the relevant variables for analysis and removing the NA values, @fig-apgar2 shows that the observations are overly clustered around high Apgar5 scores.
 
@@
 
-All of the predictor variables used in the analysis are classified under the same category, which is the 'Characteristics of labor and delivery' [@natality2023].
+All the predictor variables used in the analysis are classified under the same category, which is the 'Characteristics of labor and delivery' [@natality2023].
 This item, which contains 6 separate checkboxes that the respondent can choose from, allows for the reporting of more than one characteristic and includes a choice of "None of the above" [@cdcFacilityGuide].
@@
 
-@fig-numtreat shows that the majority of observations involve 0 to 2 treatments during labor and delivery, with a steep decline in counts for 3 or more treatments.
+@fig-numtreat shows that many observations involve 0 to 2 treatments during labor and delivery, with a steep decline in counts for 3 or more treatments.
 Most births reported in this dataset occur with minimal medical intervention, and higher number of treatments are relatively rare.
@@
 
-@fig-typetreat shows that the distribution of treatment types varies significantly, with anesthesia being the most commonly recorded intervention, while chorioamnionitis is rare.
+@fig-typetreat shows that the distribution of treatment types varies significantly, with anesthesia being the most recorded intervention, while chorioamnionitis is rare.
 Treatments like anesthesia and antibiotics, with higher observation counts, may have a greater impact on the observed outcomes, whereas less frequent interventions, such as steroids or chorioamnionitis, may require careful consideration to avoid biases due to smaller sample sizes [@selectstatistics2024].
@@
 
-### Clinical chorioamnionitis diagnosed during labor or maternal temperature over 38 degrees celcius (100.4 degrees fahrenheit)
+### Clinical chorioamnionitis diagnosed during labor or maternal temperature over 38 degrees Celcius (100.4 degrees Fahrenheit)
 
@@
 #| label: fig-corr1
-#| fig-cap: Mosaic plot showing the correlation between induction and augmentation of labor (**Note:** the shading intensity (from light green to dark green) represents the standardized residuals of the chi-squared test for the contingency table.)
+#| fig-cap: The correlation between induction and augmentation of labor (**Note:** the shading intensity (from light green to dark green) represents the standardized residuals of the chi-squared test for the contingency table.)
 #| echo: false
@@
 #| label: fig-corr2
-#| fig-cap: Mosaic plot showing the correlation between usage of steroids and antibiotics (**Note:** the shading intensity (from light green to dark green) represents the standardized residuals of the chi-squared test for the contingency table.)
+#| fig-cap: The correlation between usage of steroids and antibiotics (**Note:** the shading intensity (from light green to dark green) represents the standardized residuals of the chi-squared test for the contingency table.)
 #| echo: false
@@
 Conversely, the smaller and lighter areas, especially for "Chorioamnionitis:1" and "Antibiotics:1", suggest that when chorioamnionitis occurs, antibiotics are often used.
-The pattern reflects a likely positive correlation [@friendly1992] between chorioamnionitis and anitbiotics usage.
+The pattern reflects a likely positive correlation [@friendly1992] between chorioamnionitis and antibiotics usage.
 
@@
 #| label: fig-corr3
-#| fig-cap: Mosaic plot showing the correlation between usage of chorioamnionitis and antibiotics (**Note:** the shading intensity (from light green to dark green) represents the standardized residuals of the chi-squared test for the contingency table.)
+#| fig-cap: The correlation between usage of chorioamnionitis and antibiotics (**Note:** the shading intensity (from light green to dark green) represents the standardized residuals of the chi-squared test for the contingency table.)
 #| echo: false
@@
 
-The Random Forest model is a non-parametic method that predicts $y_i$ using a collection of decision trees [@mlnuggets2024]. The predicted 5-minute Apgar score is modeled as:
+The Random Forest model is a non-parametric method that predicts $y_i$ using a collection of decision trees [@mlnuggets2024]. The predicted 5-minute Apgar score is modeled as:
 
@@
 The model is implemented using the `stan_glm` function in the `rstanarm` package [@rstanarm].
-Markov Chian Monte Carlo (MCMC) sampling was conducted with 4 chains and 2000 iterations per chain.
+Markov Chain Monte Carlo (MCMC) sampling was conducted with 4 chains and 2000 iterations per chain.
 Convergence diagnostics were verified, and posterior predictive checks were conducted to assess model fit, as shown in @sec-model-details.
@@
 #| label: fig-result2
-#| fig-cap: Heatmap of the overall correlation between treatments administered during labor/delivery and Apgar5 scores
+#| fig-cap: The overall correlation between treatments administered during labor/delivery and Apgar5 scores
 #| echo: false
@@
 @fig-chorscore shows that the observation count peaks sharply at a score of 1, then stabilizes between scores 2 and 5, followed by a gradual decline from scores 6 through 10.
-This distribution suggests that infants that were diagnosed clinical chorioamnionitis during labor or maternal temperature over 38 degrees celcius tend to achieve lower-to-mid range Apgar5 scores.
+This distribution suggests that infants that were diagnosed clinical chorioamnionitis during labor or maternal temperature over 38 degrees Celcius tend to achieve lower-to-mid range Apgar5 scores.
 
@@
 It is followed by usage of chorioamnionitis (22.44%) and anesthesia (13.99%), which also contributes substantially.
-Meanwhile, the exposure to antibiotics (0.00%) shows no measurable impact in this context, and suggests its irrelevance to the model's predictions.
+Meanwhile, the exposure to antibiotics (0.00%) shows no measurable impact and suggests its irrelevance to the model's predictions.
 
@@
 #| label: fig-rfsumbar
-#| fig-cap: Horizontal bar plot showing the variable importance of treatments during labor and delivery on infant health
+#| fig-cap: Variable importance of treatments during labor and delivery on infant health, expressed in horizontal bar plots
 #| warning: false
@@
 
-## Optimization of anetenatal corticosteroid formulations {#sec-first-point}
+## Optimization of antenatal corticosteroid formulations {#sec-first-point}
 
 @sec-randfor-result reveals that maternal and neonatal treatments during labor, such as steroids and antibiotics, have differential impacts on infant health outcomes.
-Usage of steroids, which showed positive correlations with the Apgar5 score in @fig-sterscore, was actually affirmed their critical role in improving neonatal survival, especially in extremely preterm infants [@nih_prenatal_steroid_treatment].
+Usage of steroids, which showed positive correlations with the Apgar5 score in @fig-sterscore, was affirmed their critical role in improving neonatal survival, especially in extremely preterm infants [@nih_prenatal_steroid_treatment].
 In particular, antenatal corticosteroids stand out as the core intervention for preterm births under the gestation period of 22 weeks, significantly reducing complications like respiratory distress syndrome by accelerating fetal lung development [@nih_prenatal_steroid_treatment].
 
-In accordance to these analyses results, pharmaceutical companies can focus on optimizing steroid formulations for safer and more effective administration during earlier gestations [@nih_prenatal_steroid_treatment].
+In accordance with these analyses results, pharmaceutical companies can focus on optimizing steroid formulations for safer and more effective administration during earlier gestations [@nih_prenatal_steroid_treatment].
 Current formulations of antenatal corticosteroids, like betamethasone, have shown strong efficacy in improving neonatal outcomes but may carry risks when administered too early or in repeated doses [@nichd_prenatal_steroid_treatment].
@@
 
-In order to accurately contextualize the findings of this analysis, several limitations must be acknowledged.
+To accurately contextualize the findings of this analysis, several limitations must be acknowledged.
 First, the observational nature of the study may introduce confounding variables that were not accounted for, potentially biasing the results [@biostatistics_confounding].
@@
 The 2023 Natality Dataset [@rawdata] provides detailed records of maternal and neonatal health outcomes across diverse populations in the United States.
-This obervational dataset aggregates data from hospital records, birth certificates, and surveys completed by healthcare providers [@natality2023].
+This observational dataset aggregates data from hospital records, birth certificates, and surveys completed by healthcare providers [@natality2023].
 Its comprehensive scope includes variables like maternal age, gestational age, socio-economic status, and the application of specific labor-related treatments [@cdc_facility_worksheet_2016].
@@
 The 'Infant Health Prediction API' provides predictions for infant health status based on maternal and delivery-related factors.
-By accepting structured datasets in CSV format, it offers researchers a powerful tool to generate real-time predictions using advanced maching learning models.
+By accepting structured datasets in CSV format, it offers researchers a powerful tool to generate real-time predictions using advanced matching learning models.
 
